### 2016-01-28 AM+draft.tex
28/01 First commit of draft paper

\documentclass{article}

\begin{document}

\title{Making multiple-locus association mapping on a genome-wide scale routine}
\author{Andrew W. George and Joshua Bowden}

\maketitle

\abstract{
In this paper, we present our method that makes multiple locus association mapping routine. It is only a little more complicated than performing a linear mixed model analysis. Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying which marker-trait associations are true. There are no thresholds to be set, no multiple testing issues,  and no parameters to fine tune. We will show that our method is capable of  finding marker-trait associations more accurately, and in many cases, orders of magnitude faster than competing approaches. Our multiple locus method is implemented in a freely available, fully documented and maintained R package called am+.
}

Gwas for many are now the design of choice when interested in elucidating the genetic underpinnings  of complex traits.  Over the past decade, gwas have changed considerably in focus and scope. Early gwas were performed predominantly as case control studies. Their focus was on identifying the casual variants underlying common disorders in humans. Despite many promising findings, reproducibility of these findings proved an issue (refs). With the collection of dense whole genome data challenging, early gwas lacked power (refs).  Today, gwas are performed not just in humans but animals and plants that may have highly structured pedigrees. They can involve hundreds to thousands of individuals with genotypes collected from thousands to hundreds of thousands of snps. Where early gwas were interested in disease traits, modern gwas are also  interested in categorical and quantitative traits.

Over the past decade, there have also been impressive advancements in genome wide association mapping methods. Association mapping methods are statistical techniques for measuring the strength of association between a marker locus and trait. It has long been recognised that spurious marker-trait associations can occur due to population structure. This problem has now been solved (refs). A linear mixed model framework has also been shown to be well suited to analysing data from gwas. Linear mixed models can accommodate multiple sources of variation simultaneously, including population structure and familial relatedness (refs). Fitting lmm's to gwas data though was computationally challenging. This limitation has also been solved via spectral decomposition (ref) and approximation (ref) leading to several very fast lmm implementations (refs). Despite these impressive advancements, a core feature of mainstream  association mapping has remained unchanged. The strength of association between a maker locus and trait is assessed via a single locus model.  That is, the trait and marker data from gwas are analysed for each marker locus separately.

Interestingly, it isn't a lack of computational tools that is restricting the adoption of multiple locus association mapping.  Several  computer programs are available. Glmnet, lmmlasso, and lars are R packages that implement shrinkage methods for regression problems.   All marker-trait associations are fitted simultaneously with the coefficients of loci with spurious association to a trait shrunk to zero. Another R package that was purpose built for the multiple locus analysis of data from gwas is the bigRR package.  It implements generalised ridge regression. Similar to shrinkage methods, all marker trait associations are fitted simultaneously. However, coefficients are not shrunk and  determining their significance  can be challenging. PiMASS is a command line driven program that implements Bayesian variable selection for multiple locus association mapping. The strength of the marker-locus association is measured by a posterior probability. Parameter inferences are obtained via Markov Chain Monte Carlo.  MLMM is an R package .. linear mixed model that is implemented extremely well.

\end{document}

### 2016-01-29 AM+draft.tex
2901 Changed paragraph in intro on advancements in gwa mapping methods

@@
 \abstract{
-In this paper, we present our method that makes multiple locus association mapping routine.
-It is only a little more complicated than performing a linear mixed model analysis.
-Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying which marker-trait associations are true.
-There are no thresholds to be set, no multiple testing issues,  and no parameters to fine tune.
-We will show that our method is capable of  finding marker-trait associations more accurately, and in many cases, orders of magnitude faster than competing approaches.
-Our multiple locus method is implemented in a freely available, fully documented and maintained R package called am+.
+In this paper, we present our method that makes multiple locus association mapping routine. It is only a little more complicated than 
+performing a linear mixed model analysis. Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying
+ which marker-trait associations are true. There are no thresholds to be set, no multiple testing issues,  
+ and no parameters to fine tune. We will show that our method is capable of  finding marker-trait associations more accurately, 
+ and in many cases, orders of magnitude faster than competing approaches. Our multiple locus method is implemented 
+ in a freely available, fully documented and maintained R package called am+.
 }
 
-Gwas for many are now the design of choice when interested in elucidating the genetic underpinnings  of complex traits.
-Over the past decade, gwas have changed considerably in focus and scope.
-Early gwas were performed predominantly as case control studies.
-Their focus was on identifying the casual variants underlying common disorders in humans.
-Despite many promising findings, reproducibility of these findings proved an issue (refs).
-With the collection of dense whole genome data challenging, early gwas lacked power (refs).
+\section{Introduction}
+
+Gwas for many are now the design of choice when interested in elucidating the genetic underpinnings  
+of complex traits.  Over the past decade, gwas have changed considerably in focus and scope. Early gwas 
+were performed predominantly as case control studies. Their focus was on identifying the casual variants 
+underlying common disorders in humans. Despite many promising findings, reproducibility of these findings 
+proved an issue (refs). With the collection of dense whole genome data challenging, early gwas lacked power (refs).  
 Today, gwas are performed not just in humans but animals and plants that may have highly structured pedigrees. 
-They can involve hundreds to thousands of individuals with genotypes collected from thousands to hundreds of thousands of snps.
-Where early gwas were interested in disease traits, modern gwas are also  interested in categorical and quantitative traits.
+They can involve hundreds to thousands of individuals with genotypes collected from thousands to hundreds of
+ thousands of snps. Where early gwas were interested in disease traits, modern gwas are also  interested in 
+ categorical and quantitative traits.
+
+Over the past decade, there have also been impressive advancements in genome wide association mapping methods, the statistical techniques for measuring the strength of association between a marker locus and trait.
+Traditionally, association data were analysed with simple contingency tests or Cochran-armitage  trend tests.
+Today, far more statistically sophisticated methods are available.
+One such method is the  QK approach, a linear mixed model framework for genome wide association mapping (refs).
+Several highly efficient
+computer programs for performing QK-based association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
+XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected on thousands of subjects. However, not all 
+areas of genome-wide association mapping have advanced. It still remains standard practice to analyse genome-wide association data 
+by measuring the strength of association between a marker locus and trait for each locus separately.
 

### 2016-01-29 AM+draft.tex
2901 Finalized version of changed paragraph in Indroduction

@@
 XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected on thousands of subjects. However, not all 
-areas of genome-wide association mapping have advanced. It still remains standard practice to analyse genome-wide association data 
-by measuring the strength of association between a marker locus and trait for each locus separately.
-
-Over the past decade, there have also been impressive advancements in genome wide association mapping methods.
-Association mapping methods are statistical techniques for measuring the strength of association between a marker locus and trait.
-It has long been recognised that spurious marker-trait associations can occur due to population structure.
-This problem has now been solved (refs).
-A linear mixed model framework has also been shown to be well suited to analysing data from gwas.
-Linear mixed models can accommodate multiple sources of variation simultaneously, including population structure and familial relatedness (refs).
-Fitting lmm's to gwas data though was computationally challenging.
-This limitation has also been solved via spectral decomposition (ref) and approximation (ref) leading to several very fast lmm implementations (refs).
-Despite these impressive advancements, a core feature of mainstream  association mapping has remained unchanged.
-The strength of association between a maker locus and trait is assessed via a single locus model.
-That is, the trait and marker data from gwas are analysed for each marker locus separately.
+areas of genome-wide association mapping have advanced.
+It still remains standard practice to analyse genome-wide association data with a single locus model.
+That is,  the strength of association between a marker locus and trait is measured
+for each locus separately.
 

### 2016-02-10 AM+draft.tex
10/02/16 Working on Results:computing implementation

@@
 
+\addtolength{\oddsidemargin}{-2.5cm}
+%\addtolength{\evensidemargin}{-3cm}
+\addtolength{\textwidth}{2.5cm}
 \begin{document}
@@
 \abstract{
-In this paper, we present our method that makes multiple locus association mapping routine. It is only a little more complicated than 
-performing a linear mixed model analysis. Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying
- which marker-trait associations are true. There are no thresholds to be set, no multiple testing issues,  
- and no parameters to fine tune. We will show that our method is capable of  finding marker-trait associations more accurately, 
- and in many cases, orders of magnitude faster than competing approaches. Our multiple locus method is implemented 
+In this paper, we present our method that makes multiple locus association mapping routine. 
+It is only a little more complicated than performing a linear mixed model analysis. 
+Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying
+ which marker-trait associations are true. There are no thresholds to be set, no multiple 
+ testing issues,  and no parameters to fine tune. We will show that our method is capable 
+ of  finding marker-trait associations more accurately,  and in many cases, orders of 
+ magnitude faster than competing approaches. Our multiple locus method is implemented 
  in a freely available, fully documented and maintained R package called am+.
@@
 
-Gwas for many are now the design of choice when interested in elucidating the genetic underpinnings  
-of complex traits.  Over the past decade, gwas have changed considerably in focus and scope. Early gwas 
-were performed predominantly as case control studies. Their focus was on identifying the casual variants 
-underlying common disorders in humans. Despite many promising findings, reproducibility of these findings 
-proved an issue (refs). With the collection of dense whole genome data challenging, early gwas lacked power (refs).  
-Today, gwas are performed not just in humans but animals and plants that may have highly structured pedigrees. 
-They can involve hundreds to thousands of individuals with genotypes collected from thousands to hundreds of
- thousands of snps. Where early gwas were interested in disease traits, modern gwas are also  interested in 
- categorical and quantitative traits.
+Gwas for many are now the design of choice when interested in elucidating the genetic 
+underpinnings  of complex traits.  Over the past decade, gwas have changed 
+considerably in focus and scope. Early gwas 
+were performed predominantly as case control studies. Their focus 
+was on identifying the casual variants underlying common disorders in humans. 
+Despite many promising findings, reproducibility of these findings 
+proved an issue (refs). With the collection of dense whole genome data challenging, 
+early gwas lacked power (refs).  Today, gwas are performed not just in humans but 
+animals and plants that may have highly structured pedigrees. 
+They can involve hundreds to thousands of individuals with genotypes collected 
+from thousands to hundreds of thousands of snps. Where early gwas were interested
+ in disease traits, modern gwas are also  interested in  categorical and quantitative traits.
 
-Over the past decade, there have also been impressive advancements in genome wide association mapping methods, the statistical techniques for measuring the strength of association between a marker locus and trait.
-Traditionally, association data were analysed with simple contingency tests or Cochran-armitage  trend tests.
-Today, far more statistically sophisticated methods are available.
-One such method is the  QK approach, a linear mixed model framework for genome wide association mapping (refs).
-Several highly efficient
-computer programs for performing QK-based association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
-XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected on thousands of subjects. However, not all 
-areas of genome-wide association mapping have advanced.
-It still remains standard practice to analyse genome-wide association data with a single locus model.
-That is,  the strength of association between a marker locus and trait is measured
-for each locus separately.
+Over the past decade, there have also been impressive advancements in genome wide 
+association mapping methods, the statistical techniques for measuring the strength of 
+association between a marker locus and trait. Traditionally, association data were 
+analysed with simple contingency tests or Cochran-armitage  trend tests. Today, far 
+more statistically sophisticated methods are available. One such method is the  QK 
+approach, a linear mixed model framework for genome wide association 
+mapping (refs).  Several highly efficient computer programs for performing QK-based 
+association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
+XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected 
+on thousands of subjects. However, not all areas of genome-wide association 
+mapping have advanced. It still remains standard practice to analyse genome-wide 
+association data with a single locus model. That is,  it is common for the strength of 
+association between a marker locus and trait is measured for each locus separately.
 
-Interestingly, it isn't a lack of computational tools that is restricting the adoption of multiple locus association mapping.
-Several  computer programs are available.
-Glmnet, lmmlasso, and lars are R packages that implement shrinkage methods for regression problems.
-All marker-trait associations are fitted simultaneously with the coefficients of loci with spurious association to a trait shrunk to zero.
-Another R package that was purpose built for the multiple locus analysis of data from gwas is the bigRR package.
-It implements generalised ridge regression.
-Similar to shrinkage methods, all marker trait associations are fitted simultaneously.
-However, coefficients are not shrunk and  determining their significance  can be challenging.
-PiMASS is a command line driven program that implements Bayesian variable selection for multiple locus association mapping.
-The strength of the marker-locus association is measured by a posterior probability.
-Parameter inferences are obtained via Markov Chain Monte Carlo.
-MLMM is an R package ..
-linear mixed model that is implemented extremely well.
+Methods for association mapping with multiple locus models have been developed 
+for some time. Computer implementations of these methods have also been available. 
+Thus, it is not a lack of methodology  or computational tools preventing the uptake of 
+multiple locus association mapping. The reasons for their lack of acceptance are many. 
+Multiple locus association mapping methods are often cased in high level statistical theory, 
+making their true value difficult to appreciate.  Computationally, they can scale poorly with 
+sample size and/or number of genetic markers. Some methods require tuning parameters 
+to be specified. These methods also report their results as coefficients whose significance 
+are unknown. There are multiple locus methods that instead report their results as probabilities 
+but the difficulty here is determining  an appropriate threshold for inferring a statistically 
+significant finding. The computer implementations can also be difficult to run in some cases. 
+All these reasons contribute to making multiple locus association mapping unappealing as 
+an analysis method.
+
+In this paper, we present our new am+ approach for multiple locus association mapping. 
+Our approach is based on linear mixed models and uses the same parameter estimation 
+procedure as EMMA. Through a clever dimension reduction step, our approach scales 
+with number of subjects instead of number of marker loci. We build the multiple locus 
+model iteratively. At each step, only a single linear mixed model is fitted to the data. 
+This model fits all marker loci and their association with the trait simultaneously. There 
+are no parameters to be tuned, no thresholds to be set. Our approach automatically 
+finds the best model. Once our approach is run, the set of statistically significant 
+marker loci in strongest association with the trait is returned as its findings.
+
+Our am+ approach is implemented in a fully documented and maintained R package. 
+We are able to achieve impressive performance by moving the computationally 
+components of our algorithm to Cpp. Here, we make full use of distributed 
+computing, through multiple cpu and multiple glue processing when available. 
+Our R package is easy to use, even for non  R users. It consists of a single function
+ to read in the data and a single function for analysis. We will demonstrate that 
+ our approach returns more accurate results faster and more efficiently than 
+ competing multiple locus association mapping approaches. Our am+ approach 
+ makes multiple locus association mapping of genome wide data routine.
+
+\section{Results}
+
+\subsection{Computer implementations for Multiple Locus Association Mapping}
+
+We have chosen six computer programs/packages whose performance we  compared  to our R package, am+.
+These are .., .., .., .., ..
+We chose these six programs/packages because a.
+they were purpose-built for multiple-locus association mapping or contained general methods from which multiple-locus association mapping techniques had been developed, b.
+they could handle analysis on a genome wide scale, c.
+they were freely available, and d.
+they are state-of-the-art.
+These programs/packages implement multiple locus association mapping within different model frameworks ( ....) and by fitting these models to data in different ways ( ...).
+A summary of the key features of the programs/packages whose performance we evaluated in this paper (including our package) are given in table X.
+
+These programs/packages cast the association mapping problem in different modelling frameworks, different was of fitting the models to the data, and 
+A summary of the key features of the programs/packages whose performance we evaluated in this paper (including our package) are given in table X.
+These programs/packages employ different modelling frameworks (linear, linear mixed, Bayesian, Decision trees) and ways of fitting these models to the association data (lasso, generalised ridge regression, am+, bayesian variable selection, forward stepwise, and random forests).
+The majority of computer implementations were written in R but there was also a package written in python and a stand alone computer program.
+
+Together, these programs/packages represent a good cross section of modelling frameworks (....) and ways of fitting these models (....).
+
+Various modelling frameworks (linear, linear mixed, Bayesian, Decision trees) are employed.
+These models are also fit to the association data in different ways (lasso, generalised ridge regression, am+, bayesian variable selection, forward stepwise, and random forests).
+
+A summary of the key features of the programs/packages whose performance we evaluate in this paper (including our package) is given in table X.
+Three of the packages are based on linear models ( ).
+Three make use of  more sophisticated linear mixed models (.
+).
+The other two are based on Bayesian models (.
+) or decision tree models ( ).
+The models were fitted to the data via lasso which is a statistical technique where parameter estimation and variable selection occur simultaneously, generalised ridge regression which is a a regularisation method,
+
+We compared  the performance of our R package, am+,  to seven other implementations.
+All seven implementations are based on published methods.
+These seven implementations allowed us to compare our approach to what currently is the state-of-the-art in multiple locus association mapping methodology.
+
+The implementations evaluated are as follows.
+Glmnet is a general purpose R package.
+It implements lasso, an estimation technique where parameter estimation and variable selection occur simultaneously, for linear models.
+It's value for multiple locus association mapping has been demonstrated (ref).
+LARS in another general purpose R package implementing methods amenable for  multiple locus association mapping (ref).
+It relies on least angle regression, a form of forward selection, for model selection with linear models.
+BigRR is purpose built R package for the genome wide analysis of association data.
+It is based on linear models where parameter estimation is accomplished with ridge regression, a regularisation and shrinkage technique.
+Lmm-lasso and MLMM are two implementations purpose built for genome wide association mapping.
+Lmm-lasso uses lasso to identify the set of loci in association with the trait.
+MLMM instead implements a forward selection strategy with backward elimination.
+Both are based on linear mixed models.
+Pimass and randfor implement two very different approaches for multiple locus association mapping.
+In pimass, association mapping is performed within a Bayesian framework through Bayesian variable selection.
+In ranfor, assoc mapping is performed in a machine learning framework through random forests.
+See table X for the  key features of the implementations evaluated in this paper.
 

### 2016-02-11 AM+draft.tex
12/02/16 Changes to Results - Computer programs

@@
-\documentclass{article}
-
+\documentclass[12pt]{article}
+\linespread{1.5}
 \addtolength{\oddsidemargin}{-2.5cm}
@@
 
-We have chosen six computer programs/packages whose performance we  compared  to our R package, am+.
-These are .., .., .., .., ..
-We chose these six programs/packages because a.
-they were purpose-built for multiple-locus association mapping or contained general methods from which multiple-locus association mapping techniques had been developed, b.
-they could handle analysis on a genome wide scale, c.
-they were freely available, and d.
-they are state-of-the-art.
-These programs/packages implement multiple locus association mapping within different model frameworks ( ....) and by fitting these models to data in different ways ( ...).
-A summary of the key features of the programs/packages whose performance we evaluated in this paper (including our package) are given in table X.
+We have chosen six computer implementations whose performance we  
+compared  to our R package, am+. These are bigRR, glmnet, , lmm-lasso,
+MLMM,  piMASS, and randfor.  We chose 
+these six implementations because a. they were purpose-built for 
+multiple-locus association mapping or contained general methods from 
+which multiple-locus association mapping techniques had been developed, 
+b. they could handle analysis on a genome wide scale, c.  they were freely 
+available, and d. they are state-of-the-art.
 
-These programs/packages cast the association mapping problem in different modelling frameworks, different was of fitting the models to the data, and 
-A summary of the key features of the programs/packages whose performance we evaluated in this paper (including our package) are given in table X.
-These programs/packages employ different modelling frameworks (linear, linear mixed, Bayesian, Decision trees) and ways of fitting these models to the association data (lasso, generalised ridge regression, am+, bayesian variable selection, forward stepwise, and random forests).
-The majority of computer implementations were written in R but there was also a package written in python and a stand alone computer program.
+A summary of the key 
+features of the programs/packages whose 
+performance we evaluated in this paper are given in table X.
+Even from this small set of 
+packages and computer programs, the diversity of implementation is apparent. 
+BigRR and glmnet are based 
+on linear models, lmm-lasso and MLMM on linear mixed models, piMASS on Bayesian models, 
+and randfor on decision tree models. BigRR fit its models to data with ridge regression, a regularisation technique. 
+Glmnet and lmm-lasso instead rely on lasso where parameter estimation and variable selection occur simultaneously. 
+PiMass fits its models with bayesian variable selection, MLMM through stepwise selection, and randfor with random forests. 
+Some of the implementations evaluated here fit all snp simultaneously where others do not. Although the underlying statistical 
+problem was the same, to find the "best" set of predictions (snps) to explain a quantitative response (trait), the implementations differed greatly.
 
-Together, these programs/packages represent a good cross section of modelling frameworks (....) and ways of fitting these models (....).
+\subsection{Simulation study to evaluate computational performance}
 
-Various modelling frameworks (linear, linear mixed, Bayesian, Decision trees) are employed.
-These models are also fit to the association data in different ways (lasso, generalised ridge regression, am+, bayesian variable selection, forward stepwise, and random forests).
+\subsection{Simulation study to evaluate statistical performance}
 
-A summary of the key features of the programs/packages whose performance we evaluate in this paper (including our package) is given in table X.
-Three of the packages are based on linear models ( ).
-Three make use of  more sophisticated linear mixed models (.
-).
-The other two are based on Bayesian models (.
-) or decision tree models ( ).
-The models were fitted to the data via lasso which is a statistical technique where parameter estimation and variable selection occur simultaneously, generalised ridge regression which is a a regularisation method,
-
-We compared  the performance of our R package, am+,  to seven other implementations.
-All seven implementations are based on published methods.
-These seven implementations allowed us to compare our approach to what currently is the state-of-the-art in multiple locus association mapping methodology.
-
-The implementations evaluated are as follows.
-Glmnet is a general purpose R package.
-It implements lasso, an estimation technique where parameter estimation and variable selection occur simultaneously, for linear models.
-It's value for multiple locus association mapping has been demonstrated (ref).
-LARS in another general purpose R package implementing methods amenable for  multiple locus association mapping (ref).
-It relies on least angle regression, a form of forward selection, for model selection with linear models.
-BigRR is purpose built R package for the genome wide analysis of association data.
-It is based on linear models where parameter estimation is accomplished with ridge regression, a regularisation and shrinkage technique.
-Lmm-lasso and MLMM are two implementations purpose built for genome wide association mapping.
-Lmm-lasso uses lasso to identify the set of loci in association with the trait.
-MLMM instead implements a forward selection strategy with backward elimination.
-Both are based on linear mixed models.
-Pimass and randfor implement two very different approaches for multiple locus association mapping.
-In pimass, association mapping is performed within a Bayesian framework through Bayesian variable selection.
-In ranfor, assoc mapping is performed in a machine learning framework through random forests.
-See table X for the  key features of the implementations evaluated in this paper.
+\subsection{Application to Arabidopsis data}
 

### 2016-03-14 AM+draft.tex
150316 Working on results - sim - performance - elapse time

@@
 \documentclass[12pt]{article}
+
 \linespread{1.5}
@@
 
-\subsection{Computer implementations for Multiple Locus Association Mapping}
+\subsection{Computer programs for   multiple locus association mapping}
 
-We have chosen six computer implementations whose performance we  
-compared  to our R package, am+. These are bigRR, glmnet, , lmm-lasso,
-MLMM,  piMASS, and randfor.  We chose 
-these six implementations because a. they were purpose-built for 
-multiple-locus association mapping or contained general methods from 
-which multiple-locus association mapping techniques had been developed, 
-b. they could handle analysis on a genome wide scale, c.  they were freely 
-available, and d. they are state-of-the-art.
+We compared the performance of our R package, am, to six other computer programs for multiple locus association mapping.
+These programs are ..., ..., ...
+We decided on these six because they implement published methods for multiple locus association mapping.
+are capable of analysing association data on a genome wide scale, and are freely available.
+A summary of each of these programs along with am+ is given in table X.
+In preparing this table, we assumed the purpose of analysing data from a genome wide association study is to identify those snp in true association with a trait.
+From table X, the computer programs vary greatly in methodology, implementation, requirements, and utility.
+For example, some programs have methodology that fits all marker-trait associations simultaneously where others  fit only a subset of marker-trait associations jointly.
+Some programs have tuning parameters where others have none.
+Several cannot handle additional fixed effects such as covariates to account for population structure.
+Also, the majority of programs considered here yield results that require additional computation before a "best" set of snp is realised.
+Only am+  fits all marker trait associations simultaneously, handles different input data, accommodates additional fixed effects, has no parameters to be fine tuned, returns the "best" set of snp in joint association with a trait without additional computation, and is easy to use.
 
-A summary of the key 
-features of the programs/packages whose 
-performance we evaluated in this paper are given in table X.
-Even from this small set of 
-packages and computer programs, the diversity of implementation is apparent. 
-BigRR and glmnet are based 
-on linear models, lmm-lasso and MLMM on linear mixed models, piMASS on Bayesian models, 
-and randfor on decision tree models. BigRR fit its models to data with ridge regression, a regularisation technique. 
-Glmnet and lmm-lasso instead rely on lasso where parameter estimation and variable selection occur simultaneously. 
-PiMass fits its models with bayesian variable selection, MLMM through stepwise selection, and randfor with random forests. 
-Some of the implementations evaluated here fit all snp simultaneously where others do not. Although the underlying statistical 
-problem was the same, to find the "best" set of predictions (snps) to explain a quantitative response (trait), the implementations differed greatly.
+\subsection{Computational performance}
+
+To assess the computational performance of am+, we conducted a simulation study.
+Our performance measures are computing time and memory usage.
+We also compared the performance of am+ to six other computer programs for multiple locus association mapping (table X).
+
+Data were generated under five different scenarios.
+These scenarios are motivated by the type and size of gwas currently being performed.
+The five scenarios are  unsequenced crop plants (UCP, 150 plants and 5k snp), model plants ( 4555 plants and 34544k snp), .....
+Our marker data are from the 1000 Genomes project (Phase 3).
+Quantitative trait data are generated from these data by assigning effect sizes to a randomly selected set of snp loci along with a gaussian deviate for the error.
+The size of the error is  set to yield a heritability of 0.5.
 

### 2016-09-02 AM+draft.tex
02/09/16 Its been 6 months since I looked at this paper. Comitting prior to making major changes to paper

@@
 \subsection{Computational performance}
+We conducted a simulation study to assess  the computational performance of am+.
+For comparison, we assessed six other computer programs (table X) that  also implement multiple locus association mapping.
+The marker data in our simulation study comes from phase 3 of the 1000 genomes project.
+Here, 2315 individuals were sequenced and their genotype information made available.
+We generated trait data by randomly selecting snp, assigning allelic  effects to the selected snp, and aggregating these effects along with random error.
+The number of snp  selected  was sampled from a Poisson distribution with mean 30.
+The effect sizes were sampled from a beta distribution.
+The errors were drawn from a normal distribution with zero mean and whose variance yielded a quantitative trait with a heritability of 0.5.
 
-To assess the computational performance of am+, we conducted a simulation study.
-Our performance measures are computing time and memory usage.
-We also compared the performance of am+ to six other computer programs for multiple locus association mapping (table X).
+We were interested in how study size impacts performance.
+We generated replicates under five different study sizes.
+These were 150 individuals and 4000 snp (150 X 4K), 300 individuals and 50000 snp (300 X 50k), ...
+We based these five study sizes on the size of study being performed in unsequenced crop plants such as wheat, ......, respectively.
+For those study sizes that exceeded the number of genotyped individuals in the 1000 genomes project,  we simulated additional data by forming genotype mosaics from the real snp data.
+We generated 250 replicates for each scenario.
 
-Data were generated under five different scenarios.
-These scenarios are motivated by the type and size of gwas currently being performed.
-The five scenarios are  unsequenced crop plants (UCP, 150 plants and 5k snp), model plants ( 4555 plants and 34544k snp), .....
-Our marker data are from the 1000 Genomes project (Phase 3).
-Quantitative trait data are generated from these data by assigning effect sizes to a randomly selected set of snp loci along with a gaussian deviate for the error.
-The size of the error is  set to yield a heritability of 0.5.
+Performance was measured by recording elapse  time and maximum memory usage.
+Analyses were performed on a high end desk top which was a dell intra quad core processor with 120G of RAM and a Intel gpu.
+Programs were run for a maximum of 100 hours.
+All programs except piMass  made use of distributed computing, either implicitly via multithreaded blas and lapac linear algebra libraries or as in the case of ranfor, explicitly.
+For ranfor, we wrote R code to parallelise the random forest analysis across multiple cpu.
+All scripts for performing these analyses are available from the Csiro-data acces portal.
+
+In figure X we compare the elapse time of  am+ against the other computer programs.
+On the y axis, we have the ratio of each computer program's elapse time to the elapse time of cpu-based am+.
+On the X axis, we have the number of genotypes and study size.
+Both axes are on log scale .
+A positive (negative) log ratio means am+ had a shorter (longer) elapse time  than the other computer program.
+The magnitude (value) of the ratio is (the magnitude of)  how many times shorter or longer am+ is to the other computer program.
+Since performance varies with replicate, for a given study size, a point in figure X is the log median of 250  ratios of elapse times.
+
+From figure X, the superior performance of am+ to the other computer programs is evident. Describe ...
 

### 2016-09-06 AM+draft.tex
06/09 Playing around with the intro (first paragraph). Have near finalised abstract.

@@
 \abstract{
-In this paper, we present our method that makes multiple locus association mapping routine. 
-It is only a little more complicated than performing a linear mixed model analysis. 
-Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying
- which marker-trait associations are true. There are no thresholds to be set, no multiple 
- testing issues,  and no parameters to fine tune. We will show that our method is capable 
- of  finding marker-trait associations more accurately,  and in many cases, orders of 
- magnitude faster than competing approaches. Our multiple locus method is implemented 
- in a freely available, fully documented and maintained R package called am+.
+The methods of choice for genome wide association mapping all share a common shortcoming. 
+They measure the strength of marker-trait associations on a locus by locus basis. 
+More sophisticated multi-locus methods that map multiple marker-trait associations jointly are available. 
+However, these methods are computationally demanding, are not always easy to implement, and can give results that are difficult to interpret.
+The aim of this paper is to present our method that makes the multi-locus analysis of data from genome wide association studies routine.
+It is formulated within a linear mixed model framework.
+We call our method AMplus.
+AMplus produces results faster than competing multi locus methods and often with greater statistical power.
+In terms of speed, AMplus is just a little slower than the fastest single-locus linear mixed model implementations.
+AMplus is freely available as a fully documented R package which we port as a docker binary for platform independent portability.
 }
@@
 
+Over the past decade, gwas have changed 
+considerably in scope and method of analysis. Early gwas were performed as case-control studies and their data analysed with simple
+contingency table tests or logistic regressions.
+Differences in marker allele frequency between cases (affecteds) and controls (unaffecteds) was assumed to be due to a disease variant.
+It was soon realised though that
+unrecognised population stratification had the capacity to cause spurious associations with non-causal genes. This led to a rise in 
+popularity for family-based designs. Family-based designs are robust to population stratrification. 
+Consequently, there has been a myriad of association studies performed with affected sibling pairs, 
+parent-offspring trios, and nuclear families.
+All have an accompanying plethora of parametric (refs) and non-parametric (refs) analysis methods.
+However, as the popularity of GWAS as the study design of choice grew, so did the desire to perform GWAS in general
+populations which may have complex, possibly unknown, structure.
+Here, we could no longer rely on study design to control population stratification.
+There began a push to develop methods which were capable of accounting for the confounding affects of population
+stratification. Early methods include genomic control and structured association. However, it has been development of a linear mixed 
+modelling framework for GWAS that has the greatest flexibility for controlling for popuation structure and familial relatedness in general population studies.
+
+so to did the desire to implement GWAS in species with structured populations butw where the family structure is unknown
+
+ the statistical significance of differences in marker allele frequency determined with 
+contingency table tests or logistic regression. It was soon realised though that unrecognised population heterogeniety 
+had the potential to confound findings. This led to the rise in popularity of family-based designs based on 
+sibling pairs, trios, and nuclear families.  These designs are robust to deleterious affects of population stratification.
+
+ along with a plethora of
+different score and frequency tests being developed.
+
+and 
+and their data analysed either 
+with simple contingency table tests or logistic regression. The deleterious affects of hidden population heterogeniety though on findings was soon recognised. 
+This led to the preponderance of family-based studies because they are robust to potential population heterogeniety.
+
+ Whether performed with trios, nuclear families, 
+sibling pairs, or extended families, family-based studies correctly controlled population stratification.
+
+controlling for spurious associations 
+due to latent population heterogeniety soon became a focus. This led to family based designs becoming popular. These designs
+
+ based on 
+trios,
+
+involving trios, nuclear families, sibling pairs, or ext
+
+were performed predominantly as case control or family based studies. Their focus 
+was on identifying the casual variants underlying common disorders in humans. As the popularity of GWAS grew, 
+they started to be performed in populations of unrelated individuals and even in structured populations with complicated pedigrees. 
+Here, the traits of interest also included quantitative traits. Today, gwas are being performed at the population level, with studies 
+involving thousands of individuals with genotypes collected from hundreds of thousands of snps.
+
 Gwas for many are now the design of choice when interested in elucidating the genetic 
@@
 considerably in focus and scope. Early gwas 
-were performed predominantly as case control studies. Their focus 
+were performed predominantly as case control or family based studies. Their focus 
 was on identifying the casual variants underlying common disorders in humans. 
 Despite many promising findings, reproducibility of these findings 
-proved an issue (refs). With the collection of dense whole genome data challenging, 
-early gwas lacked power (refs).  Today, gwas are performed not just in humans but 
-animals and plants that may have highly structured pedigrees. 
-They can involve hundreds to thousands of individuals with genotypes collected 
-from thousands to hundreds of thousands of snps. Where early gwas were interested
+proved an issue (refs). Early gwas  lacked statistical power because of the 
+challenges of collecting dense whole genome data across a sufficiently large number of samples (refs).
+Today, gwas are being performed in a wide variety of species. These studies can involve thousands
+ of individuals with genotypes collected from thousands to hundreds of thousands of snps. 
+Where early gwas were interested
  in disease traits, modern gwas are also  interested in  categorical and quantitative traits.
@@
 association mapping methods, the statistical techniques for measuring the strength of 
-association between a marker locus and trait. Traditionally, association data were 
+association between a marker locus and trait.
+(Needs more detail - what are some of the early assoc study methods (give a single sentence review)....
+TDT, genomic control, structured association, score tests????
+Traditionally, association data were 
 analysed with simple contingency tests or Cochran-armitage  trend tests. Today, far 
@@
 approach, a linear mixed model framework for genome wide association 
-mapping (refs).  Several highly efficient computer programs for performing QK-based 
+mapping (refs).  The QK approach
+
+Several highly efficient computer programs for performing QK-based 
 association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
-XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected 
-on thousands of subjects. However, not all areas of genome-wide association 
+XXX(ref) make feasible the analysis of data from large GWAS. 
+ However, not all areas of genome-wide association 
 mapping have advanced. It still remains standard practice to analyse genome-wide 

### 2016-09-20 AM+draft.tex
21/09/16  about to replace introduction with ipad intro

--- 2016-09-02 AM+draft.tex
@@
 \abstract{
-In this paper, we present our method that makes multiple locus association mapping routine. 
-It is only a little more complicated than performing a linear mixed model analysis. 
-Yet, we are able to fit all marker-trait associations simultaneously while still clearly identifying
- which marker-trait associations are true. There are no thresholds to be set, no multiple 
- testing issues,  and no parameters to fine tune. We will show that our method is capable 
- of  finding marker-trait associations more accurately,  and in many cases, orders of 
- magnitude faster than competing approaches. Our multiple locus method is implemented 
- in a freely available, fully documented and maintained R package called am+.
+The methods of choice for genome wide association mapping all share a common shortcoming. 
+They measure the strength of marker-trait associations on a locus by locus basis. 
+More sophisticated multi-locus methods that map multiple marker-trait associations jointly are available. 
+However, these methods are computationally demanding, are not always easy to implement, and can give results that are difficult to interpret.
+The aim of this paper is to present our method that makes the multi-locus analysis of data from genome wide association studies routine.
+It is formulated within a linear mixed model framework.
+We call our method AMplus.
+AMplus produces results faster than competing multi locus methods and often with greater statistical power. 
+In terms of speed, AMplus is just a little slower than the fastest single-locus linear mixed model implementations. 
+AMplus is freely available as a fully documented R package which we port as a docker binary for platform independent portability.
 }
@@
 
-Gwas for many are now the design of choice when interested in elucidating the genetic 
-underpinnings  of complex traits.  Over the past decade, gwas have changed 
-considerably in focus and scope. Early gwas 
-were performed predominantly as case control studies. Their focus 
-was on identifying the casual variants underlying common disorders in humans. 
-Despite many promising findings, reproducibility of these findings 
-proved an issue (refs). With the collection of dense whole genome data challenging, 
-early gwas lacked power (refs).  Today, gwas are performed not just in humans but 
-animals and plants that may have highly structured pedigrees. 
-They can involve hundreds to thousands of individuals with genotypes collected 
-from thousands to hundreds of thousands of snps. Where early gwas were interested
- in disease traits, modern gwas are also  interested in  categorical and quantitative traits.
+Over the past decade, g-w association mapping methods have changed considerably. Early mapping methods relied upon 
+simple contingency tables or logistic regression for case-control designs and score tests for family-based designs.
+
+Over the past decade, genome-wide association mapping methods, have changed considerably. Early association mapping 
+methods were no more complicated than contingency table tests or logistic regression. This is because early association 
+studies followed case-control designs. A significant difference in the allele frequency at a locus, between the cases (affecteds) and 
+controls (unaffecteds), was interpreted as the locus being in linkage disequilibrium with a casual gene. 
+It became known though that differences in allele frequency could also arise due to unrecognised population stratification. 
+This prompted a shift towards family-based designs. Family-based designs do not suffer from 
+the confounding affects of population stratification. A myriad of new family-based association mapping methods were developed, 
+the most popular based on score tests (ref for TDT S-TDT Spielman and Ewens 1998, AJHG 62:450-458, FBAT Naird, NM; Horvath, S; Xu, X 2000, genetic epi V19, Suppl 1 S36-S42).
+
+Over the past decade, genome-wide association mapping, the statistical techniques for analysing data from association studies, has advanced considerably.
+Early genome-wide association studies (GWAS) began as case-control designs.
+The premise was that casual genes could be detected via differences in allele frequency between cases (affecteds) and controls (unaffecteds).
+Association mapping methods for these designs were no more sophisticated than contingency table tests or logistic regression.
+Unfortunately, differences
+in allele frequency can also arise due to unrecognised population stratification. Consequently, there was a shift towards family-based 
+designs. Findings from these designs are
+not susceptible to the confounding affects of population stratification. With interest growing in genome-wide association studies, 
+a myriad of different association mapping methods for family-based designs were developed.
 

### 2016-11-01 AM+draft.tex
2/11/2017 amout to begin writing the results section

@@
 
-Over the past decade, g-w association mapping methods have changed considerably. Early mapping methods relied upon 
-simple contingency tables or logistic regression for case-control designs and score tests for family-based designs.
+Over the past decade, Genome wide association studies have changed considerably in both their analysis and design.
+Early gwas followed a case-control design.
+Association mapping methods were no more complicated than contingency table tests or regression.
+These designs though had a tendency to yield spurious findings if there was unrecognised population stratification.
+This prompted a shift towards family-based designs and score tests, such as the tdt test and its variants (refs), became popular.
+Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification.
+This has meant that data can be collected from general populations, even if these populations are highly structured.
+Analysis via sophisticated association mapping methods based on linear mixed models is now almost routine.
 
-Over the past decade, genome-wide association mapping methods, have changed considerably. Early association mapping 
-methods were no more complicated than contingency table tests or logistic regression. This is because early association 
-studies followed case-control designs. A significant difference in the allele frequency at a locus, between the cases (affecteds) and 
-controls (unaffecteds), was interpreted as the locus being in linkage disequilibrium with a casual gene. 
-It became known though that differences in allele frequency could also arise due to unrecognised population stratification. 
-This prompted a shift towards family-based designs. Family-based designs do not suffer from 
-the confounding affects of population stratification. A myriad of new family-based association mapping methods were developed, 
-the most popular based on score tests (ref for TDT S-TDT Spielman and Ewens 1998, AJHG 62:450-458, FBAT Naird, NM; Horvath, S; Xu, X 2000, genetic epi V19, Suppl 1 S36-S42).
+What has not changed over the past decade is that it remains common practice to analyse gwas data on a locus by locus basis.
+This is despite there being several significant problems with analysing the data in this way.
+First, the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait.
+The identification of these regions from these analyses is not always straightforward.
+Gwas results are reported, typically, via Manhattan plots that plot the -log10 of the p value for each locus against the map position of the locus.
+The location of peaks in this plot signify regions of interest.
+Inferring the exact number of regions of interest can be difficult If the peaks are not well separated.
+Second, when multiple statistical tests are performed, the probability of wrongly accepting a significant result (type 1 error) is inflated.
+This is known as the multiple testing problem (refs).
+Many different solutions have been offered (refs).
+Yet, there is still no well accepted way of correcting for multiple testing in the context of genome- wide association mapping.
+Third, many of the traits whose genetic secretes we are trying to discover are complex.
+There will be multiple loci in linkage equilibrium with genes that influence the trait.
+Yet, A locus by locus mapping approach only assesses the evidence for association between a single locus and trait.
 
-Over the past decade, genome-wide association mapping, the statistical techniques for analysing data from association studies, has advanced considerably.
-Early genome-wide association studies (GWAS) began as case-control designs.
-The premise was that casual genes could be detected via differences in allele frequency between cases (affecteds) and controls (unaffecteds).
-Association mapping methods for these designs were no more sophisticated than contingency table tests or logistic regression.
-Unfortunately, differences
-in allele frequency can also arise due to unrecognised population stratification. Consequently, there was a shift towards family-based 
-designs. Findings from these designs are
-not susceptible to the confounding affects of population stratification. With interest growing in genome-wide association studies, 
-a myriad of different association mapping methods for family-based designs were developed.
+It is somewhat surprising then that multiple locus methods haven't attracted more attention.
+It is possible to measure all locus-trait associations jointly with methods based on the regularisation techniques, ridge regression and lasso.
+Here, multiple testing is not an issue.
+These techniques though are computationally demanding.
+Also, the results are not easy to interpret.
+The strength of association is not measured by a p-value but by the size of the regression coefficient for a locus in the model.
+More recently, associations have started to be mapped with random forests (refs).
+Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings (refs).
+An multiple locus method that does show promise is the multi-locus linear mixed model method (ref).
+The best multiple locus model is built with simple variable selection.
+Results are immediately interpretable but here, computation can be a challenge for large data.
 
-Over the past decade, there have also been impressive advancements in genome wide 
-association mapping methods, the statistical techniques for measuring the strength of 
-association between a marker locus and trait.
-(Needs more detail - what are some of the early assoc study methods (give a single sentence review)....
-TDT, genomic control, structured association, score tests????
-Traditionally, association data were 
-analysed with simple contingency tests or Cochran-armitage  trend tests. Today, far 
-more statistically sophisticated methods are available. One such method is the  QK 
-approach, a linear mixed model framework for genome wide association 
-mapping (refs).  The QK approach
-
-Several highly efficient computer programs for performing QK-based 
-association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
-XXX(ref) make feasible the analysis of data from large GWAS. 
- However, not all areas of genome-wide association 
-mapping have advanced. It still remains standard practice to analyse genome-wide 
-association data with a single locus model. That is,  it is common for the strength of 
-association between a marker locus and trait is measured for each locus separately.
-
-Methods for association mapping with multiple locus models have been developed 
-for some time. Computer implementations of these methods have also been available. 
-Thus, it is not a lack of methodology  or computational tools preventing the uptake of 
-multiple locus association mapping. The reasons for their lack of acceptance are many. 
-Multiple locus association mapping methods are often cased in high level statistical theory, 
-making their true value difficult to appreciate.  Computationally, they can scale poorly with 
-sample size and/or number of genetic markers. Some methods require tuning parameters 
-to be specified. These methods also report their results as coefficients whose significance 
-are unknown. There are multiple locus methods that instead report their results as probabilities 
-but the difficulty here is determining  an appropriate threshold for inferring a statistically 
-significant finding. The computer implementations can also be difficult to run in some cases. 
-All these reasons contribute to making multiple locus association mapping unappealing as 
-an analysis method.
-
-In this paper, we present our new am+ approach for multiple locus association mapping. 
-Our approach is based on linear mixed models and uses the same parameter estimation 
-procedure as EMMA. Through a clever dimension reduction step, our approach scales 
-with number of subjects instead of number of marker loci. We build the multiple locus 
-model iteratively. At each step, only a single linear mixed model is fitted to the data. 
-This model fits all marker loci and their association with the trait simultaneously. There 
-are no parameters to be tuned, no thresholds to be set. Our approach automatically 
-finds the best model. Once our approach is run, the set of statistically significant 
-marker loci in strongest association with the trait is returned as its findings.
+Here, we present our new multiple locus method for genome wide association mapping, which we are calling AMplus.
+AMplus combines the strength of regularisation techniques (being able to fit all locus-trait associations jointly), with simple variable selection (having easy to interpret results).
+Our method does not require a significance threshold to be set nor regularisation parameters to be fine tuned.
+Through a clever dimension reduction step, we are able to achieve a computational performance similar to the fastest single locus linear mixed model implementations.
+AMplus is statistically more powerful than single locus association mapping and is as and often more powerful than
+most multiple locus methods. Our aim is to make multiple locus association mapping on a genome wide scale routine. To this end, we have created an R package ...
 
@@
 \section{Results}
+
+\subsection{Time and Memory Performance}
 
@@
 Only am+  fits all marker trait associations simultaneously, handles different input data, accommodates additional fixed effects, has no parameters to be fine tuned, returns the "best" set of snp in joint association with a trait without additional computation, and is easy to use.
+
+\subsection{Run times}
+
+\subsection{Memory}
 

### 2016-11-15 AM+draft.tex
15/11/16 Working on results section - association mapping methods - in train

@@
 
-\abstract{
-The methods of choice for genome wide association mapping all share a common shortcoming. 
-They measure the strength of marker-trait associations on a locus by locus basis. 
-More sophisticated multi-locus methods that map multiple marker-trait associations jointly are available. 
-However, these methods are computationally demanding, are not always easy to implement, and can give results that are difficult to interpret.
-The aim of this paper is to present our method that makes the multi-locus analysis of data from genome wide association studies routine.
+\abstract{Since the earliest of genome-wide association studies, a key shortcoming in how their data have been analysed has persisted.
+The strength of association between a marker locus and trait is measured for each locus separately, on a locus-by-locus basis.
+Multiple-locus methods that map multiple locus-trait associations simultaneously have been available for some time.
+However, they have attracted little attention.
+They can be demanding, computationally, and their results are not always easy to interpret.
+Yet, it is widely accepted that multiple-locus methods are superior, statistically, to locus-by-locus methods.
+Here, we present our method that makes the multiple-locus analysis of data from genome wide association studies routine.
 It is formulated within a linear mixed model framework.
 We call our method AMplus.
-AMplus produces results faster than competing multi locus methods and often with greater statistical power. 
-In terms of speed, AMplus is just a little slower than the fastest single-locus linear mixed model implementations. 
-AMplus is freely available as a fully documented R package which we port as a docker binary for platform independent portability.
+AMplus produces results faster than competing multiple-locus methods and often with greater statistical power.
+Also, it is just a little slower than the fastest single-locus linear mixed model implementations.
+AMplus is freely available as a fully documented R package. 
 }
@@
 These designs though had a tendency to yield spurious findings if there was unrecognised population stratification.
-This prompted a shift towards family-based designs and score tests, such as the tdt test and its variants (refs), became popular.
+This prompted a shift towards family-based designs and score tests, such as the tdt test and its variants (refs).
 Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification.
@@
 AMplus is statistically more powerful than single locus association mapping and is as and often more powerful than
-most multiple locus methods. Our aim is to make multiple locus association mapping on a genome wide scale routine. To this end, we have created an R package ...
-
-Our am+ approach is implemented in a fully documented and maintained R package. 
-We are able to achieve impressive performance by moving the computationally 
-components of our algorithm to Cpp. Here, we make full use of distributed 
-computing, through multiple cpu and multiple glue processing when available. 
-Our R package is easy to use, even for non  R users. It consists of a single function
- to read in the data and a single function for analysis. We will demonstrate that 
- our approach returns more accurate results faster and more efficiently than 
- competing multiple locus association mapping approaches. Our am+ approach 
- makes multiple locus association mapping of genome wide data routine.
+most multiple locus methods.
+Our aim is to make multiple locus association mapping on a genome wide scale routine.
+To this end, we have created a fully documented R package, that is easy to use, even for non R users.
+Our package accepts marker data of different
+formats and can handle data larger than a computer's  memory capacity. It includes detailed error checking and makes heavy use of 
+distributed computing for computation when available.  The purpose of this work was to make multiple-locus association mapping on 
+a genome-wide scale, for large data sets,  practical and we have built AMplus accordingly.
 
@@
 
-\subsection{Time and Memory Performance}
+\subsection{Association Mapping Methods}
 
-\subsection{Computer programs for   multiple locus association mapping}
+We compared AMplus, in terms of computational and statistical performance, against seven other association mapping methods. 
+We chose methods that had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer software.
+Two of the methods are based on single locus (or locus-by-locus) models and five are based on multiple locus models.
+Of the many ways of performing single locus association mapping, we chose GEMMA and FaST-LMM because of their popularity and computational speed.
+For multiple locus association mapping, we chose bigRR, glmnet, LMM-Lasso, MLMM, and r2VIM.
+Each takes a different approach to mapping multiple locus-trait associations jointly.
 
-We compared the performance of our R package, am, to six other computer programs for multiple locus association mapping.
-These programs are ..., ..., ...
-We decided on these six because they implement published methods for multiple locus association mapping.
-are capable of analysing association data on a genome wide scale, and are freely available.
-A summary of each of these programs along with am+ is given in table X.
-In preparing this table, we assumed the purpose of analysing data from a genome wide association study is to identify those snp in true association with a trait.
+In table 1, we compare the association mapping software with respect to  seven features that we found to be important. 
+Only AMplus had all seven features.
+Some of the implementations come with user manuals which detail the format of the input data and how to perform genome-wide analyses.
+Most could handle additional fixed effects, which is important when seeking to account for hidden population structure.
+Few of the association mapping implementations could cope with marker data larger than the memory capacity of a computer.
+Similarly, few returned the best set of marker loci in strongest association with a trait as their result.
+Almost all the software returned results that required further post processing, either by the setting of significance thresholds or further analysis to determine the significance of the findings.
+Few of the association mapping software checked for errors in the input data.
+
+From a practical perspective, how the methods were implemented, in terms of usability, varied greatly.
+In Table 1, we list seven important features for increasing the usability of computer software for association mapping.
+We also identified which software has what features.
+Only AMplus had all seven features. 
+Some of the implementations have user manuals which detail the format of the input data and how to perform genome-wide analyses.
+Most could handle additional fixed effects, which is important when accounting for hidden population structure.
+Few could cope with marker data larger than the memory capacity of a computer.
+Similarly, few returned the best set of marker loci in strongest association with a trait as their result.
+Almost all the software returned results that required further post processing, either by the setting of significance thresholds or further analysis to determine the significance of the findings.
+Few of the association mapping software checked for errors in the input data.
+
+Computer Programs for  Single-locus and Multiple-locus Association Mapping OR Association Mapping Methods}
+We began by comparing the performance of our package, AMplus,  against eight other freely available implementations for association 
+mapping. Two of the implementations we chose are based on single locus (or locus-by-locus) association mapping and the other 
+six are based on multiple-locus association mapping.
+
+computer programs for association mapping.
+In choosing these computer programs, we focused on single- and multiple-locus methods that are based on a linear (mixed) model framework.
+Two of the programs we chose implement single locus (or locus-by-locus) association mapping methods and the other six implement multiple-locus methods.
+Of the many computer programs available for performing single-locus association mapping, we decided on GEMMA and FaST-LMM because of their popularity and computational speed.
+For multiple-locus association mapping, we compared AMplus against bigRR, glmnet, LMM-Lasso, MLMM, piMASS, and random forests.
+These six programs are based on published methods.
+They
+implement different ways of performing association mapping. They can handle data on a genome-wide scale. They are also freely 
+available.
+
+In Table X, we compare,  across the eight computer programs  and AMplus, those features which are important for genome-wide 
+association mapping. In forming this list of features, we assumed that the primary purpose of a genome-wide association study is 
+to identify the set of marker loci in true and strongest association with a trait.
 From table X, the computer programs vary greatly in methodology, implementation, requirements, and utility.
@@
 
-\subsection{Run times}
+\subsection{Computational Performance: run times and memory usage}
 
-\subsection{Memory}
+To asses the computational performance of AMplus against  we conducted a large simulation study.
 
-\subsection{Computational performance}
 We conducted a simulation study to assess  the computational performance of am+.

### 2016-11-16 AM+draft.tex
17/11/16 working on results - performance

@@
 
-In table 1, we compare the association mapping software with respect to  seven features that we found to be important. 
-Only AMplus had all seven features.
-Some of the implementations come with user manuals which detail the format of the input data and how to perform genome-wide analyses.
-Most could handle additional fixed effects, which is important when seeking to account for hidden population structure.
-Few of the association mapping implementations could cope with marker data larger than the memory capacity of a computer.
+From a practical perspective, how the methods were implemented, in terms of usability, varied greatly.
+In Table 1, we list seven important features for ensuring the usability of computer software for association mapping.
+We also identify which software has what features.
+In forming this list of features, we assumed that the primary purpose of a genome-wide association study is 
+to identify the set of marker loci in true and strongest association with a trait.  The level of documentation varied across the implementations but only 
+a few came with user manuals detailing the format of the input data and how to perform analyses. 
+Most implementations could handle additional fixed effects which is important when accounting for hidden population structure. 
+Few though could cope with marker data larger than the memory capacity of the computer.
 Similarly, few returned the best set of marker loci in strongest association with a trait as their result.
-Almost all the software returned results that required further post processing, either by the setting of significance thresholds or further analysis to determine the significance of the findings.
-Few of the association mapping software checked for errors in the input data.
-
-From a practical perspective, how the methods were implemented, in terms of usability, varied greatly.
-In Table 1, we list seven important features for increasing the usability of computer software for association mapping.
-We also identified which software has what features.
-Only AMplus had all seven features. 
-Some of the implementations have user manuals which detail the format of the input data and how to perform genome-wide analyses.
-Most could handle additional fixed effects, which is important when accounting for hidden population structure.
-Few could cope with marker data larger than the memory capacity of a computer.
-Similarly, few returned the best set of marker loci in strongest association with a trait as their result.
-Almost all the software returned results that required further post processing, either by the setting of significance thresholds or further analysis to determine the significance of the findings.
-Few of the association mapping software checked for errors in the input data.
-
-Computer Programs for  Single-locus and Multiple-locus Association Mapping OR Association Mapping Methods}
-We began by comparing the performance of our package, AMplus,  against eight other freely available implementations for association 
-mapping. Two of the implementations we chose are based on single locus (or locus-by-locus) association mapping and the other 
-six are based on multiple-locus association mapping.
-
-computer programs for association mapping.
-In choosing these computer programs, we focused on single- and multiple-locus methods that are based on a linear (mixed) model framework.
-Two of the programs we chose implement single locus (or locus-by-locus) association mapping methods and the other six implement multiple-locus methods.
-Of the many computer programs available for performing single-locus association mapping, we decided on GEMMA and FaST-LMM because of their popularity and computational speed.
-For multiple-locus association mapping, we compared AMplus against bigRR, glmnet, LMM-Lasso, MLMM, piMASS, and random forests.
-These six programs are based on published methods.
-They
-implement different ways of performing association mapping. They can handle data on a genome-wide scale. They are also freely 
-available.
-
-In Table X, we compare,  across the eight computer programs  and AMplus, those features which are important for genome-wide 
-association mapping. In forming this list of features, we assumed that the primary purpose of a genome-wide association study is 
-to identify the set of marker loci in true and strongest association with a trait.
-From table X, the computer programs vary greatly in methodology, implementation, requirements, and utility.
-For example, some programs have methodology that fits all marker-trait associations simultaneously where others  fit only a subset of marker-trait associations jointly.
-Some programs have tuning parameters where others have none.
-Several cannot handle additional fixed effects such as covariates to account for population structure.
-Also, the majority of programs considered here yield results that require additional computation before a "best" set of snp is realised.
-Only am+  fits all marker trait associations simultaneously, handles different input data, accommodates additional fixed effects, has no parameters to be fine tuned, returns the "best" set of snp in joint association with a trait without additional computation, and is easy to use.
+Almost all the software returned results that required further processing, either
+by identifying the best set of marker loci via the setting of significance thresholds or having to do further analysis to calculate the significance of the results.
+Almost none of the association mapping software checked for errors in the input data.
+Only AMplus has all seven features.
 
@@
 
-To asses the computational performance of AMplus against  we conducted a large simulation study.
+To asses the computational performance of AMplus against the above discussed software, we conducted a large simulation study. 
+In this study, we were interested in how study size impacts performance. We generated data under five different scenarios. 
+These are a study of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
+50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
+1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  These scenarios reflect some of the differences in 
+study size of GWAS being performed in plants, animals, and humans.
+
+Under each scenario, we generated 100 replicates of data. A replicate consisted of snp and trait data. 
+The snp data comes from the publically available 1000 genomes project (phase 3).
+The quantitative trait data were generated by randomly selecting a set of snp loci, assigning allelic effects to the selected snp, and aggregating these effects for each individual along with random error.
+The number of snp selected for a replicate is random, having been drawn from a Poisson distribution with mean 30.
+The quantitative trait was generated to have a heritability of 50\%.
+
+Analyses were performed on a high end desktop computer with dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM. 
+All implementations except, Gemma, made use of distributed computing for their linear algebra operations through multithreaded blas and lapac libraries.
+
+In figure ...
 

### 2016-11-16 AM+draft.tex
Revising results - about to start In Figure 1

@@
 
-To asses the computational performance of AMplus against the above discussed software, we conducted a large simulation study. 
-In this study, we were interested in how study size impacts performance. We generated data under five different scenarios. 
-These are a study of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
+To asses the computational performance of AMplus, we conducted a large simulation study. 
+We were interested in  the impact of study size on  performance so we generated data under five different scenarios. 
+These are a GWAS of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
 50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
-1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  These scenarios reflect some of the differences in 
-study size of GWAS being performed in plants, animals, and humans.
-
-Under each scenario, we generated 100 replicates of data. A replicate consisted of snp and trait data. 
-The snp data comes from the publically available 1000 genomes project (phase 3).
-The quantitative trait data were generated by randomly selecting a set of snp loci, assigning allelic effects to the selected snp, and aggregating these effects for each individual along with random error.
-The number of snp selected for a replicate is random, having been drawn from a Poisson distribution with mean 30.
+1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  
+We chose these scenarios to reflect some of the different sized GWAS being performed in animals, plants, and humans. 
+For each scenario, we generated 100 replicates of data. A single replicate consists of snp genotype data and quantitative trait data. 
+We obtained the snp data from the publicly available 1000 genome project (phase 3). The quantitative trait data we generated from the 
+snp data by selecting a set of snp loci, assigning allelic effects to these snp, and aggregating these effects for each individual along with 
+random error. The number of snp selected per replicate follows a Poisson distribution with mean 30.
  The quantitative trait was generated to have a heritability of 50\%.
-
-Analyses were performed on a high end desktop computer with dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM. 
-All implementations except, Gemma, made use of distributed computing for their linear algebra operations through multithreaded blas and lapac libraries.
+Analyses were performed on a high end desktop computer.
+It had dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM.
+All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS and LAPACK libraries.
 

### 2016-11-17 AM+draft.tex
181116 Working on results - performance - figure 1

@@
 \documentclass[12pt]{article}
+\usepackage{graphicx}
 
@@
 It had dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM.
-All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS and LAPACK libraries.
+All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
 
-In figure ...
+\begin{figure}
+\label{fig_runtimes}
+\caption{Run times}
+\includegraphics[width=10cm]{time_memory}
+\end{figure}
 
-We conducted a simulation study to assess  the computational performance of am+.
-For comparison, we assessed six other computer programs (table X) that  also implement multiple locus association mapping.
-The marker data in our simulation study comes from phase 3 of the 1000 genomes project.
-Here, 2315 individuals were sequenced and their genotype information made available.
-We generated trait data by randomly selecting snp, assigning allelic  effects to the selected snp, and aggregating these effects along with random error.
-The number of snp  selected  was sampled from a Poisson distribution with mean 30.
-The effect sizes were sampled from a beta distribution.
-The errors were drawn from a normal distribution with zero mean and whose variance yielded a quantitative trait with a heritability of 0.5.
+The run times of AMplus against the other software programs, for the five scenarios, is shown in figure X. 
+The axes in both plots are on the log scale. In the top plot, the y-axis is a ratio of elapse times. In the bottom plot, 
+the y-axis is absolute elapse times. 
+ In the top plot,  the y axis is the ratio of elapse times of a multiple-locus implementation to 
+AMplus. In the bottom plot, the y-axis is the absolute elapse times of the single 
+a multiple-locus implementation to the elapse time of AMplus.
 
-We were interested in how study size impacts performance.
-We generated replicates under five different study sizes.
-These were 150 individuals and 4000 snp (150 X 4K), 300 individuals and 50000 snp (300 X 50k), ...
-We based these five study sizes on the size of study being performed in unsequenced crop plants such as wheat, ......, respectively.
-For those study sizes that exceeded the number of genotyped individuals in the 1000 genomes project,  we simulated additional data by forming genotype mosaics from the real snp data.
-We generated 250 replicates for each scenario.
+In the bottom plot, on the y axis, we have the absolute elapse times, averaged 
+across the 100 replicates for a scenario.
 
-Performance was measured by recording elapse  time and maximum memory usage.
-Analyses were performed on a high end desk top which was a dell intra quad core processor with 120G of RAM and a Intel gpu.
-Programs were run for a maximum of 100 hours.
-All programs except piMass  made use of distributed computing, either implicitly via multithreaded blas and lapac linear algebra libraries or as in the case of ranfor, explicitly.
-For ranfor, we wrote R code to parallelise the random forest analysis across multiple cpu.
-All scripts for performing these analyses are available from the Csiro-data acces portal.
+In figure X we compare the elapse time of  AMplus against the other software programs for association mapping. 
+In the top plot, we plot
 
-In figure X we compare the elapse time of  am+ against the other computer programs.
+In the top plot, on the y axis, we have the ratio of elapse times of the multiple-locus implementations to the elapse time of AMplus.
+
 On the y axis, we have the ratio of each computer program's elapse time to the elapse time of cpu-based am+.

### 2016-11-18 AM+draft.tex
181116 Working on results - performance - about to start memory but need plot

@@
 \caption{Run times}
-\includegraphics[width=10cm]{time_memory}
+\includegraphics[width=10cm]{time}
 \end{figure}
 
-The run times of AMplus against the other software programs, for the five scenarios, is shown in figure X. 
-The axes in both plots are on the log scale. In the top plot, the y-axis is a ratio of elapse times. In the bottom plot, 
-the y-axis is absolute elapse times. 
- In the top plot,  the y axis is the ratio of elapse times of a multiple-locus implementation to 
-AMplus. In the bottom plot, the y-axis is the absolute elapse times of the single 
-a multiple-locus implementation to the elapse time of AMplus.
+The run times for AMplus against the other software programs, across the five scenarios, is shown in figure \ref{fig_runtimes} To help with interpretibility of the results, in both plots, we have taken the x and y axes to be on a log scale.
+This means a unit change on the x or y axes is equivalent to a change in the order of magnitude.
+In the top plot, a point is the median of the ratios of elapse times of the multiple locus method to AMplus for a given scenario.
+The median is over the 100 replicates.
+Here, since the median of the ratios are all positive on the log scale, it means that AMplus had a shorter run time than than all the other multiple locus methods.
+In fact, in some cases, AMplus was over a hundred times, or over two orders of magnitude, faster.
+Unlike AMplus, the size of data under scenario 10000 X 1.5M was beyond the memory constraints of the other multiple locus implementations.
+This was also the case for LMM-Lasso, bigRR, and glmnet, for scenario 4000 x 1.5M.
 
-In the bottom plot, on the y axis, we have the absolute elapse times, averaged 
-across the 100 replicates for a scenario.
+In the bottom plot of figure \ref{fig_runtimes}, we compare the median run time of AMplus against the median run times of the single locus methods, FaST-LMM and GEMMA. 
+FaST-LMM was run in two ways. It was run where the genetic similarity matrix was built with all the marker data (FaST-LMM$^all$) or with data on every 500th snp (FaST-LMM$^few$). 
+AMplus was also run in two ways.  The default behavior for AMplus is to make use of CPUs for computing. However, AMplus also has the capacity to harness multiple GPUs (AMplus$^GPU$). 
+From the bottom plot, all the implementations have short run times when analysing data from scenario 150x5K. However, for the other scenarios, AMplus and AMplus$GPU$ have 
+significantly shorter run times than GEMMA and FaST-LMM$^all$. FaST-LMM$^few$ has a shorter run time than AMplus and AMplus$^GPU$ but only for scenarios 1500x50K and 
+10000x1.5M.  Furthermore, for scenario 10000x1.5M, the median run time for GEMMA is 4071 minutes, for AMplus is 699 minutes, for AMplus$^GPU$ is 447 minutes, and for 
+FaST-LMM$^few$ is 346 minutes. The very fast single-locus method FaST-LMM$^few$ is only 29\% faster than our multiple-locus method AMplus$^GPU$. It is worth noting though that 
+there is a setup cost to accessing GPU computing, making AMplus$^GPU$ most efficient on the larger data.
 
-In figure X we compare the elapse time of  AMplus against the other software programs for association mapping. 
-In the top plot, we plot
+The memory usage of the software programs is shown in Figure \ref{fig_memory}.
 
-In the top plot, on the y axis, we have the ratio of elapse times of the multiple-locus implementations to the elapse time of AMplus.
+\begin{figure}
+\label{fig_memory}
+\caption{Memory usage}
+\includegraphics[width=10cm]{memory}
+\end{figure}
 
-On the y axis, we have the ratio of each computer program's elapse time to the elapse time of cpu-based am+.
-On the X axis, we have the number of genotypes and study size.
-Both axes are on log scale .
-A positive (negative) log ratio means am+ had a shorter (longer) elapse time  than the other computer program.
-The magnitude (value) of the ratio is (the magnitude of)  how many times shorter or longer am+ is to the other computer program.
-Since performance varies with replicate, for a given study size, a point in figure X is the log median of 250  ratios of elapse times.
-
-From figure X, the superior performance of am+ to the other computer programs is evident. Describe ...
-
-\subsection{Simulation study to evaluate computational performance}
-
-\subsection{Simulation study to evaluate statistical performance}
+\subsection{Statistical Performance: recall and precision}
 

### 2016-11-18 Table1.tex
added table to repos

\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

\begin{table}
\begin{tabular}{lccccccccc}
                                &  \multicolumn{9}{c}{Computer Software for Association Mapping} \\ \cline{2-10}
                                 & \multicolumn{6}{c}{Multiple-locus methods}  & & \multicolumn{2}{c}{Single-locus methods} \\  \cline{2-7}   \cline{9-10}
 Features                 & AMplus  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
purpose built$^\dagger$    &   \cmark     &    \cmark      &  \xmark   &  \cmark  &  \cmark  &  \cmark && \cmark & \cmark           \\ [0.2cm]
 user  manual           & \cmark &   \xmark   & \cmark   & \xmark     &   \cmark     & \xmark          && \cmark    &    \cmark    \\  [0.2cm]
all snps fitted  &    \cmark     &    \cmark        &   \cmark          &      \cmark             &  \xmark          &  \xmark              &&   \xmark   &  \xmark     \\  
simultaneously      &         &            &             &                   &            &                &&      &      \\  [0.2cm]
capacity to include          &   \cmark   &      \cmark      &     \cmark    &  \xmark                 &      \cmark      &      \cmark          &&      \cmark & \cmark     \\  
other fixed effects$^\ddagger$                  &         &              &          &                    &            &                &&      &      \\  [0.2cm]
limited  memory$^\star$                   &   \cmark    &      \xmark      & \xmark          &  \xmark    &  \xmark    &   \xmark     && \cmark      & \xmark    \\  [0.2cm]
best set of snp           &    \cmark     &  \xmark    &  \xmark         &    \xmark               &     \cmark       &  \xmark              &&  \xmark   & \xmark      \\  
returned as result$^\uparrow$               &                           &        & & & & & & & \\  [0.2cm]
error checking           &  \cmark      &     \xmark        &      \xmark     &           \xmark        &      \xmark       &         \xmark       && \cmark  & \xmark \\ \hline

\end{tabular}
{$^\dagger$ \scriptsize{computer software specifically designed for the analysis of data from GWAS}}\\
{$^\ddagger$ \scriptsize{ allows  for the effect of population structure to be included as additional fixed effects}}\\
{\small $^\star$ \scriptsize{able to deal with data larger than the memory capacity of the computer}} \\
{\small $^\uparrow$ \scriptsize{no need to determine which 
marker loci are significant relative to some threshold or further analysis to calculate the significance of the findings. }}
\end{table}

\end{document}

### 2016-11-29 AM+/ChickenResults/am+GPU_FCR/am+.R
29/11/16 added AM+ with plots for paper to repository

@@ -1,0 +1,63 @@
+# trait read in from command line 
+library(MWAM)
+library(rcppMagmaSYEVD)
+
+
+## function to accept command line arguments
+args <- commandArgs()
+#if (length(args)!=11) {cat("error: not enough input variables") break}
+for (e in args) {
+  ta = strsplit(e,"=",fixed=TRUE)
+  if(! is.na(ta[[1]][2])) {
+    temp = ta[[1]][2]
+    if(substr(ta[[1]][1],nchar(ta[[1]][1]),nchar(ta[[1]][1])) == "I") {
+      temp = as.integer(temp)
+    }
+    if(substr(ta[[1]][1],nchar(ta[[1]][1]),nchar(ta[[1]][1])) == "N") {
+      temp = as.numeric(temp)
+    }
+    assign(substr(ta[[1]][1],2,nchar(ta[[1]][1])),temp)
+    cat("assigned ",ta[[1]][1]," the value of |",temp,"|\n")
+  } else {
+    assign(substr(ta[[1]][1],2,nchar(ta[[1]][1])),TRUE)
+    cat("assigned ",ta[[1]][1]," the value of TRUE\n")
+  }
+}
+
+
+
+## read genotypes
+DIR <- "./"    ## could use $MEMDIR but need to get this. 
+geno <- read.genotypes(path=DIR,
+                      bin_path=DIR , AA=0,  AB=1, BB=2,
+               file_genotype="geno.txt",
+                availmemGb=120)
+## read map
+map <- read.map(path = DIR , file_map = "map.txt", csv = FALSE)
+
+## phenotypes 
+pheno <- read.phenotypes(path = DIR , file_phenotype = "pheno.txt" ,
+                         header = TRUE, csv = FALSE )
+
+# remove any birds with NA's
+pheno$CG <- as.factor(pheno$CG)
+pheno[pheno==-9999] <- NA
+
+## Perform AM+ analysis
+## set args for multiple_locus_am function
+argu <- list(numcores = 16, availmemGb = 120 , colname.trait = trait, 
+          map = map , pheno = pheno, geno = geno,
+            colname.feffects=c("CG"), 
+           gpu=FALSE,
+          alpha = 0.05, error_checking = FALSE , maxit = 20, verbose=TRUE)
+
+## run AM+
+am_res <- do.call( multiple_locus_am, argu)
+
+save(am_res, "res.RData")
+
+
+cat(" Finished script .... \n")
+
+
+

### 2016-11-29 AM+/ChickenResults/am+GPU_FCR/am+GPU.R

@@ -1,0 +1,58 @@
+# trait read in from command line 
+library(MWAM)
+library(rcppMagmaSYEVD)
+
+
+## function to accept command line arguments
+args <- commandArgs()
+#if (length(args)!=11) {cat("error: not enough input variables") break}
+for (e in args) {
+  ta = strsplit(e,"=",fixed=TRUE)
+  if(! is.na(ta[[1]][2])) {
+    temp = ta[[1]][2]
+    if(substr(ta[[1]][1],nchar(ta[[1]][1]),nchar(ta[[1]][1])) == "I") {
+      temp = as.integer(temp)
+    }
+    if(substr(ta[[1]][1],nchar(ta[[1]][1]),nchar(ta[[1]][1])) == "N") {
+      temp = as.numeric(temp)
+    }
+    assign(substr(ta[[1]][1],2,nchar(ta[[1]][1])),temp)
+    cat("assigned ",ta[[1]][1]," the value of |",temp,"|\n")
+  } else {
+    assign(substr(ta[[1]][1],2,nchar(ta[[1]][1])),TRUE)
+    cat("assigned ",ta[[1]][1]," the value of TRUE\n")
+  }
+}
+
+
+
+## read genotypes
+DIR <- "./"    ## could use $MEMDIR but need to get this. 
+geno <- read.genotypes(path=DIR,
+                      bin_path=DIR , AA=0,  AB=1, BB=2,
+               file_genotype="geno.txt",
+                availmemGb=120)
+## read map
+map <- read.map(path = DIR , file_map = "map.txt", csv = FALSE)
+
+## phenotypes 
+pheno <- read.phenotypes(path = DIR , file_phenotype = "pheno.txt" ,
+                         header = TRUE, csv = FALSE )
+
+# remove any birds with NA's
+pheno$CG <- as.factor(pheno$CG)
+pheno[pheno==-9999] <- NA
+
+## Perform AM+ analysis
+## set args for multiple_locus_am function
+argu <- list(numcores = 16, availmemGb = 120 , colname.trait = trait, 
+          map = map , pheno = pheno, geno = geno,
+            colname.feffects=c("CG"), 
+           gpu=TRUE,
+          alpha = 0.05, error_checking = FALSE , maxit = 20, verbose=TRUE)
+
+## run AM+
+am_res <- do.call( multiple_locus_am, argu)
+
+save(am_res, "res.RData")
+

### 2016-11-29 AM+/ChickenResults/match.R

@@ -1,0 +1,45 @@
+# match am+ results to genes using map.txt file from Toni
+# Last modified:  21/07/2016
+# performed on my mac through RStudio
+
+
+# Summary
+# Results do not look very good. Some good genes but lots of noise as well. 
+# How to move forward:  inlude PC into the analysis
+snps <- c(
+  "Gga_rs14707919",
+ # "GGaluGA204479",
+  "Gga_rs16469122",
+  #"GGaluGA300550",
+  "GGaluGA108350",
+  "Gga_rs13595487",
+  "GGaluGA283510",
+  "Gga_rs14027776")
+  
+
+  
+## map with genes file
+map <- read.table("/Users/geo047/AM+/ChickenResults/map.txt", header=FALSE)
+names(map) <- c("SNP", "Pos", "Chr", "dist_to_gene", "Gene")
+
+
+##----------------
+## Trait
+##----------------
+## which  genes have been found by am+
+indx <- match(snps, map$SNP)
+genes_found <- map$Gene[indx]
+
+## other genes not found
+genes_not_found <- unique(map$Gene)
+indx <- match(genes_found, genes_not_found)
+genes_not_found <- genes_not_found[-indx]  ## removing found genes
+genes_not_found <- genes_not_found[-grep("LOC", genes_not_found)]
+# write two lists to file for GO enrichment analysis
+write.table(data.frame(genes_found), file="/Users/geo047/AM+/ChickenResults/genes_found.txt", col.names = F, row.names = F, quote = F)
+
+write.table(data.frame(genes_not_found), file="/Users/geo047/AM+/ChickenResults/genes_not_found.txt", col.names = F, row.names = F, quote = F)
+
+
+print(genes_found)
+

### 2016-11-29 AM+/Plots_for_Paper/ChickenPlots/result.R

@@ -1,0 +1,334 @@
+## forming results matrix for am+GPU by hand from
+## /flush1/geo047/RealDataAnalysis/ChickenAnalysis/AM_GSCali
+##
+## Doing annotated manhattan plots (annotation are results of the other programs)
+
+
+chr = "CHR"
+bp = "BP"
+p = "P"
+snp = "SNP"
+col = c("gray10", "gray60")
+chrlabs = NULL
+suggestiveline = -log10(0.00001) 
+genomewideline = -log10(0.00000005)
+highlight = NULL
+logp = TRUE
+trait="Ultra"
+
+
+
+
+
+
+
+manhattan_mine <- function (x, chr = "CHR", bp = "BP", p = "P", snp = "SNP", col = c("gray10", 
+                                                                   "gray60"), chrlabs = NULL, suggestiveline = -log10(0.00001), 
+          genomewideline = -log10(0.00000005), highlight = NULL, logp = TRUE, cindx=NULL, trait="Ultra") 
+{
+  CHR = BP = P = index = NULL
+  if (!(chr %in% names(x))) 
+    stop(paste("Column", chr, "not found!"))
+  if (!(bp %in% names(x))) 
+    stop(paste("Column", bp, "not found!"))
+  if (!(p %in% names(x))) 
+    stop(paste("Column", p, "not found!"))
+  if (!(snp %in% names(x))) 
+    warning(paste("No SNP column found. OK unless you're trying to highlight."))
+  if (!is.numeric(x[[chr]])) 
+    stop(paste(chr, "column should be numeric. Do you have 'X', 'Y', 'MT', etc? If so change to numbers and try again."))
+  if (!is.numeric(x[[bp]])) 
+    stop(paste(bp, "column should be numeric."))
+  if (!is.numeric(x[[p]])) 
+    stop(paste(p, "column should be numeric."))
+  d = data.frame(CHR = x[[chr]], BP = x[[bp]], P = x[[p]])
+  if (!is.null(x[[snp]])) 
+    d = transform(d, SNP = x[[snp]])
+  d <- subset(d, (is.numeric(CHR) & is.numeric(BP) & is.numeric(P)))
+  d <- d[order(d$CHR, d$BP), ]
+  if (logp) {
+    d$logp <- -log10(d$P)
+  }  else {
[…]

### 2016-11-29 AM+/Plots_for_Paper/WithOldStoppingRule/plot_paper.R

@@ -1,0 +1,267 @@
+## Results are obtained from bragg
+## input matrices cputimes.dat and memory.dat
+## Input obtained from /home/geo047/MWAM/SimStudy/Timing/Results  on Bragg cluster. 
+## See README file on Bragg for how these input files were created. 
+##
+## amGPU just mention these results in the paper. Doesn't make sense on the figure. Doesn't add anything. 
+##-----------------
+##  Main Program 
+##-------------------
+
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library("reshape2")
+library(RColorBrewer)
+##----------------------
+##  CPU times 
+##-----------------------
+
+## raw plots and logdf plots do not look that good. am+ is better but not by much.
+## going to try am+ as benchmark and comparing everything against it. -ve values mean 
+## other approach is better. 
+
+setwd("/Users/geo047/AM+/Plots_for_Paper")
+df <- read.table(file="cputimes.dat", header=TRUE)
+sizefn <- 16
+
+## times - ratios
+
+mlmm <- with(df, mlmm)
+piMASS <- with(df, piMASS)
+glmnet <- with(df, glmnet)
+rf <- with(df, rf)
+bigRR <- with(df, bigRR)
+lasso <- with(df, lasso)
+
+dfnew <- data.frame("mlmm"=mlmm, "piMASS"=piMASS, "glmnet"=glmnet, "rf"=rf,  
+                    "bigRR"=bigRR, "lasso"=lasso, check.names=FALSE)
+
+dfnew$Pop <- df$Pop
+#dfnew$Pop <- factor(dfnew$Pop,levels(dfnew$Pop)[c(2,3,5,4,6,1)])
+dfnew$Pop <- factor(dfnew$Pop,levels(dfnew$Pop)[c(6,4,5,3,1,2)])
+dfnew$Pop <- log10(c(150*5, 1500*50, 350*400, 2000*500, 4000*1500, 10000*1500))
+dfnew$size <- 2
+
+##-------------------
+##  Main plot
+##--------------------
+p1 <- ggplot(dfnew, aes(Pop, group=1, size=size)) + 
+  geom_line(aes(y = mlmm, colour = "mlmm")) +
+  geom_line(aes(y = piMASS, colour = "piMASS")) + 
[…]

### 2016-11-29 AM+/Plots_for_Paper/WithOldStoppingRule/power_plots.R

@@ -1,0 +1,163 @@
+## plotting precision verse recall. 
+## still have to code up ROC plot
+## input data resmat.RData ftp'ed from home directory on bragg in
+## /home/geo047/MWAM/SimStudy/Timing/Results/resmat$FAM.RData
+## contains columns
+## c("n_am", "nQTL_am",
+##  "n_amGPU", "nQTL_amGPU",
+##  "n_mlmm", "nQTL_mlmm",
+##  "n_glmnet", "nQTL_glmnet",
+##  "n_bigRR",  "nQTL_bigRR",
+##  "n_rf",  "nQTL_rf", "n_lasso", "nQTL_lasso",
+##  "n_piMASS", "nQTL_piMASS", "nQTL_true")
+
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library(RColorBrewer)
+DIR <- "~/AM+/Plots_for_Paper/"
+
+
+## vector initialisation
+fam <- c("W", "S",  "L","HS","A","HL")
+names.of.methods <- c("am", "mlmm","glmnet","piMASS","lasso","rf","bigRR")
+
+
+## list initialisation
+precision <- list()
+recall <- list()
+h <- list()
+
+
+for(ff in fam){
+  precision[[ff]] <- list()
+  recall[[ff]] <- list()
+  h[[ff]] <- list()
+  ## Load Data RData Objects
+  filename <- paste(DIR,"res",ff,".RData", sep="")
+  load(filename)   ## loads res_mat
+
+  
+  cat("forming Precision and REcall values \n")
+  for(ii in names.of.methods){
+    n_method <- paste("n_",ii, sep="")
+    nQTL_method <- paste("nQTL_",ii, sep="")
+    ## ---- Set PRec and Recall for AM+  -----
+    precision[[ff]][[ii]] <- res_mat[, eval(nQTL_method)]/res_mat[, eval(n_method)]
+    recall[[ff]][[ii]] <- res_mat[,eval(nQTL_method)] / res_mat[, "nQTL_true"]
+  }  ## end for ii
+  
+} ## end for ff
+
+
+## form results structure
+## method,  fam,  rep,  precision
+dfres <- data.frame()
+for(ff in names(precision))
+{
+  for(mm in names(precision[[1]]))
+  {
+    df <- data.frame(method=mm, fam=ff, rep=1:length(precision[[1]][[1]]),
+                     precision=precision[[ff]][[mm]],
[…]

### 2016-11-29 AM+/Plots_for_Paper/oldplot.R

@@ -1,0 +1,170 @@
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library(reshape)
+library(cowplot)
+library(gtable)
+
+
+##--------------------
+## ggplot2 plots
+##-----------------
+
+setwd("/Users/geo047/AM+/Plots_for_Paper")
+df1 <- read.table(file="cputimes.dat", header=TRUE)
+
+startc  <- 2
+finishc <- 7
+
+# log10 scale
+nms <- names(df)
+logdf <- log10(df1[,startc:finishc])
+logdf[logdf<1] <- 0   ## initializing values to 0 (less than a minute)
+logdf <- cbind.data.frame(df[,1:1], logdf)
+names(logdf) <- nms
+names(logdf) <- c("Method", "150.x.5K",    "1500.x.50K",  "350.x.400K", 
+                 "2000.x.500K" ,"4000.x.1.5M", "10000 x 1.5M")
+
+# Generate basic parallel coordinate plot
+p1 <- ggparcoord(data = logdf,                 
+                # Which columns to use in the plot
+                columns = startc:finishc,
+                # Which column to use for coloring data
+                groupColumn = 1,                 
+                # Do not show points
+                showPoints = TRUE,                
+                # Turn on alpha blending for dense plots
+                alphaLines = 0.6,                
+                # Turn off box shading range
+                shadeBox = NULL,                
+                # Will normalize each column's values to [0, 1]
+                scale = "globalminmax"  # try "std" also
+                
+) 
+
+
+p1 <- p1 + theme_hc()   ## nice but no x axes
+
+
+
+
+## add y labels  
+p1 <- p1 + ylab(expression( bold(log[10] ( Computing~Time) )))
+
+
+## manually label x axis values
+p1 <- p1 + scale_y_continuous(breaks = c(0,1,2,3,log10(100*60)), labels=c("0", "1","2","3",
+                                       "beyond \n machine \n capacity"))
+
+#  increase tick marker lenght (x and y)
+p1 <- p1 + theme(axis.ticks.length=grid::unit(0.3,"cm"))
+
+##  change y labels size and bold
+p1 <- p1 + theme(axis.title.y = element_text(angle=90, vjust=1, size=14))
+
[…]

### 2016-11-29 AM+/Plots_for_Paper/plot_aligned_series.R

@@ -1,0 +1,41 @@
+#' When plotting multiple data series that share a common x axis but different y axes,
+#' we can just plot each graph separately. This suffers from the drawback that the shared axis will typically
+#' not align across graphs due to different plot margins.
+#' One easy solution is to reshape2::melt() the data and use ggplot2's facet_grid() mapping. However, there is
+#' no way to label individual y axes.
+#' facet_grid() and facet_wrap() were designed to plot small multiples, where both x- and y-axis ranges are
+#' shared acros all plots in the facetting. While the facet_ calls allow us to use different scales with
+#' the \code{scales = "free"} argument, they should not be used this way.
+#' A more robust approach is to the grid package grid.draw(), rbind() and ggplotGrob() to create a grid of 
+#' individual plots where the plot axes are properly aligned within the grid.
+#' Thanks to https://rpubs.com/MarkusLoew/13295 for the grid.arrange() idea.
+
+library(ggplot2)
+library(grid)
+library(dplyr)
+
+#' Create some data to play with. Two time series with the same timestamp.
+df <- data.frame(DateTime = ymd("2010-07-01") + c(0:8760) * hours(2), series1 = rnorm(8761), series2 = rnorm(8761, 100))
+
+#' Create the two plots.
+plot1 <- df %>%
+  select(DateTime, series1) %>%
+  na.omit() %>%
+  ggplot() +
+  geom_point(aes(x = DateTime, y = series1), size = 0.5, alpha = 0.75) +
+  ylab("Red dots / m") +
+  theme_minimal() +
+  theme(axis.title.x = element_blank())
+
+plot2 <- df %>%
+  select(DateTime, series2) %>%
+  na.omit() %>%
+  ggplot() +
+  geom_point(aes(x = DateTime, y = series2), size = 0.5, alpha = 0.75) +
+  ylab("Blue drops / L") +
+  theme_minimal() +
+  theme(axis.title.x = element_blank())
+
+grid.newpage()
+grid.draw(rbind(ggplotGrob(plot1), ggplotGrob(plot2), size = "last"))
+

### 2016-11-29 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -1,0 +1,324 @@
+## Results are obtained from bragg
+## input matrices cputimes.dat and memory.dat, and abstimes.dat
+## Input obtained from /home/geo047/MWAM/SimStudy/Timing/Results  on Bragg cluster. 
+## See README file on Bragg for how these input files were created. 
+##
+## Last modified:  19/07/2016
+
+##-----------------
+##  Main Program 
+##-------------------
+
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library("reshape2")
+library(RColorBrewer)
+##----------------------
+##  CPU times 
+##-----------------------
+
+## raw plots and logdf plots do not look that good. am+ is better but not by much.
+## going to try am+ as benchmark and comparing everything against it. -ve values mean 
+## other approach is better. 
+
+setwd("/Users/geo047/AM+/Plots_for_Paper")
+df <- read.table(file="cputimes.dat", header=TRUE)
+sizefn <- 14
+
+df<- melt(df)
+names(df) <- c("Pop", "method","ratio")
+
+df$ratio[df$ratio<1] <- 1  ## setting times less than 1 to 1
+levels(df$Pop) <- c(10000*1500000, 150*5000, 1500*50000,2000*500000, 350*400000,  4000*1500000)
+df$Pop <- as.numeric(as.character(df$Pop))
+
+pps <- sort(unique(df$Pop))
+names(df) <- c("Pop", "method","ratio")
+##--------------------
+## times - ratios
+##--------------------
+
+## data frame for line segments
+gsdf <- data.frame(x1=(pps), y1=rep(1,length(pps)), 
+                   x2 = (pps), y2=c(700,500,700,500,700,500))
+
+
+##-------------------
+##  Main plot
+##--------------------
+
+## maybe draw vertical lines first to get around problem.
+
+p <- ggplot(df, aes(x=Pop, y=ratio, group=method,shape=method, colour=method)) + 
+  geom_line(size=2) + geom_point(aes(colour=method),size=6) +
+  scale_shape_manual(breaks=c("mlmm","glmnet","lasso","rf_ranger","bigRR"), 
+                     labels=c("MLMM","glmnet", "LMM-Lasso",
+                              "r2VIM","bigRR"),
+                     values=c(rep(19,6),rep(17,3)))+
+  scale_colour_manual(breaks=c("mlmm","glmnet","lasso","rf_ranger","bigRR"),
[…]

### 2016-12-19 MISC.tex
20/12/2016 major commit. Commiting all files in directory for paper

\ might need this for methods section. ...

The implementations evaluated are as follows. Glmnet is a general purpose R package. It implements lasso, an estimation technique where parameter estimation and variable selection occur simultaneously, for linear models. It's value for multiple locus association mapping has been demonstrated (ref). LARS in another general purpose R package implementing methods amenable for  multiple locus association mapping (ref). It relies on least angle regression, a form of forward selection, for model selection with linear models. BigRR is purpose built R package for the genome wide analysis of association data. It is based on linear models where parameter estimation is accomplished with ridge regression, a regularisation and shrinkage technique. Lmm-lasso and MLMM are two implementations purpose built for genome wide association mapping. Lmm-lasso uses lasso to identify the set of loci in association with the trait. MLMM instead implements a forward selection strategy with backward elimination. Both are based on linear mixed models. Pimass and randfor implement two very different approaches for multiple locus association mapping. In pimass, association mapping is performed within a Bayesian framework through Bayesian variable selection.  In ranfor, assoc mapping is performed in a machine learning framework through random forests. See table X for the  key features of the implementations evaluated in this paper.

### 2016-12-19 AM+/Plots_for_Paper/power_plots.R

@@ -1,4 +1,5 @@
 ## plotting power and FDR for multi-locus mehtods and single-locus methods (fast, fastALL, gemma)
+## Last revised:   29/11/2016  MAjor change. Plotting Power vs FDR 
 ## 
 ## Note: don't know why but this error would come up a lot when printing final plots. 
 ##
@@ -8,8 +9,10 @@
 ##       Solution: just repeat the printing of the plot a couple of times and it will 
 ##                 correct itself. 
 ##
-## input data res?.RData ftp'ed from home directory on bragg in
-## /home/geo047/MWAM/SimStudy/Timing/Results/res$FAM.RData
+## input data resx_y.RData ftp'ed from home directory on bragg in
+## /home/geo047/MWAM/SimStudy/Timing/Results/res$FAM_$indx.RData
+## where $FAM is W S L A HS HL and $indx is the index 1, 2, 3, etc over the thresholds used in
+## results.R on bragg 
 ## contains columns
 ## c("n_am", "nQTL_am",
 ##  "n_amGPU", "nQTL_amGPU",
@@ -23,7 +26,7 @@
 ##   n_fastALL, nQTL_fastALL,
 ##  "nQTL_true")
 ## Note 
-## moved from smoothed power curves to lines based on median. Smoothed curves weird.
+## moved from smoothed power curves to lines based on mean. Smoothed curves weird.
 require(ggplot2)
 library(GGally)
 library(ggthemes)
@@ -33,7 +36,7 @@ library(extrafont)
 DIR <- paste(getwd(),"/", sep="")
 
 sizefn <- 16
-
+thresh_indx <- 1:500
 
 
 ## vector initialisation
@@ -44,109 +47,76 @@ names.of.methods <- c("am", "mlmm","glmnet","lasso","r2VIM","bigRR", "gemma", "f
 ## list initialisation
 FDR <- list()
 recall <- list()  ## == power
-FDRp <- list()  ## FDR points
-recallp <- list()  ## recall points   == power
-h <- list()
+dfres <- list()
 
-
+##----------------------------------------
+## Forming list with FDR and recall(power)
+## results over families and threshold indexes
+##------------------------------------------
[…]

### 2016-12-19 AM+/Plots_for_Paper/power_plots_old.R

@@ -1,0 +1,416 @@
+## plotting power and FDR for multi-locus mehtods and single-locus methods (fast, fastALL, gemma)
+## 
+## Note: don't know why but this error would come up a lot when printing final plots. 
+##
+##       Error in grid.Call(L_textBounds, as.graphicsAnnot(x$label), x$x, x$y,  : 
+##       polygon edge not found
+## 
+##       Solution: just repeat the printing of the plot a couple of times and it will 
+##                 correct itself. 
+##
+## input data res?.RData ftp'ed from home directory on bragg in
+## /home/geo047/MWAM/SimStudy/Timing/Results/res$FAM.RData
+## contains columns
+## c("n_am", "nQTL_am",
+##  "n_amGPU", "nQTL_amGPU",
+##  "n_mlmm", "nQTL_mlmm",
+##  "n_glmnet", "nQTL_glmnet",
+##  "n_bigRR",  "nQTL_bigRR",
+##  "n_ranger",  "nQTL_ranger", 
+##  "n_lasso", "nQTL_lasso",
+##   n_gemma, nQTL_gemma, 
+##   n_fast, nQTL_fast, 
+##   n_fastALL, nQTL_fastALL,
+##  "nQTL_true")
+## Note 
+## moved from smoothed power curves to lines based on median. Smoothed curves weird.
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library(RColorBrewer)
+library(extrafont)
+DIR <- paste(getwd(),"/", sep="")
+
+sizefn <- 16
+
+
+
+## vector initialisation
+fam <- c("W", "S",  "L","HS","A","HL")
+names.of.methods <- c("am", "mlmm","glmnet","lasso","r2VIM","bigRR", "gemma", "fastALL", "fast")
+
+
+## list initialisation
+FDR <- list()
+recall <- list()  ## == power
+FDRp <- list()  ## FDR points
+recallp <- list()  ## recall points   == power
+h <- list()
+
+
+for(ff in fam){
+  FDR[[ff]] <- list()
+  recall[[ff]] <- list()
+  FDRp[[ff]] <- list()
+  recallp[[ff]] <- list()
+  
+  ## Load Data RData Objects
+  filename <- paste(DIR,"res",ff,".RData", sep="")
+  load(filename)   ## loads res_mat
+
+  
[…]

### 2016-12-19 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -23,7 +23,7 @@ library(RColorBrewer)
 ## going to try am+ as benchmark and comparing everything against it. -ve values mean 
 ## other approach is better. 
 
-setwd("/Users/geo047/AM+/Plots_for_Paper")
+setwd("/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper")
 df <- read.table(file="cputimes.dat", header=TRUE)
 sizefn <- 14
 
@@ -52,16 +52,20 @@ gsdf <- data.frame(x1=(pps), y1=rep(1,length(pps)),
 ## maybe draw vertical lines first to get around problem.
 
 p <- ggplot(df, aes(x=Pop, y=ratio, group=method,shape=method, colour=method)) + 
-  geom_line(size=2) + geom_point(aes(colour=method),size=6) +
+  geom_line(size=2) + geom_point(aes(colour=method),size=5) +
   scale_shape_manual(breaks=c("mlmm","glmnet","lasso","rf_ranger","bigRR"), 
                      labels=c("MLMM","glmnet", "LMM-Lasso",
                               "r2VIM","bigRR"),
-                     values=c(rep(19,6),rep(17,3)))+
+                     values=c(rep(19,5)))+
   scale_colour_manual(breaks=c("mlmm","glmnet","lasso","rf_ranger","bigRR"),
                       labels=c("MLMM","glmnet", "LMM-Lasso",
                                "r2VIM","bigRR"),
-                      values=brewer.pal(9, "Paired")[c(2,3,5,6,4,1,7,8,9)]) + 
-  scale_size(guide = 'none') ## suppressed extra legend for size
+                      values=brewer.pal(9, "Paired")[c(2,3,9, 4,5)]) + 
+  scale_size(guide = 'none') 
+p
+
+
+## suppressed extra legend for size
 ## the reordered colors is complicated but it is based on the original colors 
 ## being of the order 1,2,5,3,4 (don't know why ggplot2 would do it this way) so
 ## I had to put color 2 in position 1, 
@@ -150,7 +154,6 @@ p <- p + theme(legend.position = c(0.82, 0.8),
  p <- p + theme(plot.margin = grid::unit(c(1,5,1,1), "lines"))
  
  
- 
 p_time <- p
 p_time
 
@@ -189,8 +192,8 @@ p_time
   ##--------------------
   
   
-  p <- ggplot(df, aes(x=Pop, y=time, group=method,shape=method, colour=method)) + 
[…]

### 2016-12-19 Precision and recall - Wikipedia, the free encyclopedia_files/load_002.js

@@ -1,0 +1,71 @@
+var mediaWikiLoadStart=(new Date()).getTime(),mwPerformance=(window.performance&&performance.mark)?performance:{mark:function(){}};mwPerformance.mark('mwLoadStart');function isCompatible(str){var ua=str||navigator.userAgent;return!!('querySelector'in document&&'localStorage'in window&&'addEventListener'in window&&!(ua.match(/webOS\/1\.[0-4]/)||ua.match(/PlayStation/i)||ua.match(/SymbianOS|Series60|NetFront|Opera Mini|S40OviBrowser|MeeGo/)||(ua.match(/Glass/)&&ua.match(/Android/))));}(function(){var NORLQ,script;if(!isCompatible()){document.documentElement.className=document.documentElement.className.replace(/(^|\s)client-js(\s|$)/,'$1client-nojs$2');NORLQ=window.NORLQ||[];while(NORLQ.length){NORLQ.shift()();}window.NORLQ={push:function(fn){fn();}};window.RLQ={push:function(){}};return;}function startUp(){mw.config=new mw.Map(true);mw.loader.addSource({"local":"/w/load.php","metawiki":"//meta.wikimedia.org/w/load.php"});mw.loader.register([["site","1t3wrg7",[1]],["site.styles","0idn2vw"
[…]

### 2016-12-20 AM+/Plots_for_Paper/threshold_plots.R
Daily backup 21/12/2016

@@ -1,0 +1,289 @@
+## plotting power and FDR for multi-locus mehtods and single-locus methods (fast, fastALL, gemma)
+## Last revised:   29/11/2016  MAjor change. Plotting Power vs FDR 
+## 
+## Note: don't know why but this error would come up a lot when printing final plots. 
+##
+##       Error in grid.Call(L_textBounds, as.graphicsAnnot(x$label), x$x, x$y,  : 
+##       polygon edge not found
+## 
+##       Solution: just repeat the printing of the plot a couple of times and it will 
+##                 correct itself. 
+##
+## input data resx_y.RData ftp'ed from home directory on bragg in
+## /home/geo047/MWAM/SimStudy/Timing/Results/res$FAM_$indx.RData
+## where $FAM is W S L A HS HL and $indx is the index 1, 2, 3, etc over the thresholds used in
+## results.R on bragg 
+## contains columns
+## c("n_am", "nQTL_am",
+##  "n_amGPU", "nQTL_amGPU",
+##  "n_mlmm", "nQTL_mlmm",
+##  "n_glmnet", "nQTL_glmnet",
+##  "n_bigRR",  "nQTL_bigRR",
+##  "n_ranger",  "nQTL_ranger", 
+##  "n_lasso", "nQTL_lasso",
+##   n_gemma, nQTL_gemma, 
+##   n_fast, nQTL_fast, 
+##   n_fastALL, nQTL_fastALL,
+##  "nQTL_true")
+## Note 
+## moved from smoothed power curves to lines based on mean. Smoothed curves weird.
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library(RColorBrewer)
+library(extrafont)
+DIR <- paste(getwd(),"/", sep="")
+
+sizefn <- 16
+thresh_indx <- 1:500
+
+
+## vector initialisation
+fam <- c("W", "S",  "L","HS","A","HL")
+names.of.methods <- c("am", "mlmm","glmnet","lasso","r2VIM","bigRR", "gemma", "fastALL", "fast")
+
+
+## list initialisation
+FDR <- list()
+recall <- list()  ## == power
+dfres <- list()
+
+##----------------------------------------
+## Forming list with FDR and recall(power)
+## results over families and threshold indexes
+##------------------------------------------
[…]

### 2017-01-21 AM+draft.tex
29/11/16 added AM+ with plots for paper to repository
Regular update

@@
 
-The memory usage of the software programs is shown in Figure \ref{fig_memory}.
+We also examined the memory usage of the different software programs (figure \ref{fig_memory} supplementary materials). 
+AMplus is comparable to the most efficient single locus implementation, FaST-LMM$^few$. 
+AMplus is also the only multiple locus program able to analyse data under scenarios ..... and ..... . 
+AMplus can process data larger than the memory capacity of the computer.
 
@@
 \caption{Memory usage}
-\includegraphics[width=10cm]{memory}
+\includegraphics[width=10cm]{mem}
 \end{figure}
 
-\subsection{Statistical Performance: recall and precision}
+\subsection{Statistical Performance: power, FDR, and significance thresholds}
+
+As part of assessing the performance of AMplus against the other association mapping methods, 
+we calculated their statistical power and false discovery rates. 
+We want a method to have high statistical power (probability of finding a true positive finding) but 
+low false discovery rate (proportion of findings that are false positives). 
+All the methods, besides AMplus and MLMM, required a threshold to be set in order to identify significant findings. 
+A conservative threshold guards against false discoveries but reduces power. 
+A anti-conservative threshold increases power at the cost of also increasing the false discovery rate.
+Only AMplus and mlmm avoid thresholds by treating association mapping as a model selection problem where the best model is found by minimising the model selection statistic, extBIC.
+
+Figures \ref{fig_power_multiple}  and \ref{fig_power_single} show the relationship, calculated empirically, between the power and fdr for each method.
+The multiple-locus (Figure \ref{fig_power_multiple})   and the single-locus (Figure \ref{fig_power_single})  methods are plotted separately for clarity.
+AMplus features in both plots for comparison.
+By varying the significance thresholds for the different methods, we were able to calculate the change in power and fdr.
+Since AMplus and MLMM do not rely on thresholds, their power and fdr appear as single points in the plots.
+In practice, the true power and fdr is unknown.
+However, because we generated data where we knew which loci were acting as quantitative trait loci, we were able to calculate the true power and fdr for the methods.
+
+From figures \ref{fig_power_multiple}  and \ref{fig_power_single}, the superior performance of AMplus is apparent.
+When the fdr is low, AMplus has the highest power with Mlmm also performing well.
+AMplus is noticeably more powerful than the single locus methods.
+It is possible to set a significance threshold for these other methods that results in higher power. 
+However, this comes at the cost of an intolerably high fdr.
+
+For the methods evaluated here that require a significance threshold, figure  \ref{fig_threshold}  shows how 
+threshold size impacts fdr. Each plot corresponds to the results from a different method. 
+A curve within a plot is formed from analysing data with the method and using different threshold values to identify locus-trait associations. 
+The true fdr is then calculated from these findings.
+We can see from figure \ref{fig_threshold}  that the threshold impacts the fdr differently across methods.
+The relationship between threshold value and fdr even differs when data are analysed with the same method but for different scenarios.
+This demonstrates one of the challenges in using association mapping methods that are reliant upon thresholds.
+With AMplus, we do not have this challenge.
+For completeness, for the small sampled size scenarios ...
+And ....., the fdr for AMplus was 0.07 and 0.09, respectively.
+With the larger study sizes of ..., ..., and ...., the fdr drops to ..., .., and .....
+respectively.
+
+\begin{figure}
+\label{fig_power_multiple}
+\caption{Power curves for  multiple locus methods}
+\includegraphics[width=10cm]{powerMultiple}
+\end{figure}
+
+\begin{figure}
+\label{fig_power_single}
+\caption{Power curves for single-locus methods -- Put into sup methods}
+\includegraphics[width=10cm]{powerSingle}
+\end{figure}
+
+\begin{figure}
+\label{fig_threshold}
+\caption{The impact of threshold on FDR }
+\includegraphics[width=10cm]{threshold}
+\end{figure}
 

### 2017-01-21 AM+/Plots_for_Paper/threshold_plots.R

@@ -45,7 +45,8 @@ recall <- list()
 ## bragg. It sits in the results.R file on /home/geo047/MWAM/SimStudy/Timing/RScripts/
 pthresh  <- seq(0.01,0.99, length.out=500)
 pthresh_r2VIM <- seq(0.05, 100, length.out=500)
-alpha <- 10**-seq(-log10(1e-35), -log10(1e-2), length.out=500)
+#alpha <- 10**-seq(-log10(1e-35), -log10(1e-2), length.out=500)
+alpha <- seq(-log10(1e-35), -log10(1e-2), length.out=500)
 
 
 
@@ -75,7 +76,7 @@ for(ff in fam){
       ## ---- Set Power (recall) and FDR 
       tmp <-  1 - (mat[, eval(nQTL_method)]/
                     mat[, eval(n_method)])
-      #tmp[is.nan(tmp)] <- 0
+      tmp[is.nan(tmp)] <- 0
           FDR[[ff]][[indx]][[ii]] <- mean(tmp, na.rm=TRUE)
       # capturing case where there may not be any results 
      if(is.nan(FDR[[ff]][[indx]][[ii]]))
@@ -98,14 +99,17 @@ dfres <- data.frame()
 
 ## looping over families
 for(ff in names(FDR))
-{
+{ 
+   
+  dftmp <- data.frame()
   cat(" Reading in family", ff, "\n")
   ## looping over threshold indexes
   for(indx in thresh_indx)
   {
     ## looping over methods
     for(mm in names(FDR[[1]][[1]])){
-      if(mm=="glmnet"){
+      #for(mm in "gemma"){
+        if(mm=="glmnet"){
         df <- data.frame(method=mm, fam=ff, rep=1:length(FDR[[1]][[1]][[1]]),
                          FDR=FDR[[ff]][[indx]][[mm]] ,
                          threshold = pthresh[indx])    
@@ -156,23 +160,33 @@ for(ff in names(FDR))
         
       }
         
-      dfres <- rbind.data.frame(dfres , df)
+      dftmp <- rbind.data.frame(dftmp , df)
       
 
   } ## end for thresh_indx  
   }  ## end for mm
+  dfres <- rbind.data.frame(dfres, dftmp)
 }  ## end for ff
 
 
 
-
+## checking if all values are 0 for a curve
+for(mm in levels(dfres$method)){
[…]

### 2017-09-13 AM+/Plots_for_Paper/power_plots.R
14Sep17 About to begin writing methods section

@@ -34,7 +34,7 @@ library(gridExtra)
 library(RColorBrewer)
 library(extrafont)
 DIR <- paste(getwd(),"/", sep="")
-
+DIR <- "/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper/"
 sizefn <- 16
 thresh_indx <- 1:500
 

### 2017-09-26 AM+draft.tex
14Sep17 About to begin writing methods section
122717 Working on Methods - Overview

@@
 This has meant that data can be collected from general populations, even if these populations are highly structured.
-Analysis via sophisticated association mapping methods based on linear mixed models is now almost routine.
+Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine.
 
 What has not changed over the past decade is that it remains common practice to analyse gwas data on a locus by locus basis.
-This is despite there being several significant problems with analysing the data in this way.
+This is despite there being several significant problems with analysing data in this way.
 First, the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait.
@@
 There will be multiple loci in linkage equilibrium with genes that influence the trait.
-Yet, A locus by locus mapping approach only assesses the evidence for association between a single locus and trait.
+Yet, A locus by locus mapping approach only assesses the evidence for association between a single marker locus and trait.
 
-It is somewhat surprising then that multiple locus methods haven't attracted more attention.
-It is possible to measure all locus-trait associations jointly with methods based on the regularisation techniques, ridge regression and lasso.
+It is somewhat surprising then that multiple locus association mapping methods haven't attracted more attention. Methods based on 
+regularisation techniques, such as ridge regression and lasso, measure all locus-trait associations simultaneously. 
 Here, multiple testing is not an issue.
 These techniques though are computationally demanding.
-Also, the results are not easy to interpret.
+Also, their results can be difficult to interpret.
 The strength of association is not measured by a p-value but by the size of the regression coefficient for a locus in the model.
@@
 
-Here, we present our new multiple locus method for genome wide association mapping, which we are calling AMplus.
-AMplus combines the strength of regularisation techniques (being able to fit all locus-trait associations jointly), with simple variable selection (having easy to interpret results).
+In this paper, we present our new multiple locus method for genome wide association mapping, which we are calling Eagle.
+Eagle combines the strength of regularisation techniques (being able to fit all locus-trait associations jointly), with simple variable selection (having easy to interpret results).
 Our method does not require a significance threshold to be set nor regularisation parameters to be fine tuned.
 Through a clever dimension reduction step, we are able to achieve a computational performance similar to the fastest single locus linear mixed model implementations.
-AMplus is statistically more powerful than single locus association mapping and is as and often more powerful than
+Eagle is statistically more powerful than single locus association mapping and is as and often more powerful than
 most multiple locus methods.
@@
 
-\subsection{Application to Arabidopsis data}
+\subsection{Mouse Study}
+
+\section{Methods}
+
+\subsection{Outbred Mice Study}
+
+Phenotypic and genotypic data were collected on a large study population of commercially 
+available outbred mice. The aim of the study was to map, through association, the genetic basis
+of complex traits in mice. The construction of the population and collection of data are described in detail in 
+\cite{Nicod:2016aa}.  Briefly, the population consists of 2073 adult outbred mice from Charles River Laboratories. Phenotypic data
+were obtained from 200 traits. These traits fell into the categories of behaviour, tissue, and physiology. Genotype 
+data were collected from 7, 073, 398 SNPs.  A reduced set of 359, 559 SNP across the 19 
+autosomes and sex chromosome was then identified, upon which all genome-wide analyses were based.  These SNP tagged all other SNP with MAF $> 0.1\%$
+and LD $r^2 > 0.98$.  Missing data?
+
+\subsection{Our Multiple-Locus Association Mapping Approach}
+\subsubsection{Overview}
+
+Eagle is an approach that identifies genomic regions of interest for a trait via the building of a (linear mixed) model.
+The model is built iteratively, without need of significance levels or thresholds.
+The model has two parts, a fixed effects part and a random effects part.
+The random effects
+part models the error and association across the entire genome. It is from the random part that we are able to identify 
+the snp in strongest association with the trait.  We then move this snp out from the random effects part of the model and 
+into the fixed effects part.  This serves two purposes. First, by moving the snp into the fixed effects part of the model, we are accounting
+ for its effect and the effect of its neighbours in the random part of the model. So, when we come an look for the next snp in association 
+ with the trait, we do not find the neighbour of a snp already identified. Second, it allows us to discover snp-trait associations that would 
+ otherwise be masked.  Our iterative process is to fit the model to the data,  identify the snp in strongest association with the trait from the 
+ random effects part, move the snp into the fixed effects part and remove it from the random effects part, fit the new model, calculate if the new 
+ model is a "better" fit to the data.  If so, continue the process.  The snps that have been moved into the fixed effects part are the findings from the 
+ Eagle approach. Each snp identifies a separate region of the genome that houses genes that are influencing the trait. 
+ Below, we present the details of the Eagle algorithm.
+
+\subsubsection{The specifics}
+
+Our linear mixed model for multiple-locus association mapping, given trait data $y^{(n \times 1)}$ which are collected on $n$ individuals, 
+is of the form 
+\begin{equation}
+y = X \tau + u_g + e
+\end{equation}
+where $X^{(n \times p)}$ is a known design matrix of full rank, $\tau^{(p \times 1)}$ is a vector of fixed effects parameters, $u_g^{(n \times 1)}$ is a vector 
+of genetic effects, and $e^{(n \times 1)}$ is a vector of residuals whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \times n)})$.
+
+Suppose $s$ SNP have been already been found. Let $S={ S_1, S_2, \ldots, S_s}$ be the set of column indexes
+
+\bibliographystyle{nature}
+\bibliography{bibcitelibrary}
 

### 2017-09-27 AM+draft.tex
270917 Finished first draft of Methods - Overview - happy with this.

@@
 
-\subsection{Our Multiple-Locus Association Mapping Approach}
+\subsection{Eagle Approach for Multiple-locus Association Mapping}
 \subsubsection{Overview}
@@
 The model has two parts, a fixed effects part and a random effects part.
-The random effects
-part models the error and association across the entire genome. It is from the random part that we are able to identify 
-the snp in strongest association with the trait.  We then move this snp out from the random effects part of the model and 
-into the fixed effects part.  This serves two purposes. First, by moving the snp into the fixed effects part of the model, we are accounting
- for its effect and the effect of its neighbours in the random part of the model. So, when we come an look for the next snp in association 
- with the trait, we do not find the neighbour of a snp already identified. Second, it allows us to discover snp-trait associations that would 
- otherwise be masked.  Our iterative process is to fit the model to the data,  identify the snp in strongest association with the trait from the 
- random effects part, move the snp into the fixed effects part and remove it from the random effects part, fit the new model, calculate if the new 
- model is a "better" fit to the data.  If so, continue the process.  The snps that have been moved into the fixed effects part are the findings from the 
- Eagle approach. Each snp identifies a separate region of the genome that houses genes that are influencing the trait. 
- Below, we present the details of the Eagle algorithm.
+The random effects part contains the error and an effect for the snp-trait association across the entire genome.
+It is from this effect
+ that we are able to identify which snp is in strongest association with the trait.  
+ Our model building process is as follows.
+First we fit the current model to the data.
+Second, from the random effect for association, we identify which snp is in strongest association  with the trait.
+Third, we move this snp from the random effects to the fixed effects.
+We do this to stop the neighbours of the chosen snp from being selected in subsequent iterations and 
+to give opportunity to snp in linkage disequilibrium with genes having lesser impact on the trait to be discovered. 
+ Fourth, we fit the  new model.
+Fifth, we calculate if the new model fits the data "better" than the current model.
+If the new model is a better fit,  then we repeat the process.
+Once the model building process is complete, the findings from Eagle are the set of snp that have been moved into the fixed effects part of the model.
+These snp map the seperate regions of the genome that house genes that are influencing the trait.
 

### 2017-10-06 AM+draft.tex
061017 Working on methods section - Eagle approach

@@
 to give opportunity to snp in linkage disequilibrium with genes having lesser impact on the trait to be discovered. 
- Fourth, we fit the  new model.
-Fifth, we calculate if the new model fits the data "better" than the current model.
-If the new model is a better fit,  then we repeat the process.
+ Fourth, we calculate if the new model is "better" than the current model. 
+If so,  the new model becomes the current model and we repeat the process.
 Once the model building process is complete, the findings from Eagle are the set of snp that have been moved into the fixed effects part of the model.
@@
 \subsubsection{The specifics}
+The Eagle approach for multiple-locus association mapping is based on building a linear mixed model, iteratively, via forward 
+selection.  Suppose $s$ iterations of our model building process have already been performed. 
+The next iteration is performed as follows. First, we fit the current model to the data. The model is of the form 
+\begin{equation}
+y = X \tau + u_g + e
+\end{equation}
+where $y^{(n \times 1)}$ is a vector containing the quantitative trait values observed from $n$ individuals, 
+$X^{(n \times p)}$ is a known design matrix of full rank, 
+$\tau^{(p \times 1)}$ is a vector of $p$ fixed effects parameters, 
+$u_g^{(n \times 1)}$ is a vector 
+of genetic effects, and 
+$e^{(n \times 1)}$ is a vector of residuals whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \times n)})$.
+
+Suppose phenotypic and genotypic data are collected from a genome-wide association study involving $n$ individuals. 
+Let $Y^{(n \times 1)}$ be a vector containing the quantitative trait data.
+Also, let $M^{(n \times L)}$ be a matrix of snp genotypes which have been collected from the $n$ individuals across $L$ snp loci.
+$M$ contains values -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
+
+To identify genomic regions which contain genes that influence the trait
+
+We map genomic regions containing genes that influence the trait
+
+Suppose $s$ iterations have been performed of our multiple locus association mapping algorithm. This means $s$ snp-trait 
+associations of significance have been identified. It also means $s$ separate regions containing putative genes 
+that influence the trait have been mapped.  To find perform iteration $s+1$ and the next snp, we perform the following steps.  First,
+
+Let $M^{(n \times L} $be a matrix of snp genotypes which have been collected from the n individuals across L snp loci.
+These L loci are assumed to span the entire genome.
+M contains values -1, 0, and 1 corresponding to snp genotypes aa, ab, and bb, respectively.
+
+S={S1,...Ss} be
+
+Our model building process is as follows.  Suppose $s$ SNP have already been found. First we fit the current model to the data.
+
+Suppose our iterative model building procedure has already found $s$ snp. Then
 
@@
 
-Suppose $s$ SNP have been already been found. Let $S={ S_1, S_2, \ldots, S_s}$ be the set of column indexes
+Now, let $M^{(n \times L)}$ be a matrix of genotypes collected from $L$ snp.  The 
+matrix contains the values -1, 0, and 1, corresponding to the snp genotypes AA, AB, and BB, respectively.  It is assumed that 
+missing genotypes have been imputed.
+
+Suppose $s$ SNP have already been found.
+
+$S={ S_1, S_2, \ldots, S_s}$ be the set of column indexes
 

### 2017-10-06 AM+draft.tex
061017 Just some minor tweaks to methods section - end of day

@@
 
+Suppose a genome-wide association study is conducted. Phenotypic and genotype data are collected.
+
 Suppose phenotypic and genotypic data are collected from a genome-wide association study involving $n$ individuals. 
-Let $Y^{(n \times 1)}$ be a vector containing the quantitative trait data.
-Also, let $M^{(n \times L)}$ be a matrix of snp genotypes which have been collected from the $n$ individuals across $L$ snp loci.
-$M$ contains values -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
-
-To identify genomic regions which contain genes that influence the trait
-
-We map genomic regions containing genes that influence the trait
+Let $y^{(n \times 1)}$ be a vector containing the quantitative trait data. 
+Also, let $M^{(n \times L)} = [m_1 m_2 \ldots m_L]$ be a matrix of snp genotypes which have been collected from $L$ snp loci 
+where $m_j^{(n \times 1)}$ is the vector of snp genotypes collected on the $j$ snp. 
+The vector $m_j$  contains values -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
 

### 2017-10-16 AM+draft.tex
161017 Working on methods  section

--- 2017-10-06 AM+draft.tex
@@
 \subsubsection{The specifics}
+
+%%Suppose data are collected from a genome-wide association study.  The study involves $n$ individuals. 
+%Let $y^{(n \times 1)}$ be the vector containing the quantitative trait data. 
+%Let $M^{(n \times L)} = [m_1 m_2 \ldots m_L]$ be the matrix containing the genotype data. Snp genotypes 
+%are collected from $L$ loci. The vector $m_j^{(n \times 1)}$ contains the snp genotypes for the $j$th snp. The genotypes 
+%are recoded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
+
 The Eagle approach for multiple-locus association mapping is based on building a linear mixed model, iteratively, via forward 
-selection.  Suppose $s$ iterations of our model building process have already been performed. 
-The next iteration is performed as follows. First, we fit the current model to the data. The model is of the form 
+selection.  Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
+associations which identify $s$ separate genomic regions of interest have been found.  To perform the $s+1$ iteration of 
+our model building procedure, we perform the following.
+
+First, we fit the current model to the data. 
+Let $y^{(n \times 1)}$ be the vector containing the quantitative trait data. 
+Let $M^{(n \times L)} = [m_1 m_2 \ldots m_L]$ be the matrix containing the genotype data which has been collected 
+from $L$ loci that span the genome. Here, the vector $m_j^{(n \times 1)}$ contains the $j$th snp genotypes for the $n$ individuals. 
+The genotypes are coded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
+
+The model is of the form 
 \begin{equation}
@@
 \end{equation}
-where $y^{(n \times 1)}$ is a vector containing the quantitative trait values observed from $n$ individuals, 
-$X^{(n \times p)}$ is a known design matrix of full rank, 
-$\tau^{(p \times 1)}$ is a vector of $p$ fixed effects parameters, 
-$u_g^{(n \times 1)}$ is a vector 
-of genetic effects, and 
-$e^{(n \times 1)}$ is a vector of residuals whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \times n)})$.
+where 
+$X^{(n \times p)}$ is a known design matrix of full rank and may contain data on non-genetic effects such as age and height 
+or principal component values to account for population structure (ref). The vector 
+$\tau^{(p \times 1)}$ has $p$ fixed effects parameters. The vector 
+$u_g^{(n \times 1)}$ contains the 
+genetic effects. The vector of residuals is 
+$e^{(n \times 1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \times n)})$.
 
-Suppose phenotypic and genotypic data are collected from a genome-wide association study involving $n$ individuals. 
-Let $Y^{(n \times 1)}$ be a vector containing the quantitative trait data.
-Also, let $M^{(n \times L)}$ be a matrix of snp genotypes which have been collected from the $n$ individuals across $L$ snp loci.
-$M$ contains values -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
-
-To identify genomic regions which contain genes that influence the trait
-
-We map genomic regions containing genes that influence the trait
+Up to this point,  this model differs little from standard linear mixed models for association mapping (refs). 
+However, 
+it is how we specify $u_g$ that distinguishes our model from others. The genetic effects $u_g$ are modelled as
 

### 2017-10-18 AM+draft.tex
181017 Working on methods

@@
 selection.  Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
-associations which identify $s$ separate genomic regions of interest have been found.  To perform the $s+1$ iteration of 
-our model building procedure, we perform the following.
+associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.
 
+To perform the $s+1$th  iteration of 
+our model building procedure, we do the following. 
 First, we fit the current model to the data. 
-Let $y^{(n \times 1)}$ be the vector containing the quantitative trait data. 
-Let $M^{(n \times L)} = [m_1 m_2 \ldots m_L]$ be the matrix containing the genotype data which has been collected 
-from $L$ loci that span the genome. Here, the vector $m_j^{(n \times 1)}$ contains the $j$th snp genotypes for the $n$ individuals. 
+Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered snp that was 
+selected in the $k$th iteration. For example, if three iterations of our model building procedure 
+have been performed and say the 500023rd, 15th, and 420th, 
+snp were selected, then $S=\{500023, 15, 420\}$. 
+Let $y^{(n \; \times \;1)}$ be a vector containing the quantitative trait data. 
+Let $M^{(n \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
+from $L$ loci that span the genome.  The columns of this matrix are most commonly placed in map ordered but 
+this is not a requirement. 
+The vector $m_j^{(n \; \times \; 1)}$ contains the genotypes, from the $n$ individuals,  for the $j$th snp. 
 The genotypes are coded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
 
-The model is of the form 
+gocard123
+
+The (current) model is of the form 
 \begin{equation}
+\label{eq1}
 y = X \tau + u_g + e
@@
 where 
-$X^{(n \times p)}$ is a known design matrix of full rank and may contain data on non-genetic effects such as age and height 
-or principal component values to account for population structure (ref). The vector 
-$\tau^{(p \times 1)}$ has $p$ fixed effects parameters. The vector 
-$u_g^{(n \times 1)}$ contains the 
+$X^{(n \; \times \; p)}$ is a known design matrix of full rank and may contain data on non-genetic effects such as age and height 
+and/or principal component values to account for population structure (ref). The vector 
+$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters. The vector 
+$u_g^{(n \; \times \; 1)}$ contains the 
 genetic effects. The vector of residuals is 
-$e^{(n \times 1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \times n)})$.
+$e^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \; \times \; n)})$. 
+So far,  this model differs little from standard linear mixed models for association mapping (refs). 
+However, 
+it is how we specify $u_g$ that distinguishes our model from others.
 
-Up to this point,  this model differs little from standard linear mixed models for association mapping (refs). 
-However, 
-it is how we specify $u_g$ that distinguishes our model from others. The genetic effects $u_g$ are modelled as
+The genetic effects $u_g$ are modelled as 
+\begin{equation}
+\label{eq2}
+u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + M_{-S} a_{-S}
+\end{equation}
+where $m_{S_k}^{(n \; \times \; 1)}$ is the vector of genotypes for the $S_k$th snp locus which is the $k$th selected snp, 
+$a_{S_k}$ is the additive effect of the $S_k$th snp locus, $M_{-S}^{(n \; \times \; L-s)}$ is the matrix of  snp genotypes 
+with the data for the selected snp in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
+$a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
+The sum of terms on the left hand side are fixed effects.  The other term is a random effect.  The sum of terms accounts 
+for the additive effects of genes that are in linkage disequilibrium with the selected snp. The other term models 
+snp-trait associations along the entire genome, except for those snp that have already been selected. This is a very simple genetic model but it 
+is effective for discovering snp-trait associations. It also reduces the computational cost of our model building procedure.
 
-Suppose $s$ iterations have been performed of our multiple locus association mapping algorithm. This means $s$ snp-trait 
-associations of significance have been identified. It also means $s$ separate regions containing putative genes 
-that influence the trait have been mapped.  To find perform iteration $s+1$ and the next snp, we perform the following steps.  First,
+%The simplicity of our genetic model also reduces the cost of computation.
 
-Let $M^{(n \times L} $be a matrix of snp genotypes which have been collected from the n individuals across L snp loci.
-These L loci are assumed to span the entire genome.
-M contains values -1, 0, and 1 corresponding to snp genotypes aa, ab, and bb, respectively.
+Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood estimation.
 
-S={S1,...Ss} be
+Third, we identify the $(s+1)$th snp that is in strongest association with the trait, based on the maximum score statistic
+$t_j^2 = \frac{ \widetilde{a} _j^2}{\textrm{var}(\widetilde{a}_j)}$ where $\widetilde{a}_j$ is the best linear unbiased predictor (BLUP), 
+and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
+identify a snp based on its (random) effect size and accuracy, but is theoretically justified.  It follows from outlier detection 
+in linear and linear mixed models (ref).
 
-Our model building process is as follows.  Suppose $s$ SNP have already been found. First we fit the current model to the data.
+Fourth, we determine the importance of the $(s+1)$th selected snp via a model selection strategy. 
+We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected snp.  We then fit this new model to the data
+via maximum likelihood and calculate its extended Bayesian information criteria \cite{chen2008extended}.
 
-Suppose our iterative model building procedure has already found $s$ snp. Then
+We reform (\ref{eq2}) where 
+$S$ contains $s+1$ selected snp. We then fit (\ref{eq1}) and (\ref{eq2}) to the data via maximum likelihood and calculate $extBIC$.
+We found this model selection criterium to perform better for association mapping problems than standard model selection 
+criteria such as BIC and IAC, as was also found by XXXXX.  If the extBIC value is less than the extBIC for the previous model, 
+then we add the $s+1$th snp  to the model and repeat the process. If  the extBIC value is greater than or equal to the extBIC for the 
+previous model, then our iterative procedure is finished.  The set of snp in strongest association with the trait is the $s$ snp
+previously identified.
 
-Our linear mixed model for multiple-locus association mapping, given trait data $y^{(n \times 1)}$ which are collected on $n$ individuals, 
-is of the form 
-\begin{equation}
-y = X \tau + u_g + e
-\end{equation}
-where $X^{(n \times p)}$ is a known design matrix of full rank, $\tau^{(p \times 1)}$ is a vector of fixed effects parameters, $u_g^{(n \times 1)}$ is a vector 
-of genetic effects, and $e^{(n \times 1)}$ is a vector of residuals whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \times n)})$.
-
-Now, let $M^{(n \times L)}$ be a matrix of genotypes collected from $L$ snp.  The 
-matrix contains the values -1, 0, and 1, corresponding to the snp genotypes AA, AB, and BB, respectively.  It is assumed that 
-missing genotypes have been imputed.
-
-Suppose $s$ SNP have already been found.
-
-$S={ S_1, S_2, \ldots, S_s}$ be the set of column indexes
-
-\bibliographystyle{nature}
+\bibliographystyle{naturemag}
 \bibliography{bibcitelibrary}

### 2017-10-20 AM+draft.tex
211017  Methods

@@
-\documentclass[12pt]{article}
-\usepackage{graphicx}
+\documentclass{nature}
+\bibliographystyle{naturemag}
 
-\linespread{1.5}
-\addtolength{\oddsidemargin}{-2.5cm}
+%%\usepackage{graphicx}
+
+%%\linespread{1.5}
+%%\addtolength{\oddsidemargin}{-2.5cm}
 %\addtolength{\evensidemargin}{-3cm}
-\addtolength{\textwidth}{2.5cm}
+%%\addtolength{\textwidth}{2.5cm}
 \begin{document}
@@
 \title{Making multiple-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George and Joshua Bowden}
+\author{Andrew W. George$^1$ and Joshua Bowden$^2$}
 
@@
 
-\abstract{Since the earliest of genome-wide association studies, a key shortcoming in how their data have been analysed has persisted.
+\begin{affiliations}
+\item Data61, CSIRO, Brisbane, Australia.
+\item IM \&T, CSIRO, Brisbane, Australia.
+\end{affiliations}
+
+\begin{abstract}
+Since the earliest of genome-wide association studies, a key shortcoming in how their data have been analysed has persisted.
 The strength of association between a marker locus and trait is measured for each locus separately, on a locus-by-locus basis.
@@
 AMplus is freely available as a fully documented R package. 
-}
-
-\section{Introduction}
+\end{abstract}
 
@@
 Let $M^{(n \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome.  The columns of this matrix are most commonly placed in map ordered but 
-this is not a requirement. 
+from $L$ loci that span the genome.  It is common for the columns of this matrix to be in map order but this is not a requirement. 
 The vector $m_j^{(n \; \times \; 1)}$ contains the genotypes, from the $n$ individuals,  for the $j$th snp. 
@@
 $a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
-The sum of terms on the left hand side are fixed effects.  The other term is a random effect.  The sum of terms accounts 
+The terms in the summation on the left hand side are fixed effects.  The other term is a random effect.  The terms 
+in the summation account 
 for the additive effects of genes that are in linkage disequilibrium with the selected snp. The other term models 
-snp-trait associations along the entire genome, except for those snp that have already been selected. This is a very simple genetic model but it 
+snp-trait associations along the entire genome, except for those snp that have already been selected. 
+This is a simple genetic model but it 
 is effective for discovering snp-trait associations. It also reduces the computational cost of our model building procedure.
@@
 We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected snp.  We then fit this new model to the data
-via maximum likelihood and calculate its extended Bayesian information criteria \cite{chen2008extended}.
+via maximum likelihood and calculate its extended Bayesian information criteria (ExtBIC) \cite{chen2008extended}.
+The ExtBIC is a model selection measure that takes into account the number of unknown parameters and the complexity
+of the model space.
+It is especially well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
+If this new model has a larger ExtBIC than the current model, then the $s+1$th selected snp is added to
+the current model and the above process is repeated. If this new model has a smaller ExtBIC than the current model, then the 
+model building process is complete. The set of snp in strongest association with the trait is the $s$ snp previously identified.
 
-We reform (\ref{eq2}) where 
-$S$ contains $s+1$ selected snp. We then fit (\ref{eq1}) and (\ref{eq2}) to the data via maximum likelihood and calculate $extBIC$.
-We found this model selection criterium to perform better for association mapping problems than standard model selection 
-criteria such as BIC and IAC, as was also found by XXXXX.  If the extBIC value is less than the extBIC for the previous model, 
-then we add the $s+1$th snp  to the model and repeat the process. If  the extBIC value is greater than or equal to the extBIC for the 
-previous model, then our iterative procedure is finished.  The set of snp in strongest association with the trait is the $s$ snp
-previously identified.
+The Eagle approach has its foundations in the work of XXXX and XXXXX. Here, the authors developed a more 
+general approach and have used it for mapping QTL in linkage mapping studies in plants (refs).
 
 \bibliographystyle{naturemag}
-\bibliography{bibcitelibrary}
+\bibliography{biblibrary}
+
+\begin{addendum}
+ \item Put acknowledgements here.
+ \item[Competing Interests] The authors declare that they have no
+competing financial interests.
+ \item[Correspondence] Correspondence and requests for materials
+should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
+\end{addendum}
 

### 2017-10-24 AM+draft.tex
25Oct17 Methods - Dimension reduction

@@
 
-%%\usepackage{graphicx}
-
-%%\linespread{1.5}
-%%\addtolength{\oddsidemargin}{-2.5cm}
-%\addtolength{\evensidemargin}{-3cm}
-%%\addtolength{\textwidth}{2.5cm}
 \begin{document}
@@
 of complex traits in mice. The construction of the population and collection of data are described in detail in 
-\cite{Nicod:2016aa}.  Briefly, the population consists of 2073 adult outbred mice from Charles River Laboratories. Phenotypic data
+\cite{nicod2016genome}.  Briefly, the population consists of 2073 adult outbred mice from Charles River Laboratories. Phenotypic data
 were obtained from 200 traits. These traits fell into the categories of behaviour, tissue, and physiology. Genotype 
@@
 The Eagle approach for multiple-locus association mapping is based on building a linear mixed model, iteratively, via forward 
-selection.  Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
+selection.  It is a simplification of a whole-genome method for mapping quantitative trait loci in classic linkage studies \cite{verbyla2007analysis}.
+Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
 associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
-
 To perform the $s+1$th  iteration of 
 our model building procedure, we do the following.
+
 First, we fit the current model to the data. 
@@
 
-The Eagle approach has its foundations in the work of XXXX and XXXXX. Here, the authors developed a more 
-general approach and have used it for mapping QTL in linkage mapping studies in plants (refs).
+\subsubsection{Reducing the dimension of the model}
 
-\bibliographystyle{naturemag}
+In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
+The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
+$L$, the number of snp, can be extremely large.  An alternative model is given by 
+Verbyla. They show how to reformulate (\ref{eq2}) as 
+.....
+where ..... and XXX is calculated via s. v. d.  Although it may not be obvious, the two models are equivalent, 
+having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
+model \ref{eq2}, in most modern studies,  is much less because the sample size $n$ is 
+often much smaller than the number of snp $L$.
+
+Verbyla go on to show how to recover $\widetilde{a}$ from estimates from model (3) with 
+\begin{equation}
+\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
+\end{equation}
+and its variance matrix with 
+\begin{equation}
+\label{eq4}
+\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
+\end{equation}
+These values are needed to calculate the score statistic $t_j^2$ for identifying the snp in strongest association with the trait. 
+Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation. 
+of (\ref{eq4}).
+
+Fortunately
+
+Since 
+$\widetilde{a}$ and $\textrm{var} (\widetilde{a}) $ are needed  to calculate $t_j^2$, as described in 
+Verbyla \cite{verbyla2012rwgaim, verbyla2014whole}, 
+\begin{equation}
+\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
+\end{equation}
+with variance matrix
+\begin{equation}
+\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
+\end{equation}
+Fortunately, only the diagonal elements of the variance matrix are needed which simplifies its calculation.
+
+It is not uncommon in genome wide association studies for the number of 
+snp ($L$) to be many orders of magnitude larger than the sample size ($n$). 
+As a result, estimating $a$ in (\ref{eq2}) can be demanding, computationally. 
+We avoid having to estimate $a$ directly by instead reformulating (\ref{eq2}) as
+the dimension reduced model 
+\begin{equation}
+\label{eq3}
+u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + (M_{-S} M_{-S}^T)^{1/2} a^*_{-S}
+\end{equation}
+where $a^*$ is a $n \times 1$ vector of random effects that follows $N(0, \sigma_a^2 I^{(n \; \times \;  n)})$, and 
+$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via single value decomposition (ref). 
+Models (\ref{eq2})  and (\ref{eq3}) are equivalent models 
+\cite{verbyla2012rwgaim, verbyla2014whole}. They have the same variance structure, 
+despite the random effect in (\ref{eq3}) being only of size $n$.  This allows us to use (\ref{eq3}) in place of 
+(\ref{eq2}) in the above described Eagle approach, affording us considerable computational benefits. Since 
+$\widetilde{a}$ and $\textrm{var} (\widetilde{a}) $ are needed  to calculate $t_j^2$, as described in 
+Verbyla \cite{verbyla2012rwgaim, verbyla2014whole}, 
+\begin{equation}
+\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
+\end{equation}
+with variance matrix
+\begin{equation}
+\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
+\end{equation}
+Fortunately, only the diagonal elements of the variance matrix are needed which simplifies its calculation.
+
+%%\bibliographystyle{naturemag}
 \bibliography{biblibrary}

### 2017-10-24 AM+draft.tex
25Oct2017 Finished first draft of Eagle approach - methods

--- 2017-10-20 AM+draft.tex
@@
 
-The Eagle approach has its foundations in the work of XXXX and XXXXX. Here, the authors developed a more 
-general approach and have used it for mapping QTL in linkage mapping studies in plants (refs).
+\subsubsection{Reducing the dimension of the model}
+In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
+The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
+$L$, the number of snp, can be extremely large.  An alternative model is given by 
+Verbyla. 
+They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
+\begin{equation}
+\label{eq3}
+u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + (M_{-S} M_{-S}^T)^{1/2} a^*_{-S}
+\end{equation}
+where $a^* \sim N(0, \sigma_a^2 I^{(n \; \times \;  n)})$, and 
+$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via single value decomposition (ref).  
+Although it may not be obvious, the two models are equivalent, 
+having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
+model (\ref{eq2}) is much less due to the random term having only $n$ effects needing estimating.
 
-\bibliographystyle{naturemag}
+Verbyla go on to show how to recover $\widetilde{a}$ from estimates from model (3) with 
+\begin{equation}
+\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
+\end{equation}
+where its variance matrix is
+\begin{equation}
+\label{eq4}
+\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
+\end{equation}
+These values are needed in order to calculate the score statistic $t_j^2$ for identifying the snp in strongest association with the trait. 
+Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation. 
+of (\ref{eq4}).
+
+%%\bibliographystyle{naturemag}
 \bibliography{biblibrary}

### 2017-10-25 AM+draft.tex
26Oct2017 Working on methods - comparison methods

@@
 
+\subsection{Comparison Methods}
+Overview - 
+  * Eagle compared to six other association mapping methods.  
+  * 4 of these methods are multiple-locus methods
+  * 2 of these methods are single-locus methods. 
+  * methods were chosen because they were specifically designed for association mapping and have been implemented in computer
+    software. 
+  * Core features summarized in Table X. 
+  * For each method, a summary of its underlying statistical theory is presented below.
+
+Multiple locus methods
+
+\subsubsection{BigRR}
+BigRR is an R package that implements generalised ridge regression for
+the analysis of genetic data (ref). Generalised ridge regression is an
+extension of
+ridge regression. It is based on linear models and is a regularisation
+method.  When applied to the analysis of snp data, unlike ridge
+regression
+that shrinks snp effects by a constant amount, generalised ridge
+regression allows the amount of shrinkage to vary across snp.
+The challenge though is how to turn the size estimates of the snp
+Effects into measures of significance. Xxx suggest permutation but we
+instead
+rely on stability selection (described below).
+
+  *  \cite{shen2013novel}
+
+Single locus methods
+
 %%\bibliographystyle{naturemag}

### 2017-10-27 AM+draft.tex
27Oct17 Methods - about to work on Lmm-Lasso section

--- 2017-10-24 AM+draft.tex
@@
 $L$, the number of snp, can be extremely large.  An alternative model is given by 
-Verbyla. 
+Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
 They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
@@
 
-Verbyla go on to show how to recover $\widetilde{a}$ from estimates from model (3) with 
+Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model (3) with 
 \begin{equation}
@@
 of (\ref{eq4}).
+
+\subsection{Comparison Packages and Methods}
+
+The performance of the Eagle package is compared against seven other genome-wide association packages.
+Five of these packages, BigRR, glmnet, FaST-LASSO, MMLM, r2VIM implement methods that have been applied to the multiple locus analysis of data from genome-wide association studies. 
+The other two packages, GEMMA, and FaST-LMM implement single-locus association mapping methods. 
+The packages 
+differ in their features (summarized in Table X).  The packages also differ in their underlying methods (summarized below).
+
+\subsubsection{Multiple locus methods}
+
+\textbf{BigRR:} The BigRR package implements generalised ridge regression  for
+the analysis of genetic data \cite{shen2013novel}.  It is a regularisation method in that it produces regression solutions 
+that have higher bias than ordinary least squares but reduced variance for improved prediction accuracy. The underlying 
+model can be formulated as  $y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of genetic effects, and each genetic 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.  Here, the 
+amount of shrinkage is allowed to vary across the genetic (snp) effects which differs to 
+standard ridge regression where the shrinkage is constant. 
+Generalised ridge regression is apt at estimating the effect sizes of snp on a genome-wide scale but  obtaining their 
+statistical significance is not straightforward. This makes using the BigRR package for association mapping a challenge. 
+Shen \cite{shen2013novel} suggest permutation as a way of identifying the genomic regions of interest 
+but we instead employ stability selection (described below).
+
+\textbf{glmnet:}  The glmnet package is a general purpose statistical package that implements elastic net, a regularisation 
+method for generalised linear models  \cite{zou2005regularization}.  The model used for association mapping is 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the regression coefficient for the $l$th snp. 
+It's penalty function for regularisation is a mixture of $\ell_1$ (lasso) and $\ell_2$ (ridge regression).  This allows elastic net to avoid some of the
+limitations of lasso (ref) while still being able to perform shrinkage and variable selection, simultaneously.  The glmnet package 
+implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. However, when 
+applied to the analysis of data from genome-wide association studies we found this yielded findings containing a high proportion of 
+false positives.  We instead fixed the shrinkage parameter to a suitable value 
+and employed stability selection to estimate the significance of the snp regression coefficients, empirically (see below).
+
+  MMLM
+
+  r2VIM
+
+  LMM-lasso
+
+Single locus methods
+
+gemma
+
+Fast-LMM
 

### 2017-10-31 AM+draft.tex
31Oct Still working on methods - glmnet

--- 2017-10-24 AM+draft.tex
@@
 of (\ref{eq4}).
+
+\subsection{Comparison Methods}
+
+We compared the performance of Eagle against seven other association mapping applications. 
+Five of these applications, BigRR, glmnet, FaST-LASSO, MMLM, r2VIM implement methods that have been applied to the multiple locus analysis of data from genome-wide association studies. 
+The other two applications, GEMMA, and FaST-LMM implement single-locus association mapping methods. 
+The packages 
+differ in their features (summarized in Table X).  The packages also differ in their underlying methods (summarized below).
+
+\subsubsection{Multiple locus methods}
+
+\textbf{BigRR:}  In the BigRR package,  generalised ridge regression is implemented  for
+the analysis of genetic data \cite{shen2013novel}.  Generalised ridge regression 
+ is a regularisation method in that it produces regression solutions 
+that have higher bias than ordinary least squares but reduced variance for improved prediction accuracy. The underlying 
+model is  $y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of genetic effects, and each genetic 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.  The 
+amount of shrinkage is allowed to vary across the genetic (snp) effects. This differs to 
+standard ridge regression where the shrinkage is constant. 
+Generalised ridge regression is apt at estimating the effect sizes of snp on a genome-wide scale but  obtaining their 
+statistical significance analytically is difficult. Shen \cite{shen2013novel} suggested permutation as a way of calculating significance 
+empirically. We instead opted for stability selection (see below of details).
+
+\textbf{glmnet:}  Glmnet is an R package where elastic net regression is implemented. Elastic net is a regularisation method that 
+has been applied to generalised linear models \cite{zou2005regularization}.
+
+Glmnet is a general purpose R package for statistical analysis. It implements elastic net regression, 
+a regularisation method that has been applied to generalised linear models \cite{zou2005regularization}.   
+Glmnet was not specifically built for association mapping. However, we included it amongst the applications to benchmark 
+Eagle against because of the interest surrounding its use for analysing data from genetic
+studies \cite{cho2010joint,huang2015empirical}.  The regression model we used for analysis is 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
+Elastic net allows us to obtain parameter estimates from this model via a penalty function that is a mixture of 
+$\ell_1$ (lasso) and $\ell_2$ (ridge regression).
+
+Although glmnet has not been specifically built for association mapping, interest surrounds its use for the analysis of data from 
+an association study.  \cite{cho2010joint,huang2015empirical}. 
+The analysis model for association mapping 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the regression coefficient for the $l$th snp. 
+It's penalty function for regularisation is a mixture of $\ell_1$ (lasso) and $\ell_2$ (ridge regression).  
+This allows elastic net to avoid some of the
+limitations of lasso (ref) while still being able to perform shrinkage and variable selection, simultaneously.  The glmnet package 
+implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. However, 
+from exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
+We instead fixed the shrinkage parameter to a suitable value 
+and employed stability selection to estimate the significance of the snp regression coefficients, empirically (see below).
+
+  MMLM
+
+  r2VIM
+
+  LMM-lasso
+
+Single locus methods
+
+gemma
+
+Fast-LMM
 

### 2017-10-31 AM+draft.tex
31Oct17 end of day updates

@@
 
-Glmnet is a general purpose R package for statistical analysis. It implements elastic net regression, 
-a regularisation method that has been applied to generalised linear models \cite{zou2005regularization}.   
-Glmnet was not specifically built for association mapping. However, we included it amongst the applications to benchmark 
+Glmnet is a general purpose R package for regression analysis. It implements elastic net, 
+a regularisation method where the penalty function 
+is a mixture of 
+$\ell_1$ (lasso) and $\ell_2$ (ridge regression) normalisation  \cite{zou2005regularization}. 
+This allows elastic net to avoid the limitations of lasso (ref) 
+while still being able to perform shrinkage and variable selection, simultaneously. 
+Glmnet was not specifically built for association mapping. 
+However, it is included in our list of comparison applications due to its solid performance for genomic prediction (refs).
+The model upon which our analyses are based is 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
+The glmnet package 
+implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. 
+However, from exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
+We instead fixed the shrinkage parameter to a suitable value 
+and employed stability selection to estimate the significance of the snp coefficients, empirically (see below).
+
+we included it amongst the applications to benchmark 
 Eagle against because of the interest surrounding its use for analysing data from genetic
@@
 $\ell_1$ (lasso) and $\ell_2$ (ridge regression). 
+The glmnet package 
+implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. 
+However, from exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
+We instead fixed the shrinkage parameter to a suitable value 
+and employed stability selection to estimate the significance of the snp coefficients, empirically (see below).
 

### 2017-11-01 AM+draft.tex
1Nov17 morning commit - methods

@@
 
-\textbf{glmnet:}  Glmnet is an R package where elastic net regression is implemented. Elastic net is a regularisation method that 
-has been applied to generalised linear models \cite{zou2005regularization}.
-
-Glmnet is a general purpose R package for regression analysis. It implements elastic net, 
+\textbf{glmnet:}  Glmnet is a general purpose R package for regression analysis. It implements elastic net, 
 a regularisation method where the penalty function 

### 2017-11-02 AM+draft.tex
2Nov2017 Methods under development - comparison methods

--- 2017-10-31 AM+draft.tex
@@
 
-\textbf{glmnet:}  Glmnet is an R package where elastic net regression is implemented. Elastic net is a regularisation method that 
-has been applied to generalised linear models \cite{zou2005regularization}.
+\textbf{glmnet:}  Glmnet is a general purpose R package for regression analysis  \cite{Friedman2010glmnet}. 
+The model used for analysis is 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
+All snp effects are modelled, simultaneously. 
+Glmnet implements elastic net for the estimation of the parameters. Elastic net is a regularisation method where the 
+penalty function is a mixture of 
+$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
+This allows elastic net to avoid the limitations of lasso 
+while still being able to perform shrinkage and variable selection, simultaneously \cite{zou2005regularization}. 
+Glmnet was not specifically built for association mapping. 
+However, it is included in our list of comparison applications due to its solid performance for genomic prediction 
+\cite{heslot2012genomic,ogutu2012genomic,}.
+The glmnet package 
+implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. 
+From exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
+We instead fixed the shrinkage parameter to a suitable value 
+and employed stability selection to estimate the significance of the snp coefficients, empirically (see below).
 
-Glmnet is a general purpose R package for statistical analysis. It implements elastic net regression, 
-a regularisation method that has been applied to generalised linear models \cite{zou2005regularization}.   
-Glmnet was not specifically built for association mapping. However, we included it amongst the applications to benchmark 
-Eagle against because of the interest surrounding its use for analysing data from genetic
-studies \cite{cho2010joint,huang2015empirical}.  The regression model we used for analysis is 
-$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
-Elastic net allows us to obtain parameter estimates from this model via a penalty function that is a mixture of 
-$\ell_1$ (lasso) and $\ell_2$ (ridge regression).
+\textbf{LMM-Lasso:} LMM-Lasso is a stand-alone computer program for association mapping within a linear 
+mixed model framework \cite{rakitsch2013lasso}. The (full) linear mixed model used for analysis is 
+$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
+$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
+All snp effects are modelled, simultaneously. However, a limitation of this model is its inability to accommodate 
+additional fixed effects (there is no $X \beta$ term). 
+The parameters are estimated via lasso \cite{Tibshirani1996lasso}. Lasso is a regularisation method with the capacity to perform parameter estimation and variable selection, simultaneously.  
+Lasso is computationally intensive, especially when applied to linear mixed models. To avoid this problem, a two-step 
+estimation procedure is adopted.  In the first step, the above model but without any snp effects is fitted via 
+maximum likelihood.
+In the second step, the unknown variance components in the full model are replaced by their estimates that were calculated in the first step.
+This turns the full model from a linear mixed model into a simpler linear model.
+Here, lasso is far more tractable, computationally.
+To assess the significance of the snp coefficients, as suggested \cite{rakitsch2013lasso}, we implemented stability selection (see below).
 
-Although glmnet has not been specifically built for association mapping, interest surrounds its use for the analysis of data from 
-an association study.  \cite{cho2010joint,huang2015empirical}. 
-The analysis model for association mapping 
-$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the regression coefficient for the $l$th snp. 
-It's penalty function for regularisation is a mixture of $\ell_1$ (lasso) and $\ell_2$ (ridge regression).  
-This allows elastic net to avoid some of the
-limitations of lasso (ref) while still being able to perform shrinkage and variable selection, simultaneously.  The glmnet package 
-implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. However, 
-from exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
-We instead fixed the shrinkage parameter to a suitable value 
-and employed stability selection to estimate the significance of the snp regression coefficients, empirically (see below).
-
-  MMLM
+\textbf{MLMM:}  Of the multiple-locus applications considered in this paper, 
+MLMM bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
+association mapping by building the "best" linear mixed model.  The model used for analysis 
+is the same as (\ref{eq1}) except that the genetic effect is 
+\begin{equation}
+\label{eqmlmm}
+u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + g
+\end{equation}
+where 
+$g^{(n ; \times 1)}$ is a vector of random polygenic effects with distribution 
+$g \sim N(0, \sigma^2_g K)$. The model is built iteratively, via forward-backward variable selection.  In Eagle, at each iteration, 
+the snp in strongest association with the trait is found by calculating the score statistic $t_j^2$.  In MLMM, at each iteration, 
+a separate linear mixed model analysis is performed for each snp not already selected. The snp yielding the most 
+significant result (lowest p-value) is selected for inclusion in the fixed effects part of (\ref{eqmlmm}).  
+Two selection 
+criteria are available for the model building process: 
+the EBIC and the multiple-Bonferroni criterion. We used only the EBIC.
 
   r2VIM
-
-  LMM-lasso
 

### 2017-11-02 AM+draft.tex
02Nov17 end of day updates - methods

@@
 Early gwas followed a case-control design.
-Association mapping methods were no more complicated than contingency table tests or regression.
+Association mapping methods were no more complicated than contingency table tests or simple
+linear regression.
 These designs though had a tendency to yield spurious findings if there was unrecognised population stratification.
@@
 However, it is included in our list of comparison applications due to its solid performance for genomic prediction 
-\cite{heslot2012genomic,ogutu2012genomic,}.
+\cite{heslot2012genomic,ogutu2012genomic}.
 The glmnet package 
@@
 
-  r2VIM
+\textbf{r2VIM:} r2VIM is an R package, specifically designed for genome-wide association mapping \cite{szymczak2016r2vim}.
+It implements a modified form of random forests. When random forests are applied to association study data in a standard way, 
+the result is a ranking of the snp according to their importance values. The difficulty is in knowing how many of the top ranked 
+snp are associated, significantly, with the trait. In r2VIM, multiple separate random forest analyses of the data are performed.
 
-Single locus methods
+In r2VIM, a strategy is implemented for identifying the snp in strongest association with the trait. 
+Here, multiple random forest analyses of the data are performed. For
+each analysis, a relative importance score is calculated for each snp based on the minimum observed importance value.  Only 
+those snp with large relative importance scores are determined to be in significant association with the trait.
+
+\subsubsection{Single locus methods}
 

### 2017-11-06 AM+draft.tex
06Nov2017

--- 2017-11-02 AM+draft.tex
@@
 
-  r2VIM
+\textbf{r2VIM:} R2RIM is an R package specifically designed for association mapping \cite{szymczak2016r2vim}. 
+It implements random forests and a new way of measuring the importance of a snp.
+In random forests, the worth of a predictor is measured, empirically, by calculating its importance score.
+It is from these importance scores that snp can be ordered in terms of their strength of association with a trait.
+The challenge though is in knowing what proportion of the
+highest ordered snp are in true association. R2VIM addresses this by calculating a relative importance score for a snp. Briefly,
+$m$ random forest analyses of the data are performed. Each analysis is reliant upon a different random number seed. 
+From an analysis, a relative importance score for a snp is obtained by taking the ratio of its importance score to the absolute 
+minimum of the importance scores across all the snp.  Those 
+snp with high relative scores across all $m$ analyses are deemed to be of interest \cite{szymczak2016r2vim}.
 
-Single locus methods
+\subsubsection{Single locus methods}
 

### 2017-11-06 AM+draft.tex
07 Nov, 2017

@@
 
-gemma
+\textbf{FaST-LMM:}  FaST-LMM is a highly efficient stand along computer program for association mapping.
+It is based on a standard single-locus model (refs).
+Its computational efficiencies come through rotating the data with spectral decomposition.
+This
+transforms the linear mixed 
+model into a much simpler linear model.  In doing so, FaST-LMM inherits the favourable property of having its analyses 
+scale linearly, in both run time and memory, with sample size ($n$).
+
+\textbf{gemma}
 

### 2017-11-08 AM+draft.tex
08Nov17 Morning - methods

--- 2017-11-06 AM+draft.tex
@@
 
+\subsubsection{Multiple locus methods}
+
+We compared the performance of Eagle against five other multiple locus association mapping methods: BigRR, 
+glmnet, FaST-LASSO, MMLM, and r2VIM.  These five methods were selected because of their demonstrated value 
+for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
+either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
+features is given in Table X.
+
+BigRR, FaST-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
+are well suited to solving problems where the number of predictors (p) is far greater than the number of samples (n).  It 
+should come as no surprise then that multiple locus association mapping are dominated by regularization-based approaches.
+Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
+bias against reduced variance.  BigRR, an R package, implements gen
+
+BigRR is an R package that implements
+
+$------$
+
 We compared the performance of Eagle against seven other association mapping applications. 
@@
 
-gemma
+\textbf{FaST-LMM:}  FaST-LMM is a highly efficient stand along computer program for association 
+mapping \cite{lippert2011fast}. It is based on a standard single-locus model 
+\cite{yu2006unified,zhao2007arabidopsis}.
+Its computational efficiencies come through factorising $K$ with spectral decomposition so that $K=U (S + \delta I) U^T$ where $U$ contains the eigenvectors and $\delta$ is the ratio of the genetic variance to the residual variance.
+By rotating the data with $U$, the log likelihood of the linear mixed model can be expressed as the log likelihood for a simple 
+linear model.  This greatly reduces the computational cost of performing lmm-based  association mapping. With FaST-LMM, 
+analyses scale linearly with sample size $n$, both in terms of run-time and memory usage.
+
+\textbf{gemma}
 

### 2017-11-08 AM+draft.tex
8nov17 night

@@
 
-BigRR, FaST-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
+Of the five multiple locus methods considered, BigRR, LMM-Lasso, and glmnet,  implement regularization methods for parameter
+estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
+number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
+genomic prediction (refs) but also association mapping (refs). BigRR is an R package that implements generalized ridge regression
+which is an improvement on standard ridge regression. In standard ridge regression, coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
+a more realistic model to be specified. The model used in BigRR is 
+$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
+
+BigRR, LMM-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
 are well suited to solving problems where the number of predictors (p) is far greater than the number of samples (n).  It 
-should come as no surprise then that multiple locus association mapping are dominated by regularization-based approaches.
+is not surprising that multiple locus association mapping are dominated by regularization-based approaches. Regularisation-based 
+approaches model the effect of all the snp on a trait simultaneously. 
 Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
-bias against reduced variance.  BigRR, an R package, implements gen
+bias against reduced variance.  BigRR, an R package, implements generalised ridge regression. Here, the underlying model 
+can be formulated as a linear mixed model that is
+ $y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$. The size of the random 
+snp effect is a measure of the strength of association between the snp and trait. The parameters are estimated via an
+algorithm that makes clever use of the equivalence between the above model and animal models \ref{}.
+
+Lmm-Lasso  \cite{rakitsch2013lasso}, a stand-alone computer program, implements an estimation 
+procedure based on lasso for the (full) model 
+$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
+$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
+Lasso is popular due to its capacity to perform parameter estimation and variable selection, simultaneously. Lasso though is demanding, computationally. To reduce this problem, a two-step 
+estimation procedure is adopted.
+First, the above model without any snp effects is fitted.
+Second, the unknown variance components in the full model are replaced by their estimates that were calculated in the first step.
+This turns the model into a simpler linear model for which lasso is far more tractable, computationally.
+
+Glmnet, a general purpose R package, uses elastic net for parameter estimation.
+
+makes use of elastic net. 
+The model used for analysis is 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
+Glmnet implements elastic net for the estimation of the parameters. Elastic net is a regularisation method where the 
+penalty function is a mixture of 
+$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
+This allows elastic net to avoid the limitations of lasso 
+while still being able to perform shrinkage and variable selection, simultaneously \cite{zou2005regularization}.
+
+$------$
+
+Here, 
+the 
+amount of shrinkage is allowed to vary across the predictors (snp effects) where in standard ridge regression, the amount of 
+shrinkage is constant. FaST-LMM, a stand-alone computer program, implements a two-step lasso approach.
 
 BigRR is an R package that implements
-
-$------$
 

### 2017-11-08 AM+draft.tex
09Nov17 morning - about to work on methods

@@
 
-We compared the performance of Eagle against five other multiple locus association mapping methods: BigRR, 
-glmnet, FaST-LASSO, MMLM, and r2VIM.  These five methods were selected because of their demonstrated value 
+We compared the computational and statistical performance of Eagle against five other multiple locus applications for 
+association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
+BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
+the development of multiple locus approaches.  They give opportunity for the effect of all the snp to be modelled simultaneously. 
+By introducing a penalty function that balances bias and variance, regularisation methods can 
+handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
+MLMM is similar in philosophy to Eagle in that it treats association mapping as a model building problem. Here, a linear mixed 
+model is built through a forward-backward selected strategy.  
+R2VIM is different to the others in that it implements 
+random forests for association mapping. 
+Random forests is a machine learning approach that forms an ensemble of decision trees for regression and 
+classification. 
+These five methods were selected because of their demonstrated value 
 for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
 either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
-features is given in Table X.
+features is given in Table X. Each method is now described in a little more detail.
 
-Of the five multiple locus methods considered, BigRR, LMM-Lasso, and glmnet,  implement regularization methods for parameter
+Regression based regularization methods 
+are well suited to solving problems where the number of predictors (p) is much greater than the number of samples (as in
+genome-wide association studies).  Unsurprisingly,  regularization methods have dominated the development of 
+multiple locus assciation mapping approaches.  Regularisation-based 
+approaches model the effect of all the snp on a trait simultaneously. 
+Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
+bias against reduced variance. The fourth application implements a forward-backward selection approach for building a linear 
+mixed model. The fifth application uses random forests. 
+These five methods were selected because of their demonstrated value 
+for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
+either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
+features is given in Table X. Each method is now described briefly.
+
+BigRR is an R package that implements generalized ridge regression. Generalized ridge regression is 
+which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
+a more realistic model to be specified. The model used in BigRR is 
+$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
+
+BigRR, LMM-Lasso, and glmnet all implement regularization methods for parameter
 estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
 number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
-genomic prediction (refs) but also association mapping (refs). BigRR is an R package that implements generalized ridge regression
-which is an improvement on standard ridge regression. In standard ridge regression, coefficients of the predictors are shrunk 
+genomic prediction (refs) but also association mapping (refs). 
+BigRR implements generalized ridge regression, LMM-Lasso implements a two-step lasso approach, and glmnet makes use of 
+elastic net for parameter estimation.
+
+BigRR is an R package that implements generalized ridge regression
+which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
 to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 

### 2017-11-09 AM+draft.tex
09Nov17 11am commit of methods under development

--- 2017-11-08 AM+draft.tex
@@
 
-We compared the performance of Eagle against five other multiple locus association mapping methods: BigRR, 
-glmnet, FaST-LASSO, MMLM, and r2VIM.  These five methods were selected because of their demonstrated value 
+We compared the computational and statistical performance of Eagle against five other multiple locus applications for 
+association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
+BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
+the development of multiple locus approaches.  They give opportunity for the effect of all the snp to be modelled simultaneously. 
+By introducing a penalty function that balances bias and variance, regularisation methods can 
+handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
+MLMM is similar in philosophy to Eagle in that it treats association mapping as a model building problem. Here, a linear mixed 
+model is built through a forward-backward selected strategy.  
+R2VIM is different to the others in that it implements 
+random forests for association mapping. 
+Random forests is a machine learning approach that forms an ensemble of decision trees for regression and 
+classification. 
+These five methods were selected because of their demonstrated value 
 for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
 either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
-features is given in Table X.
+features is given in Table X. Each method is now described in a little more detail.
 
-BigRR, FaST-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
-are well suited to solving problems where the number of predictors (p) is far greater than the number of samples (n).  It 
-should come as no surprise then that multiple locus association mapping are dominated by regularization-based approaches.
-Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
-bias against reduced variance.  BigRR, an R package, implements gen
+BigRR is an R package that implements generalized ridge regression. Generalized ridge regression is 
+ an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
+a more realistic model to be specified. The model used in BigRR is 
+$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
 
-BigRR is an R package that implements
+BigRR is an R package that implements generalized ridge regression
+which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
+a more realistic model to be specified. The model used in BigRR is 
+$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
+
+BigRR, LMM-Lasso, and glmnet all implement regularization methods for parameter
+estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
+number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
+genomic prediction (refs) but also association mapping (refs). 
+BigRR implements generalized ridge regression, LMM-Lasso implements a two-step lasso approach, and glmnet makes use of 
+elastic net for parameter estimation.
+
+BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
+In ridge regression, the coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary, giving opportunity for 
+a more realistic model to be specified. The model used in BigRR can be formulated as a linear mixed model that is 
+$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
+effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$. 
+The parameters are estimated via a 
+algorithm that makes clever use of the Cholesky decomposition of the genomic relationship matrix, 
+and the equivalence between the above model and animal models \ref{stranden2009animalmodel}. 
+One of the challenges in using BigRR for association mapping is that while it has the capacity to estimate the effect sizes of all the snp simultaneously, 
+obtaining their statistical significance is difficult. 
+Permutation has been suggested as a way of calculating significance 
+empirically \cite{shen2013novel} but we instead opted for stability selection.
+
+Lmm-Lasso  \cite{rakitsch2013lasso} is a stand-alone computer program that implements lasso within a linear mixed model 
+framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
+coefficients of "unimportant" predictors are shrunk to exactly zero.
+The (full) model is 
+$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
+$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
+It is challenging, computationally,  to apply lasso directly to the full model for parameter estimation. Instead, a two step procedure 
+is adopted. In the first step, the above model without any snp effects is fitted. In the second step, 
+the unknown variance components in the full model are replaced by their estimates 
+that were calculated in the first step. Lasso is then applied to this simpler model making lasso more tractable, computationally. 
+As with BigRR,  the statistical significance of the snp effects are not calculated. We use stability selection to obtain the 
+significance empirically.
+
+Glmnet  \cite{Friedman2010glmnet}, unlike BigRR and LMM-Lasso, was not purpose built for 
+association mapping.  It is a general R package for 
+regression analysis. We have included it in our list of applications for comparison with Eagle because the 
+regularization method, elastic net, upon which the package is based has been found to be superior to ridge regression and 
+lasso (ref needed???).  Elastic net uses a penalty function that is a mixture of 
+$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
+This allows elastic net to avoid the limitations of lasso 
+while still being able to perform shrinkage and variable selection, simultaneously. 
+Glmnet can handle a range of different models. For the analyses performed in this paper, the model is 
+$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
+We used stability selection to obtain the significance of the coefficients (snp effects).
 

### 2017-11-09 AM+draft.tex
09 Nov 2017  lunch time commit

@@
 
+Of the multiple-locus applications considered in this paper, 
+MLMM \cite{segura2012efficient} bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
+association mapping by building the "best" linear mixed model.  The model used for analysis 
+is the same as (\ref{eq1}) except that the genetic effect is 
+\begin{equation}
+\label{eqmlmm}
+u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + g
+\end{equation}
+where 
+$g^{(n ; \times 1)}$ is a vector of random polygenic effects with distribution 
+$g \sim N(0, \sigma^2_g K)$. The model is built iteratively, via forward-backward variable selection.  In Eagle, at each iteration, 
+the snp in strongest association with the trait is found by calculating the score statistic $t_j^2$.  In MLMM, at each iteration, 
+a separate linear mixed model analysis is performed for each snp not already selected. The snp yielding the most 
+significant result (lowest p-value) is selected for inclusion in the fixed effects part of (\ref{eqmlmm}).  
+Two selection 
+criteria are available for the model building process: 
+the EBIC and the multiple-Bonferroni criterion. We used only the EBIC.
+
+\textbf{r2VIM:} R2VIM is an R package specifically designed for association mapping \cite{szymczak2016r2vim}. 
+It implements random forests and a new way of measuring the importance of a snp.
+In random forests, the worth of a predictor is measured, empirically, by calculating its importance score.
+It is from these importance scores that snp can be ordered in terms of their strength of association with a trait.
+The challenge though is in knowing what proportion of the
+highest ordered snp are in true association. R2VIM addresses this by calculating a relative importance score for a snp. Briefly,
+$m$ random forest analyses of the data are performed. Each analysis is reliant upon a different random number seed. 
+From an analysis, a relative importance score for a snp is obtained by taking the ratio of its importance score to the absolute 
+minimum of the importance scores across all the snp.  Those 
+snp with high relative scores across all $m$ analyses are deemed to be of interest \cite{szymczak2016r2vim}.
+Applying r2VIM to the analysis of GWAS data is not without its challenges.  First, it is not possible to force every 
+decision tree to have important (fixed) effects without a change to the internal workings of the r2VIM package. 
+So we adopted a two step 
+strategy. In the first step, the fixed effects are regressed on to the trait and the residuals recorded.  In the second step, 
+r2VIM is employed were the trait data are the residuals obtained from the first step. This is not ideal (ref to problems with two
+step approaches).  Second, there is no relationship between relative importance scores and significance levels. Consequently, 
+setting an appropriate genome-wide threshold is problematic. We avoid this problem by focusing, in the simulation study, 
+only on the relationship between threshold level and its false discovery rate and power.
+
 $------$

### 2017-11-09 AM+draft.tex
10Nov17 Morning commit

@@
 
-BigRR is an R package that implements generalized ridge regression. Generalized ridge regression is 
- an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
-to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
-a more realistic model to be specified. The model used in BigRR is 
-$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
-effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
-
-BigRR is an R package that implements generalized ridge regression
-which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
-to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
-a more realistic model to be specified. The model used in BigRR is 
-$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
-effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
-
-BigRR, LMM-Lasso, and glmnet all implement regularization methods for parameter
-estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
-number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
-genomic prediction (refs) but also association mapping (refs). 
-BigRR implements generalized ridge regression, LMM-Lasso implements a two-step lasso approach, and glmnet makes use of 
-elastic net for parameter estimation.
-
-BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
+\textbf{BigRR:} BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
 In ridge regression, the coefficients of the predictors are shrunk 
@@
 
-Lmm-Lasso  \cite{rakitsch2013lasso} is a stand-alone computer program that implements lasso within a linear mixed model 
+\textbf{LMM-Lasso:} LMM-Lasso  \cite{rakitsch2013lasso} is a stand-alone computer program that implements lasso within a linear mixed model 
 framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
@@
 
-Glmnet  \cite{Friedman2010glmnet}, unlike BigRR and LMM-Lasso, was not purpose built for 
+\textbf{Glmnet:} Glmnet  \cite{Friedman2010glmnet}, unlike BigRR and LMM-Lasso, was not purpose built for 
 association mapping.  It is a general R package for 
@@
 
-Of the multiple-locus applications considered in this paper, 
+\textbf{MLMM:} Of the multiple-locus applications considered in this paper, 
 MLMM \cite{segura2012efficient} bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
@@
 only on the relationship between threshold level and its false discovery rate and power.
+
+\subsubsection{Single locus methods}
+
+We were also interested in comparing Eagle to single locus methods to measure the difference in run times and statistical power.
+Even by limiting our focus to LMM-based methods, there were a number of efficient applications to choose from.
+We decided on GEMMA and FaST-LMM \cite{lippert2011fast}.
+These two applications have the same computational complexity (ref), produce exact instead of approximate results, and are highly efficient computationally.
+They were developed at the same time, independently, but are very similar theoretically.
+Both perform a single spectral decomposition of the relationship matrix K.
+Both use the eigenvector matrix to rotate the data.
+Both reformulate the log likelihood and reml log likelihood into a sum of $n$ terms that are easier to compute.
+The difference lies in their estimation procedure.
+FaST-LMM implement's the Brent algorithm to optimise $\delta$.
+GEMMA instead implement's the Newton Raphson algorithm.
+Newton Raphson is more complicated in that it also requires the first and second derivatives of a function to be calculated.
+However, it is superior in terms of its convergence properties to the Brent algorithm (quadratic verse super-linear).
+Both applications are stand along computer programs, popular, and in common use.
+
+\section{Stability Selection}
+
+\section{Simulation Study}
+
+\section{Implementation}
 
@@
 
-\textbf{gemma}
-
-Fast-LMM
-
 %%\bibliographystyle{naturemag}

### 2017-11-10 AM+draft.tex
10nov17 night

@@
 Even by limiting our focus to LMM-based methods, there were a number of efficient applications to choose from.
-We decided on GEMMA and FaST-LMM \cite{lippert2011fast}.
-These two applications have the same computational complexity (ref), produce exact instead of approximate results, and are highly efficient computationally.
+We decided on GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
+These two applications have the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.
 They were developed at the same time, independently, but are very similar theoretically.
-Both perform a single spectral decomposition of the relationship matrix K.
+Both perform a single spectral decomposition of the relationship matrix $K$.
 Both use the eigenvector matrix to rotate the data.
-Both reformulate the log likelihood and reml log likelihood into a sum of $n$ terms that are easier to compute.
+Both reformulate the log likelihood and
+restricted/residual maximum-likelihood (REML) 
+log likelihood into a sum of $n$ terms that are easier to compute.
 The difference lies in their estimation procedure.
-FaST-LMM implement's the Brent algorithm to optimise $\delta$.
-GEMMA instead implement's the Newton Raphson algorithm.
-Newton Raphson is more complicated in that it also requires the first and second derivatives of a function to be calculated.
-However, it is superior in terms of its convergence properties to the Brent algorithm (quadratic verse super-linear).
-Both applications are stand along computer programs, popular, and in common use.
+FaST-LMM implement's the Brent's algorithm to optimise $\delta$.
+GEMMA instead implement's the Newton-Raphson algorithm.
+Newton-Raphson is more complicated in that it also requires the first and second derivatives of a function to be calculated.
+However, it is superior in terms of its convergence properties to the Brent algorithm.
+Both applications are stand-alone computer programs, popular, and in common use.
 
 \section{Stability Selection}
+When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 
+BigRR has an approximate way of setting each snp's separate regularization parameter (ref). However, the optimal value of 
+$\lambda$ must be found when using LMM-Lasso and glmnet. Also, 
+all three applications yield the effect sizes of the snp across the entire genome but not their statistical significance. To address 
+these issues, we made use of stability selection \cite{meinshausen2010stability}. 
+Stability selection is a resampling strategy.  With stability selection, we were able to avoid having to optimise $\lambda$ while 
+still being able to calculate, empirically, the statistical significance of the snp effects.
+
+We performed stability selection as follows. For LMM-Lasso and glmnet, we performed a preliminary analysis to find an 
+appropriate value for the regularization parameter.  We adjusted $\lambda$ so that they yielded 
+between 10 to 30 snp with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
+but limiting LMM-Lasso and glmnet to only 10 to 30 non-zero effect sizes was straightforward. We then repeatedly subsampled, 
+without replacement, the data. As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$. We 
+analysed the subsample, with $\lambda$ fixed for LMM-Lasso and glmnet.  To calculate, empirically,  the statistical significance of each 
+snp across the genome, we counted the number of times a snp had a non-zero effect size over all the replicates to the total 
+number of replicates (which was 100).
+
+For BigRR, we modified our stability selection procedure slightly. 
+There was no need to find an appropriate value for the regularization 
+parameters as the BigRR package has an internal procedure for their 
+calculation. We draw subsamples as above and analyzed the data with BigRR. However, within an analysis, we ordered 
+the snp according to the absolute size of their snp effects and recorded the top 20 snp. We then measured the significance 
+of the snp across the entire genome as above.
+
+\cite{meinshausen2010stability}
 
 \section{Simulation Study}
-
+Replicates are generated with data perturbation (Zhao et al.
+2007).
+Data perturbation makes use of the phenotypic and genotypic data observed in a real study to create a simu- lated (quantitative) trait.
+It affords us the opportunity to generate replicates whose analysis more closely mirrors the complexities of analyzing real data.
+For computational expediency, our simulation study is based on data observed from the Biloela site in 2006.
+In generating trait data for a replicate, we assume a broad based heritability of 0.6, there is a single major QTL, the polygenic component consists of 30 polygenes where the effect of a polygene is formed from an unmapped SNP which was chosen randomly, and the variance structure of the simulated trait closely follows that of the loaf volume trait.
+The marker data consists of those genotypes collected in the association study from the 3129 SNP across the 21 homologous chromosomes.
+Each replicate has the same marker data.
+It is the simulated trait data that varies across replicates.
+Our simulation study consists of three parts.
+First, we conduct a null study to investigate whether our Monte Carlo sampling approach controls correctly the genome- wide type I error rate.
+We also evaluate the impact of assuming a block diagonal structure for V.
+We do this by
+also calculating the genome-wide type I error rates when the variance matrix for the joint distribution of the test sta- tistic is formed from loci pairings across the entire genome.
+Ten thousand replicates are generated where the QTL has no effect.
+Results are reported for QK-based association mapping implemented as a single-stage and weighted two- stage analysis.
+Second, a power study is performed to investigate if there is a difference in performance between implementing QK-based association mapping as a single-stage analysis versus a two-stage analysis.
+Data are generated under QTL of different sizes and with varying percentages of missing marker data (1.1, 5, 9, 11 %).
+Missing marker genotypes are imputed for a marker locus by coding its SNP geno- types as ?1 and 1, normalizing its marker data, and replac- ing missing genotypes with zero values.
+One thousand replicates are generated for each of the four different per- centages of missing marker data and each differently sized QTL.
+A genome-wide signi cance level of 5 % is assumed.
+Third, a study to evaluate our heuristic procedure for choosing between single-stage and two-stage analysis is performed.
+Data are generated under a QTL of no, mod- erate, and large effect.
+One thousand replicates are gen- erated for each differently sized QTL.
+The performance of our heuristic procedure is measured by calculating the proportion of replicates for which the analysis type is inferred correctly.
+To determine if a single-stage analy- sis is truly needed, it is necessary to analyze fully the genome-wide data via single-stage and weighted two-stage analysis.
+Then, for each marker locus j, perform the test
+| ? log padj ? log qadj| > | ? log padj ? 1.3|, where we have jjj
+assumed a genome-wide signi cance level of 5 %. If true, an incorrect  nding occurs if our heuristic procedure deter- mines that it is valid to implement two-stage analysis.
 \section{Implementation}

### 2017-11-19 AM+draft.tex
19Nov 2017 Sunday comits

@@
 Let $y^{(n \; \times \;1)}$ be a vector containing the quantitative trait data. 
-Let $M^{(n \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome.  It is common for the columns of this matrix to be in map order but this is not a requirement. 
+Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
+from $L$ loci that span the genome on $n_g$ groups/lines/strains.  In this paper, $n$ and $n_g$ are equal. 
+ It is common for the columns of this matrix to be in map order but this is not a requirement. 
 The vector $m_j^{(n \; \times \; 1)}$ contains the genotypes, from the $n$ individuals,  for the $j$th snp. 
@@
 \label{eq1}
-y = X \tau + u_g + e
+y = X \tau + Z u_g + e
 \end{equation}
 where 
-$X^{(n \; \times \; p)}$ is a known design matrix of full rank and may contain data on non-genetic effects such as age and height 
-and/or principal component values to account for population structure (ref). The vector 
-$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters. The vector 
-$u_g^{(n \; \times \; 1)}$ contains the 
+$X^{(n \; \times \; p)}$ and $Z^{( n \; \times \; n_g)}$ are known design matrices with $X$ being of full rank and $Z$ 
+containing zeros and ones that assign the appropriate genetic effect to each observation. 
+The vector 
+$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
+$u_g^{(n_g \; \times \; 1)}$ contains the 
 genetic effects. The vector of residuals is 
@@
 \end{equation}
-where $m_{S_k}^{(n \; \times \; 1)}$ is the vector of genotypes for the $S_k$th snp locus which is the $k$th selected snp, 
-$a_{S_k}$ is the additive effect of the $S_k$th snp locus, $M_{-S}^{(n \; \times \; L-s)}$ is the matrix of  snp genotypes 
+where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the $S_k$th snp locus which is the $k$th selected snp, 
+$a_{S_k}$ is the additive effect of the $S_k$th snp locus, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  snp genotypes 
 with the data for the selected snp in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
@@
 \end{equation}
-where $a^* \sim N(0, \sigma_a^2 I^{(n \; \times \;  n)})$, and 
+where $a^* \sim N(0, \sigma_a^2 I^{(n_g \; \times \;  n_g)})$, and 
 $(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via single value decomposition (ref).  
@@
 \section{Simulation Study}
+
+To explore the computational and statistical performance of Eagle, we conducted a large simulation data.  Genome-wide data 
+were generated under five different scenarios. These scenarios were data where the sample size was X and the number of snp Y (X x Y) ,
+
+We generated data under five different scenarios.
+
+We generated 
+replicates via data perturbation (ref: Zhao et al. 2007).
+
+These are a GWAS of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
+50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
+1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  
+We chose these scenarios to reflect some of the different sized GWAS being performed in animals, plants, and humans. 
+For each scenario, we generated 100 replicates of data. A single replicate consists of snp genotype data and quantitative trait data. 
+We obtained the snp data from the publicly available 1000 genome project (phase 3). The quantitative trait data we generated from the 
+snp data by selecting a set of snp loci, assigning allelic effects to these snp, and aggregating these effects for each individual along with 
+random error. The number of snp selected per replicate follows a Poisson distribution with mean 30.
+ The quantitative trait was generated to have a heritability of 50\%.
+Analyses were performed on a high end desktop computer.
+It had dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM.
+All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
+
 Replicates are generated with data perturbation (Zhao et al.

### 2018-02-05 AM+draft.tex
2Feb18 Large commit

@@
 
+\section{New section written on iPhone added 5 Feb 2018}
+\section{Simulation Study}
+We conduct a large simulation study to examine the performance characteristics of Eagle.
+Genome-wide data are generated under five different scenarios.
+These are a gwas of size 150 individuals and 5000 snp (150 x 5K), 350 individuals and 400000 snp (350 x 400k), 1500 individuals and 50000 snp (1500 x 50k), 200 individuals and .....
+We chose these five scenarios to reflect some of the different sized studies being run in animals, plants, and humans.
+For each scenario, 100 replicates are generated.
+
+To generate a replacate, we used data perturbation (ref). Data perturbation makes use of real genotype data from which a (quantitative) trait is created.
+With data perturbation, we are able to preserve some of the challenges of analysing real data while still having trait data whose genetic architecture is known.
+
+The genotype data comes from the 1000 genome project, version 3 (ref).
+The original data contains X individuals with XX XX snp.
+From these data, according to the specifications of the scenario, the individuals and the snp loci are identified at random.
+What about polymorphic affect numbers?
+Chech).
+
+To generate the quantitative trait data, first the number of snp that are to be in complete with an underlying qtl (x) is drawn from a poison distribution with mean equal to 30.
+Second, x snp loci are select randomly from a scenario?s set of genotype data.
+Third, the snp values are summed across the selected loci for each individual to generate a vector of n genetic values.
+Here, each snp is having, roughly, the same effect on the trait, remembering that the maf of a snp also impacts a snps influence on a trait.
+Fourth, n deviates are drawn from a $N(0,sigma^2_e)$ where sigma_e^2 has been set to yield a trait with a heritability of 0.6.
+(Put in formula).
+Fifth, the quantitative trait data is generate as $y = g + e$.
+In this way, the same genotype data are used, within a scenario, for all replicates.
+Only the quantitative trait data varies across replicates.
+
+\section{Implementation}
+
+Eagle has been implemented as an R package of the same name.
+However, much of the calculation is performed through c++ functions for improved performance.
+Eagle has been specially designed to harness the multithreaded math libraries blast and lapack but it will run on a single processor computer.
+It can also deal with data larger than the memory capacity of a computer.
+Eagle can be run as a standard R package or via the web based GUI.
+The GUI relinquishes the need to know R to use Eagle.
+
+Eagle can be downloaded from CRAN and installed in the usual way.
+A containerised version of Eagle is also available.
+It is built with singularity (ref).
+This container already has Eagle installed on a multi-threaded version of R.
+The container is downloaded as a single file from ccxx
+
+A quick start guide, how to videos, installation instructions, and answers to frequently asked questions are available at the Eagle website (http://...)
+
 $------$
@@
 
+\texbf{GEMMA}
+
+\section{Discussion}
+
 %%\bibliographystyle{naturemag}

### 2018-02-05 AM+/Plots_for_Paper/power_plots.R

@@ -218,6 +218,8 @@ postscript("~/Papers/AM-Paper/powerMultiple.eps", width=10, height=10, fonts=c("
 p
 dev.off()
 
+p1 <- p
+
 ##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 ## power vs fdr for single-locus models
 ##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
@@ -281,9 +283,11 @@ p <- p+ theme(legend.key.width=grid:::unit(1.5,"cm"))
 
 
 
+
 postscript("~/Papers/AM-Paper/powerSingle.eps", width=10, height=10, fonts=c("sans", fonts()),
            horizontal=FALSE)
 p
 dev.off()
 
 
+p2 <- p

### 2018-05-08 AM+draft.tex
08 May 18 Prior to working on results section - tidy up

@@
 
-\title{Making multiple-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George$^1$ and Joshua Bowden$^2$}
+\title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
+\author{Andrew W. George$^1$, Joshua Bowden$^2$, need to sort out the other authors}
 
@@
 \subsection{Mouse Study}
+Still to be written.
 
@@
 
-The Eagle approach for multiple-locus association mapping is based on building a linear mixed model, iteratively, via forward 
-selection.  It is a simplification of a whole-genome method for mapping quantitative trait loci in classic linkage studies \cite{verbyla2007analysis}.
+The Eagle approach for multiple-locus association mapping follows from methods developed for QTL mapping in plant 
+populations \cite{verbyla2007analysis} (and some other WGAIM references). 
+It is based on building a linear mixed model, iteratively, via forward 
+selection.
+
 Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
@@
 The genotypes are coded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
-
-gocard123
 
@@
 Here, each snp is having, roughly, the same effect on the trait, remembering that the maf of a snp also impacts a snps influence on a trait.
-Fourth, n deviates are drawn from a $N(0,sigma^2_e)$ where sigma_e^2 has been set to yield a trait with a heritability of 0.6.
+Fourth, n deviates are drawn from a $N(0,sigma^2_e)$ where $\sigma_e^2$ has been set to yield a trait with a heritability of 0.6.
 (Put in formula).
@@
 
-\texbf{GEMMA}
+\textbf{GEMMA:}  to be done.
 

### 2018-05-08 AM+/Plots_for_Paper/power_plots.R

@@ -37,14 +37,27 @@ DIR <- paste(getwd(),"/", sep="")
 DIR <- "/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper/"
 sizefn <- 16
 thresh_indx <- 1:500
-
+#thresh_indx <- 1:2
 
 ## vector initialisation
 fam <- c("W", "S",  "L","HS","A","HL")
+#fam <- NULL
+#for(ii in c(700,800,900,1000))
+#{
+#  for(jj in c(5,50,500))
+#  {
+#    nme <- paste0("X",ii,"new",jj)
+#    fam <- c(fam, nme)
+#  }
+#  
+#}
+
+
 names.of.methods <- c("am", "mlmm","glmnet","lasso","r2VIM","bigRR", "gemma", "fastALL", "fast")
 
 
 ## list initialisation
+
 FDR <- list()
 recall <- list()  ## == power
 dfres <- list()
@@ -132,8 +145,8 @@ dfres$method <- factor(dfres$method, levels=c("am", "mlmm",   "glmnet", "lasso",
 levels(dfres$method) <- c("AMplus", "MLMM", "glmnet", "LMM-Lasso", "r2VIM", "bigRR", "GEMMA", "FaST-LMM^all", 
                       "FaST-LMM^few")
 
-levels(dfres$fam) <- c("150 x 5K", "350 x 500K", "1500 x 50K", "2000 x 500K", 
-                       "4000 x 1.5M", "10000 x 1.5M")
+    levels(dfres$fam) <- c("150 x 5K", "350 x 500K", "1500 x 50K", "2000 x 500K", 
+                      "4000 x 1.5M", "10000 x 1.5M")
 
 ## change family labels to simulation labels
 #  W  150 x 5 K           750
@@ -157,7 +170,7 @@ df1 <- subset(subset(dfres, !(method=="AMplus" | method=="MLMM" | method=="GEMMA
 
 df2 <- subset(subset(dfres, method=="AMplus" | method=="MLMM"),  !(fam=="4000 x 1.5M" | fam=="10000 x 1.5M"))
 
-
+library(RColorBrewer)
 
 p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_line(size=1)  +
   geom_point(data=df2, aes(FDR, recall), size=2) +
[…]

### 2018-05-08 AM+/Plots_for_Paper/power_plots_GammaSetting.R

@@ -1,0 +1,306 @@
+## plotting power and FDR for multi-locus mehtods and single-locus methods (fast, fastALL, gemma)
+## Last revised:   29/11/2016  MAjor change. Plotting Power vs FDR 
+## 
+## Note: don't know why but this error would come up a lot when printing final plots. 
+##
+##       Error in grid.Call(L_textBounds, as.graphicsAnnot(x$label), x$x, x$y,  : 
+##       polygon edge not found
+## 
+##       Solution: just repeat the printing of the plot a couple of times and it will 
+##                 correct itself. 
+##
+## input data resx_y.RData ftp'ed from home directory on bragg in
+## /home/geo047/MWAM/SimStudy/Timing/Results/res$FAM_$indx.RData
+## where $FAM is W S L A HS HL and $indx is the index 1, 2, 3, etc over the thresholds used in
+## results.R on bragg 
+## contains columns
+## c("n_am", "nQTL_am",
+##  "n_amGPU", "nQTL_amGPU",
+##  "n_mlmm", "nQTL_mlmm",
+##  "n_glmnet", "nQTL_glmnet",
+##  "n_bigRR",  "nQTL_bigRR",
+##  "n_ranger",  "nQTL_ranger", 
+##  "n_lasso", "nQTL_lasso",
+##   n_gemma, nQTL_gemma, 
+##   n_fast, nQTL_fast, 
+##   n_fastALL, nQTL_fastALL,
+##  "nQTL_true")
+## Note 
+## moved from smoothed power curves to lines based on mean. Smoothed curves weird.
+require(ggplot2)
+library(GGally)
+library(ggthemes)
+library(gridExtra)
+library(RColorBrewer)
+library(extrafont)
+DIR <- paste(getwd(),"/", sep="")
+DIR <- "/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper/"
+sizefn <- 16
+##thresh_indx <- 1:500
+thresh_indx <- 1:2
+
+## vector initialisation
+fam <- c("W", "S",  "L","HS","A","HL")
+#fam <- NULL
+#for(ii in c(700,800,900,1000))
+#{
+#  for(jj in c(5,50,500))
+#  {
+#    nme <- paste0("X",ii,"new",jj)
+#    fam <- c(fam, nme)
+#  }
+#  
+#}
+
+
+names.of.methods <- c("am", "mlmm","glmnet","lasso","r2VIM","bigRR", "gemma", "fastALL", "fast")
+
+
[…]

### 2018-05-08 Precision and recall - Wikipedia, the free encyclopedia_files/load_002.js

@@ -1,71 +1,0 @@
-var mediaWikiLoadStart=(new Date()).getTime(),mwPerformance=(window.performance&&performance.mark)?performance:{mark:function(){}};mwPerformance.mark('mwLoadStart');function isCompatible(str){var ua=str||navigator.userAgent;return!!('querySelector'in document&&'localStorage'in window&&'addEventListener'in window&&!(ua.match(/webOS\/1\.[0-4]/)||ua.match(/PlayStation/i)||ua.match(/SymbianOS|Series60|NetFront|Opera Mini|S40OviBrowser|MeeGo/)||(ua.match(/Glass/)&&ua.match(/Android/))));}(function(){var NORLQ,script;if(!isCompatible()){document.documentElement.className=document.documentElement.className.replace(/(^|\s)client-js(\s|$)/,'$1client-nojs$2');NORLQ=window.NORLQ||[];while(NORLQ.length){NORLQ.shift()();}window.NORLQ={push:function(fn){fn();}};window.RLQ={push:function(){}};return;}function startUp(){mw.config=new mw.Map(true);mw.loader.addSource({"local":"/w/load.php","metawiki":"//meta.wikimedia.org/w/load.php"});mw.loader.register([["site","1t3wrg7",[1]],["site.styles","0idn2vw"
[…]

### 2018-05-08 AM+draft.tex
9 May 18 Morning update

@@
 We compared AMplus, in terms of computational and statistical performance, against seven other association mapping methods. 
-We chose methods that had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer software.
+We chose methods that, mostly, had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer software.
 Two of the methods are based on single locus (or locus-by-locus) models and five are based on multiple locus models.
@@
 
-From a practical perspective, how the methods were implemented, in terms of usability, varied greatly.
-In Table 1, we list seven important features for ensuring the usability of computer software for association mapping.
-We also identify which software has what features.
-In forming this list of features, we assumed that the primary purpose of a genome-wide association study is 
-to identify the set of marker loci in true and strongest association with a trait.  The level of documentation varied across the implementations but only 
-a few came with user manuals detailing the format of the input data and how to perform analyses. 
-Most implementations could handle additional fixed effects which is important when accounting for hidden population structure. 
-Few though could cope with marker data larger than the memory capacity of the computer.
-Similarly, few returned the best set of marker loci in strongest association with a trait as their result.
-Almost all the software returned results that required further processing, either
-by identifying the best set of marker loci via the setting of significance thresholds or having to do further analysis to calculate the significance of the results.
-Almost none of the association mapping software checked for errors in the input data.
-Only AMplus has all seven features.
+ In Table 1, 
+we list some of the important features that make computer software for association mapping useful. We also identify which of these 
+features are present in Eagle and the other seven implementations. For example, all but glmnet were purpose-built for 
+association mapping.
+Not all the implementations come with a user manual.
+Most could accomodate data on additional fixed effects but only Eagle and FaST-LMM could deal with data larger than the memory capacity of the computer.
+Also, if
+the input data contained errors, we found only Eagle and FaST-LMM identified the errors and provided advice on how to fix 
+the errors.
+
+\subsection{Simulation Study}
+
+We were interested in answering two questions. 
+First, is Eagle faster, computationally, than other competing implementations? Second, is Eagle better, statistically, at 
+identifying the snp closest to the genes that are influencing a trait? To answer these questions, we conducted a large 
+simulation study.
+
+First, how does Eagle compare, computationally, to other competing implementations? Second, 
+how accurate is Eagle for finding
 

### 2018-05-10 AM+draft.tex
10 May, 18 Results section changes

@@
 \subsection{Simulation Study}
+We performed a large simulation study where we sought to  answer two questions. 
+First, how does Eagle compare, in terms of run time and memory usage, to 
+competing implementations? Second, how well does Eagle find true associations (power) and avoid 
+false associations (type 1 errors)? We generated data under five different scenarios; a study of size 150 individuals 
+and 500 snp (150 x 5K),  350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
+50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
+1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).   
+We chose these scenarios to mirror some of the different sized GWAS being performed in animals, plants, and humans.
 
-We were interested in answering two questions. 
-First, is Eagle faster, computationally, than other competing implementations? Second, is Eagle better, statistically, at 
-identifying the snp closest to the genes that are influencing a trait? To answer these questions, we conducted a large 
-simulation study.
+For each scenario, we generated 100 replicates of data. A single replicate consists of snp data and quantitative trait data. 
+To introduce some of the complexities of dealing with real genotypes into the study, we obtained the snp genotypes from 
+the publicly available 1000 genome project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
+by selecting randomly a set of snp loci, assigning allelic effects to these loci, and then aggregating these effects for 
+each individual along with a random error.  
+ The number of snp selected per replicate follows a Poisson distribution with mean 30. The sizes of the allelic effects 
+ across the selected loci are equal.  A heritability of  50\% is assumed for the trait.
 
-First, how does Eagle compare, computationally, to other competing implementations? Second, 
-how accurate is Eagle for finding
+\subsection{Run times and memory usage}
 
-\subsection{Computational Performance: run times and memory usage}
+Analyses were performed on a high end desktop computer.
+It had dual 8-core Xeon processors, and 128 Giggabytes of RAM.
+All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
 

### 2018-05-11 AM+draft.tex
11 May, 2018 Change to plots for paper + change to sim section of paper

@@
 For multiple locus association mapping, we chose bigRR, glmnet, LMM-Lasso, MLMM, and r2VIM.
-Each takes a different approach to mapping multiple locus-trait associations jointly.
+Each takes a different approach to multiple-locus association mapping.
 
@@
 
+We analysed the simulated data with Eagle and the other implementations, recording their run (or elapse) times and 
+memory usage. The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM.
+Not all implementations could be run under all scenarios.  Of the multiple-locus methods, bigRR and 
+LMM-Lasso could only deal with the smaller sized data sets generated under scenarios 150x5K, 1500x50K, and 350x400K.
+This was because of their large memory footprint.   The implementations glmnet, MLMM, and r2VIM could analyse larger data sets. 
+However, of the multiple-locus methods, only Eagle, due to its capacity to handle data larger than the memory capacity of a computer, could produce results for data generated under scenario 10000x1.5M.
+Of the single-locus methods,
+it was feasible to analyse data under any of the scenarios with GEMMA and 
+FaST-LMM$^few$ (where the covariance matrix for the polygenic effect is calculated from a subset of the available snp data) . 
+However, the memory requirements for FaST-LMM$^all$ (all snp data used in formation of covariance matrix for polygenic effect) 
+were too large for data generated under 4000x1.5M and 10000x1.5M.
+
+The goal of association mapping is to identify those snp closest to the genes controlling a trait. Glmnet, r2VIM, and ... 
+produced results relatively quickly. However, these results were an ordering of the snp from most to least important. 
+Additional computation, in the form of stability selection (see Methods), was required before the statistical significance 
+of the snp could be determined. Only Eagle and MLMM produced results where no further computation was required.
+
+Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM.
+Our goal in performing these analyses was to not just order the snp from most to least important but to identify those snp which are closest to the genes controlling the trait.
+Although glmnet, r2VIM, and ....
+produced an ordering of the snp
+relatively quickly, additional computation in the form of stability selection (see Methods)  was required to assign significance to the results.
+
+In performing an analysis, our goal was not merely to order the snp from most important to least important. If this was the goal, 
+then the implementations 
+glmnet, r2VIM, ...
+
 Analyses were performed on a high end desktop computer.
 It had dual 8-core Xeon processors, and 128 Giggabytes of RAM.
-All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
-
-To asses the computational performance of AMplus, we conducted a large simulation study. 
-We were interested in  the impact of study size on  performance so we generated data under five different scenarios. 
-These are a GWAS of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
-50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
-1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  
-We chose these scenarios to reflect some of the different sized GWAS being performed in animals, plants, and humans. 
-For each scenario, we generated 100 replicates of data. A single replicate consists of snp genotype data and quantitative trait data. 
-We obtained the snp data from the publicly available 1000 genome project (phase 3). The quantitative trait data we generated from the 
-snp data by selecting a set of snp loci, assigning allelic effects to these snp, and aggregating these effects for each individual along with 
-random error. The number of snp selected per replicate follows a Poisson distribution with mean 30.
- The quantitative trait was generated to have a heritability of 50\%.
-Analyses were performed on a high end desktop computer.
-It had dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM.
 All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
@@
 
-The run times for AMplus against the other software programs, across the five scenarios, is shown in figure \ref{fig_runtimes} To help with interpretibility of the results, in both plots, we have taken the x and y axes to be on a log scale.
-This means a unit change on the x or y axes is equivalent to a change in the order of magnitude.
+The run times for Eagle against the other computer implementations, across the five scenarios, is shown in figure \ref{fig_runtimes} 
+The x and y axes are on a log scale due to the large difference in run times and number of genotypes across the different study 
+scenarios.  This means a unit change on the x or y axes is equivalent to a change in the order of magnitude. 
 In the top plot, a point is the median of the ratios of elapse times of the multiple locus method to AMplus for a given scenario.

### 2018-05-11 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -25,19 +25,20 @@ library(RColorBrewer)
 
 setwd("/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper")
 df <- read.table(file="cputimes.dat", header=TRUE)
-sizefn <- 14
+
+sizefn <- 12
 
 df<- melt(df)
-names(df) <- c("Pop", "method","ratio")
+names(df) <- c("Pop", "method","time")
 
-df$ratio[df$ratio<1] <- 1  ## setting times less than 1 to 1
+df$time[df$time<1] <- 1  ## setting times less than 1 to 1
 levels(df$Pop) <- c(10000*1500000, 150*5000, 1500*50000,2000*500000, 350*400000,  4000*1500000)
 df$Pop <- as.numeric(as.character(df$Pop))
 
 pps <- sort(unique(df$Pop))
-names(df) <- c("Pop", "method","ratio")
+names(df) <- c("Pop", "method","time")
 ##--------------------
-## times - ratios
+## times - times
 ##--------------------
 
 ## data frame for line segments
@@ -51,16 +52,16 @@ gsdf <- data.frame(x1=(pps), y1=rep(1,length(pps)),
 
 ## maybe draw vertical lines first to get around problem.
 
-p <- ggplot(df, aes(x=Pop, y=ratio, group=method,shape=method, colour=method)) + 
-  geom_line(size=2) + geom_point(aes(colour=method),size=5) +
-  scale_shape_manual(breaks=c("mlmm","glmnet","lasso","rf_ranger","bigRR"), 
-                     labels=c("MLMM","glmnet", "LMM-Lasso",
+p <- ggplot(df, aes(x=Pop, y=time, group=method,shape=method, colour=method)) + 
+  geom_line(size=2) + geom_point(aes(colour=method),size=4) +
+  scale_shape_manual(breaks=c("am.", "mlmm","glmnet","lasso","rf_ranger","bigRR"), 
+                     labels=c("Eagle", "MLMM","glmnet", "LMM-Lasso",
                               "r2VIM","bigRR"),
-                     values=c(rep(19,5)))+
-  scale_colour_manual(breaks=c("mlmm","glmnet","lasso","rf_ranger","bigRR"),
-                      labels=c("MLMM","glmnet", "LMM-Lasso",
+                     values=c(rep(19,6)))+
+  scale_colour_manual(breaks=c("am.", "mlmm","glmnet","lasso","rf_ranger","bigRR"),
[…]

### 2018-05-14 AM+draft.tex
14 May 18,  Working on simulation results section

@@
 
-We compared AMplus, in terms of computational and statistical performance, against seven other association mapping methods. 
-We chose methods that, mostly, had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer software.
+We compared Eagle, in terms of computational and statistical performance, against seven other association mapping methods. 
+We chose methods that (mostly) had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs.
 Two of the methods are based on single locus (or locus-by-locus) models and five are based on multiple locus models.
@@
  In Table 1, 
-we list some of the important features that make computer software for association mapping useful. We also identify which of these 
+we list some of the key features that make computer programs for association mapping useful. We also identify which of these 
 features are present in Eagle and the other seven implementations. For example, all but glmnet were purpose-built for 
 association mapping.
-Not all the implementations come with a user manual.
+Not all the computer programs come with a user manual.
 Most could accomodate data on additional fixed effects but only Eagle and FaST-LMM could deal with data larger than the memory capacity of the computer.
 Also, if
-the input data contained errors, we found only Eagle and FaST-LMM identified the errors and provided advice on how to fix 
-the errors.
+the input data contained errors, we found only Eagle and FaST-LMM tried to diagnose the problem and 
+provided advice on how to fix the errors.
 
@@
 
-We analysed the simulated data with Eagle and the other implementations, recording their run (or elapse) times and 
-memory usage. The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM.
+We analysed the simulated data with Eagle and the other computer programs, recording their memory usage and 
+run (or elapse) times.
+The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM.
+Not all data generated under the five scenarios could be analysed by all implementations.
+Memory usage
+for many of the computer programs was the limiting factor (See Sup figure 1).  The single-locus program GEMMA was by 
+far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
+excess of the 128 Gbytes of available RAM for the analysis of data generated under 4000x1.5M and 10000x1.5M.  Even FaST-LMM, a single locus
+implementation, required more than 128 Gbytes of memory for the analysis of the larger data sets 
+when all the marker data was used to calculate the relationship matrix. Of the multiple-locus programs, only Eagle,  
+with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
+for data from scenario 10000x1.5M.
+
+As the data sets became large, (4000x1.5m and 10000x1.5m), the multiple locus implementations memory footprint become excessive.
+Some even struggled with memory for data generated under 2000 x 500k.
+GEMMA, a single locus implementation had the lowest memory usage (sup figure 1).
+However, Eagle, with its ability to handle data larger than the memory capacity of the computer, was the only multiple locus implementation capable of producing findings for data from 10000x1.5m.
+
+In Figure 1, the median run times is displayed for the multiple locus implementations (top plot) and single locus implementations (bottom plot) compared to Eagle.
+
 Not all implementations could be run under all scenarios.  Of the multiple-locus methods, bigRR and 
-LMM-Lasso could only deal with the smaller sized data sets generated under scenarios 150x5K, 1500x50K, and 350x400K.
+LMM-Lasso were confined to the smaller sized data sets generated under scenarios 150x5K, 1500x50K, and 350x400K.
 This was because of their large memory footprint.   The implementations glmnet, MLMM, and r2VIM could analyse larger data sets. 
-However, of the multiple-locus methods, only Eagle, due to its capacity to handle data larger than the memory capacity of a computer, could produce results for data generated under scenario 10000x1.5M.
+However, of the multiple-locus methods, only Eagle, due to its ability to analyse data larger than the memory capacity of a computer, could produce results for data generated under scenario 10000x1.5M.
 Of the single-locus methods,

### 2018-05-15 AM+draft.tex
15 May added supmaterials

@@
 with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
-for data from scenario 10000x1.5M.
-
-As the data sets became large, (4000x1.5m and 10000x1.5m), the multiple locus implementations memory footprint become excessive.
-Some even struggled with memory for data generated under 2000 x 500k.
-GEMMA, a single locus implementation had the lowest memory usage (sup figure 1).
-However, Eagle, with its ability to handle data larger than the memory capacity of the computer, was the only multiple locus implementation capable of producing findings for data from 10000x1.5m.
+for data from our largest scenario 10000x1.5M.
 

### 2018-05-15 supmaterials.tex
15 May added supmaterials

\documentclass{article}
\usepackage{graphicx}
\usepackage{authblk}

\begin{document}

\title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
\author[1]{Andrew W. George}
\author[2]{Joshua Bowden}
\author[1]{Some other authors}

\affil[1]{Data61, CSIRO, Brisbane, Australia.}
\affil[2]{IM \&T, CSIRO, Brisbane, Australia.}

\maketitle

\begin{figure}
\caption{Memory usage (in Gigabytes) of Eagle and the other association mapping programs across 
the six simulation scenarios. The maximum amount of memory on the computer is 128 Gigabytes. 
The x-axis is on the log scale. GEMMA, a single-locus implementation, had the lowest memory usage. 
Of the multiple-locus implementations, Eagle had the lowest memory usage. Also, it 
was the only multiple-locus 
implementation able to produce results for data under  scenario 10000x1.5M. This is due to its ability 
to handle data larger than the available memory of a computer. FaST-LMM was run where all the snp data are used 
to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth snp are used to 
estimate the relationship matrix (FaST-LMM$^{few}$)}.
\begin{center}
\includegraphics[width=15cm, height=15cm]{mem.eps}
\end{center}
\end{figure}

\begin{table}
\caption{The median run times (in minutes) of Eagle and the other association mapping programs across the six simulation scenarios. The }
\begin{tabular}{llcccccc}
              &           &  \multicolumn{6}{c}{Simulation Scenarios} \\ \cline{3-8}
 Method & Name & 150x5K & 1500x50K & 350x400K & 2000x500K & 4000x1.5M & 10000x1.5M \\ \hline
  Multiple & Eagle 	&	   0.08  &   1.62   &  \bf{2.7}1  &   \bf{13.65}   & \bf{127.63}  &   699.55  \\
               & MLMM 	&	   0.15    &  2.91    & 19.04  &  143.01  &  870.84  &     \\
               & glmnet 	&	  0.11     & 3.95    & 14.06    & 74.03    &        &    \\
               & r2VIM 	&	   0.09    &  3.66    &  5.51    & 50.59    & 380.52  &   \\ 
               & bigRR 	&	    1.01   & 113.35   &  54.99   & 1030.61  &        &     \\
               & LMM-Lasso 	&     0.57  &   52.08 &    92.20  & 1031.85 &           &     \\ \\
Single     &  GEMMA 	&      0.02  &   5.02   &   6.17  &   84.83   & 723.33  & 4071.60 \\
               & FaST-LMM$^{few}$ 	&    \bf{0.01}   &  \bf{0.80}   &   7.07   &  20.16   & 193.90   & \bf{346.19} \\ 
               & FaST-LMM$^{all}$   	&   0.03   & 2.96  &    7.90   &  41.27  &            &   \\ \hline
\end{tabular}

\end{table}

\end{document}

### 2018-05-15 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -344,10 +344,11 @@ p_time
   
   
   
-  p <- ggplot(df, aes(x=Pop, y=mem, group=method, shape=method, colour=method)) + 
+  p <- ggplot(df, aes(x=Pop, y=mem, group=method, colour=method)) + 
     geom_line(size=2) + geom_point(aes(colour=method),size=5) +
+   # scale_shape_manual(values=1:nlevels(df$method)) +
     scale_colour_manual(breaks=c("am.", "bigRR", "glmnet", "lasso", "mlmm", "r2VIM", "gemma", "fastALL", "fast"),
-                       labels=c("AMplus","bigRR", "glmnet", "LMM-Lasso", "MLMM", "r2VIM", "GEMMA", bquote("FaST-LMM"^all), 
+                       labels=c("Eagle","bigRR", "glmnet", "LMM-Lasso", "MLMM", "r2VIM", "GEMMA", bquote("FaST-LMM"^all), 
                                 bquote("FaST-LMM"^few)),
                        values=brewer.pal(9, "Paired")) +
     
@@ -356,15 +357,15 @@ p_time
                                 bquote("FaST-LMM"^few)),
                        values = c(19,19,19,19,19,17,17,17,19)) +
 
-    scale_size(guide = 'none')  + scale_y_continuous(limits = c(0, 120))
+    scale_size(guide = 'none')  + scale_y_continuous(limits = c(0, 130))
   
   
-  p  <- p +  geom_segment(aes(x=7.5e5, y=0,xend=7.5e5,yend=115), colour="grey", cex=0.5) +
-    geom_segment(aes(x=7.5e7, y=0,xend=7.5e7,yend=110), colour="grey", cex=0.5) +     
-    geom_segment(aes(x=1.4e8, y=0,xend=1.4e8,yend=115), colour="grey", cex=0.5) +
-    geom_segment(aes(x=1.0e9, y=0,xend=1.0e9,yend=110), colour="grey", cex=0.5) +
-    geom_segment(aes(x=6.0e9, y=0,xend=6.0e9,yend=115), colour="grey", cex=0.5) +
-    geom_segment(aes(x=1.5e10,y=0,xend=1.5e10,yend=110), colour="grey", cex=0.5) 
+  p  <- p +  geom_segment(aes(x=7.5e5, y=0,xend=7.5e5,yend=127), colour="grey", cex=0.5) +
[…]

### 2018-05-16 AM+draft.tex
17 May morning commit - sim section

@@
 for data from our largest scenario 10000x1.5M.
+
+The median run times for Eagle and the other computer programs across the six scenarios are shown in Figure 1. 
+Eagle was significantly faster, sometimes by orders of magnitude,  than the other multiple-locus
+ implementations, especially for the larger data sets. For a simulation study with 150 individuals and 
+ 5000 snp, Eagle produced results in seconds.  For the larger simulation scenarios 1500x50K and 350x400K, 
+ analyses with Eagle took under two minutes. Even for data from a couple of thousand individuals and half a million 
+ snp (2000x500K), the median run time of Eagle was under 14 minutes. For our scenarios where there 
+ were thousands of individuals and 1.5 million snp, Eagle took just over two hours for the analysis of data from 
+ 4000x1.5M and just under 12 hours for the analysis of data from 10000x1.5M. However, the long run time for 10000x1.5M 
+ is due, largely, to the extra processing required for having data larger than the memory capacity of the computer. 
+ Towards the end of our simulation study, we gained access to a high memory computer with 512 Gigabytes of RAM. 
+ We reran Eagle on data from the largest scenario 10000x1.5M to measure the impact on run time.  By increasing 
+ the memory, the median run time dropped by almost half from 699.55 minutes to 413.23 minutes.
 

### 2018-06-13 AM+draft.tex
13June18 plots for paper update

--- 2018-05-15 AM+draft.tex
@@
 the publicly available 1000 genome project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
-by selecting randomly a set of snp loci, assigning allelic effects to these loci, and then aggregating these effects for 
+by selecting randomly a set of snp loci, assigning additive allelic effects to these loci, and then aggregating these effects for 
 each individual along with a random error.  
@@
 
-In Figure 1, the median run times is displayed for the multiple locus implementations (top plot) and single locus implementations (bottom plot) compared to Eagle.
+The median run times for Eagle and the other computer programs across the six scenarios are shown in Figure \ref{fig_time}. 
+The x- and y-axes are on a log scale.
+This means a unit change on the x- or y-axes is equivalent to a change in the order of magnitude.
+In answer to our question of how does Eagle compare, in terms of run time, to competing implementations,
+Eagle was significantly faster, sometimes by orders of magnitude,  than the other multiple-locus
+ implementations and is comparable to the single-locus implementations. For a simulation study with 150 individuals and 
+ 5000 snp, Eagle produced results in seconds.  For the larger simulation scenarios 1500x50K and 350x400K, 
+ analyses with Eagle took under two minutes. Even for data from a couple of thousand individuals and half a million 
+ snp (2000x500K), the median run time of Eagle was under 14 minutes. For our scenarios where there 
+ were thousands of individuals and 1.5 million snp, Eagle took just over two hours for the analysis of data from 
+ 4000x1.5M and just under 12 hours for the analysis of data from 10000x1.5M. 
+ Towards the final stages of writing this paper, 
+ we gained access to a XXXXXXXXXX, a faster desktop computer. We reran Eagle on data from the largest
+  scenario 10000x1.5M to measure the impact, if any, on run time. The median run time dropped by more than 50\% 
+  from just under 12 hours to 4.56 hours.
 
-Not all implementations could be run under all scenarios.  Of the multiple-locus methods, bigRR and 
-LMM-Lasso were confined to the smaller sized data sets generated under scenarios 150x5K, 1500x50K, and 350x400K.
-This was because of their large memory footprint.   The implementations glmnet, MLMM, and r2VIM could analyse larger data sets. 
-However, of the multiple-locus methods, only Eagle, due to its ability to analyse data larger than the memory capacity of a computer, could produce results for data generated under scenario 10000x1.5M.
-Of the single-locus methods,
-it was feasible to analyse data under any of the scenarios with GEMMA and 
-FaST-LMM$^few$ (where the covariance matrix for the polygenic effect is calculated from a subset of the available snp data) . 
-However, the memory requirements for FaST-LMM$^all$ (all snp data used in formation of covariance matrix for polygenic effect) 
-were too large for data generated under 4000x1.5M and 10000x1.5M.
-
-The goal of association mapping is to identify those snp closest to the genes controlling a trait. Glmnet, r2VIM, and ... 
-produced results relatively quickly. However, these results were an ordering of the snp from most to least important. 
-Additional computation, in the form of stability selection (see Methods), was required before the statistical significance 
-of the snp could be determined. Only Eagle and MLMM produced results where no further computation was required.
-
-Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM.
-Our goal in performing these analyses was to not just order the snp from most to least important but to identify those snp which are closest to the genes controlling the trait.
-Although glmnet, r2VIM, and ....
-produced an ordering of the snp
-relatively quickly, additional computation in the form of stability selection (see Methods)  was required to assign significance to the results.
-
-In performing an analysis, our goal was not merely to order the snp from most important to least important. If this was the goal, 
-then the implementations 
-glmnet, r2VIM, ...
-
-Analyses were performed on a high end desktop computer.
-It had dual 8-core Xeon processors, and 128 Giggabytes of RAM.
-All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
+ It is also worth noting that an increase in the number of snp genotypes in a study does not necessitate, automatically,  
+ an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
+ increasing behaviour of the curves in the memory (Sup Figure 1) and run-time plots (Figure \ref{fig_time}).  It 
+it is the study dimensions, not number of genotypes, that drives the computing resources required by 
+association mapping programs.
 
 \begin{figure}
-\label{fig_runtimes}
-\caption{Run times}
-\includegraphics[width=10cm]{time}
+\caption{Median run times, in minutes, for the analysis of association study data from the six simulation scenarios. 
+Eagle is compared against five other multiple-locus programs (top left plot) and two single-locus programs (top right plot). 
+The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
+programs, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
+The actual median run times for the programs across the scenarios are given in the table. The entries in a bold font 
+correspond to the lowest run-time for a scenario. 
+ FaST-LMM$^{all}$ is where the single-locus results are based on all the available 
+snp data. FaST-LMM$^{few}$ is where the single-locus results as based on a subset of the available snp data. }
+
+\label{fig_time}
+figure 2 goes here {Figure2_time.pdf}
+
 \end{figure}
 
-The run times for Eagle against the other computer implementations, across the five scenarios, is shown in figure \ref{fig_runtimes} 
-The x and y axes are on a log scale due to the large difference in run times and number of genotypes across the different study 
-scenarios.  This means a unit change on the x or y axes is equivalent to a change in the order of magnitude. 
-In the top plot, a point is the median of the ratios of elapse times of the multiple locus method to AMplus for a given scenario.
-The median is over the 100 replicates.
-Here, since the median of the ratios are all positive on the log scale, it means that AMplus had a shorter run time than than all the other multiple locus methods.
-In fact, in some cases, AMplus was over a hundred times, or over two orders of magnitude, faster.
-Unlike AMplus, the size of data under scenario 10000 X 1.5M was beyond the memory constraints of the other multiple locus implementations.
-This was also the case for LMM-Lasso, bigRR, and glmnet, for scenario 4000 x 1.5M.
+\subsection{Power and false discovery rates}
 
-In the bottom plot of figure \ref{fig_runtimes}, we compare the median run time of AMplus against the median run times of the single locus methods, FaST-LMM and GEMMA. 
-FaST-LMM was run in two ways. It was run where the genetic similarity matrix was built with all the marker data (FaST-LMM$^all$) or with data on every 500th snp (FaST-LMM$^few$). 
-AMplus was also run in two ways.  The default behavior for AMplus is to make use of CPUs for computing. However, AMplus also has the capacity to harness multiple GPUs (AMplus$^GPU$). 
-From the bottom plot, all the implementations have short run times when analysing data from scenario 150x5K. However, for the other scenarios, AMplus and AMplus$GPU$ have 
-significantly shorter run times than GEMMA and FaST-LMM$^all$. FaST-LMM$^few$ has a shorter run time than AMplus and AMplus$^GPU$ but only for scenarios 1500x50K and 
-10000x1.5M.  Furthermore, for scenario 10000x1.5M, the median run time for GEMMA is 4071 minutes, for AMplus is 699 minutes, for AMplus$^GPU$ is 447 minutes, and for 
-FaST-LMM$^few$ is 346 minutes. The very fast single-locus method FaST-LMM$^few$ is only 29\% faster than our multiple-locus method AMplus$^GPU$. It is worth noting though that 
-there is a setup cost to accessing GPU computing, making AMplus$^GPU$ most efficient on the larger data.
+To answer how well Eagle finds true associations (power) and avoids 
+false associations (type 1 errors), we calculated the statistical power and false discovery rate of the different 
+association mapping methods, across the six simulation scenarios.  We want Eagle to have a high statistical power 
+(probability of finding a true positive finding) and low false discovery rate (proportion of findings that are false positives).
 
-We also examined the memory usage of the different software programs (figure \ref{fig_memory} supplementary materials). 
-AMplus is comparable to the most efficient single locus implementation, FaST-LMM$^few$. 
-AMplus is also the only multiple locus program able to analyse data under scenarios ..... and ..... . 
-AMplus can process data larger than the memory capacity of the computer.
-
-\begin{figure}
-\label{fig_memory}
-\caption{Memory usage}
-\includegraphics[width=10cm]{mem}
-\end{figure}
-
-\subsection{Statistical Performance: power, FDR, and significance thresholds}
-
-As part of assessing the performance of AMplus against the other association mapping methods, 
-we calculated their statistical power and false discovery rates. 
-We want a method to have high statistical power (probability of finding a true positive finding) but 
-low false discovery rate (proportion of findings that are false positives). 
 All the methods, besides AMplus and MLMM, required a threshold to be set in order to identify significant findings. 

### 2018-06-13 AM+/Plots_for_Paper/power_plots.R

@@ -1,5 +1,5 @@
-## plotting power and FDR for multi-locus mehtods and single-locus methods (fast, fastALL, gemma)
 ## Last revised:   29/11/2016  MAjor change. Plotting Power vs FDR 
+## plotting power and FDR for multi-locus mehtods and single-locus methods (fast, fastALL, gemma)
 ## 
 ## Note: don't know why but this error would come up a lot when printing final plots. 
 ##
@@ -36,8 +36,8 @@ library(extrafont)
 DIR <- paste(getwd(),"/", sep="")
 DIR <- "/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper/"
 sizefn <- 16
-thresh_indx <- 1:500
-#thresh_indx <- 1:2
+#thresh_indx <- 1:499
+thresh_indx <- 1:100
 
 ## vector initialisation
 fam <- c("W", "S",  "L","HS","A","HL")
@@ -153,16 +153,23 @@ levels(dfres$method) <- c("AMplus", "MLMM", "glmnet", "LMM-Lasso", "r2VIM", "big
 #  S  350 x 400K         140,000
 #  L  1500 x 50K          75,000
 #  HS 2000 x 500K         1,000,000
-#  A  4000 x 1.5M       600,000,000
+#  A  400 x 1.5M       600,000,000
 #  HL 10000 x 1.5M      1.5*e10
 
 # "am"      "mlmm"    "glmnet"  "lasso"   "r2VIM"   "bigRR"   "gemma"   "fastALL" "fast"  
 
-##------------------------------------
+# remove any 0's 
+    indx <- which(dfres$FDR==0  )
+    dfres <- dfres[-indx,]
+   # stop()
+    indx <- which(dfres$fam=="150 x 5K" & dfres$method=="glmnet")
+
+    
+    ##------------------------------------
 ## Draw plot for multi-locus methods
 ##-------------------------------------
 ## dropping MLMM recall values by a little bit because they are lower and so that we can see it in the plot
-dfres[which(dfres$method=="MLMM"), "recall"] <-  dfres[which(dfres$method=="MLMM"), "recall"] * 0.95
+#dfres[which(dfres$method=="MLMM"), "recall"] <-  dfres[which(dfres$method=="MLMM"), "recall"] * 0.95
 
 
[…]

### 2018-06-13 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -61,7 +61,7 @@ p <- ggplot(df, aes(x=Pop, y=time, group=method,shape=method, colour=method)) +
   scale_colour_manual(breaks=c("am.", "mlmm","glmnet","lasso","rf_ranger","bigRR"),
                       labels=c("Eagle", "MLMM","glmnet", "LMM-Lasso",
                                "r2VIM","bigRR"),
-                      values=brewer.pal(9, "Paired")[c(2,3,9, 4,5, 1)]) + 
+                      values=brewer.pal(10, "Paired")[c(2,1,3, 4,5, 6)]) + 
   scale_size(guide = 'none') 
 p
 
@@ -78,12 +78,14 @@ p
 
 
 
-p  <- p +  geom_segment(aes(x=7.5e5, y=0,xend=7.5e5,yend=700), colour="grey", cex=0.5) +
-  geom_segment(aes(x=7.5e7, y=0,xend=7.5e7,yend=500), colour="grey", cex=0.5) +     
-  geom_segment(aes(x=1.4e8, y=0,xend=1.4e8,yend=700), colour="grey", cex=0.5) +
-  geom_segment(aes(x=1.0e9, y=0,xend=1.0e9,yend=500), colour="grey", cex=0.5) +
-  geom_segment(aes(x=6.0e9, y=0,xend=6.0e9,yend=700), colour="grey", cex=0.5) +
-  geom_segment(aes(x=1.5e10,y=0,xend=1.5e10,yend=500), colour="grey", cex=0.5) 
+
+p  <- p +  geom_segment(aes(x=7.5e5, y=0,xend=7.5e5,yend=6000), colour="grey", cex=0.5) +
+  geom_segment(aes(x=7.5e7, y=0,xend=7.5e7,yend=4000), colour="grey", cex=0.5) +     
+  geom_segment(aes(x=1.4e8, y=0,xend=1.4e8,yend=6000), colour="grey", cex=0.5) +
+  geom_segment(aes(x=1.0e9, y=0,xend=1.0e9,yend=4000), colour="grey", cex=0.5) +
+  geom_segment(aes(x=6.0e9, y=0,xend=6.0e9,yend=6000), colour="grey", cex=0.5) +
+  geom_segment(aes(x=1.5e10,y=0,xend=1.5e10,yend=4000), colour="grey", cex=0.5) 
+
 
 
 ## log log scale
@@ -91,13 +93,13 @@ p  <- p + scale_x_continuous(trans="log10", limits=c(500000, 1e11),
[…]

### 2018-06-18 AM+/Plots_for_Paper/power_plots.R
18Jun18 Morning update - working on results sim section

@@ -36,7 +36,7 @@ library(extrafont)
 DIR <- paste(getwd(),"/", sep="")
 DIR <- "/Users/geo047/Papers/AM-Paper/AM+/Plots_for_Paper/"
 sizefn <- 16
-#thresh_indx <- 1:499
+#thresh_indx <- 1:100
 thresh_indx <- 1:100
 
 ## vector initialisation
@@ -179,6 +179,13 @@ df2 <- subset(subset(dfres, method=="AMplus" | method=="MLMM"),  !(fam=="4000 x 
 
 library(RColorBrewer)
 
+
+#p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_line(size=1)  +
+#  geom_point(data=df2, aes(FDR, recall), size=1.5) +
+#  facet_wrap(~fam, ncol=2) + 
+#  theme(aspect.ratio = 1) # try with and without
+
+
 p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_smooth(size=1, se=FALSE, method="loess")  +
   geom_point(data=df2, aes(FDR, recall), size=1.5) +
   facet_wrap(~fam, ncol=2) + 
@@ -187,13 +194,16 @@ p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_smooth(size=1, se=F
 
 
 p <- p + scale_color_manual(
-  breaks=c("AMplus","MLMM","bigRR","glmnet","LMM-Lasso","r2VIM"),
-values=RColorBrewer:::brewer.pal(9, "Paired")[c(1,4,3,5,2,9)], 
+  breaks=c("AMplus", "MLMM","glmnet","LMM-Lasso","r2VIM","bigRR"),
+  labels=c("Eagle", "MLMM","glmnet", "LMM-Lasso",
+           "r2VIM","bigRR"),
+    values=brewer.pal(10, "Paired")[c(2,5,3,6,1,4)],
+  
 guide = guide_legend(override.aes = list(
   linetype = c("blank", "blank", "solid", "solid", "solid", "solid"),
   shape=c(16, 16, NA, NA, NA, NA))))
 
-p
+
 
 
 
@@ -264,9 +274,10 @@ insert_plot <- function(df1, df2, family){
 
    
    p <- p + scale_color_manual(
-     breaks=c("AMplus","MLMM","bigRR","glmnet","LMM-Lasso","r2VIM"),
[…]

### 2018-06-18 AM+draft.tex
19June18 morning commit - results - simulation

@@
 \subsection{Power and false discovery rates}
+
+In answer to the question of how well Eagle finds true snp-trait associations and avoids false snp-trait associations, we calculated, empirically, 
+the statistical power and false discovery rate of the different methods across the six scenarios. We did this by recording the set of snp that had been selected when generating a 
+replicate's trait data. These snp are in true association with the trait. Also, each replicate had a different set of selected snp. 
+When the data were analysed via a particular method, 
+we were able to determine for that method which of its snp-trait findings were true and which were false.
+A snp finding was deemed to be true if it was positioned within 40kb of the location of a true snp.
+
+We did this by recording, at the time a replicate was generated, those snp which had been selected to 
+have an additive effect on the quantitative trait. These snp are the true snp in association with the trait. When we analysed a
+replicate by each of the methods, we also recorded the number of snp reported to be in significant association with the trait. 
+Those snp findings that were positioned within 40 kb of the location of a true snp were deemed as true findings.  Hence, 
+for each method for a given replicate, we were able to record the number of true and false findings.
 

### 2018-06-18 AM+/Plots_for_Paper/power_plots.R

@@ -185,8 +185,14 @@ library(RColorBrewer)
 #  facet_wrap(~fam, ncol=2) + 
 #  theme(aspect.ratio = 1) # try with and without
 
-
+require(mgcv)
 p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_smooth(size=1, se=FALSE, method="loess")  +
+  geom_point(data=df2, aes(FDR, recall), size=1.5) +
+  facet_wrap(~fam, ncol=2) + 
+  theme(aspect.ratio = 1) # try with and without
+
+
+ggplot(data=df1, aes(FDR, recall, color=method)) + geom_line(size=1)  +
   geom_point(data=df2, aes(FDR, recall), size=1.5) +
   facet_wrap(~fam, ncol=2) + 
   theme(aspect.ratio = 1) # try with and without

### 2018-06-19 AM+draft.tex
19June18 morning commit

--- 2018-06-13 AM+draft.tex
@@
 \subsection{Power and false discovery rates}
+
+In answer to the question of how well Eagle finds true snp-trait associations and avoids false snp-trait associations, we calculated, empirically, 
+the statistical power and false discovery rate of the different methods across the six scenarios.
+We did this by recording the set of snp that had been selected to have an additive allelic effect when generating a
+replicate's trait data. These snp are in true association with the trait. Each replicate had a different set of selected snp. 
+When the data were analysed via a particular method, 
+we were able to determine for that method which of its snp-trait findings were true and which were false.
+A snp finding was deemed to be true if it was positioned within 40 kb of the location of a true snp.
+Hence,
+for each method for a given replicate, we were able to record the number of true and false findings.
+
+To calculate the power of a method, we averaged across the 100 replicates, the number of snp that were found by the method and true, divided by the number of true snp used in generating the quantitative trait.
+To calculate the false discovery rate of a method, we averaged across the 100 replicates, the number of snp that were found by the method and true divided by the number of (true and false) snp that were found by the method.
+
+In Figure X, we plot how power changes with fdr for the multiple locus methods. See sup Figure X for how power changes with fdr for Eagle and 
+the single-locus methods.
+A plot contains a mix of single points and power curves.
+The single points are the power and corresponding fdrs for Eagle and mlmm.
+These two methods treat association mapping as a model selection problem.
+Their are no significance thresholds to be set.
+The other
+methods all require a significance threshold to be specified before a finding can be deemed as
+
+In contrast, the other methods treat association mapping as a parameter estimation problem. As such, a threshold value is required before a snp 
+can be deemed as being in significant association with a trait.  The challenge is in setting this significance threshold. The power curves in Figure 
+X and Sup Figure X  are obtained by varying the significance level. A high (low) significance level results in low (high) power but a high (low) fdr.
+
+ in order to determine which are significant findings.
+
+before the truth of a snp result can be assessed.
+
+This is not true for the other methods.
+They all require a significance threshold upon which to determine the truth of a finding.
+The power curves are formed from varying the threshold so that the significance level ranges from 99.9% to 0.1%.
 

### 2018-06-19 AM+draft.tex
19June18 lunch time update

@@
 
-In answer to the question of how well Eagle finds true snp-trait associations and avoids false snp-trait associations, we calculated, empirically, 
-the statistical power and false discovery rate of the different methods across the six scenarios.
-We did this by recording the set of snp that had been selected to have an additive allelic effect when generating a
-replicate's trait data. These snp are in true association with the trait. Each replicate had a different set of selected snp. 
+In answer to the question of how well Eagle finds true and avoids false snp-trait associations, we calculated, empirically, 
+the statistical power and false discovery rates of the different methods across the six scenarios.
+We did this by recording the set of snp that had been used in generating a replicate's trait data.
+These snp are in true association with the trait.
+Each replicate had a different set of selected snp.
 When the data were analysed via a particular method, 
@@
 
-To calculate the power of a method, we averaged across the 100 replicates, the number of snp that were found by the method and true, divided by the number of true snp used in generating the quantitative trait.
-To calculate the false discovery rate of a method, we averaged across the 100 replicates, the number of snp that were found by the method and true divided by the number of (true and false) snp that were found by the method.
+To calculate the power of a method, for each replicate,  the number of true snp that were found by a method is divided by the number 
+of (true) snp used in generating the quantitative trait data. We then averaged across the 100 replicates. Similarly, to calculate the fdr, for 
+each replicate, the number of true snp that were found is divided by the number of (true and false) snp findings for the method. We then 
+averaged these power and fdr values across the 100 replicates.
+
+To calculate the power of a method, we averaged across the 100 replicates of a scenario, the number of snp that were found by the method and true, divided by the number of true snp used in generating the quantitative trait.
+To calculate the false discovery rate of a method, we averaged across the 100 replicates, the number of snp that were found and true divided by the number of (true and false) snp that were found.
 
@@
 Their are no significance thresholds to be set.
-The other
-methods all require a significance threshold to be specified before a finding can be deemed as
-
 In contrast, the other methods treat association mapping as a parameter estimation problem. As such, a threshold value is required before a snp 
-can be deemed as being in significant association with a trait.  The challenge is in setting this significance threshold. The power curves in Figure 
-X and Sup Figure X  are obtained by varying the significance level. A high (low) significance level results in low (high) power but a high (low) fdr.
+can be deemed as being in significant association with a trait.  The challenge is in setting this significance threshold. The curves in Figure 
+X and Sup Figure X  show how power changes with the fdr as the significance level (and resulting significance threshold) is allowed to vary.
 

### 2018-06-19 AM+draft.tex
19June18 rev 3

@@
 the statistical power and false discovery rates of the different methods across the six scenarios.
-We did this by recording the set of snp that had been used in generating a replicate's trait data.
-These snp are in true association with the trait.
-Each replicate had a different set of selected snp.
-When the data were analysed via a particular method, 
-we were able to determine for that method which of its snp-trait findings were true and which were false.
-A snp finding was deemed to be true if it was positioned within 40 kb of the location of a true snp.
-Hence,
-for each method for a given replicate, we were able to record the number of true and false findings.
+We did this by recording the set of snp that had been selected for generating a replicate's trait data.
+These snp are in true association with the trait. Each replicate had a different set of selected snp. We then determined for a method
+ which of its findings were true and which were false. Those snp findings that were located within 40 kb of the location of a true snp were 
+ deemed to be true.
 
-To calculate the power of a method, for each replicate,  the number of true snp that were found by a method is divided by the number 
-of (true) snp used in generating the quantitative trait data. We then averaged across the 100 replicates. Similarly, to calculate the fdr, for 
-each replicate, the number of true snp that were found is divided by the number of (true and false) snp findings for the method. We then 
-averaged these power and fdr values across the 100 replicates.
-
-To calculate the power of a method, we averaged across the 100 replicates of a scenario, the number of snp that were found by the method and true, divided by the number of true snp used in generating the quantitative trait.
-To calculate the false discovery rate of a method, we averaged across the 100 replicates, the number of snp that were found and true divided by the number of (true and false) snp that were found.
+To calculate the power of a method, we averaged across the 100 replicates of a scenario where for each replicate, we divided the number of snp findings that were true by the number of selected (true) 
+snp used in generating the trait data.  Similarly, to calculate a method's fdr, we averaged across the 100 replicates of a scenario where 
+for each replicate, we divide the number of true snp found by the number of (true and false) snp found.
 

### 2018-06-19 AM+draft.tex
20June18 rev1 Power and FDR section

@@
 
+We calculated, empirically, the statistical power and false discovery rates of Eagle and the other methods across the six scenarios. 
+We were interested in answering the question of how well Eagle finds true snp-trait associations and avoids false snp-trait associations. 
+For each replicate, we knew which snp had been used in creating the quantitative trait data. These snp are in true association with 
+the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these snp. 
+ Based on these snp, we were able to assess the validity of a  method's findings. When a replicated was analysed, 
+ a method's findings were counted as true if the snp were located within 40 kb of snp in true 
+association with the trait.
+
 In answer to the question of how well Eagle finds true and avoids false snp-trait associations, we calculated, empirically, 
-the statistical power and false discovery rates of the different methods across the six scenarios.
-We did this by recording the set of snp that had been selected for generating a replicate's trait data.
-These snp are in true association with the trait. Each replicate had a different set of selected snp. We then determined for a method
+the statistical power and false discovery rates of Eagle and the other methods across the six scenarios.
+We did this by recording the set of snp that had been used for generating a replicate's trait data.
+These selected snp are in true association with the trait and these are the snp that each of the association mapping methods are 
+trying to identify. Each replicate had a different set of selected snp. Based on these snp, 
+we then determined for a method
  which of its findings were true and which were false. Those snp findings that were located within 40 kb of the location of a true snp were 
  deemed to be true.
-
 To calculate the power of a method, we averaged across the 100 replicates of a scenario where for each replicate, we divided the number of snp findings that were true by the number of selected (true)
@@
 
-In Figure X, we plot how power changes with fdr for the multiple locus methods. See sup Figure X for how power changes with fdr for Eagle and 
-the single-locus methods.
-A plot contains a mix of single points and power curves.
-The single points are the power and corresponding fdrs for Eagle and mlmm.
-These two methods treat association mapping as a model selection problem.
-Their are no significance thresholds to be set.
+In Figure X, the power and fdrs of Eagle are compared to the other multiple locus methods. For a comparison with single-locus methods, 
+see Sup Figure X. Each plot contains a mix of single points and power curves. The single points are the power and fdrs for Eagle and MLMM. 
+These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
 In contrast, the other methods treat association mapping as a parameter estimation problem. As such, a threshold value is required before a snp 

### 2018-06-20 AM+draft.tex
20June18 rev2

@@
 the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these snp. 
- Based on these snp, we were able to assess the validity of a  method's findings. When a replicated was analysed, 
+ Based on these snp, we were able to assess the validity of a  method's findings. When a replicate was analysed, 
  a method's findings were counted as true if the snp were located within 40 kb of snp in true 
@@
 
-In answer to the question of how well Eagle finds true and avoids false snp-trait associations, we calculated, empirically, 
-the statistical power and false discovery rates of Eagle and the other methods across the six scenarios.
-We did this by recording the set of snp that had been used for generating a replicate's trait data.
-These selected snp are in true association with the trait and these are the snp that each of the association mapping methods are 
-trying to identify. Each replicate had a different set of selected snp. Based on these snp, 
-we then determined for a method
- which of its findings were true and which were false. Those snp findings that were located within 40 kb of the location of a true snp were 
- deemed to be true.
+To calculate the power of a method, for each replicate, we divided the number of true snp findings by the number of snp that had been used 
+in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the fdr of a method, for each replicate, 
+we divided the number of true snp findings by the number of (true and false) snp findings. We then averaged across the 100 replicates.
+
  To calculate the power of a method, we averaged across the 100 replicates of a scenario where for each replicate, we divided the number of snp findings that were true by the number of selected (true)

### 2018-06-20 AM+draft.tex
20June18 rev3

@@
 
- To calculate the power of a method, we averaged across the 100 replicates of a scenario where for each replicate, we divided the number of snp findings that were true by the number of selected (true)
-snp used in generating the trait data.  Similarly, to calculate a method's fdr, we averaged across the 100 replicates of a scenario where 
-for each replicate, we divide the number of true snp found by the number of (true and false) snp found.
+The power and fdr of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure X.
+We restricted our attention to these scenarios for the multiple-locus methods because for scenario 4000 x 1.5M,
+the data could only be analysed by Eagle, MLMM, and r2VIM. For scenario 10000 x 1.5M, the data could only be analysed by Eagle. 
+The power and fdr of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in Sup Figure X.  Each plot contains 
+single points and power curves. The single points are the power and fdrs for Eagle and MLMM.
+These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
+The power curves are for those methods that treat association mapping as a parameter estimation problem. Here, the 
+significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
+the fdr as the significance threshold (and underlying significance level) is varied.
 
-In Figure X, the power and fdrs of Eagle are compared to the other multiple locus methods. For a comparison with single-locus methods, 
+In Figure X, the power and fdr of Eagle are compared to the other multiple locus methods. For a comparison with single-locus methods, 
 see Sup Figure X. Each plot contains a mix of single points and power curves. The single points are the power and fdrs for Eagle and MLMM. 
@@
 X and Sup Figure X  show how power changes with the fdr as the significance level (and resulting significance threshold) is allowed to vary.
-
- in order to determine which are significant findings.
-
-before the truth of a snp result can be assessed.
-
-This is not true for the other methods.
-They all require a significance threshold upon which to determine the truth of a finding.
-The power curves are formed from varying the threshold so that the significance level ranges from 99.9% to 0.1%.
 

### 2018-06-20 AM+draft.tex
21June18 rev1 sim section

@@
 For each replicate, we knew which snp had been used in creating the quantitative trait data. These snp are in true association with 
-the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these snp. 
- Based on these snp, we were able to assess the validity of a  method's findings. When a replicate was analysed, 
+the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these snp. By knowing which snp are in true 
+association with the trait, we were able to assess the validity of a  method's findings. When a replicate was analysed, 
  a method's findings were counted as true if the snp were located within 40 kb of snp in true 
-association with the trait.
-
-To calculate the power of a method, for each replicate, we divided the number of true snp findings by the number of snp that had been used 
+association with the trait. To calculate the power of a method, for each replicate, we divided the number of true snp findings by the number of snp that had been used 
 in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the fdr of a method, for each replicate, 
@@
 significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
-the fdr as the significance threshold (and underlying significance level) is varied.
+the fdr as the significance threshold (and underlying significance level) is adjusted.
 
-In Figure X, the power and fdr of Eagle are compared to the other multiple locus methods. For a comparison with single-locus methods, 
-see Sup Figure X. Each plot contains a mix of single points and power curves. The single points are the power and fdrs for Eagle and MLMM. 
-These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
-In contrast, the other methods treat association mapping as a parameter estimation problem. As such, a threshold value is required before a snp 
-can be deemed as being in significant association with a trait.  The challenge is in setting this significance threshold. The curves in Figure 
-X and Sup Figure X  show how power changes with the fdr as the significance level (and resulting significance threshold) is allowed to vary.
+At first glance, Eagle appears underpowered. Several of the other methods have higher power than Eagle. The goal of association 
+mapping though is not only to find true snp-trait associations but to also avoid false snp-trait associations.
+For those methods with higher power than Eagle, the cost of this higher power is much higher false discovery rates.
+In other words, these methods are finding more true results
+but they also find many more false results. If we restrict our attention to the power of these methods when the (genome-wide) fdr is less 
+than 5\% (see inset plots in Figure X and Sup Figure X), we see that Eagle, closely followed by MLMM,  has the highest power for the lowest 
+fdr.
+This is especially true when Eagle is compared against single-locus methods.
+For single-locus methods to achieve the same level of power as Eagle, a threshold has to be set where the fdr is extremely high.
 
-To answer how well Eagle finds true associations (power) and avoids 
-false associations (type 1 errors), we calculated the statistical power and false discovery rate of the different 
-association mapping methods, across the six simulation scenarios.  We want Eagle to have a high statistical power 
-(probability of finding a true positive finding) and low false discovery rate (proportion of findings that are false positives).
+If the goal of association mapping was purely to implemented methods with high power, hence having high probability of finding 
+snp in true association with a trait, then from Fig ..., it is possible to do this.
 
-All the methods, besides AMplus and MLMM, required a threshold to be set in order to identify significant findings. 
-A conservative threshold guards against false discoveries but reduces power. 
-A anti-conservative threshold increases power at the cost of also increasing the false discovery rate.
-Only AMplus and mlmm avoid thresholds by treating association mapping as a model selection problem where the best model is found by minimising the model selection statistic, extBIC.
+If the goal of association mapping was purely to find snp in true association with a trait, then from Figure X and Sup Figure X,  for those 
+methods based on parameter estimation, to 
+set a threshold that results in higher power of the method than Eagle.
 
-Figures \ref{fig_power_multiple}  and \ref{fig_power_single} show the relationship, calculated empirically, between the power and fdr for each method.
-The multiple-locus (Figure \ref{fig_power_multiple})   and the single-locus (Figure \ref{fig_power_single})  methods are plotted separately for clarity.
-AMplus features in both plots for comparison.
-By varying the significance thresholds for the different methods, we were able to calculate the change in power and fdr.
-Since AMplus and MLMM do not rely on thresholds, their power and fdr appear as single points in the plots.
-In practice, the true power and fdr is unknown.
-However, because we generated data where we knew which loci were acting as quantitative trait loci, we were able to calculate the true power and fdr for the methods.
+From Figure X and Sup Figure X, across the scenarios, it is possible to
 

### 2018-06-22 AM+draft.tex
22June18 rev1 real data

@@
 figure 2 goes here {Figure2_time.pdf}
-
 \end{figure}
@@
 significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
-the fdr as the significance threshold (and underlying significance level) is adjusted.
+the fdr as the significance threshold  is adjusted.
 
@@
 mapping though is not only to find true snp-trait associations but to also avoid false snp-trait associations.
-For those methods with higher power than Eagle, the cost of this higher power is much higher false discovery rates.
+For those methods with higher power than Eagle, the cost is much higher false discovery rates.
 In other words, these methods are finding more true results
-but they also find many more false results. If we restrict our attention to the power of these methods when the (genome-wide) fdr is less 
-than 5\% (see inset plots in Figure X and Sup Figure X), we see that Eagle, closely followed by MLMM,  has the highest power for the lowest 
+but they are also finding many more false results. If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than 
+5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest 
 fdr.
-This is especially true when Eagle is compared against single-locus methods.
+This is especially true when the power and fdr of Eagle is compared to the single-locus methods.
 For single-locus methods to achieve the same level of power as Eagle, a threshold has to be set where the fdr is extremely high.
-
-If the goal of association mapping was purely to implemented methods with high power, hence having high probability of finding 
-snp in true association with a trait, then from Fig ..., it is possible to do this.
-
-If the goal of association mapping was purely to find snp in true association with a trait, then from Figure X and Sup Figure X,  for those 
-methods based on parameter estimation, to 
-set a threshold that results in higher power of the method than Eagle.
-
-From Figure X and Sup Figure X, across the scenarios, it is possible to
-
-From figures \ref{fig_power_multiple}  and \ref{fig_power_single}, the superior performance of AMplus is apparent.
-When the fdr is low, AMplus has the highest power with Mlmm also performing well.
-AMplus is noticeably more powerful than the single locus methods.
-It is possible to set a significance threshold for these other methods that results in higher power. 
-However, this comes at the cost of an intolerably high fdr.
-
-For the methods evaluated here that require a significance threshold, figure  \ref{fig_threshold}  shows how 
-threshold size impacts fdr. Each plot corresponds to the results from a different method. 
-A curve within a plot is formed from analysing data with the method and using different threshold values to identify locus-trait associations. 
-The true fdr is then calculated from these findings.
-We can see from figure \ref{fig_threshold}  that the threshold impacts the fdr differently across methods.
-The relationship between threshold value and fdr even differs when data are analysed with the same method but for different scenarios.
-This demonstrates one of the challenges in using association mapping methods that are reliant upon thresholds.
-With AMplus, we do not have this challenge.
-For completeness, for the small sampled size scenarios ...
-And ....., the fdr for AMplus was 0.07 and 0.09, respectively.
-With the larger study sizes of ..., ..., and ...., the fdr drops to ..., .., and .....
-respectively.
 
@@
 \subsection{Mouse Study}
+
+To demonstrate the strengths of Eagle over single-locus association mapping, we analysed data from a large study in outbred mice.
+We chose this study for several reasons; the data are publicly available, measurements were taken on a large number of traits, and many of the snp-trait associations were validated through other published results.
+The primary reason for choosing this study though was that its analyses were based on snp dosages (continuous values from zero to one of expected allele counts) rather than snp genotypes.
+Analyses based on dosages rather than genotypes have been shown to have increased power \cite{zheng2011comparison}.  
+By repeating their single-locus analyses but with dosages converted into genotypes, we obtained a subset of the original snp-trait findings. 
+Whether additional findings by Eagle were true could be determined, if they were found in the original study.
+
+The performance of Eagle
+
+We were able to evaluate the performance of Eagle against single-locus association mapping with 
+real data.
+
+Here, we had the unique opportunity of being able to evaluate the performance of Eagle against single-locus association mapping with 
+real data. The truth of new findings from Eagle, if found in the original study but now found when the dosage data was converted into 
+genotypes, could be assessed.
+
+Any new findings from Eagle not found by single-locus mapping
+
+The performance of Eagle could then be compared to single-locus association mapping for real data without the delema of knowing 
+whether new findings from Eagle were valid.  We were able to confirm the validity of Eagle's findings through the original findings.
+
+This then gave us the opportunity that any new findings found by Eagle
+
+When Eagle analyses were performed on the
+
+By repeating  the analyses of REF but with genotypes converted from the dosage data, we obtained single-locus results that were a 
+subset of the results found and confirmed in ref. When Eagle was run on the same genotype data and new results not found by the single-locus 
+method found, we had the unique opportunity of evaluating the validity of these new results against the original findings for a real data set.
+
+We were then afforded the opportunity to
+
+Since our 
+focus is on the more commonly used genotypes, the dosages were converted into the discrete values AA, AB, and BB.
+
+was that it use of snp 
+dosages rather than genotypes.
+
+though was that this study differed from most other studies
+in that the analyses were based on snp dosages (continuous values from zero to one of expected allele counts) rather than snp genotypes (Discrete values AA,AB, BB).
+Analyses based on dosages rather than genotypes have been shown to have increased power \cite{zheng2011comparison}.
+When we repeated the single-locus analyses of ref, by focusing on genoytpes, we found a subset of the previously
+found results due to a loss in information.
+
+Here, we have a unique opportunity.
+
+By focusing on genotypes, we had a unique opportunity.
+
+Since the use of genotypes is more common in association studies than dosages,
+
 Still to be written.

### 2018-06-25 AM+draft.tex
25June18 rev1 results almost there but needs some details

@@
 
-To demonstrate the strengths of Eagle over single-locus association mapping, we analysed data from a large study in outbred mice.
-We chose this study for several reasons; the data are publicly available, measurements were taken on a large number of traits, and many of the snp-trait associations were validated through other published results.
-The primary reason for choosing this study though was that its analyses were based on snp dosages (continuous values from zero to one of expected allele counts) rather than snp genotypes.
-Analyses based on dosages rather than genotypes have been shown to have increased power \cite{zheng2011comparison}.  
-By repeating their single-locus analyses but with dosages converted into genotypes, we obtained a subset of the original snp-trait findings. 
-Whether additional findings by Eagle were true could be determined, if they were found in the original study.
+We were interested in comparing results from Eagle with those from single locus association mapping for a real data set.
+ We chose to focus on a large outbred mouse study.
+This study was unusual in that it collected and analysed snp dosages (continuous values from zero to one of expected allele counts)  instead of the more common snp genotypes.
+Analyses based on dosages rather than discrete genotypes have been shown to have greater power of detection \cite{zheng2011comparison}.
+By converting the dosages into genotypes and analysing the data with the single locus program FaSt-LMM, we obtained a subset of those findings reported in the original study.
+We then analysed the data with Eagle.
+Due to Eagles increased power, we found snp-trait associations not found with the FaST-LMM.
+However, we were
+ able to confirm the validity of these new findings as they matched what was found in the original study.
 
-The performance of Eagle
+For the single locus analysis of the data, we followed the same procedure as xxxx. 
+The only exceptions were that we focused on the autosomal snp and it was necessary to 
+increase the number of permutations for the controlling of the fdr from 100 to 500.
 
-We were able to evaluate the performance of Eagle against single-locus association mapping with 
-real data.
+Eagle was run in two ways; under its default settings ($Eagle^{default}$) and where we specified the regularisation parameter for model selection ($Eagle^{optimal}$ ) Eagle chooses the best model via the extended BIC criteria (ref).
+The conservativeness of the extended BIC can be adjusted by a single regularisation parameter that ranges from 0 to 1.
+In the simulation study, this parameter was set to 1, its most conservative and default setting.
+However, there is also opportunity to set the parameter to a value less than 1.
+This increases power but also increases the fdr.
+For each trait, we used permutation to set the regularisation parameter to give a fdr of 5\% .
 
-Here, we had the unique opportunity of being able to evaluate the performance of Eagle against single-locus association mapping with 
-real data. The truth of new findings from Eagle, if found in the original study but now found when the dosage data was converted into 
-genotypes, could be assessed.
+The genome wide results from the analyses of the mouse data are shown in Figure X. The mouse study took
+measurements on 200 traits.  However, we focused our analyses on the 43 traits whose snp-trait associations could be
+corroborated by prior published evidence.
+Each plot contains the number of snp-trait associations that were found and in agreement with the original findings.
+Neither method found snp not identified in the original mouse study.
+As we saw in the simulation study, there is a notable difference in the two methods capacity to discover snp-trait associations.
+$Eagle^{default}$, under its default settings, for x traits found the same number of findings as FaST-LMM and for y traits found more findings.
+$Eagle^{optimal}$,
+with its regularisation parameter fine tuned to the trait, for x traits found the same number of findings as FaST-LMM and for y traits 
+found more findings. Overall, $Eagle^{default}$ and $Eagle^{optimal}$ found xx\% and yy\% more snp-trait associations than a 
+single-locus analysis of the mouse data.
 
-Any new findings from Eagle not found by single-locus mapping
-
-The performance of Eagle could then be compared to single-locus association mapping for real data without the delema of knowing 
-whether new findings from Eagle were valid.  We were able to confirm the validity of Eagle's findings through the original findings.
-
-This then gave us the opportunity that any new findings found by Eagle
-
-When Eagle analyses were performed on the
-
-By repeating  the analyses of REF but with genotypes converted from the dosage data, we obtained single-locus results that were a 
-subset of the results found and confirmed in ref. When Eagle was run on the same genotype data and new results not found by the single-locus 
-method found, we had the unique opportunity of evaluating the validity of these new results against the original findings for a real data set.
-
-We were then afforded the opportunity to
-
-Since our 
-focus is on the more commonly used genotypes, the dosages were converted into the discrete values AA, AB, and BB.
-
-was that it use of snp 
-dosages rather than genotypes.
-
-though was that this study differed from most other studies
-in that the analyses were based on snp dosages (continuous values from zero to one of expected allele counts) rather than snp genotypes (Discrete values AA,AB, BB).
-Analyses based on dosages rather than genotypes have been shown to have increased power \cite{zheng2011comparison}.
-When we repeated the single-locus analyses of ref, by focusing on genoytpes, we found a subset of the previously
-found results due to a loss in information.
-
-Here, we have a unique opportunity.
-
-By focusing on genotypes, we had a unique opportunity.
-
-Since the use of genotypes is more common in association studies than dosages,
-
-Still to be written.
+ß
 

### 2018-06-26 AM+draft.tex
26June18 methods v1

@@
 \bibliographystyle{naturemag}
+\renewcommand{\baselinestretch}{1}
 
@@
 
-\subsection{Mouse Study}
+\subsection{Mouse Data Analysis}
 
 We were interested in comparing results from Eagle with those from single locus association mapping for a real data set.
- We chose to focus on a large outbred mouse study.
+ We chose to focus on data from a large outbred mouse study.
 This study was unusual in that it collected and analysed snp dosages (continuous values from zero to one of expected allele counts)  instead of the more common snp genotypes.
@@
 
-ß
-
 \section{Methods}
@@
 \subsection{Outbred Mice Study}
+\subsection{Mouse Data}
 
-Phenotypic and genotypic data were collected on a large study population of commercially 
-available outbred mice. The aim of the study was to map, through association, the genetic basis
-of complex traits in mice. The construction of the population and collection of data are described in detail in 
-\cite{nicod2016genome}.  Briefly, the population consists of 2073 adult outbred mice from Charles River Laboratories. Phenotypic data
-were obtained from 200 traits. These traits fell into the categories of behaviour, tissue, and physiology. Genotype 
-data were collected from 7, 073, 398 SNPs.  A reduced set of 359, 559 SNP across the 19 
-autosomes and sex chromosome was then identified, upon which all genome-wide analyses were based.  These SNP tagged all other SNP with MAF $> 0.1\%$
-and LD $r^2 > 0.98$.  Missing data?
+The mouse data were obtained from a large genome-wide association study which was performed in outbred mice. 
+The study is described in detail in \cite{nicod2016genome}. Phenotypic and genotypic data were available on 1,887 adult mice. 
+The phenotypic data consisted of measurements from 200 behavioural, tissue based, and physiological traits.  Of these traits, 
+43 yielded snp-trait associations that could be corroborated through other independent published work. It was these 
+43 traits which were the focus of our real data analyses. Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
+form of marker dosages (expected allele counts that range from zero to one). All missing data had been imputed. 
+We converted these  dosages into discrete genotypes 
+by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively.
 
 \subsection{Eagle Approach for Multiple-locus Association Mapping}
-\subsubsection{Overview}
 
-Eagle is an approach that identifies genomic regions of interest for a trait via the building of a (linear mixed) model.
-The model is built iteratively, without need of significance levels or thresholds.
-The model has two parts, a fixed effects part and a random effects part.
-The random effects part contains the error and an effect for the snp-trait association across the entire genome.
-It is from this effect
- that we are able to identify which snp is in strongest association with the trait.  
- Our model building process is as follows.
-First we fit the current model to the data.
-Second, from the random effect for association, we identify which snp is in strongest association  with the trait.
-Third, we move this snp from the random effects to the fixed effects.
-We do this to stop the neighbours of the chosen snp from being selected in subsequent iterations and 
-to give opportunity to snp in linkage disequilibrium with genes having lesser impact on the trait to be discovered. 
- Fourth, we calculate if the new model is "better" than the current model. 
-If so,  the new model becomes the current model and we repeat the process.
-Once the model building process is complete, the findings from Eagle are the set of snp that have been moved into the fixed effects part of the model.
-These snp map the seperate regions of the genome that house genes that are influencing the trait.
+Eagle is a method for multiple locus association mapping on a genome-wide scale.
+It is based on linear mixed models.
+It differs from most other single- and multiple-locus association mapping methods.
+Eagle treats association mapping as a model selection instead of parameter estimation problem.
+Consequently,  we do not have to content with multiple testing issues or having to construct significance thresholds.
+The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
+quantitative trait loci in experimental crosses \cite{verbyla2007analysis}
 
-\subsubsection{The specifics}
-
-%%Suppose data are collected from a genome-wide association study.  The study involves $n$ individuals. 
-%Let $y^{(n \times 1)}$ be the vector containing the quantitative trait data. 
-%Let $M^{(n \times L)} = [m_1 m_2 \ldots m_L]$ be the matrix containing the genotype data. Snp genotypes 
-%are collected from $L$ loci. The vector $m_j^{(n \times 1)}$ contains the snp genotypes for the $j$th snp. The genotypes 
-%are recoded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
-
-The Eagle approach for multiple-locus association mapping follows from methods developed for QTL mapping in plant 
-populations \cite{verbyla2007analysis} (and some other WGAIM references). 
-It is based on building a linear mixed model, iteratively, via forward 
-selection.
-
-Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
-associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
-To perform the $s+1$th  iteration of 
-our model building procedure, we do the following.
-
-First, we fit the current model to the data. 
 Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered snp that was 
@@
 snp were selected, then $S=\{500023, 15, 420\}$. 
-Let $y^{(n \; \times \;1)}$ be a vector containing the quantitative trait data. 
+Let $y^{(n \; \times \;1)}$ be a vector containing $n$ quantitative trait measurements. 
 Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome on $n_g$ groups/lines/strains.  In this paper, $n$ and $n_g$ are equal. 
- It is common for the columns of this matrix to be in map order but this is not a requirement. 
-The vector $m_j^{(n \; \times \; 1)}$ contains the genotypes, from the $n$ individuals,  for the $j$th snp. 
+from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that several trait measurements 
+may be taken of the same group/line/strain. 
+ It is common for the columns of $M$ to be in map order but this is not a requirement. 
+The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th snp. 
 The genotypes are coded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
 
+The specifics of the Eagle method are as follows. 
+Eagle builds the "best" model iteratively, via forward selection. 
+Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
+associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
+To perform the $s+1$th  iteration, we first fit the current model to the data. 
 The (current) model is of the form 
@@
 $X^{(n \; \times \; p)}$ and $Z^{( n \; \times \; n_g)}$ are known design matrices with $X$ being of full rank and $Z$ 
-containing zeros and ones that assign the appropriate genetic effect to each observation. 
+containing zeros and ones that assign the appropriate genetic effect to each measurement. 
 The vector 
@@
 is effective for discovering snp-trait associations. It also reduces the computational cost of our model building procedure.
-
-%The simplicity of our genetic model also reduces the cost of computation.
 

### 2018-06-26 AM+draft.tex
26June18 methods v2

@@
 
-We compared the computational and statistical performance of Eagle against five other multiple locus applications for 
+We compared the computational and statistical performance of Eagle against five other multiple locus methods for 
 association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
+These five methods were selected because of their demonstrated value 
+for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
+either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
+features is given in Table X.
+
+BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. 
+Regularisation methods can handle parameter estimation problems where the number of predictors is far greater than the number of 
+samples. When applied to association mapping, they allow the effect of all snp to be modelled simultaneously. 
+BigRR implements generalized ridge regression.
+In ridge regression, the coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary. LMM-Lasso 
+implements lasso within a linear mixed model 
+framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
+coefficients of "unimportant" predictors are shrunk to exactly zero.
+Glmnet is based on elastic net, a regularisation method that is an improvement over ridge regression and lasso (ref needed). 
+All three methods are sophisticated pieces of statistical machinery. The challenge is that while they have the capacity to estimate the 
+effect sizes of snp across an entire genome simultaneously, obtaining the statistical significance of these snp is problematic. 
+Permutation has been suggested as a way of calculating significance 
+empirically \cite{shen2013novel} but we instead opted for stability selection (see below).
+
+Regularisation methods have dominated 
+the development of multiple locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
+By introducing a penalty function that balances bias and variance, regularisation methods can 
+handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
+One of the challenges in using BigRR for association mapping is that while it has the capacity to estimate the effect sizes of all the snp simultaneously, 
+obtaining their statistical significance is difficult. 
+Permutation has been suggested as a way of calculating significance 
+empirically \cite{shen2013novel} but we instead opted for stability selection.
+
+MLMM, of the five multiple locus methods, is the closest in philosophy to Eagle. It too is based on a model building philosophy 
+and builds the best linear mixed model via a forward selection strategy. However, MLMM does not make use of dimension 
+reduction which adds to its computational cost. Also, its model building process is different. Eagle 
+identifies the $(s+1)$th snp that is in strongest association with the trait by maximising a score statistic. MLMM 
+identifies this snp by constructing a separate linear mixed model for each unidentified snp and calculating its statistical significance in
+the fixed effects part of the model. The snp with the highest significance is included in the next iteration of the model building process.
+
+R2VIM is quite different to the other association mapping methods. It implements random forests for association mapping, 
+a machine learning algorithm to measure the 
+importance of the snp-trait associations. 
+There is no relationship between relative importance scores and significance levels. Consequently, 
+setting an appropriate genome-wide threshold is problematic. We avoid this problem by focusing, in the simulation study, 
+only on the relationship between threshold level and its false discovery rate and power.
+
+BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
+In ridge regression, the coefficients of the predictors are shrunk 
+to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary. 
+The model used in BigRR can be formulated as a linear mixed model.
+One of the challenges in using BigRR for association mapping is that while it can estimate the effect sizes of all the snp simultaneously, 
+obtaining their statistical significance is difficult. 
+Permutation has been suggested as a way of calculating significance 
+empirically \cite{shen2013novel} but we instead opted for stability selection (see below).
+
 BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
-the development of multiple locus approaches.  They give opportunity for the effect of all the snp to be modelled simultaneously. 
+the development of multiple locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
 By introducing a penalty function that balances bias and variance, regularisation methods can 

### 2018-06-26 AM+draft.tex
27June18 methods v2

@@
 
-BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. 
+BigRR, LMM-Lasso, and glmnet  implement three different regression-based regularisation methods; generalized ridge regression, 
+lasso, and elastic net, respectively. 
 Regularisation methods can handle parameter estimation problems where the number of predictors is far greater than the number of 
-samples. When applied to association mapping, they allow the effect of all snp to be modelled simultaneously. 
-BigRR implements generalized ridge regression.
-In ridge regression, the coefficients of the predictors are shrunk 
-to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary. LMM-Lasso 
-implements lasso within a linear mixed model 
-framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
-coefficients of "unimportant" predictors are shrunk to exactly zero.
-Glmnet is based on elastic net, a regularisation method that is an improvement over ridge regression and lasso (ref needed). 
-All three methods are sophisticated pieces of statistical machinery. The challenge is that while they have the capacity to estimate the 
-effect sizes of snp across an entire genome simultaneously, obtaining the statistical significance of these snp is problematic. 
-Permutation has been suggested as a way of calculating significance 
-empirically \cite{shen2013novel} but we instead opted for stability selection (see below).
-
-Regularisation methods have dominated 
-the development of multiple locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
-By introducing a penalty function that balances bias and variance, regularisation methods can 
-handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
-One of the challenges in using BigRR for association mapping is that while it has the capacity to estimate the effect sizes of all the snp simultaneously, 
-obtaining their statistical significance is difficult. 
-Permutation has been suggested as a way of calculating significance 
-empirically \cite{shen2013novel} but we instead opted for stability selection.
+samples. This is particularly advantageous in the context of association mapping as it allows the effects of snp across an entire 
+genome to be estimated simultaneously. 
+All three methods are sophisticated pieces of statistical machinery. The limitation though is that the significance of an effect, which 
+is needed in association mapping,  can not be calculated easily. 
+ A possible solution is to estimate the significance of the effects via permutation  \cite{shen2013novel}  
+but we have instead opted for stability selection (see below).
 
@@
 and builds the best linear mixed model via a forward selection strategy. However, MLMM does not make use of dimension 
-reduction which adds to its computational cost. Also, its model building process is different. Eagle 
-identifies the $(s+1)$th snp that is in strongest association with the trait by maximising a score statistic. MLMM 
-identifies this snp by constructing a separate linear mixed model for each unidentified snp and calculating its statistical significance in
-the fixed effects part of the model. The snp with the highest significance is included in the next iteration of the model building process.
+reduction which adds to its computational cost. Also, its model building process differs from that of Eagle's. Eagle 
+identifies the $(s+1)$th snp that is in strongest association with the trait by finding the snp effect with the maximum score statistic. 
+MLMM instead identifies the snp, that when included as a fixed effect in the model, has the highest statistical 
+significance. This neccesitates the construction and fitting of a separate linear mixed model for each unidentified snp. MLMM does this efficiently.
 
-R2VIM is quite different to the other association mapping methods. It implements random forests for association mapping, 
-a machine learning algorithm to measure the 
-importance of the snp-trait associations. 
-There is no relationship between relative importance scores and significance levels. Consequently, 
-setting an appropriate genome-wide threshold is problematic. We avoid this problem by focusing, in the simulation study, 
-only on the relationship between threshold level and its false discovery rate and power.
+R2VIM implements random forests and a new way of measuring the importance of a snp.
+In random forests, the worth of a predictor is measured, empirically, by calculating its importance score.
+It is from these importance scores that snp can be ordered in terms of their strength of association with a trait.
+The challenge though is in knowing what proportion of the
+highest ordered snp are in true association. R2VIM addresses this, partially, by calculating a relative importance score for a snp. However, 
+there is no relationship between relative importance scores and significance levels, making the setting of a threshold difficult.
+
+======================= The above needs a bit more detail may need to look up papers ================
 

### 2018-06-27 AM+draft.tex
27June18 methods v3

@@
 
-The mouse data were obtained from a large genome-wide association study which was performed in outbred mice. 
+The data were obtained from a large genome-wide association study which was performed in outbred mice. 
 The study is described in detail in \cite{nicod2016genome}. Phenotypic and genotypic data were available on 1,887 adult mice. 
-The phenotypic data consisted of measurements from 200 behavioural, tissue based, and physiological traits.  Of these traits, 
+The phenotypic data consisted of measurements from 200 behavioural, tissue, and physiological traits.  Of these traits, 
 43 yielded snp-trait associations that could be corroborated through other independent published work. It was these 
-43 traits which were the focus of our real data analyses. Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
-form of marker dosages (expected allele counts that range from zero to one). All missing data had been imputed. 
-We converted these  dosages into discrete genotypes 
+43 traits that were the focus of our real data analyses. Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
+form of marker dosages (expected allele counts that ranged from zero to one). All missing data had been imputed. 
+We converted the dosages into discrete genotypes 
 by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively.
@@
 Eagle treats association mapping as a model selection instead of parameter estimation problem.
-Consequently,  we do not have to content with multiple testing issues or having to construct significance thresholds.
+Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
+Eagle also 
+reports as its findings only those snp that are in strongest linkage disequilibrium, and hence closest, to the genes influencing a trait. 
 The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
@@
 Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered snp that was 
-selected in the $k$th iteration. For example, if three iterations of our model building procedure 
+selected in the $k$th iteration. Suppose three iterations of our model building procedure 
 have been performed and say the 500023rd, 15th, and 420th, 
 snp were selected, then $S=\{500023, 15, 420\}$. 
-Let $y^{(n \; \times \;1)}$ be a vector containing $n$ quantitative trait measurements. 
+Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
 Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that several trait measurements 
+from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
 may be taken of the same group/line/strain. 
@@
 in the summation account 
-for the additive effects of genes that are in linkage disequilibrium with the selected snp. The other term models 
-snp-trait associations along the entire genome, except for those snp that have already been selected. 
+for the additive effects of the selected snp that are in linkage disequilibrium with genes that are influencing the trait. The other term models 
+snp-trait associations along the entire genome, simultaneously, except for those snp that have already been selected. 
 This is a simple genetic model but it 
-is effective for discovering snp-trait associations. It also reduces the computational cost of our model building procedure.
+is effective for discovering snp-trait associations.
 
@@
 having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
-model (\ref{eq2}) is much less due to the random term having only $n$ effects needing estimating.
+model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
+effects needing estimating.
 
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model (3) with 
+Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model  (\ref{eq3})  with 
 \begin{equation}
@@
 for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
-either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
-features is given in Table X.
+either stand-alone computer programs or as  packages. See Table X for a summary of their features.
 
 BigRR, LMM-Lasso, and glmnet  implement three different regression-based regularisation methods; generalized ridge regression, 
-lasso, and elastic net, respectively. 
-Regularisation methods can handle parameter estimation problems where the number of predictors is far greater than the number of 
-samples. This is particularly advantageous in the context of association mapping as it allows the effects of snp across an entire 
-genome to be estimated simultaneously. 
-All three methods are sophisticated pieces of statistical machinery. The limitation though is that the significance of an effect, which 
-is needed in association mapping,  can not be calculated easily. 
- A possible solution is to estimate the significance of the effects via permutation  \cite{shen2013novel}  
+lasso, and elastic net, respectively. Regularisation methods can handle parameter estimation problems where the number of predictors is
+ far greater than the number of  samples. This is made possible through a penalty function that balances bias against variance. 
+The application of regularisation methods to association mapping has appeal because it means that the snp effects across an 
+entire genome can be estimated, simultaneously. 
+All three methods are sophisticated pieces of statistical machinery but the significance of an effect, which 
+is needed in association mapping,  can not be calculated easily. Permutation as been suggest as an empirical solution to this 
+limitation \cite{shen2013novel} 
 but we have instead opted for stability selection (see below).
 
-MLMM, of the five multiple locus methods, is the closest in philosophy to Eagle. It too is based on a model building philosophy 
-and builds the best linear mixed model via a forward selection strategy. However, MLMM does not make use of dimension 
-reduction which adds to its computational cost. Also, its model building process differs from that of Eagle's. Eagle 
-identifies the $(s+1)$th snp that is in strongest association with the trait by finding the snp effect with the maximum score statistic. 
-MLMM instead identifies the snp, that when included as a fixed effect in the model, has the highest statistical 
-significance. This neccesitates the construction and fitting of a separate linear mixed model for each unidentified snp. MLMM does this efficiently.
+MLMM, of the five multiple locus methods, is the closest in philosophy to Eagle. It too is based on  building the 
+best  linear mixed model via forward selection using the ExtBIC criterion. 
+However, there are several differences between the two methods.
+ MLMM does not make use of dimension reduction and they are implemented quite differently leading to substantial 
+ differences  in run time and memory usage (see Results).  Also, they differ in how a model is built. 
+ Eagle uses a score statistic formed from parameter estimates of the random genetic effect in which to identify the next snp to enter 
+ the model.  MLMM uses the statistical significance of a snp when it is treated as a fixed effect in the model. 
+ This necessitates fitting a separate linear mixed model for each unidentified snp. Fortunately,  MLMM does do this 
+ efficiently.
 
+R2VIM is quite different to the other four methods. Here, a non-parametric model-free rather than model-based statistical approach is adopted. 
 R2VIM implements random forests and a new way of measuring the importance of a snp.  
@@
 It is from these importance scores that snp can be ordered in terms of their strength of association with a trait. 
-The challenge though is in knowing what proportion of the
-highest ordered snp are in true association. R2VIM addresses this, partially, by calculating a relative importance score for a snp. However, 
-there is no relationship between relative importance scores and significance levels, making the setting of a threshold difficult.
+The difficulty is in knowing what proportion of the 
+highest ordered snp are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
+random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score (f) is calculated for 
+each snp. The greater f is than one, the more evidence there is that the snp is in association with a trait.  A snp-trait association is 
+declared if f for the snp is above some threshold value across all parallel runs. The difficulty is that the relationship between 
+this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
+this would be computationally challenging for all but the smallest of association studies.
 

### 2018-06-27 AM+draft.tex
27June18 methods v4

@@
 \documentclass{nature}
+\usepackage{xr}
 \bibliographystyle{naturemag}
 \renewcommand{\baselinestretch}{1}
+
+%% external tex files
+\externaldocument{Figure2_time}
 
@@
 \title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George$^1$, Joshua Bowden$^2$, need to sort out the other authors}
+\author{Andrew W. George$^1$, Arunas Verbyla$^1$, Joshua Bowden$^2$, need to sort out the other authors}
 
@@
 \begin{affiliations}
-\item Data61, CSIRO, Brisbane, Australia.
-\item IM \&T, CSIRO, Brisbane, Australia.
+\item Data61, CSIRO, Australia.
+\item IM \&T, CSIRO, Australia.
 \end{affiliations}
@@
 The strength of association between a marker locus and trait is measured for each locus separately, on a locus-by-locus basis.
-Multiple-locus methods that map multiple locus-trait associations simultaneously have been available for some time.
+Multiple-locus methods that map multiple-locus trait associations simultaneously have been available for some time.
 However, they have attracted little attention.
@@
 
-Over the past decade,  Genome wide association studies have changed considerably in both their analysis and design.
-Early gwas followed a case-control design.
-Association mapping methods were no more complicated than contingency table tests or simple
+Over the past decade,  Genome wide association studies (GWAS) have changed considerably in both their analysis and design. Early GWAS
+ followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
 linear regression.
@@
 The identification of these regions from these analyses is not always straightforward.
-Gwas results are reported, typically, via Manhattan plots that plot the -log10 of the p value for each locus against the map position of the locus.
+GWAS results are reported, typically, via Manhattan plots that plot the -log10 of the p value for each locus against the map position of the locus.
 The location of peaks in this plot signify regions of interest.
@@
 
-It is somewhat surprising then that multiple locus association mapping methods haven't attracted more attention. Methods based on 
+It is somewhat surprising then that multiple-locus association mapping methods haven't attracted more attention. Methods based on 
 regularisation techniques, such as ridge regression and lasso, measure all locus-trait associations simultaneously. 
@@
 Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings (refs).
-An multiple locus method that does show promise is the multi-locus linear mixed model method (ref).
-The best multiple locus model is built with simple variable selection.
+An multiple-locus method that does show promise is the multi-locus linear mixed model method (ref).
+The best multiple-locus model is built with simple variable selection.
 Results are immediately interpretable but here, computation can be a challenge for large data.
 
-In this paper, we present our new multiple locus method for genome wide association mapping, which we are calling Eagle.
+In this paper, we present our new multiple-locus method for genome wide association mapping, which we are calling Eagle.
 Eagle combines the strength of regularisation techniques (being able to fit all locus-trait associations jointly), with simple variable selection (having easy to interpret results).
 Our method does not require a significance threshold to be set nor regularisation parameters to be fine tuned.
-Through a clever dimension reduction step, we are able to achieve a computational performance similar to the fastest single locus linear mixed model implementations.
-Eagle is statistically more powerful than single locus association mapping and is as and often more powerful than
-most multiple locus methods.
-Our aim is to make multiple locus association mapping on a genome wide scale routine.
+Through a clever dimension reduction step, we are able to achieve a computational performance similar to the fastest single-locus linear mixed model implementations.
+Eagle is statistically more powerful than single-locus association mapping and is as and often more powerful than
+most multiple-locus methods.
+Our aim is to make multiple-locus association mapping on a genome wide scale routine.
 To this end, we have created a fully documented R package, that is easy to use, even for non R users.
@@
 We chose methods that (mostly) had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs.
-Two of the methods are based on single locus (or locus-by-locus) models and five are based on multiple locus models.
-Of the many ways of performing single locus association mapping, we chose GEMMA and FaST-LMM because of their popularity and computational speed.
-For multiple locus association mapping, we chose bigRR, glmnet, LMM-Lasso, MLMM, and r2VIM.
+Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multiple-locus models.
+Of the many ways of performing single-locus association mapping, we chose
+GEMMA \cite{zhou2012genome}  and FaST-LMM \cite{lippert2011fast} because of their popularity and computational speed. 
+For multiple-locus association mapping, we chose bigRR \cite{shen2013novel}, glmnet \cite{Friedman2010glmnet}, 
+LMM-Lasso \cite{rakitsch2013lasso}, MLMM \cite{segura2012efficient} , and r2VIM \cite{szymczak2016r2vim}.  
 Each takes a different approach to multiple-locus association mapping.
 
- In Table 1, 
+ In Table \ref{suptabsummary}, 
 we list some of the key features that make computer programs for association mapping useful. We also identify which of these 
@@
 
+{\em Table 1 goes around here}
+
 \subsection{Simulation Study}
@@
 false associations (type 1 errors)? We generated data under five different scenarios; a study of size 150 individuals 
-and 500 snp (150 x 5K),  350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
-50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
-1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).   
+and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
+50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
+1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).   
 We chose these scenarios to mirror some of the different sized GWAS being performed in animals, plants, and humans.
 
-For each scenario, we generated 100 replicates of data. A single replicate consists of snp data and quantitative trait data. 
-To introduce some of the complexities of dealing with real genotypes into the study, we obtained the snp genotypes from 
-the publicly available 1000 genome project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
-by selecting randomly a set of snp loci, assigning additive allelic effects to these loci, and then aggregating these effects for 
+For each scenario, we generated 100 replicates of data. A single replicate consists of SNP data and quantitative trait data. 
+To introduce some of the complexities of dealing with real genotypes into the study, we obtained the SNP genotypes from 
+the publicly available 1000 Genome Project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
+by selecting, randomly, a set of SNP loci, assigning additive allelic effects to these loci, and then aggregating these effects for 
 each individual along with a random error.  
- The number of snp selected per replicate follows a Poisson distribution with mean 30. The sizes of the allelic effects 
+ The number of SNPs selected per replicate follows a Poisson distribution with mean 30. The sizes of the allelic effects 
  across the selected loci are equal.  A heritability of  50\% is assumed for the trait.
@@
 run (or elapse) times.
-The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 Giggabytes of RAM.
+The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM.
 Not all data generated under the five scenarios could be analysed by all implementations.
 Memory usage
-for many of the computer programs was the limiting factor (See Sup figure 1).  The single-locus program GEMMA was by 
+for many of the computer programs was the limiting factor (See Sup Figure 1).  The single-locus program GEMMA was by 
 far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
-excess of the 128 Gbytes of available RAM for the analysis of data generated under 4000x1.5M and 10000x1.5M.  Even FaST-LMM, a single locus
+excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
+Even FaST-LMM, a single-locus
 implementation, required more than 128 Gbytes of memory for the analysis of the larger data sets 
@@
 with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
-for data from our largest scenario 10000x1.5M.
+for data from our largest scenario 10000 x 1.5M.
 
@@
 The x- and y-axes are on a log scale.
-This means a unit change on the x- or y-axes is equivalent to a change in the order of magnitude.
-In answer to our question of how does Eagle compare, in terms of run time, to competing implementations,
+This means a unit change on the x- or y-axis is equivalent to a change in the order of magnitude.
+In answer to our question of how does Eagle compare in terms of run time to competing implementations,
 Eagle was significantly faster, sometimes by orders of magnitude,  than the other multiple-locus
  implementations and is comparable to the single-locus implementations. For a simulation study with 150 individuals and 
- 5000 snp, Eagle produced results in seconds.  For the larger simulation scenarios 1500x50K and 350x400K, 
+ 5000 SNPs, Eagle produced results in seconds.  For the larger simulation scenarios 1500 x  50K and 350 x4 00K, 
  analyses with Eagle took under two minutes. Even for data from a couple of thousand individuals and half a million 
- snp (2000x500K), the median run time of Eagle was under 14 minutes. For our scenarios where there 
- were thousands of individuals and 1.5 million snp, Eagle took just over two hours for the analysis of data from 
- 4000x1.5M and just under 12 hours for the analysis of data from 10000x1.5M. 
+ SNPs (2000 x 500K), the median run time of Eagle was under 14 minutes. For our scenarios where there 
+ were thousands of individuals and 1.5 million SNPs, Eagle took just over two hours for the analysis of data from 
+ 4000 x 1.5M and just under 12 hours for the analysis of data from 10000 x 1.5M. 
  Towards the final stages of writing this paper, 
  we gained access to a XXXXXXXXXX, a faster desktop computer. We reran Eagle on data from the largest
-  scenario 10000x1.5M to measure the impact, if any, on run time. The median run time dropped by more than 50\% 
+  scenario 10000 x 1.5M to measure the impact, if any, on run time. The median run time dropped by more than 50\% 
   from just under 12 hours to 4.56 hours.
 
- It is also worth noting that an increase in the number of snp genotypes in a study does not necessitate, automatically,  
+ It is also worth noting that an increase in the number of SNP genotypes in a study does not necessitate, automatically,  
  an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
@@
 
-\begin{figure}
-\caption{Median run times, in minutes, for the analysis of association study data from the six simulation scenarios. 
-Eagle is compared against five other multiple-locus programs (top left plot) and two single-locus programs (top right plot). 
-The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
-programs, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
-The actual median run times for the programs across the scenarios are given in the table. The entries in a bold font 
-correspond to the lowest run-time for a scenario. 
- FaST-LMM$^{all}$ is where the single-locus results are based on all the available 
-snp data. FaST-LMM$^{few}$ is where the single-locus results as based on a subset of the available snp data. }
-
-\label{fig_time}
-figure 2 goes here {Figure2_time.pdf}
-\end{figure}
+\em{ Figure 2 goes here}
 
@@
 We calculated, empirically, the statistical power and false discovery rates of Eagle and the other methods across the six scenarios. 
-We were interested in answering the question of how well Eagle finds true snp-trait associations and avoids false snp-trait associations. 
-For each replicate, we knew which snp had been used in creating the quantitative trait data. These snp are in true association with 
-the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these snp. By knowing which snp are in true 
+We were interested in answering the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations. 
+For each replicate, we knew which SNPs had been used in creating the quantitative trait data. These SNPs are in true association with 
+the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these SNPs. By knowing which SNPs are in true 
 association with the trait, we were able to assess the validity of a  method's findings. When a replicate was analysed, 
- a method's findings were counted as true if the snp were located within 40 kb of snp in true 
-association with the trait. To calculate the power of a method, for each replicate, we divided the number of true snp findings by the number of snp that had been used 
+ a method's findings were counted as true if the SNPs were located within 40 kb of SNP in true 
+association with the trait. To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
 in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the fdr of a method, for each replicate, 
-we divided the number of true snp findings by the number of (true and false) snp findings. We then averaged across the 100 replicates.
+we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.
 
@@
 At first glance, Eagle appears underpowered. Several of the other methods have higher power than Eagle. The goal of association 
-mapping though is not only to find true snp-trait associations but to also avoid false snp-trait associations.
+mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
 For those methods with higher power than Eagle, the cost is much higher false discovery rates.
@@
 \label{fig_power_multiple}
-\caption{Power curves for  multiple locus methods}
+\caption{Power curves for  multiple-locus methods}
 \includegraphics[width=10cm]{powerMultiple}
@@
 
-We were interested in comparing results from Eagle with those from single locus association mapping for a real data set.
+We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
  We chose to focus on data from a large outbred mouse study.
-This study was unusual in that it collected and analysed snp dosages (continuous values from zero to one of expected allele counts)  instead of the more common snp genotypes.
+This study was unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
 Analyses based on dosages rather than discrete genotypes have been shown to have greater power of detection \cite{zheng2011comparison}.
-By converting the dosages into genotypes and analysing the data with the single locus program FaSt-LMM, we obtained a subset of those findings reported in the original study.
+By converting the dosages into genotypes and analysing the data with the single-locus program FaSt-LMM, we obtained a subset of those findings reported in the original study.
 We then analysed the data with Eagle.
-Due to Eagles increased power, we found snp-trait associations not found with the FaST-LMM.
+Due to Eagles increased power, we found SNP-trait associations not found with the FaST-LMM.
 However, we were
@@
 
-For the single locus analysis of the data, we followed the same procedure as xxxx. 
-The only exceptions were that we focused on the autosomal snp and it was necessary to 
+For the single-locus analysis of the data, we followed the same procedure as xxxx. 
+The only exceptions were that we focused on the autosomal SNP and it was necessary to 
 increase the number of permutations for the controlling of the fdr from 100 to 500.
@@
 The genome wide results from the analyses of the mouse data are shown in Figure X. The mouse study took
-measurements on 200 traits.  However, we focused our analyses on the 43 traits whose snp-trait associations could be
+measurements on 200 traits.  However, we focused our analyses on the 43 traits whose SNP-trait associations could be
 corroborated by prior published evidence.
-Each plot contains the number of snp-trait associations that were found and in agreement with the original findings.
-Neither method found snp not identified in the original mouse study.
-As we saw in the simulation study, there is a notable difference in the two methods capacity to discover snp-trait associations.
+Each plot contains the number of SNP-trait associations that were found and in agreement with the original findings.
+Neither method found SNP not identified in the original mouse study.
+As we saw in the simulation study, there is a notable difference in the two methods capacity to discover SNP-trait associations.
 $Eagle^{default}$, under its default settings, for x traits found the same number of findings as FaST-LMM and for y traits found more findings.
@@
 with its regularisation parameter fine tuned to the trait, for x traits found the same number of findings as FaST-LMM and for y traits 
-found more findings. Overall, $Eagle^{default}$ and $Eagle^{optimal}$ found xx\% and yy\% more snp-trait associations than a 
+found more findings. Overall, $Eagle^{default}$ and $Eagle^{optimal}$ found xx\% and yy\% more SNP-trait associations than a 
 single-locus analysis of the mouse data.
@@
 The phenotypic data consisted of measurements from 200 behavioural, tissue, and physiological traits.  Of these traits, 
-43 yielded snp-trait associations that could be corroborated through other independent published work. It was these 
+43 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
 43 traits that were the focus of our real data analyses. Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
@@
 
-Eagle is a method for multiple locus association mapping on a genome-wide scale.
+Eagle is a method for multiple-locus association mapping on a genome-wide scale.
 It is based on linear mixed models.
@@
 Eagle also 
-reports as its findings only those snp that are in strongest linkage disequilibrium, and hence closest, to the genes influencing a trait. 
+reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest, to the genes influencing a trait. 
 The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
@@
 
-Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered snp that was 
+Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
 selected in the $k$th iteration. Suppose three iterations of our model building procedure 
 have been performed and say the 500023rd, 15th, and 420th, 
-snp were selected, then $S=\{500023, 15, 420\}$. 
+SNP were selected, then $S=\{500023, 15, 420\}$. 
 Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
@@
  It is common for the columns of $M$ to be in map order but this is not a requirement. 
-The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th snp. 
-The genotypes are coded as -1, 0, and 1 corresponding to snp genotypes AA, AB, and BB, respectively.
+The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
+The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.
 
@@
 Eagle builds the "best" model iteratively, via forward selection. 
-Suppose $s$ iterations of our model building process have already been performed. This means $s$ snp-trait 
+Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
 associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
@@
 \end{equation}
-where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the $S_k$th snp locus which is the $k$th selected snp, 
-$a_{S_k}$ is the additive effect of the $S_k$th snp locus, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  snp genotypes 
-with the data for the selected snp in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
+where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the $S_k$th SNP locus which is the $k$th selected SNP, 
+$a_{S_k}$ is the additive effect of the $S_k$th SNP locus, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
+with the data for the selected SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
 $a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
@@
 
-\subsubsection{Multiple locus methods}
+\subsubsection{Multiple-locus methods}
 
-We compared the computational and statistical performance of Eagle against five other multiple locus methods for 
+We compared the computational and statistical performance of Eagle against five other multiple-locus methods for 
 association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
@@
 
-MLMM, of the five multiple locus methods, is the closest in philosophy to Eagle. It too is based on  building the 
+MLMM, of the five multiple-locus methods, is the closest in philosophy to Eagle. It too is based on  building the 
 best  linear mixed model via forward selection using the ExtBIC criterion. 
@@
 BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
-the development of multiple locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
+the development of multiple-locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
 By introducing a penalty function that balances bias and variance, regularisation methods can 
@@
 
-\subsubsection{Single locus methods}
+\subsubsection{Single-locus methods}
 
-We were also interested in comparing Eagle to single locus methods to measure the difference in run times and statistical power.
+We were also interested in comparing Eagle to single-locus methods to measure the difference in run times and statistical power.
 Even by limiting our focus to LMM-based methods, there were a number of efficient applications to choose from.
@@
 We compared the performance of Eagle against seven other association mapping applications. 
-Five of these applications, BigRR, glmnet, FaST-LASSO, MMLM, r2VIM implement methods that have been applied to the multiple locus analysis of data from genome-wide association studies. 
+Five of these applications, BigRR, glmnet, FaST-LASSO, MMLM, r2VIM implement methods that have been applied to the Multiple-locus analysis of data from genome-wide association studies. 
 The other two applications, GEMMA, and FaST-LMM implement single-locus association mapping methods. 
@@
 
-\subsubsection{Multiple locus methods}
+\subsubsection{Multiple-locus methods}
 
@@
 
-\subsubsection{Single locus methods}
+\subsubsection{Single-locus methods}
 

### 2018-06-27 Figure2_time.tex
27June18 methods v4

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

\begin{figure}
\caption{Median run times, in minutes, for the analysis of association study data from the six simulation scenarios. 
Eagle is compared against five other multiple-locus programs (top left plot) and two single-locus programs (top right plot). 
The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
programs, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
The actual median run times for the programs across the scenarios are given in the table. The entries in a bold font 
correspond to the lowest run-time for a scenario. 
 FaST-LMM$^{all}$ is where the single-locus results are based on all the available 
SNP data. FaST-LMM$^{few}$ is where the single-locus results as based on a subset of the available SNP data. }

\label{fig_time}
\begin{center}
\includegraphics[width=12cm, height=12cm]{Figure2_time.jpg}
\end{center}

\end{figure}

\end{document}

### 2018-06-27 Table1_methodssummary.tex
27June18 methods v4

\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

\begin{table}
\caption{A summary of the features possessed by the Eagle package and the comparison implementations.}
\begin{tabular}{lccccccccc}
                                &  \multicolumn{9}{c}{Computer Software for Association Mapping} \\ \cline{2-10}
                                 & \multicolumn{6}{c}{Multiple-locus}  & & \multicolumn{2}{c}{Single-locus} \\  \cline{2-7}   \cline{9-10}
 Features                 & Eagle  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
Purpose built$^1$    &   \cmark     &    \cmark      &  \xmark   &  \cmark  &  \cmark  &  \cmark && \cmark & \cmark           \\ [0.2cm]
Well documented$^2$           & \cmark &   \xmark   & \cmark   & \xmark     &   \cmark     & \xmark          && \cmark    &    \cmark    \\  [0.2cm]
Simultaneous   &    \cmark     &    \cmark        &   \cmark          &      \cmark             &  \xmark          &  \xmark              &&   \xmark   &  \xmark     \\  
fitting of SNPs     &         &            &             &                   &            &                &&      &      \\  [0.2cm]
Additional fixed effects$^3$          &   \cmark   &      \cmark      &     \cmark    &  \xmark                 &      \cmark      &      \cmark          &&      \cmark & \cmark     \\   [0.2cm]
Data larger than RAM$^4$                   &   \cmark    &      \xmark      & \xmark          &  \xmark    &  \xmark    &   \xmark     && \cmark      & \xmark    \\  [0.2cm]
Threshold free$^5$        &    \cmark     &  \xmark    &  \xmark         &    \xmark               &     \cmark       &  \xmark              &&  \xmark   & \xmark      \\  [0.2cm]
Informative error checking           &  \cmark      &     \xmark        &      \xmark     &           \xmark        &      \xmark       &         \xmark       && \cmark  & \xmark \\ \hline

\end{tabular}
{$^1$ \scriptsize{Computer software specifically designed for the analysis of data from GWAS.}}\\
{$^2$ \scriptsize{More than just a readme file or comments in an example file. Those programs with ticks had detailed user manuals.}}\\
{$^3$ \scriptsize{Ability to accommodate additional fixed effects in the model such as age, sex, and population structure effects.}} \\
{$^4$ \scriptsize{Able to deal with data larger than the memory capacity of the computer.}} \\
{$^5$ \scriptsize{Results reported as the set of SNP closest to the genes influencing a trait. No need to construct thresholds to determine 
significance of the findings.}}\\
{$^6$  \scriptsize{All the programs terminated on errors. However, not all the programs informed the user of the cause and how to fix the errors.}}\\
\end{table}

\end{document}

### 2018-06-27 supmaterials.tex
27June18 methods v4

@@
 \usepackage{authblk}
+\usepackage{amsfonts}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
@@
 \author[1]{Andrew W. George}
+\author[1]{Arunas Verbyla}
 \author[2]{Joshua Bowden}
@@
 
-\affil[1]{Data61, CSIRO, Brisbane, Australia.}
-\affil[2]{IM \&T, CSIRO, Brisbane, Australia.}
+\affil[1]{Data61, CSIRO, Australia.}
+\affil[2]{IM \&T, CSIRO, Australia.}
 
@@
 
+\begin{table}
+\label{suptabsummary}
+\caption{A summary of the features possessed by the Eagle package and the comparison implementations.}
+\begin{tabular}{lccccccccc}
+                                &  \multicolumn{9}{c}{Computer Software for Association Mapping} \\ \cline{2-10}
+                                 & \multicolumn{6}{c}{Multiple-locus}  & & \multicolumn{2}{c}{Single-locus} \\  \cline{2-7}   \cline{9-10}
+ Features                 & Eagle  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
+Purpose built$^1$    &   \cmark     &    \cmark      &  \xmark   &  \cmark  &  \cmark  &  \cmark && \cmark & \cmark           \\ [0.2cm]
+Well documented$^2$           & \cmark &   \xmark   & \cmark   & \xmark     &   \cmark     & \xmark          && \cmark    &    \cmark    \\  [0.2cm]
+Simultaneous   &    \cmark     &    \cmark        &   \cmark          &      \cmark             &  \xmark          &  \xmark              &&   \xmark   &  \xmark     \\  
+fitting of SNPs     &         &            &             &                   &            &                &&      &      \\  [0.2cm]
+Additional fixed effects$^3$          &   \cmark   &      \cmark      &     \cmark    &  \xmark                 &      \cmark      &      \cmark          &&      \cmark & \cmark     \\   [0.2cm]
+Data larger than RAM$^4$                   &   \cmark    &      \xmark      & \xmark          &  \xmark    &  \xmark    &   \xmark     && \cmark      & \xmark    \\  [0.2cm]
+Threshold free$^5$        &    \cmark     &  \xmark    &  \xmark         &    \xmark               &     \cmark       &  \xmark              &&  \xmark   & \xmark      \\  [0.2cm]
+Informative error checking           &  \cmark      &     \xmark        &      \xmark     &           \xmark        &      \xmark       &         \xmark       && \cmark  & \xmark \\ \hline
+
+\end{tabular}
+{$^1$ \scriptsize{Computer software specifically designed for the analysis of data from GWAS.}}\\
+{$^2$ \scriptsize{More than just a readme file or comments in an example file. Those programs with ticks had detailed user manuals.}}\\
+{$^3$ \scriptsize{Ability to accommodate additional fixed effects in the model such as age, sex, and population structure effects.}} \\
+{$^4$ \scriptsize{Able to deal with data larger than the memory capacity of the computer.}} \\
+{$^5$ \scriptsize{Results reported as the set of SNP closest to the genes influencing a trait. No need to construct thresholds to determine 
+significance of the findings.}}\\
+{$^6$  \scriptsize{All the programs terminated on errors. However, not all the programs informed the user of the cause and how to fix the errors.}}\\
+\end{table}
+
 \begin{figure}
-\caption{Memory usage (in Gigabytes) of Eagle and the other association mapping programs across 
-the six simulation scenarios. The maximum amount of memory on the computer is 128 Gigabytes. 
+\caption{Memory usage (in gigabytes) of Eagle and the other association mapping programs across 
+the six simulation scenarios. The maximum amount of memory on the computer is 128 gigabytes. 
 The x-axis is on the log scale. GEMMA, a single-locus implementation, had the lowest memory usage. 
@@
 was the only multiple-locus 
-implementation able to produce results for data under  scenario 10000x1.5M. This is due to its ability 
-to handle data larger than the available memory of a computer. FaST-LMM was run where all the snp data are used 
-to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth snp are used to 
+implementation able to produce results for data under  scenario 10000 x 1.5M. This is due to its ability 
+to handle data larger than the available memory of a computer. FaST-LMM was run where all the SNP data are used 
+to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
 estimate the relationship matrix (FaST-LMM$^{few}$)}.
@@
 \begin{table}
-\caption{The median run times (in minutes) of Eagle and the other association mapping programs across the six simulation scenarios. The }
+\caption{The median run times (in minutes) of Eagle and the other association mapping programs across the six simulation scenarios. }
 \begin{tabular}{llcccccc}
               &           &  \multicolumn{6}{c}{Simulation Scenarios} \\ \cline{3-8}
- Method & Name & 150x5K & 1500x50K & 350x400K & 2000x500K & 4000x1.5M & 10000x1.5M \\ \hline
+ Method & Name & 150 x 5K & 1500 x 50K & 350 x 400K & 2000 x 500K & 4000 x 1.5M & 10000 x 1.5M \\ \hline
   Multiple & Eagle 	&	   0.08  &   1.62   &  \bf{2.7}1  &   \bf{13.65}   & \bf{127.63}  &   699.55  \\

### 2018-06-27 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -132,21 +132,21 @@ p <- p+ theme(legend.key.width=grid:::unit(1.5,"cm"))
 
 
 ## add text to plots
-p <- p +  annotate("text", x=pps, y=rep(9000,length(pps)), label= c("150x5K" , 
+p <- p +  annotate("text", x=pps, y=rep(9000,length(pps)), label= c("150 x 5K" , 
                                                                     "",
-                                                                    "350x400K" , 
+                                                                    "350 x 400K" , 
                                                                     "",  
-                                                                    "4000x1.5M",  
+                                                                    "4000 x 1.5M",  
                                                                     ""),
                    size=4)
 
 
 p <- p +  annotate("text", x=pps, y=rep(6000,length(pps)), label= c("" , 
-                                                                    "1500x50K",
+                                                                    "1500 x 50K",
                                                                     "" ,  
-                                                                    "2000x500K", 
+                                                                    "2000 x 500K", 
                                                                     "",  
-                                                                    "10000x1.5M"),
+                                                                    "10000 x 1.5M"),
                    size=4)
 
 
@@ -191,7 +191,7 @@ p_time
   
   
   
-  df[df<1] <- 1  ## so that it will plot on the log scale
+  #df[df<1] <- 1  ## so that it will plot on the log scale
   
   df<- melt(df)
   
@@ -322,9 +322,11 @@ p_time
   
   
   
- # grid.newpage()
-#  grid.draw(cbind(ggplotGrob(p_time), ggplotGrob(p_abstime), size = "last"))
+ grid.newpage()
+  grid.draw(cbind(ggplotGrob(p_time), ggplotGrob(p_abstime), size = "last"))
  # dev.off()
+  stop()
+  
   
 
   
@@ -427,21 +429,21 @@ p_time
   
   
   ## add text to plots
-  p <- p +  annotate("text", x=pps, y=rep(129,length(pps)), label= c("150x5K" , 
+  p <- p +  annotate("text", x=pps, y=rep(129,length(pps)), label= c("150 x 5K" , 
                                                                       "",
-                                                                      "350x400K" , 
+                                                                      "350 x 400K" , 
                                                                       "",  
-                                                                      "4000x1.5M",  
+                                                                      "4000 x 1.5M",  
                                                                       ""),
                      size=5)
   
   
   p <- p +  annotate("text", x=pps, y=rep(123,length(pps)), label= c("" , 
-                                                                     "1500x50K",
+                                                                     "1500 x 50K",
                                                                      "" ,  
[…]

### 2018-06-29 AM+draft.tex
27June18 prior to removing a big chunk of methods text
29June18 1st draft Results completed

@@
 \bibliographystyle{naturemag}
-\renewcommand{\baselinestretch}{1}
+\renewcommand{\baselinestretch}{1.3}
 
 %% external tex files
-\externaldocument{Figure2_time}
+\externaldocument{figure1_runtimes}
+\externaldocument{figure2_powerFDRmultiple}
+\externaldocument{figure3_mouse}
+
+\externaldocument{supmaterials}
 
@@
 
- In Table \ref{suptabsummary}, 
-we list some of the key features that make computer programs for association mapping useful. We also identify which of these 
+ We list some of the key features that make computer programs for association mapping useful in 
+Supplementary Table \ref{suptabsummary}. We also identify which of these 
 features are present in Eagle and the other seven implementations. For example, all but glmnet were purpose-built for 
@@
 provided advice on how to fix the errors.
-
-{\em Table 1 goes around here}
 
@@
 
-\subsection{Run times and memory usage}
+\subsection{Memory Usage and Run Times}
 
@@
 Memory usage
-for many of the computer programs was the limiting factor (See Sup Figure 1).  The single-locus program GEMMA was by 
+for many of the computer programs was the limiting factor (See {\bf Supplementary Figure \ref{supfigmem}}).  The single-locus program GEMMA was by 
 far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
@@
 Even FaST-LMM, a single-locus
-implementation, required more than 128 Gbytes of memory for the analysis of the larger data sets 
+implementation, required more than 128 gigabytes of memory for the analysis of the larger data sets 
 when all the marker data was used to calculate the relationship matrix. Of the multiple-locus programs, only Eagle,  
@@
  Towards the final stages of writing this paper, 
- we gained access to a XXXXXXXXXX, a faster desktop computer. We reran Eagle on data from the largest
-  scenario 10000 x 1.5M to measure the impact, if any, on run time. The median run time dropped by more than 50\% 
-  from just under 12 hours to 4.56 hours.
+ we gained access to a desktop computer with 14-core Xeon processors with 256 gigabytes of RAM. We reran Eagle on data from the largest
+  scenario 10000 x 1.5M to measure the impact on run time. The median run time dropped by more than 70\% 
+  from just under 12 hours to 3.31 hours.
 
@@
  an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
- increasing behaviour of the curves in the memory (Sup Figure 1) and run-time plots (Figure \ref{fig_time}).  It 
-it is the study dimensions, not number of genotypes, that drives the computing resources required by 
+ increasing behaviour of the curves in the memory (Supplementary Figure \ref{supfigmem}) and run-time plots (Figure \ref{fig_time}).  It 
+is the study dimensions, not number of genotypes, that drives the computing resources required by 
 association mapping programs.
 
-\em{ Figure 2 goes here}
+{\em Figure \ref{fig_time} goes around here}
 
-\subsection{Power and false discovery rates}
+\subsection{Power and False Discovery Rates}
 
@@
 association with the trait, we were able to assess the validity of a  method's findings. When a replicate was analysed, 
- a method's findings were counted as true if the SNPs were located within 40 kb of SNP in true 
+ a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of SNP in true 
 association with the trait. To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
-in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the fdr of a method, for each replicate, 
+in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the false discovery rate of a method, 
+for each replicate, 
 we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.
 
-The power and fdr of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure X.
+The power and false discovery rate of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
 We restricted our attention to these scenarios for the multiple-locus methods because for scenario 4000 x 1.5M,
 the data could only be analysed by Eagle, MLMM, and r2VIM. For scenario 10000 x 1.5M, the data could only be analysed by Eagle. 
-The power and fdr of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in Sup Figure X.  Each plot contains 
-single points and power curves. The single points are the power and fdrs for Eagle and MLMM.
+The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
+Supplementary Figure \ref{supfigpowersingle}.  Each plot contains 
+single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
 These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
@@
 significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
-the fdr as the significance threshold  is adjusted.
+the false discovery rate as the significance threshold  is adjusted.
 
-At first glance, Eagle appears underpowered. Several of the other methods have higher power than Eagle. The goal of association 
-mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
+In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
+well.  Of the multiple-locus methods, Eagle has the highest power
+while keeping its false discovery rate low (Figure \ref{figpowermultiple}). MLMM also performed well. However, it is when Eagle is compared against single-locus methods 
+that the difference in power is most noticeable.  Eagle has much greater power than single-locus methods for finding SNP in true 
+association with a trait while avoiding false associations (Supplementary Figure \ref{supfigpowersingle}).
+
+%%At first glance, Eagle appears underpowered.
+Several of the other methods have higher power than Eagle.
+The goal of association mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
 For those methods with higher power than Eagle, the cost is much higher false discovery rates.
-In other words, these methods are finding more true results
-but they are also finding many more false results. If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than 
-5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest 
-fdr.
+In other words, these methods are finding more true results  but they are also finding many more false results.
+If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than  5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest  fdr.
 This is especially true when the power and fdr of Eagle is compared to the single-locus methods.
@@
 
-\begin{figure}
-\label{fig_power_multiple}
-\caption{Power curves for  multiple-locus methods}
-\includegraphics[width=10cm]{powerMultiple}
-\end{figure}
-
-\begin{figure}
-\label{fig_power_single}
-\caption{Power curves for single-locus methods -- Put into sup methods}
-\includegraphics[width=10cm]{powerSingle}
-\end{figure}
-
-\begin{figure}
-\label{fig_threshold}
-\caption{The impact of threshold on FDR }
-\includegraphics[width=10cm]{threshold}
-\end{figure}
+{\em Figure \ref{figpowermultiple} goes here}
 
@@
 We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
- We chose to focus on data from a large outbred mouse study.
+ We chose to focus on data from a large outbred mouse study \cite{nicod2016genome}.
 This study was unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
-Analyses based on dosages rather than discrete genotypes have been shown to have greater power of detection \cite{zheng2011comparison}.
-By converting the dosages into genotypes and analysing the data with the single-locus program FaSt-LMM, we obtained a subset of those findings reported in the original study.
+Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \cite{zheng2011comparison}.
+By converting the dosages into genotypes and analysing the data with the single-locus program FaST-LMM, we obtained a subset of those findings reported in the original study.
 We then analysed the data with Eagle.
@@
 However, we were
- able to confirm the validity of these new findings as they matched what was found in the original study.
+ able to confirm the validity of these new findings as they matched what was found in the original study. Having the ability to confirm new findings 
+ in a real study was 
+ one of the primary motivators for choosing these data for analysis. 
+For the single-locus analysis of the data, we followed the same procedure as originally followed for the analysis of the mouse data \cite{nicod2016genome}.
+The only differences were that we focused on the autosomal SNP and it was necessary to
+increase the number of permutations for the controlling of the false discovery rate from 100 to 500.
 
-For the single-locus analysis of the data, we followed the same procedure as xxxx. 
-The only exceptions were that we focused on the autosomal SNP and it was necessary to 
-increase the number of permutations for the controlling of the fdr from 100 to 500.
+Eagle was run in two ways; under its default settings (Eagle$^{default}$) and where we specified the regularisation parameter for model selection (Eagle$^{optimal}$ ) Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}.
+  The conservativeness of the extBIC can be adjusted by a single regularisation parameter that ranges from zero to one.
+In the simulation study, this parameter was set to one, its most conservative and default setting.
+However, there is also opportunity to set the parameter to a value less than one.
+This increases power but also increases the false discovery rate.
+For each trait, we used permutation to set the regularisation parameter to give a false discovery rate of 5\% .
 
-Eagle was run in two ways; under its default settings ($Eagle^{default}$) and where we specified the regularisation parameter for model selection ($Eagle^{optimal}$ ) Eagle chooses the best model via the extended BIC criteria (ref).
-The conservativeness of the extended BIC can be adjusted by a single regularisation parameter that ranges from 0 to 1.
-In the simulation study, this parameter was set to 1, its most conservative and default setting.
-However, there is also opportunity to set the parameter to a value less than 1.
-This increases power but also increases the fdr.
-For each trait, we used permutation to set the regularisation parameter to give a fdr of 5\% .
-
-The genome wide results from the analyses of the mouse data are shown in Figure X. The mouse study took
-measurements on 200 traits.  However, we focused our analyses on the 43 traits whose SNP-trait associations could be
-corroborated by prior published evidence.
+The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}. The mouse study took
+measurements on 200 traits. When these traits were first analysed in the original study, findings for 45 of these traits were able to be 
+corroborated by prior published evidence. We focused our analyses here on these same 45 traits. For 39 traits, we found SNP-trait associations. 
+For the other six, neither FaST-LMM nor Eagle found any associations. 
 Each plot contains the number of SNP-trait associations that were found and in agreement with the original findings. 
-Neither method found SNP not identified in the original mouse study.
+Neither method found SNP not identified in the original mouse study so neither method found false positives. 
 As we saw in the simulation study, there is a notable difference in the two methods capacity to discover SNP-trait associations.
-$Eagle^{default}$, under its default settings, for x traits found the same number of findings as FaST-LMM and for y traits found more findings.
-$Eagle^{optimal}$,
-with its regularisation parameter fine tuned to the trait, for x traits found the same number of findings as FaST-LMM and for y traits 
-found more findings. Overall, $Eagle^{default}$ and $Eagle^{optimal}$ found xx\% and yy\% more SNP-trait associations than a 
-single-locus analysis of the mouse data.
+Eagle$^{default}$, under its default settings, for eight traits found the same number of findings as FaST-LMM and for 28 traits found more findings.
+Eagle$^{optimal}$,
+with its regularisation parameter fine tuned to the trait, for six traits found the same number of findings as FaST-LMM and for 32 traits 
+found more findings. Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, snp-trait findings respectively. 
+Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
+associations than what is the established way of analysing these data. Furthermore, these are all findings that were confirmed in the original study.
 
@@
 We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected snp.  We then fit this new model to the data
-via maximum likelihood and calculate its extended Bayesian information criteria (ExtBIC) \cite{chen2008extended}.
-The ExtBIC is a model selection measure that takes into account the number of unknown parameters and the complexity
+via maximum likelihood and calculate its extended Bayesian information criteria (extBIC) \cite{chen2008extended}.  The 
+extBIC is a model selection measure that takes into account the number of unknown parameters and the complexity 
 of the model space.
 It is especially well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
-If this new model has a larger ExtBIC than the current model, then the $s+1$th selected snp is added to
-the current model and the above process is repeated. If this new model has a smaller ExtBIC than the current model, then the 
+If this new model has a larger  than the current model, then the $s+1$th selected snp is added to
+the current model and the above process is repeated. If this new model has a smaller extBIC than the current model, then the 
 model building process is complete. The set of snp in strongest association with the trait is the $s$ snp previously identified.
@@
 ======================= The above needs a bit more detail may need to look up papers ================
-
-BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
-In ridge regression, the coefficients of the predictors are shrunk 
-to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary. 
-The model used in BigRR can be formulated as a linear mixed model.
-One of the challenges in using BigRR for association mapping is that while it can estimate the effect sizes of all the snp simultaneously, 
-obtaining their statistical significance is difficult. 
-Permutation has been suggested as a way of calculating significance 
-empirically \cite{shen2013novel} but we instead opted for stability selection (see below).
-
-BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
-the development of multiple-locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
-By introducing a penalty function that balances bias and variance, regularisation methods can 
-handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
-MLMM is similar in philosophy to Eagle in that it treats association mapping as a model building problem. Here, a linear mixed 
-model is built through a forward-backward selected strategy.  
-R2VIM is different to the others in that it implements 
-random forests for association mapping. 
-Random forests is a machine learning approach that forms an ensemble of decision trees for regression and 
-classification. 
-These five methods were selected because of their demonstrated value 
-for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
-either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
-features is given in Table X. Each method is now described in a little more detail.
-
-\textbf{BigRR:} BigRR \cite{shen2013novel}  is an R package that implements generalized ridge regression.
-In ridge regression, the coefficients of the predictors are shrunk 
-to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary, giving opportunity for 
-a more realistic model to be specified. The model used in BigRR can be formulated as a linear mixed model that is 
-$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
-effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$. 
-The parameters are estimated via a 
-algorithm that makes clever use of the Cholesky decomposition of the genomic relationship matrix, 
-and the equivalence between the above model and animal models \ref{stranden2009animalmodel}. 
-One of the challenges in using BigRR for association mapping is that while it has the capacity to estimate the effect sizes of all the snp simultaneously, 
-obtaining their statistical significance is difficult. 
-Permutation has been suggested as a way of calculating significance 
-empirically \cite{shen2013novel} but we instead opted for stability selection.
-
-\textbf{LMM-Lasso:} LMM-Lasso  \cite{rakitsch2013lasso} is a stand-alone computer program that implements lasso within a linear mixed model 
-framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
-coefficients of "unimportant" predictors are shrunk to exactly zero.
-The (full) model is 
-$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
-$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
-It is challenging, computationally,  to apply lasso directly to the full model for parameter estimation. Instead, a two step procedure 
-is adopted. In the first step, the above model without any snp effects is fitted. In the second step, 
-the unknown variance components in the full model are replaced by their estimates 
-that were calculated in the first step. Lasso is then applied to this simpler model making lasso more tractable, computationally. 
-As with BigRR,  the statistical significance of the snp effects are not calculated. We use stability selection to obtain the 
-significance empirically.
-
-\textbf{Glmnet:} Glmnet  \cite{Friedman2010glmnet}, unlike BigRR and LMM-Lasso, was not purpose built for 
-association mapping.  It is a general R package for 
-regression analysis. We have included it in our list of applications for comparison with Eagle because the 
-regularization method, elastic net, upon which the package is based has been found to be superior to ridge regression and 
-lasso (ref needed???).  Elastic net uses a penalty function that is a mixture of 
-$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
-This allows elastic net to avoid the limitations of lasso 
-while still being able to perform shrinkage and variable selection, simultaneously. 
-Glmnet can handle a range of different models. For the analyses performed in this paper, the model is 
-$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
-We used stability selection to obtain the significance of the coefficients (snp effects).
-
-\textbf{MLMM:} Of the multiple-locus applications considered in this paper, 
-MLMM \cite{segura2012efficient} bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
-association mapping by building the "best" linear mixed model.  The model used for analysis 
-is the same as (\ref{eq1}) except that the genetic effect is 
-\begin{equation}
-\label{eqmlmm}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + g
-\end{equation}
-where 
-$g^{(n ; \times 1)}$ is a vector of random polygenic effects with distribution 
-$g \sim N(0, \sigma^2_g K)$. The model is built iteratively, via forward-backward variable selection.  In Eagle, at each iteration, 
-the snp in strongest association with the trait is found by calculating the score statistic $t_j^2$.  In MLMM, at each iteration, 
-a separate linear mixed model analysis is performed for each snp not already selected. The snp yielding the most 
-significant result (lowest p-value) is selected for inclusion in the fixed effects part of (\ref{eqmlmm}).  
-Two selection 
-criteria are available for the model building process: 
-the EBIC and the multiple-Bonferroni criterion. We used only the EBIC.
-
-\textbf{r2VIM:} R2VIM is an R package specifically designed for association mapping \cite{szymczak2016r2vim}. 
-It implements random forests and a new way of measuring the importance of a snp.
-In random forests, the worth of a predictor is measured, empirically, by calculating its importance score.
-It is from these importance scores that snp can be ordered in terms of their strength of association with a trait.
-The challenge though is in knowing what proportion of the
-highest ordered snp are in true association. R2VIM addresses this by calculating a relative importance score for a snp. Briefly,
-$m$ random forest analyses of the data are performed. Each analysis is reliant upon a different random number seed. 
-From an analysis, a relative importance score for a snp is obtained by taking the ratio of its importance score to the absolute 
-minimum of the importance scores across all the snp.  Those 
-snp with high relative scores across all $m$ analyses are deemed to be of interest \cite{szymczak2016r2vim}.
-Applying r2VIM to the analysis of GWAS data is not without its challenges.  First, it is not possible to force every 
-decision tree to have important (fixed) effects without a change to the internal workings of the r2VIM package. 
-So we adopted a two step 
-strategy. In the first step, the fixed effects are regressed on to the trait and the residuals recorded.  In the second step, 
-r2VIM is employed were the trait data are the residuals obtained from the first step. This is not ideal (ref to problems with two
-step approaches).  Second, there is no relationship between relative importance scores and significance levels. Consequently, 
-setting an appropriate genome-wide threshold is problematic. We avoid this problem by focusing, in the simulation study, 
-only on the relationship between threshold level and its false discovery rate and power.
 

### 2018-06-29 ForAri/draft29June.tex
29June18 1st draft Results completed

--- 2018-06-29 AM+draft.tex
@@
 
-\subsection{Eagle Approach for Multiple-locus Association Mapping}
-
-Eagle is a method for multiple-locus association mapping on a genome-wide scale.
-It is based on linear mixed models.
-It differs from most other single- and multiple-locus association mapping methods.
-Eagle treats association mapping as a model selection instead of parameter estimation problem.
-Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
-Eagle also 
-reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest, to the genes influencing a trait. 
-The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
-quantitative trait loci in experimental crosses \cite{verbyla2007analysis}
-
-Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
-selected in the $k$th iteration. Suppose three iterations of our model building procedure 
-have been performed and say the 500023rd, 15th, and 420th, 
-SNP were selected, then $S=\{500023, 15, 420\}$. 
-Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
-Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
-may be taken of the same group/line/strain. 
- It is common for the columns of $M$ to be in map order but this is not a requirement. 
-The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
-The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.
-
-The specifics of the Eagle method are as follows. 
-Eagle builds the "best" model iteratively, via forward selection. 
-Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
-associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
-To perform the $s+1$th  iteration, we first fit the current model to the data. 
-The (current) model is of the form 
-\begin{equation}
-\label{eq1}
-y = X \tau + Z u_g + e
-\end{equation}
-where 
-$X^{(n \; \times \; p)}$ and $Z^{( n \; \times \; n_g)}$ are known design matrices with $X$ being of full rank and $Z$ 
-containing zeros and ones that assign the appropriate genetic effect to each measurement. 
-The vector 
-$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
-$u_g^{(n_g \; \times \; 1)}$ contains the 
-genetic effects. The vector of residuals is 
-$e^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \; \times \; n)})$. 
-So far,  this model differs little from standard linear mixed models for association mapping (refs). 
-However, 
-it is how we specify $u_g$ that distinguishes our model from others.
-
-The genetic effects $u_g$ are modelled as 
-\begin{equation}
-\label{eq2}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + M_{-S} a_{-S}
-\end{equation}
-where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the $S_k$th SNP locus which is the $k$th selected SNP, 
-$a_{S_k}$ is the additive effect of the $S_k$th SNP locus, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
-with the data for the selected SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
-$a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
-The terms in the summation on the left hand side are fixed effects.  The other term is a random effect.  The terms 
-in the summation account 
-for the additive effects of the selected snp that are in linkage disequilibrium with genes that are influencing the trait. The other term models 
-snp-trait associations along the entire genome, simultaneously, except for those snp that have already been selected. 
-This is a simple genetic model but it 
-is effective for discovering snp-trait associations.
-
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood estimation.
-
-Third, we identify the $(s+1)$th snp that is in strongest association with the trait, based on the maximum score statistic
-$t_j^2 = \frac{ \widetilde{a} _j^2}{\textrm{var}(\widetilde{a}_j)}$ where $\widetilde{a}_j$ is the best linear unbiased predictor (BLUP), 
-and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
-identify a snp based on its (random) effect size and accuracy, but is theoretically justified.  It follows from outlier detection 
-in linear and linear mixed models (ref).
-
-Fourth, we determine the importance of the $(s+1)$th selected snp via a model selection strategy. 
-We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected snp.  We then fit this new model to the data
-via maximum likelihood and calculate its extended Bayesian information criteria (extBIC) \cite{chen2008extended}.  The 
-extBIC is a model selection measure that takes into account the number of unknown parameters and the complexity 
-of the model space.
-It is especially well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
-If this new model has a larger  than the current model, then the $s+1$th selected snp is added to
-the current model and the above process is repeated. If this new model has a smaller extBIC than the current model, then the 
-model building process is complete. The set of snp in strongest association with the trait is the $s$ snp previously identified.
-
-\subsubsection{Reducing the dimension of the model}
-In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
-The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
-$L$, the number of snp, can be extremely large.  An alternative model is given by 
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
-They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
-\begin{equation}
-\label{eq3}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + (M_{-S} M_{-S}^T)^{1/2} a^*_{-S}
-\end{equation}
-where $a^* \sim N(0, \sigma_a^2 I^{(n_g \; \times \;  n_g)})$, and 
-$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via single value decomposition (ref).  
-Although it may not be obvious, the two models are equivalent, 
-having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
-model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
-effects needing estimating.
-
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model  (\ref{eq3})  with 
-\begin{equation}
-\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
-\end{equation}
-where its variance matrix is
-\begin{equation}
-\label{eq4}
-\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
-\end{equation}
-These values are needed in order to calculate the score statistic $t_j^2$ for identifying the snp in strongest association with the trait. 
-Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation. 
-of (\ref{eq4}).
-
-\subsection{Comparison Methods}
-
-\subsubsection{Multiple-locus methods}
-
-We compared the computational and statistical performance of Eagle against five other multiple-locus methods for 
-association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
-These five methods were selected because of their demonstrated value 
-for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
-either stand-alone computer programs or as  packages. See Table X for a summary of their features.
-
-BigRR, LMM-Lasso, and glmnet  implement three different regression-based regularisation methods; generalized ridge regression, 
-lasso, and elastic net, respectively. Regularisation methods can handle parameter estimation problems where the number of predictors is
- far greater than the number of  samples. This is made possible through a penalty function that balances bias against variance. 
-The application of regularisation methods to association mapping has appeal because it means that the snp effects across an 
-entire genome can be estimated, simultaneously. 
-All three methods are sophisticated pieces of statistical machinery but the significance of an effect, which 
-is needed in association mapping,  can not be calculated easily. Permutation as been suggest as an empirical solution to this 
-limitation \cite{shen2013novel} 
-but we have instead opted for stability selection (see below).
-
-MLMM, of the five multiple-locus methods, is the closest in philosophy to Eagle. It too is based on  building the 
-best  linear mixed model via forward selection using the ExtBIC criterion. 
-However, there are several differences between the two methods.
- MLMM does not make use of dimension reduction and they are implemented quite differently leading to substantial 
- differences  in run time and memory usage (see Results).  Also, they differ in how a model is built. 
- Eagle uses a score statistic formed from parameter estimates of the random genetic effect in which to identify the next snp to enter 
- the model.  MLMM uses the statistical significance of a snp when it is treated as a fixed effect in the model. 
- This necessitates fitting a separate linear mixed model for each unidentified snp. Fortunately,  MLMM does do this 
- efficiently.
-
-R2VIM is quite different to the other four methods. Here, a non-parametric model-free rather than model-based statistical approach is adopted. 
-R2VIM implements random forests and a new way of measuring the importance of a snp.  
-In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
-It is from these importance scores that snp can be ordered in terms of their strength of association with a trait. 
-The difficulty is in knowing what proportion of the 
-highest ordered snp are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
-random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score (f) is calculated for 
-each snp. The greater f is than one, the more evidence there is that the snp is in association with a trait.  A snp-trait association is 
-declared if f for the snp is above some threshold value across all parallel runs. The difficulty is that the relationship between 
-this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
-this would be computationally challenging for all but the smallest of association studies.
-
-======================= The above needs a bit more detail may need to look up papers ================
-
-\subsubsection{Single-locus methods}
-
-We were also interested in comparing Eagle to single-locus methods to measure the difference in run times and statistical power.
-Even by limiting our focus to LMM-based methods, there were a number of efficient applications to choose from.
-We decided on GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
-These two applications have the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.
-They were developed at the same time, independently, but are very similar theoretically.
-Both perform a single spectral decomposition of the relationship matrix $K$.
-Both use the eigenvector matrix to rotate the data.
-Both reformulate the log likelihood and
-restricted/residual maximum-likelihood (REML) 
-log likelihood into a sum of $n$ terms that are easier to compute.
-The difference lies in their estimation procedure.
-FaST-LMM implement's the Brent's algorithm to optimise $\delta$.
-GEMMA instead implement's the Newton-Raphson algorithm.
-Newton-Raphson is more complicated in that it also requires the first and second derivatives of a function to be calculated.
-However, it is superior in terms of its convergence properties to the Brent algorithm.
-Both applications are stand-alone computer programs, popular, and in common use.
-
-\section{Stability Selection}
-When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 
-BigRR has an approximate way of setting each snp's separate regularization parameter (ref). However, the optimal value of 
-$\lambda$ must be found when using LMM-Lasso and glmnet. Also, 
-all three applications yield the effect sizes of the snp across the entire genome but not their statistical significance. To address 
-these issues, we made use of stability selection \cite{meinshausen2010stability}. 
-Stability selection is a resampling strategy.  With stability selection, we were able to avoid having to optimise $\lambda$ while 
-still being able to calculate, empirically, the statistical significance of the snp effects.
-
-We performed stability selection as follows. For LMM-Lasso and glmnet, we performed a preliminary analysis to find an 
-appropriate value for the regularization parameter.  We adjusted $\lambda$ so that they yielded 
-between 10 to 30 snp with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
-but limiting LMM-Lasso and glmnet to only 10 to 30 non-zero effect sizes was straightforward. We then repeatedly subsampled, 
-without replacement, the data. As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$. We 
-analysed the subsample, with $\lambda$ fixed for LMM-Lasso and glmnet.  To calculate, empirically,  the statistical significance of each 
-snp across the genome, we counted the number of times a snp had a non-zero effect size over all the replicates to the total 
-number of replicates (which was 100).
-
-For BigRR, we modified our stability selection procedure slightly. 
-There was no need to find an appropriate value for the regularization 
-parameters as the BigRR package has an internal procedure for their 
-calculation. We draw subsamples as above and analyzed the data with BigRR. However, within an analysis, we ordered 
-the snp according to the absolute size of their snp effects and recorded the top 20 snp. We then measured the significance 
-of the snp across the entire genome as above.
-
-\cite{meinshausen2010stability}
-
-\section{Simulation Study}
-
-To explore the computational and statistical performance of Eagle, we conducted a large simulation data.  Genome-wide data 
-were generated under five different scenarios. These scenarios were data where the sample size was X and the number of snp Y (X x Y) ,
-
-We generated data under five different scenarios.
-
-We generated 
-replicates via data perturbation (ref: Zhao et al. 2007).
-
-These are a GWAS of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
-50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
-1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  
-We chose these scenarios to reflect some of the different sized GWAS being performed in animals, plants, and humans. 
-For each scenario, we generated 100 replicates of data. A single replicate consists of snp genotype data and quantitative trait data. 
-We obtained the snp data from the publicly available 1000 genome project (phase 3). The quantitative trait data we generated from the 
-snp data by selecting a set of snp loci, assigning allelic effects to these snp, and aggregating these effects for each individual along with 
-random error. The number of snp selected per replicate follows a Poisson distribution with mean 30.
- The quantitative trait was generated to have a heritability of 50\%.
-Analyses were performed on a high end desktop computer.
-It had dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM.
-All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
-
-Replicates are generated with data perturbation (Zhao et al.
-2007).
-Data perturbation makes use of the phenotypic and genotypic data observed in a real study to create a simu- lated (quantitative) trait.
-It affords us the opportunity to generate replicates whose analysis more closely mirrors the complexities of analyzing real data.
-For computational expediency, our simulation study is based on data observed from the Biloela site in 2006.
-In generating trait data for a replicate, we assume a broad based heritability of 0.6, there is a single major QTL, the polygenic component consists of 30 polygenes where the effect of a polygene is formed from an unmapped SNP which was chosen randomly, and the variance structure of the simulated trait closely follows that of the loaf volume trait.
-The marker data consists of those genotypes collected in the association study from the 3129 SNP across the 21 homologous chromosomes.
-Each replicate has the same marker data.
-It is the simulated trait data that varies across replicates.
-Our simulation study consists of three parts.
-First, we conduct a null study to investigate whether our Monte Carlo sampling approach controls correctly the genome- wide type I error rate.
-We also evaluate the impact of assuming a block diagonal structure for V.
-We do this by
-also calculating the genome-wide type I error rates when the variance matrix for the joint distribution of the test sta- tistic is formed from loci pairings across the entire genome.
-Ten thousand replicates are generated where the QTL has no effect.
-Results are reported for QK-based association mapping implemented as a single-stage and weighted two- stage analysis.
-Second, a power study is performed to investigate if there is a difference in performance between implementing QK-based association mapping as a single-stage analysis versus a two-stage analysis.
-Data are generated under QTL of different sizes and with varying percentages of missing marker data (1.1, 5, 9, 11 %).
-Missing marker genotypes are imputed for a marker locus by coding its SNP geno- types as ?1 and 1, normalizing its marker data, and replac- ing missing genotypes with zero values.
-One thousand replicates are generated for each of the four different per- centages of missing marker data and each differently sized QTL.
-A genome-wide signi cance level of 5 % is assumed.
-Third, a study to evaluate our heuristic procedure for choosing between single-stage and two-stage analysis is performed.
-Data are generated under a QTL of no, mod- erate, and large effect.
-One thousand replicates are gen- erated for each differently sized QTL.
-The performance of our heuristic procedure is measured by calculating the proportion of replicates for which the analysis type is inferred correctly.
-To determine if a single-stage analy- sis is truly needed, it is necessary to analyze fully the genome-wide data via single-stage and weighted two-stage analysis.
-Then, for each marker locus j, perform the test
-| ? log padj ? log qadj| > | ? log padj ? 1.3|, where we have jjj
-assumed a genome-wide signi cance level of 5 %. If true, an incorrect  nding occurs if our heuristic procedure deter- mines that it is valid to implement two-stage analysis.
-\section{Implementation}
-
-\section{New section written on iPhone added 5 Feb 2018}
-\section{Simulation Study}
-We conduct a large simulation study to examine the performance characteristics of Eagle.
-Genome-wide data are generated under five different scenarios.
-These are a gwas of size 150 individuals and 5000 snp (150 x 5K), 350 individuals and 400000 snp (350 x 400k), 1500 individuals and 50000 snp (1500 x 50k), 200 individuals and .....
-We chose these five scenarios to reflect some of the different sized studies being run in animals, plants, and humans.
-For each scenario, 100 replicates are generated.
-
-To generate a replacate, we used data perturbation (ref). Data perturbation makes use of real genotype data from which a (quantitative) trait is created.
-With data perturbation, we are able to preserve some of the challenges of analysing real data while still having trait data whose genetic architecture is known.
-
-The genotype data comes from the 1000 genome project, version 3 (ref).
-The original data contains X individuals with XX XX snp.
-From these data, according to the specifications of the scenario, the individuals and the snp loci are identified at random.
-What about polymorphic affect numbers?
-Chech).
-
-To generate the quantitative trait data, first the number of snp that are to be in complete with an underlying qtl (x) is drawn from a poison distribution with mean equal to 30.
-Second, x snp loci are select randomly from a scenario?s set of genotype data.
-Third, the snp values are summed across the selected loci for each individual to generate a vector of n genetic values.
-Here, each snp is having, roughly, the same effect on the trait, remembering that the maf of a snp also impacts a snps influence on a trait.
-Fourth, n deviates are drawn from a $N(0,sigma^2_e)$ where $\sigma_e^2$ has been set to yield a trait with a heritability of 0.6.
-(Put in formula).
-Fifth, the quantitative trait data is generate as $y = g + e$.
-In this way, the same genotype data are used, within a scenario, for all replicates.
-Only the quantitative trait data varies across replicates.
-
-\section{Implementation}
-
-Eagle has been implemented as an R package of the same name.
-However, much of the calculation is performed through c++ functions for improved performance.
-Eagle has been specially designed to harness the multithreaded math libraries blast and lapack but it will run on a single processor computer.
-It can also deal with data larger than the memory capacity of a computer.
-Eagle can be run as a standard R package or via the web based GUI.
-The GUI relinquishes the need to know R to use Eagle.
-
-Eagle can be downloaded from CRAN and installed in the usual way.
-A containerised version of Eagle is also available.
-It is built with singularity (ref).
-This container already has Eagle installed on a multi-threaded version of R.
-The container is downloaded as a single file from ccxx
-
-A quick start guide, how to videos, installation instructions, and answers to frequently asked questions are available at the Eagle website (http://...)
-
-$------$
-
-We compared the performance of Eagle against seven other association mapping applications. 
-Five of these applications, BigRR, glmnet, FaST-LASSO, MMLM, r2VIM implement methods that have been applied to the Multiple-locus analysis of data from genome-wide association studies. 
-The other two applications, GEMMA, and FaST-LMM implement single-locus association mapping methods. 
-The packages 
-differ in their features (summarized in Table X).  The packages also differ in their underlying methods (summarized below).
-
-\subsubsection{Multiple-locus methods}
-
-\textbf{BigRR:}  In the BigRR package,  generalised ridge regression is implemented  for
-the analysis of genetic data \cite{shen2013novel}.  Generalised ridge regression 
- is a regularisation method in that it produces regression solutions 
-that have higher bias than ordinary least squares but reduced variance for improved prediction accuracy. The underlying 
-model is  $y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of genetic effects, and each genetic 
-effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.  The 
-amount of shrinkage is allowed to vary across the genetic (snp) effects. This differs to 
-standard ridge regression where the shrinkage is constant. 
-Generalised ridge regression is apt at estimating the effect sizes of snp on a genome-wide scale but  obtaining their 
-statistical significance analytically is difficult. Shen \cite{shen2013novel} suggested permutation as a way of calculating significance 
-empirically. We instead opted for stability selection (see below of details).
-
-\textbf{glmnet:}  Glmnet is a general purpose R package for regression analysis  \cite{Friedman2010glmnet}. 
-The model used for analysis is 
-$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
-All snp effects are modelled, simultaneously. 
-Glmnet implements elastic net for the estimation of the parameters. Elastic net is a regularisation method where the 
-penalty function is a mixture of 
-$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
-This allows elastic net to avoid the limitations of lasso 
-while still being able to perform shrinkage and variable selection, simultaneously \cite{zou2005regularization}. 
-Glmnet was not specifically built for association mapping. 
-However, it is included in our list of comparison applications due to its solid performance for genomic prediction 
-\cite{heslot2012genomic,ogutu2012genomic}.
-The glmnet package 
-implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. 
-From exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
-We instead fixed the shrinkage parameter to a suitable value 
-and employed stability selection to estimate the significance of the snp coefficients, empirically (see below).
-
-\textbf{LMM-Lasso:} LMM-Lasso is a stand-alone computer program for association mapping within a linear 
-mixed model framework \cite{rakitsch2013lasso}. The (full) linear mixed model used for analysis is 
-$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
-$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
-All snp effects are modelled, simultaneously. However, a limitation of this model is its inability to accommodate 
-additional fixed effects (there is no $X \beta$ term). 
-The parameters are estimated via lasso \cite{Tibshirani1996lasso}. Lasso is a regularisation method with the capacity to perform parameter estimation and variable selection, simultaneously.  
-Lasso is computationally intensive, especially when applied to linear mixed models. To avoid this problem, a two-step 
-estimation procedure is adopted.  In the first step, the above model but without any snp effects is fitted via 
-maximum likelihood.
-In the second step, the unknown variance components in the full model are replaced by their estimates that were calculated in the first step.
-This turns the full model from a linear mixed model into a simpler linear model.
-Here, lasso is far more tractable, computationally.
-To assess the significance of the snp coefficients, as suggested \cite{rakitsch2013lasso}, we implemented stability selection (see below).
-
-\textbf{MLMM:}  Of the multiple-locus applications considered in this paper, 
-MLMM bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
-association mapping by building the "best" linear mixed model.  The model used for analysis 
-is the same as (\ref{eq1}) except that the genetic effect is 
-\begin{equation}
-\label{eqmlmm}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + g
-\end{equation}
-where 
-$g^{(n ; \times 1)}$ is a vector of random polygenic effects with distribution 
-$g \sim N(0, \sigma^2_g K)$. The model is built iteratively, via forward-backward variable selection.  In Eagle, at each iteration, 
-the snp in strongest association with the trait is found by calculating the score statistic $t_j^2$.  In MLMM, at each iteration, 
-a separate linear mixed model analysis is performed for each snp not already selected. The snp yielding the most 
-significant result (lowest p-value) is selected for inclusion in the fixed effects part of (\ref{eqmlmm}).  
-Two selection 
-criteria are available for the model building process: 
-the EBIC and the multiple-Bonferroni criterion. We used only the EBIC.
-
-\textbf{r2VIM:} R2RIM is an R package specifically designed for association mapping \cite{szymczak2016r2vim}. 
-It implements random forests and a new way of measuring the importance of a snp.
-In random forests, the worth of a predictor is measured, empirically, by calculating its importance score.
-It is from these importance scores that snp can be ordered in terms of their strength of association with a trait.
-The challenge though is in knowing what proportion of the
-highest ordered snp are in true association. R2VIM addresses this by calculating a relative importance score for a snp. Briefly,
-$m$ random forest analyses of the data are performed. Each analysis is reliant upon a different random number seed. 
-From an analysis, a relative importance score for a snp is obtained by taking the ratio of its importance score to the absolute 
-minimum of the importance scores across all the snp.  Those 
-snp with high relative scores across all $m$ analyses are deemed to be of interest \cite{szymczak2016r2vim}.
-
-\subsubsection{Single-locus methods}
-
-\textbf{FaST-LMM:}  FaST-LMM is a highly efficient stand along computer program for association 
-mapping \cite{lippert2011fast}. It is based on a standard single-locus model 
-\cite{yu2006unified,zhao2007arabidopsis}.
-Its computational efficiencies come through factorising $K$ with spectral decomposition so that $K=U (S + \delta I) U^T$ where $U$ contains the eigenvectors and $\delta$ is the ratio of the genetic variance to the residual variance.
-By rotating the data with $U$, the log likelihood of the linear mixed model can be expressed as the log likelihood for a simple 
-linear model.  This greatly reduces the computational cost of performing lmm-based  association mapping. With FaST-LMM, 
-analyses scale linearly with sample size $n$, both in terms of run-time and memory usage.
-
-\textbf{GEMMA:}  to be done.
-
-\section{Discussion}
+{\em Still working on the rest of the methods ...}
 

### 2018-06-29 ForAri/figure2_powerFDRmultiple.tex
29June18 1st draft Results completed

--- 2018-06-27 Figure2_time.tex
@@
 \documentclass[10pt]{article}
+\usepackage{xr}
+
 \usepackage{graphicx}
@@
 \addtolength{\textwidth}{2.5cm}
+
+\setcounter{figure}{1}    
 \begin{document}
@@
 \begin{figure}
-\caption{Median run times, in minutes, for the analysis of association study data from the six simulation scenarios. 
-Eagle is compared against five other multiple-locus programs (top left plot) and two single-locus programs (top right plot). 
-The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
-programs, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
-The actual median run times for the programs across the scenarios are given in the table. The entries in a bold font 
-correspond to the lowest run-time for a scenario. 
- FaST-LMM$^{all}$ is where the single-locus results are based on all the available 
-SNP data. FaST-LMM$^{few}$ is where the single-locus results as based on a subset of the available SNP data. }
-
-\label{fig_time}
+\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
+scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
+scenarios but MLMM also performs well. }
+\label{figpowermultiple}
 \begin{center}
-\includegraphics[width=12cm, height=12cm]{Figure2_time.jpg}
+\includegraphics[width=15cm, height=15cm]{power1main.eps}
 \end{center}

### 2018-06-29 ForAri/figure3_mouse.tex
29June18 1st draft Results completed

\documentclass[10pt]{article}
\usepackage{xr}

\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}

\setcounter{figure}{2}    
\begin{document}

\begin{figure}
\caption{Genome-wide association mapping results from the analysis of the mouse data for the single-locus method FaST-LMM and the 
multiple-locus method
Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
for a false discovery rate of 5\% (Eagle$^{optimal}$). 
The number of SNP-trait associations found are reported. The bar plots with boxes around them are for those traits where using 
Eagle$^{default}$ and Eagle$^{optimal}$ resulted in more findings than with single-locus analysis.} 
\label{figmouse}
\begin{center}
%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
\includegraphics[width=15cm, height=15cm]{mouseresults.eps}
\end{center}

\end{figure}

\end{document}

### 2018-06-29 ForAri/supmaterials.tex
29June18 1st draft Results completed

--- 2018-06-27 supmaterials.tex
@@
 \documentclass{article}
+\usepackage{xr}
+
 \usepackage{graphicx}
@@
 \newcommand{\xmark}{\ding{55}}%
+
+\renewcommand{\tablename}{Supplementary Table}
+\renewcommand{\figurename}{Supplementary Figure}
 
@@
 \begin{table}
+\caption{A summary of the features possessed by the Eagle package and the comparison implementations.}
 \label{suptabsummary}
-\caption{A summary of the features possessed by the Eagle package and the comparison implementations.}
+
 \begin{tabular}{lccccccccc}
@@
 estimate the relationship matrix (FaST-LMM$^{few}$)}.
+\label{supfigmem}
+
 \begin{center}
@@
 \end{center}
+\end{figure}
+
+\begin{figure}
+\caption{Power verse false discovery rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
+FaST-LMM was run where all the SNP data are used 
+to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
+estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
+
+\label{supfigpowersingle}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{power2main.eps}
+\end{center}
+
 \end{figure}

### 2018-06-29 AM+/Plots_for_Paper/power_plots.R

@@ -246,11 +246,27 @@ p <- p+ theme(legend.key.width=grid:::unit(1.5,"cm"))
 #p + theme_few()
 
 ## xlimit
-p <- p + coord_cartesian(xlim = c(0, 0.05), ylim=c(0, 1)) 
+p <- p + coord_cartesian(xlim = c(0, 0.1), ylim=c(0, 1)) 
 
-jpeg("/Users/geo047/Papers/AM-Paper/power1main.jpg", width=7, height=7, units="in", res=500)
+# change number of digits in labels on x-axis
+scaleFUN <- function(x) sprintf("%.2f", x)
+p <- p + scale_x_continuous(labels=scaleFUN)
+
+
+
+#jpeg("/Users/geo047/Papers/AM-Paper/power1main.jpg", width=7, height=7, units="in", res=500)
+#print(p)
+#dev.off()
+
+
+postscript("~/Papers/AM-Paper/power1main.eps", width=10, height=10, fonts=c("sans"),
+           horizontal=FALSE)
+# square plot
 print(p)
 dev.off()
+
+
+
 
 p1 <- p
 
@@ -310,22 +326,25 @@ insert_plot <- function(df1, df2, family){
 
 }
 
+# inset pltos 
+# decided not to use these
 
-jpeg("/Users/geo047//Papers/AM-Paper/power1sub1.jpg", width=12, height=12, units="in", res=2000)
-print(insert_plot(df1, df2, "150 x 5K"))
-dev.off()
 
-jpeg("/Users/geo047/Papers/AM-Paper/power1sub2.jpg", width=12, height=12, units="in", res=2000)
-print(insert_plot(df1, df2, "350 x 500K"))
-dev.off()
+#jpeg("/Users/geo047//Papers/AM-Paper/power1sub1.jpg", width=12, height=12, units="in", res=2000)
+#print(insert_plot(df1, df2, "150 x 5K"))
+#dev.off()
 
-jpeg("/Users/geo047/Papers/AM-Paper/power1sub3.jpg", width=12, height=12, units="in", res=2000)
-print(insert_plot(df1, df2, "1500 x 50K"))
-dev.off()
+#jpeg("/Users/geo047/Papers/AM-Paper/power1sub2.jpg", width=12, height=12, units="in", res=2000)
+#print(insert_plot(df1, df2, "350 x 500K"))
+#dev.off()
 
[…]

### 2018-06-29 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -1,3 +1,14 @@
+## Notes to self
+# 1. this script cannot be run from source. The plots do not write to file.
+# 2. this script sometimes gives plotting errors. Just keep running it with "run" until success.
+# 3. the runtime plots are printed separately. They then need to be read in to powerpoint a long with a 
+#     screen grab of the latex produced table.  This is then saved as a jpeg file to produce Figure 1 in paper.
+
+
+
+
+
+
 ## Results are obtained from bragg
 ## input matrices cputimes.dat and memory.dat, and abstimes.dat
 ## Input obtained from /home/geo047/MWAM/SimStudy/Timing/Results  on Bragg cluster. 
@@ -321,11 +332,13 @@ p_time
   dev.off()
   
   
-  
- grid.newpage()
-  grid.draw(cbind(ggplotGrob(p_time), ggplotGrob(p_abstime), size = "last"))
- # dev.off()
-  stop()
+#  postscript("~/Papers/AM-Paper/time.eps", width=10, height=10, 
+#             horizontal=FALSE)
+# grid.newpage()
+#  grid.draw(rbind(ggplotGrob(p_time), ggplotGrob(p_abstime), size = "last"))
+#  
+#dev.off()
+#  stop()
   
   
 

### 2018-06-29 AM+draft.tex
29June End of day update - new implementation section

@@
 
-======================= The above needs a bit more detail may need to look up papers ================
-
 \subsubsection{Single-locus methods}
@@
 These two applications have the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.
-They were developed at the same time, independently, but are very similar theoretically.
+They were developed at the same time, independently, but are similar theoretically.
 Both perform a single spectral decomposition of the relationship matrix $K$.
@@
 We performed stability selection as follows. For LMM-Lasso and glmnet, we performed a preliminary analysis to find an 
-appropriate value for the regularization parameter.  We adjusted $\lambda$ so that they yielded 
+appropriate value for the regularization parameter.  We adjusted $\lambda$ so that LMM-Lasso and glmnet yielded 
 between 10 to 30 snp with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
-but limiting LMM-Lasso and glmnet to only 10 to 30 non-zero effect sizes was straightforward. We then repeatedly subsampled, 
+ We then repeatedly subsampled, 
 without replacement, the data. As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$. We 
 analysed the subsample, with $\lambda$ fixed for LMM-Lasso and glmnet.  To calculate, empirically,  the statistical significance of each 
-snp across the genome, we counted the number of times a snp had a non-zero effect size over all the replicates to the total 
-number of replicates (which was 100).
+SNP across the genome, we counted the number of times a SNP had a non-zero effect size over all the replicates. We then 
+divided this number by the number of replicates (which was 100) to form an empirical probability estimate of the significance of a SNP.
 
@@
 calculation. We draw subsamples as above and analyzed the data with BigRR. However, within an analysis, we ordered 
-the snp according to the absolute size of their snp effects and recorded the top 20 snp. We then measured the significance 
-of the snp across the entire genome as above.
-
-\cite{meinshausen2010stability}
+the SNPs according to the absolute size of their SNP effects and recorded the top 20 SNPs. We then measured the significance 
+of the SNPs across the entire genome as above.
 
@@
 \section{Simulation Study}
-We conduct a large simulation study to examine the performance characteristics of Eagle.
-Genome-wide data are generated under five different scenarios.
-These are a gwas of size 150 individuals and 5000 snp (150 x 5K), 350 individuals and 400000 snp (350 x 400k), 1500 individuals and 50000 snp (1500 x 50k), 200 individuals and .....
-We chose these five scenarios to reflect some of the different sized studies being run in animals, plants, and humans.
-For each scenario, 100 replicates are generated.
+Genome-wide data are generated under five different scenarios. These five scenarios partially mimic the size of study being performed in 
+animals, plants, and humans. They are a GWAS of size 150 individuals 
+and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
+50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
+1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).
 
-To generate a replacate, we used data perturbation (ref). Data perturbation makes use of real genotype data from which a (quantitative) trait is created.
+To generate a replicate, we used data perturbation (ref).
+Data perturbation makes use of real genotype data from which a (quantitative) trait is created.
 With data perturbation, we are able to preserve some of the challenges of analysing real data while still having trait data whose genetic architecture is known.
+We obtained the genotype data from the 1000 Genome Project, version 3   \cite{10002010map} consisting of over
+78 million autosomal SNPs collected from 2504 individuals across 26 populations.  To form the genotype data, we 
+sampled randomly from the 2504 individuals and 78 million loci according to the specifications of the scenario.
 
-The genotype data comes from the 1000 genome project, version 3 (ref).
+According to the specifications of the scenario,
+we identified, randomly, the individuals and loci. We then extracted this information from the larger data set 
+to form the genotype data for a scenario. The individuals 
+and loci chosen varied across scenarios but the same genotype data were used across the 100 replicates within a scenario.
+
  The original data contains X individuals with XX XX snp.
@@
 
-Eagle has been implemented as an R package of the same name.
-However, much of the calculation is performed through c++ functions for improved performance.
-Eagle has been specially designed to harness the multithreaded math libraries blast and lapack but it will run on a single processor computer.
-It can also deal with data larger than the memory capacity of a computer.
-Eagle can be run as a standard R package or via the web based GUI.
-The GUI relinquishes the need to know R to use Eagle.
-
-Eagle can be downloaded from CRAN and installed in the usual way.
-A containerised version of Eagle is also available.
-It is built with singularity (ref).
-This container already has Eagle installed on a multi-threaded version of R.
-The container is downloaded as a single file from ccxx
-
-A quick start guide, how to videos, installation instructions, and answers to frequently asked questions are available at the Eagle website (http://...)
-
-$------$
-
-We compared the performance of Eagle against seven other association mapping applications. 
-Five of these applications, BigRR, glmnet, FaST-LASSO, MMLM, r2VIM implement methods that have been applied to the Multiple-locus analysis of data from genome-wide association studies. 
-The other two applications, GEMMA, and FaST-LMM implement single-locus association mapping methods. 
-The packages 
-differ in their features (summarized in Table X).  The packages also differ in their underlying methods (summarized below).
-
-\subsubsection{Multiple-locus methods}
-
-\textbf{BigRR:}  In the BigRR package,  generalised ridge regression is implemented  for
-the analysis of genetic data \cite{shen2013novel}.  Generalised ridge regression 
- is a regularisation method in that it produces regression solutions 
-that have higher bias than ordinary least squares but reduced variance for improved prediction accuracy. The underlying 
-model is  $y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of genetic effects, and each genetic 
-effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.  The 
-amount of shrinkage is allowed to vary across the genetic (snp) effects. This differs to 
-standard ridge regression where the shrinkage is constant. 
-Generalised ridge regression is apt at estimating the effect sizes of snp on a genome-wide scale but  obtaining their 
-statistical significance analytically is difficult. Shen \cite{shen2013novel} suggested permutation as a way of calculating significance 
-empirically. We instead opted for stability selection (see below of details).
-
-\textbf{glmnet:}  Glmnet is a general purpose R package for regression analysis  \cite{Friedman2010glmnet}. 
-The model used for analysis is 
-$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
-All snp effects are modelled, simultaneously. 
-Glmnet implements elastic net for the estimation of the parameters. Elastic net is a regularisation method where the 
-penalty function is a mixture of 
-$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
-This allows elastic net to avoid the limitations of lasso 
-while still being able to perform shrinkage and variable selection, simultaneously \cite{zou2005regularization}. 
-Glmnet was not specifically built for association mapping. 
-However, it is included in our list of comparison applications due to its solid performance for genomic prediction 
-\cite{heslot2012genomic,ogutu2012genomic}.
-The glmnet package 
-implements a modification of the LARS algorithm \cite{efron2004least} for finding the optimal regularisation (solution) path. 
-From exploratory analyses of the simulated data, we found this algorithm yielded a high proportion of false positives. 
-We instead fixed the shrinkage parameter to a suitable value 
-and employed stability selection to estimate the significance of the snp coefficients, empirically (see below).
-
-\textbf{LMM-Lasso:} LMM-Lasso is a stand-alone computer program for association mapping within a linear 
-mixed model framework \cite{rakitsch2013lasso}. The (full) linear mixed model used for analysis is 
-$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
-$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
-All snp effects are modelled, simultaneously. However, a limitation of this model is its inability to accommodate 
-additional fixed effects (there is no $X \beta$ term). 
-The parameters are estimated via lasso \cite{Tibshirani1996lasso}. Lasso is a regularisation method with the capacity to perform parameter estimation and variable selection, simultaneously.  
-Lasso is computationally intensive, especially when applied to linear mixed models. To avoid this problem, a two-step 
-estimation procedure is adopted.  In the first step, the above model but without any snp effects is fitted via 
-maximum likelihood.
-In the second step, the unknown variance components in the full model are replaced by their estimates that were calculated in the first step.
-This turns the full model from a linear mixed model into a simpler linear model.
-Here, lasso is far more tractable, computationally.
-To assess the significance of the snp coefficients, as suggested \cite{rakitsch2013lasso}, we implemented stability selection (see below).
-
-\textbf{MLMM:}  Of the multiple-locus applications considered in this paper, 
-MLMM bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
-association mapping by building the "best" linear mixed model.  The model used for analysis 
-is the same as (\ref{eq1}) except that the genetic effect is 
-\begin{equation}
-\label{eqmlmm}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + g
-\end{equation}
-where 
-$g^{(n ; \times 1)}$ is a vector of random polygenic effects with distribution 
-$g \sim N(0, \sigma^2_g K)$. The model is built iteratively, via forward-backward variable selection.  In Eagle, at each iteration, 
-the snp in strongest association with the trait is found by calculating the score statistic $t_j^2$.  In MLMM, at each iteration, 
-a separate linear mixed model analysis is performed for each snp not already selected. The snp yielding the most 
-significant result (lowest p-value) is selected for inclusion in the fixed effects part of (\ref{eqmlmm}).  
-Two selection 
-criteria are available for the model building process: 
-the EBIC and the multiple-Bonferroni criterion. We used only the EBIC.
-
-\textbf{r2VIM:} R2RIM is an R package specifically designed for association mapping \cite{szymczak2016r2vim}. 
-It implements random forests and a new way of measuring the importance of a snp.
-In random forests, the worth of a predictor is measured, empirically, by calculating its importance score.
-It is from these importance scores that snp can be ordered in terms of their strength of association with a trait.
-The challenge though is in knowing what proportion of the
-highest ordered snp are in true association. R2VIM addresses this by calculating a relative importance score for a snp. Briefly,
-$m$ random forest analyses of the data are performed. Each analysis is reliant upon a different random number seed. 
-From an analysis, a relative importance score for a snp is obtained by taking the ratio of its importance score to the absolute 
-minimum of the importance scores across all the snp.  Those 
-snp with high relative scores across all $m$ analyses are deemed to be of interest \cite{szymczak2016r2vim}.
-
-\subsubsection{Single-locus methods}
-
-\textbf{FaST-LMM:}  FaST-LMM is a highly efficient stand along computer program for association 
-mapping \cite{lippert2011fast}. It is based on a standard single-locus model 
-\cite{yu2006unified,zhao2007arabidopsis}.
-Its computational efficiencies come through factorising $K$ with spectral decomposition so that $K=U (S + \delta I) U^T$ where $U$ contains the eigenvectors and $\delta$ is the ratio of the genetic variance to the residual variance.
-By rotating the data with $U$, the log likelihood of the linear mixed model can be expressed as the log likelihood for a simple 
-linear model.  This greatly reduces the computational cost of performing lmm-based  association mapping. With FaST-LMM, 
-analyses scale linearly with sample size $n$, both in terms of run-time and memory usage.
-
-\textbf{GEMMA:}  to be done.
+Eagle has been implemented as an R package of the same name. Much of the heavy computation though is performed outside of R 
+via c++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
+mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
+calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How To guides, and 
+demos are available on the Eagle website (ref).  Eagle is available for download from the CRAN website.
 

### 2018-07-03 AM+draft.tex
refined simulation section methods. Removing a big chunk.

@@
 \maketitle
+
+%% To Add
+%%-------------------------
+%% 1. how real data was analysed
+%% 2. default settings for methods in simulation study.
+%% 3. real data analyses - how were they done fixed effects?
+%% 4. simulation study very simple in terms of design, fixed effects etc.
 
@@
 \section{New section written on iPhone added 5 Feb 2018}
-\section{Simulation Study}
-Genome-wide data are generated under five different scenarios. These five scenarios partially mimic the size of study being performed in 
-animals, plants, and humans. They are a GWAS of size 150 individuals 
-and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
-50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
-1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).
+\section{Generation of Simulation Data}
 
-To generate a replicate, we used data perturbation (ref).
-Data perturbation makes use of real genotype data from which a (quantitative) trait is created.
-With data perturbation, we are able to preserve some of the challenges of analysing real data while still having trait data whose genetic architecture is known.
-We obtained the genotype data from the 1000 Genome Project, version 3   \cite{10002010map} consisting of over
-78 million autosomal SNPs collected from 2504 individuals across 26 populations.  To form the genotype data, we 
-sampled randomly from the 2504 individuals and 78 million loci according to the specifications of the scenario.
+The data were generated via data perturbation. Data perturbation amalgamates real with simulated data to generate replicates. 
+It is a way of introducing greater realism into a simulation study. 
+ Here, the genotype data are real, the 
+quantitative trait data are simulated. 
+ The SNP genotypes were drawn, 
+according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \cite{10002010map}. 
+Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.
 
-According to the specifications of the scenario,
-we identified, randomly, the individuals and loci. We then extracted this information from the larger data set 
-to form the genotype data for a scenario. The individuals 
-and loci chosen varied across scenarios but the same genotype data were used across the 100 replicates within a scenario.
+To generate the trait data $y$, first, $q$, the number of SNP that are to be assigned a quantitative value is drawn from a Poisson distribution with 
+mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNP. The SNP genotypes AA, AB, and BB 
+are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
+generate a $g^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
+Fifth, $e^{(n \times 1)}, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
+the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $y =  g + e$.
 
- The original data contains X individuals with XX XX snp.
-From these data, according to the specifications of the scenario, the individuals and the snp loci are identified at random.
-What about polymorphic affect numbers?
-Chech).
-
-To generate the quantitative trait data, first the number of snp that are to be in complete with an underlying qtl (x) is drawn from a poison distribution with mean equal to 30.
-Second, x snp loci are select randomly from a scenario?s set of genotype data.
-Third, the snp values are summed across the selected loci for each individual to generate a vector of n genetic values.
-Here, each snp is having, roughly, the same effect on the trait, remembering that the maf of a snp also impacts a snps influence on a trait.
-Fourth, n deviates are drawn from a $N(0,sigma^2_e)$ where $\sigma_e^2$ has been set to yield a trait with a heritability of 0.6.
-(Put in formula).
-Fifth, the quantitative trait data is generate as $y = g + e$.
-In this way, the same genotype data are used, within a scenario, for all replicates.
-Only the quantitative trait data varies across replicates.
+ In forming $y$, 
+we have purposely not included any fixed effects such as age, sex, or experimental design effects. The reason for this is that 
+not all the multiple-locus methods considered could accommodate additional random effects. 
+Although a two-stage approach is often adopted to deal with this limitation, we wanted to 
+generate data that did not require further processing before analysis by any of the methods.
 

### 2018-07-09 AM+draft.tex
9July refining text v1

@@
 \title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George$^1$, Arunas Verbyla$^1$, Joshua Bowden$^2$, need to sort out the other authors}
+\author{Andrew W. George$^1$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
 
@@
 
-\section{Simulation Study}
-
-To explore the computational and statistical performance of Eagle, we conducted a large simulation data.  Genome-wide data 
-were generated under five different scenarios. These scenarios were data where the sample size was X and the number of snp Y (X x Y) ,
-
-We generated data under five different scenarios.
-
-We generated 
-replicates via data perturbation (ref: Zhao et al. 2007).
-
-These are a GWAS of size 150 individuals and 5000 snp (150 X 5K), 350 individuals and 400000 snp (350 X 400K),  1500 individuals and 
-50000 snp (1500 x 50K), 2000 individuals and 500000 snp (2000 x 500K), 4000 individuals and 
-1500000 snp (4000 x 1.5M), and 10000 individuals and 1500000 snp (10000 x 1.5M).  
-We chose these scenarios to reflect some of the different sized GWAS being performed in animals, plants, and humans. 
-For each scenario, we generated 100 replicates of data. A single replicate consists of snp genotype data and quantitative trait data. 
-We obtained the snp data from the publicly available 1000 genome project (phase 3). The quantitative trait data we generated from the 
-snp data by selecting a set of snp loci, assigning allelic effects to these snp, and aggregating these effects for each individual along with 
-random error. The number of snp selected per replicate follows a Poisson distribution with mean 30.
- The quantitative trait was generated to have a heritability of 50\%.
-Analyses were performed on a high end desktop computer.
-It had dual 8-core Xeon processors, three Kepler Tesla GPUs, and 128 Giggabytes of RAM.
-All implementations except, GEMMA, made use of distributed computing, either explicitly or implicitly through multi-threaded BLAS/LAPACK libraries.
-
-Replicates are generated with data perturbation (Zhao et al.
-2007).
-Data perturbation makes use of the phenotypic and genotypic data observed in a real study to create a simu- lated (quantitative) trait.
-It affords us the opportunity to generate replicates whose analysis more closely mirrors the complexities of analyzing real data.
-For computational expediency, our simulation study is based on data observed from the Biloela site in 2006.
-In generating trait data for a replicate, we assume a broad based heritability of 0.6, there is a single major QTL, the polygenic component consists of 30 polygenes where the effect of a polygene is formed from an unmapped SNP which was chosen randomly, and the variance structure of the simulated trait closely follows that of the loaf volume trait.
-The marker data consists of those genotypes collected in the association study from the 3129 SNP across the 21 homologous chromosomes.
-Each replicate has the same marker data.
-It is the simulated trait data that varies across replicates.
-Our simulation study consists of three parts.
-First, we conduct a null study to investigate whether our Monte Carlo sampling approach controls correctly the genome- wide type I error rate.
-We also evaluate the impact of assuming a block diagonal structure for V.
-We do this by
-also calculating the genome-wide type I error rates when the variance matrix for the joint distribution of the test sta- tistic is formed from loci pairings across the entire genome.
-Ten thousand replicates are generated where the QTL has no effect.
-Results are reported for QK-based association mapping implemented as a single-stage and weighted two- stage analysis.
-Second, a power study is performed to investigate if there is a difference in performance between implementing QK-based association mapping as a single-stage analysis versus a two-stage analysis.
-Data are generated under QTL of different sizes and with varying percentages of missing marker data (1.1, 5, 9, 11 %).
-Missing marker genotypes are imputed for a marker locus by coding its SNP geno- types as ?1 and 1, normalizing its marker data, and replac- ing missing genotypes with zero values.
-One thousand replicates are generated for each of the four different per- centages of missing marker data and each differently sized QTL.
-A genome-wide signi cance level of 5 % is assumed.
-Third, a study to evaluate our heuristic procedure for choosing between single-stage and two-stage analysis is performed.
-Data are generated under a QTL of no, mod- erate, and large effect.
-One thousand replicates are gen- erated for each differently sized QTL.
-The performance of our heuristic procedure is measured by calculating the proportion of replicates for which the analysis type is inferred correctly.
-To determine if a single-stage analy- sis is truly needed, it is necessary to analyze fully the genome-wide data via single-stage and weighted two-stage analysis.
-Then, for each marker locus j, perform the test
-| ? log padj ? log qadj| > | ? log padj ? 1.3|, where we have jjj
-assumed a genome-wide signi cance level of 5 %. If true, an incorrect  nding occurs if our heuristic procedure deter- mines that it is valid to implement two-stage analysis.
-\section{Implementation}
-
-\section{New section written on iPhone added 5 Feb 2018}
 \section{Generation of Simulation Data}
 
-The data were generated via data perturbation. Data perturbation amalgamates real with simulated data to generate replicates. 
+The data were generated via data perturbation (Zhao et al. 2007). Data perturbation amalgamates real with simulated data to generate replicates. 
 It is a way of introducing greater realism into a simulation study. 
@@
 generate a $g^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
-Fifth, $e^{(n \times 1)}, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
+Fifth, $e^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
 the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $y =  g + e$.

### 2018-07-10 AM+draft.tex
3Jul revised table sup.

@@
 \begin{abstract}
-Since the earliest of genome-wide association studies, a key shortcoming in how their data have been analysed has persisted.
-The strength of association between a marker locus and trait is measured for each locus separately, on a locus-by-locus basis.
-Multiple-locus methods that map multiple-locus trait associations simultaneously have been available for some time.
-However, they have attracted little attention.
-They can be demanding, computationally, and their results are not always easy to interpret.
-Yet, it is widely accepted that multiple-locus methods are superior, statistically, to locus-by-locus methods.
-Here, we present our method that makes the multiple-locus analysis of data from genome wide association studies routine.
-It is formulated within a linear mixed model framework.
-We call our method AMplus.
-AMplus produces results faster than competing multiple-locus methods and often with greater statistical power.
-Also, it is just a little slower than the fastest single-locus linear mixed model implementations.
-AMplus is freely available as a fully documented R package. 
+
+For over a decade, methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
+Yet a well know limitation has persisted, the strength of association between a locus and trait is assessed on a locus-by-locus basis.
+Superior multiple-locus methods that detect multiple SNP-trait associations, simultaneously, have been available for some time but their application has been minimal.
+Here, we present Eagle for multiple-locus association mapping.
+Our goal was to develop a method that shifts mainstream association mapping from being single-locus to multiple-locus.
+We will show through the analysis of simulated and mouse data that Eagle is more powerful than competing single- and multiple-locus methods and can be orders of magnitude faster.
+Eagle has been implemented as an R package, with an easy to use web-based Graphical User Interface (GUI) for users unfamiliar with R.
 \end{abstract}
 
-Over the past decade,  Genome wide association studies (GWAS) have changed considerably in both their analysis and design. Early GWAS
+Over the past decade,  genome-wide association studies have changed considerably in both their analysis and design. Early studies
  followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
@@
 
-What has not changed over the past decade is that it remains common practice to analyse gwas data on a locus by locus basis.
+What has not changed over the past decade is that it remains common practice to analyse genome-wide association study (GWAS) data on a locus-by-locus basis.
 This is despite there being several significant problems with analysing data in this way.
@@
 The identification of these regions from these analyses is not always straightforward.
-GWAS results are reported, typically, via Manhattan plots that plot the -log10 of the p value for each locus against the map position of the locus.
+GWAS results are reported, typically, via Manhattan plots that plot the -log$_{10}$ of the P value for each locus against the map position of the locus.
 The location of peaks in this plot signify regions of interest.
-Inferring the exact number of regions of interest can be difficult If the peaks are not well separated.
+Inferring the exact number of regions of interest can be difficult if the peaks are not well separated.
 Second, when multiple statistical tests are performed, the probability of wrongly accepting a significant result (type 1 error) is inflated.
@@
 Many different solutions have been offered (refs).
-Yet, there is still no well accepted way of correcting for multiple testing in the context of genome- wide association mapping.
+Yet, there is still no well accepted way of correcting for multiple testing in the context of genome-wide association mapping.
 Third, many of the traits whose genetic secretes we are trying to discover are complex.
 There will be multiple loci in linkage equilibrium with genes that influence the trait.
-Yet, A locus by locus mapping approach only assesses the evidence for association between a single marker locus and trait.
+Yet, a locus-by-locus mapping approach only assesses the evidence for association between a single marker locus and trait.
 
@@
 Also, their results can be difficult to interpret.
-The strength of association is not measured by a p-value but by the size of the regression coefficient for a locus in the model.
+The strength of association is not measured by a P value but by the size of the regression coefficient for a locus in the model.
 More recently, associations have started to be mapped with random forests (refs).
 Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings (refs).
-An multiple-locus method that does show promise is the multi-locus linear mixed model method (ref).
+A multiple-locus method that does show promise is the multiple-locus linear mixed model method (ref).
 The best multiple-locus model is built with simple variable selection.
-Results are immediately interpretable but here, computation can be a challenge for large data.
+Results are immediately interpretable but here, computation can be a challenge if the data set is large.
 
@@
 Eagle combines the strength of regularisation techniques (being able to fit all locus-trait associations jointly), with simple variable selection (having easy to interpret results).
-Our method does not require a significance threshold to be set nor regularisation parameters to be fine tuned.
-Through a clever dimension reduction step, we are able to achieve a computational performance similar to the fastest single-locus linear mixed model implementations.
-Eagle is statistically more powerful than single-locus association mapping and is as and often more powerful than
-most multiple-locus methods.
-Our aim is to make multiple-locus association mapping on a genome wide scale routine.
-To this end, we have created a fully documented R package, that is easy to use, even for non R users.
+Our method does not require a significance threshold to be set.
+Through a simple dimension reduction step, we are able to achieve a computational performance similar to the fastest single-locus linear mixed model implementations.
+Our aim was to make multiple-locus association mapping on a genome wide scale routine. To this end, we have implemented Eagle 
+within an R package of the same name. 
 Our package accepts marker data of different 
-formats and can handle data larger than a computer's  memory capacity. It includes detailed error checking and makes heavy use of 
-distributed computing for computation when available.  The purpose of this work was to make multiple-locus association mapping on 
-a genome-wide scale, for large data sets,  practical and we have built AMplus accordingly.
+formats,  can handle data larger than a computer's  memory capacity, and makes heavy use of 
+distributed computing for computation when available.
 
@@
 the input data contained errors, we found only Eagle and FaST-LMM tried to diagnose the problem and 
-provided advice on how to fix the errors.
+provided guidance on how to fix the errors.
 

### 2018-07-10 supmaterials.tex
3Jul revised table sup.

@@
 \documentclass{article}
+\usepackage{xr}
+\usepackage{lscape}
+
 \usepackage{graphicx}
@@
 \usepackage{pifont}
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
+
+\renewcommand{\tablename}{Supplementary Table}
+\renewcommand{\figurename}{Supplementary Figure}
+\newcommand{\rowgroup}[1]{\hspace{-1em}#1}
 
@@
 \author[2]{Joshua Bowden}
-\author[1]{Some other authors}
 
@@
 
+\begin{landscape}
+
 \begin{table}
+\caption{{\bf Implementation and methodology attributes of eight computer programs/packages for genome-wide association mapping.} For the different types of model, 
+LMM is linear mixed model.  GLM is generalised linear model., GLMM is generalised linear mixed model, and RF is random forests. 
+}
 \label{suptabsummary}
-\caption{A summary of the features possessed by the Eagle package and the comparison implementations.}
+\vspace{0.5cm}
+\begin{tabular}{lcccccccc} \hline
+                                                   Attributes                  & {\bf Eagle}                                 & {\bf bigRR}             & {\bf glmnet}            & {\bf LMM-Lasso}                    & {\bf MLMM} & {\bf r2VIM}      & {\bf FaST-LMM} & {\bf GEMMA} \\  \hline
+{\em Implementation}    &         &            &             &                   &            &                &      &      \\ 
+\hspace{1mm}  Purpose built$^a$    &   Yes     &    Yes      &  No   &   Yes  &  Yes  &  Yes  & Yes  & Yes          \\ [0.35cm]
+
+\hspace{1mm}  Language                 &  R/C++       &    R        &      R       &     Python     &  R          &    R         &  C++ and     &   C++   \\  
+                                                          &         &            &             &                   &            &                                     &   Python$^b$       &      \\  [0.35cm]
+
+\hspace{1mm} GUI                            & Yes &    No      & No          &  No    &  No    &   No     & No     & No    \\  [0.35cm]
+
+\hspace{1mm} Documentation        &  Videos,                &    R help       &  Vignettes,    &   Readme.txt,       &   Vignette,          &     R help           &   Videos,              &  User-manual,     \\
+                                                        &   user-manuals,    &                      &  R help         &    test script           &  R help              &                           &    user-manuals,  &   website  \\
+                                                        &   website,    R help          &                      &                     &                              &                           &                           &    website            &    \\   [0.35cm]
+
+\hspace{1mm} Additional fixed        &   Yes   &      Yes      &     Yes    &  No                 &      Yes      &      Yes          &      Yes & Yes     \\  
+\hspace{1mm} effects$^c$               &           &                   &                 &                   &                   &                     &              &   adsf    \\  [0.35cm]
+
+\hspace{1mm} Types of  &  Cont.      &   Cont.,          &   Cont.,           &   Cont.                 &   Cont.          &     Cont.           &   Cont.   &  Cont.    \\  
+\hspace{1mm} trait data  &                &    binary,      &      binary,         &                   &            &                &      &      \\  
+                                      &                &   count         &       count          &                   &            &                &      &      \\  [0.35cm]
+
+\hspace{1mm} Data larger                   &   Yes    &      No      & No          &  No    &  No    &   No     & Yes      & No    \\  
+\hspace{1mm} than memory                                             &         &            &             &                   &            &                &      &      \\  [0.8cm]
+
+{\em Methodology}    &         &            &             &                   &            &                &      &      \\  
+\hspace{1mm}  Model                     & LMM   &   &  GLMM &   GLM  & LMM  & RF & LMM   & LMM \\ [0.15cm]
+
+\hspace{1mm} SNPs fitted       &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.35cm]
+
+\hspace{1mm}  Selection type              & Model   & Variable  &  Variable &  Variable  &  Model  &  Variable & Variable  & Variable \\  [0.35cm]
+
+\hspace{1mm} Threshold free$^5$        &    Yes     &  No    &  No         &    No               &     Yes       &  No              &  No   & No      \\  [0.15cm]  \hline
+
+\end{tabular}
+{$^a$ \scriptsize{Specifically created for the analysis of GWAS data.}}\\
+{$^b$ \scriptsize{Separate programs, one written in Python, the other C++}} \\
+{$^c$ \scriptsize{Capacity for additional fixed effects (such as  age, sex, and/or population structure effects) to be included directly in the model.}} \\
+\end{table}
+
+\end{landscape}
+
+\begin{landscape}
+
+\begin{table}
+\caption{Some of the key features possessed by Eagle and the other seven computer programs/packages for association mapping}
+\label{suptabsummary}
+
 \begin{tabular}{lccccccccc} 
-                                &  \multicolumn{9}{c}{Computer Software for Association Mapping} \\ \cline{2-10}
+                                &  \multicolumn{9}{c}{Computer Programs/Packages for Association Mapping} \\ \cline{2-10}
                                  & \multicolumn{6}{c}{Multiple-locus}  & & \multicolumn{2}{c}{Single-locus} \\  \cline{2-7}   \cline{9-10}
  Features                 & Eagle  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
-Purpose built$^1$    &   \cmark     &    \cmark      &  \xmark   &  \cmark  &  \cmark  &  \cmark && \cmark & \cmark           \\ [0.2cm]
-Well documented$^2$           & \cmark &   \xmark   & \cmark   & \xmark     &   \cmark     & \xmark          && \cmark    &    \cmark    \\  [0.2cm]
-Simultaneous   &    \cmark     &    \cmark        &   \cmark          &      \cmark             &  \xmark          &  \xmark              &&   \xmark   &  \xmark     \\  
+Purpose built$^1$    &   Yes     &    Yes      &  No   &  Yes  &  Yes  &  Yes && Yes & Yes           \\ [0.2cm]
+Well documented$^2$           & Yes &   No   & Yes   & No     &   Yes     & No          && Yes    &    Yes    \\  [0.2cm]
+Simultaneous   &    Yes     &    Yes        &   Yes          &      Yes             &  No          &  No              &&   No   &  No     \\  
 fitting of SNPs     &         &            &             &                   &            &                &&      &      \\  [0.2cm]
-Additional fixed effects$^3$          &   \cmark   &      \cmark      &     \cmark    &  \xmark                 &      \cmark      &      \cmark          &&      \cmark & \cmark     \\   [0.2cm]
-Data larger than RAM$^4$                   &   \cmark    &      \xmark      & \xmark          &  \xmark    &  \xmark    &   \xmark     && \cmark      & \xmark    \\  [0.2cm]
-Threshold free$^5$        &    \cmark     &  \xmark    &  \xmark         &    \xmark               &     \cmark       &  \xmark              &&  \xmark   & \xmark      \\  [0.2cm]
-Informative error checking           &  \cmark      &     \xmark        &      \xmark     &           \xmark        &      \xmark       &         \xmark       && \cmark  & \xmark \\ \hline
+Additional fixed effects$^3$          &   Yes   &      Yes      &     Yes    &  No                 &      Yes      &      Yes          &&      Yes & Yes     \\   [0.2cm]
+Data larger than RAM$^4$                   &   Yes    &      No      & No          &  No    &  No    &   No     && Yes      & No    \\  [0.2cm]
+Threshold free$^5$        &    Yes     &  No    &  No         &    No               &     Yes       &  No              &&  No   & No      \\  [0.2cm]
+Informative error checking           &  Yes      &     No        &      No     &           No        &      No       &         No       && Yes  & No \\ \hline
 
 \end{tabular}
-{$^1$ \scriptsize{Computer software specifically designed for the analysis of data from GWAS.}}\\
-{$^2$ \scriptsize{More than just a readme file or comments in an example file. Those programs with ticks had detailed user manuals.}}\\
-{$^3$ \scriptsize{Ability to accommodate additional fixed effects in the model such as age, sex, and population structure effects.}} \\
+{$^1$ \scriptsize{Specifically created for the analysis of GWAS data.}}\\
+{$^2$ \scriptsize{More than just a readme file or comments in an example file. Programs/packages with a tick had a detailed user manual.}}\\
+{$^3$ \scriptsize{Ability to handle additional fixed effects such as  age, sex, and/or population structure effects.}} \\
 {$^4$ \scriptsize{Able to deal with data larger than the memory capacity of the computer.}} \\
@@
 \end{table}
+
+\end{landscape}
 

### 2018-07-10 AM+draft.tex
11Jul fine tuning v1

@@
 We compared Eagle, in terms of computational and statistical performance, against seven other association mapping methods. 
-We chose methods that (mostly) had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs.
+We chose methods that almost all had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs or packages.
 Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multiple-locus models.
@@
 LMM-Lasso \cite{rakitsch2013lasso}, MLMM \cite{segura2012efficient} , and r2VIM \cite{szymczak2016r2vim}.  
-Each takes a different approach to multiple-locus association mapping.
-
- We list some of the key features that make computer programs for association mapping useful in 
-Supplementary Table \ref{suptabsummary}. We also identify which of these 
-features are present in Eagle and the other seven implementations. For example, all but glmnet were purpose-built for 
-association mapping.
-Not all the computer programs come with a user manual.
-Most could accomodate data on additional fixed effects but only Eagle and FaST-LMM could deal with data larger than the memory capacity of the computer.
-Also, if
-the input data contained errors, we found only Eagle and FaST-LMM tried to diagnose the problem and 
-provided guidance on how to fix the errors.
+Each takes a different approach to multiple-locus association mapping. A summary of the key attributes to the different computer programs/packages 
+is given in Supplementary Table \ref{suptabsummary}. 
+Further details about the computer programs/packages and how they were used in the analysis of the simulated data can be found in Methods.
 
@@
 
-\subsection{Outbred Mice Study}
 \subsection{Mouse Data}
 
-The data were obtained from a large genome-wide association study which was performed in outbred mice. 
-The study is described in detail in \cite{nicod2016genome}. Phenotypic and genotypic data were available on 1,887 adult mice. 
+The data were obtained from a large genome-wide association study which was performed in outbred mice \cite{nicod2016genome}. 
+Phenotypic and genotypic data were available on 1,887 adult mice. 
 The phenotypic data consisted of measurements from 200 behavioural, tissue, and physiological traits.  Of these traits, 
@@
 It differs from most other single- and multiple-locus association mapping methods.
-Eagle treats association mapping as a model selection instead of parameter estimation problem.
+Eagle treats association mapping as a model selection instead of variable selection problem.
 Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
@@
 for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
-either stand-alone computer programs or as  packages. See Table X for a summary of their features.
+either stand-alone computer programs or as  packages.
 
@@
 All three methods are sophisticated pieces of statistical machinery but the significance of an effect, which 
-is needed in association mapping,  can not be calculated easily. Permutation as been suggest as an empirical solution to this 
+is needed in association mapping,  can not be calculated easily. Permutation has been suggest as an empirical solution to this 
 limitation \cite{shen2013novel} 
@@
 
-\section{Stability Selection}
+\subsection{Stability Selection}
 When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 
@@
 
-\section{Generation of Simulation Data}
+\subsection{Generation of Simulation Data}
 
@@
 the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $y =  g + e$.  
+In forming $y$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
+not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach (ref) 
+is often adopted to deal with this situation, be we chose not to introduce this complexity to the analyses.
 
- In forming $y$, 
-we have purposely not included any fixed effects such as age, sex, or experimental design effects. The reason for this is that 
-not all the multiple-locus methods considered could accommodate additional random effects. 
-Although a two-stage approach is often adopted to deal with this limitation, we wanted to 
-generate data that did not require further processing before analysis by any of the methods.
-
-\section{Implementation}
+\subsection{Implementation}
 
 Eagle has been implemented as an R package of the same name. Much of the heavy computation though is performed outside of R 
-via c++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
+via C++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
 mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
-calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How To guides, and 
-demos are available on the Eagle website (ref).  Eagle is available for download from the CRAN website.
+calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
+a link to our server for demonstrating Eagle on some test data are available on the Eagle website (ref).  
+Eagle is available for download from the CRAN website.
 
 \section{Discussion}
+
+How to deal with more complex models.
+
+Comment on cloud based FaST-LMM
+
+Performance could change with changes to default values.
 

### 2018-07-12 AM+draft.tex
12/07 working on refinement

@@
 
+%% Issue 
+%%====================
+%% Where does us doing stability selection and adding time to jobs come in.
+
 \begin{affiliations}
@@
 
-For over a decade, methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
-Yet a well know limitation has persisted, the strength of association between a locus and trait is assessed on a locus-by-locus basis.
-Superior multiple-locus methods that detect multiple SNP-trait associations, simultaneously, have been available for some time but their application has been minimal.
+Methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
+Yet a well know limitation has persisted, the strength of association between a locus and trait is assessed only on a locus-by-locus basis.
+Statistically superior multiple-locus methods that detect multiple SNP-trait associations, simultaneously, have been available for some time but they have failed to become the method-of-choice.
 Here, we present Eagle for multiple-locus association mapping.
 Our goal was to develop a method that shifts mainstream association mapping from being single-locus to multiple-locus.
-We will show through the analysis of simulated and mouse data that Eagle is more powerful than competing single- and multiple-locus methods and can be orders of magnitude faster.
-Eagle has been implemented as an R package, with an easy to use web-based Graphical User Interface (GUI) for users unfamiliar with R.
+We will show through the analysis of simulated and mouse data that Eagle finds true and avoids false
+associations better than completing single- and multiple-locus methods and does it faster.
+Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users unfamiliar with R.
 \end{abstract}
@@
 
-What has not changed over the past decade is that it remains common practice to analyse genome-wide association study (GWAS) data on a locus-by-locus basis.
+What has not changed is that it remains common practice to analyse genome-wide association study (GWAS) data on a locus-by-locus basis.
 This is despite there being several significant problems with analysing data in this way.
 First, the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
-The identification of these regions from these analyses is not always straightforward.
-GWAS results are reported, typically, via Manhattan plots that plot the -log$_{10}$ of the P value for each locus against the map position of the locus.
-The location of peaks in this plot signify regions of interest.
-Inferring the exact number of regions of interest can be difficult if the peaks are not well separated.
-Second, when multiple statistical tests are performed, the probability of wrongly accepting a significant result (type 1 error) is inflated.
+The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
+that plot the -log$_{10}$ of the P value for each locus against the map position of the locus. The P value is obtained by testing the statistical 
+significance of a SNP when treated as an effect in an appropriate model. 
+The location of peaks in this plot identify genomic 
+regions of interest.
+Inferring the exact number of regions though can be difficult if the peaks are not well separated.
+Second, when multiple statistical tests are performed, the probability of wrongly accepting a result (type 1 error) is inflated.
 This is known as the multiple testing problem (refs).
@@
 Third, many of the traits whose genetic secretes we are trying to discover are complex.
-There will be multiple loci in linkage equilibrium with genes that influence the trait.
+There will be multiple SNPs in linkage disequilibrium with genes that influence the trait.
 Yet, a locus-by-locus mapping approach only assesses the evidence for association between a single marker locus and trait.
@@
 Also, their results can be difficult to interpret.
-The strength of association is not measured by a P value but by the size of the regression coefficient for a locus in the model.
+The strength of association is not measured by a P value but by the size of the regression coefficient for the SNP in the model.
 More recently, associations have started to be mapped with random forests (refs).
@@
 A multiple-locus method that does show promise is the multiple-locus linear mixed model method (ref).
-The best multiple-locus model is built with simple variable selection.
-Results are immediately interpretable but here, computation can be a challenge if the data set is large.
+The best multiple-locus model is built with simple forward selection.
+Results are immediately interpretable but here, computation can become challenging if the datasets become large.
 
-In this paper, we present our new multiple-locus method for genome wide association mapping, which we are calling Eagle.
-Eagle combines the strength of regularisation techniques (being able to fit all locus-trait associations jointly), with simple variable selection (having easy to interpret results).
-Our method does not require a significance threshold to be set.
-Through a simple dimension reduction step, we are able to achieve a computational performance similar to the fastest single-locus linear mixed model implementations.
-Our aim was to make multiple-locus association mapping on a genome wide scale routine. To this end, we have implemented Eagle 
+In this paper, we present our new multiple-locus method for genome-wide association mapping, which we are calling Eagle. 
+Eagle combines the strength of regularisation techniques (being able to fit all SNP-trait associations jointly), with forward selection giving easy-to-interpret threshold-free results.
+We are able to achieve a computational performance similar to the fastest single-locus linear mixed model implementations
+through a dimension reduction step.
+Our aim was to make multiple-locus association mapping on a genome-wide scale routine. To this end, we have implemented Eagle 
 within an R package of the same name. 
@@
 LMM-Lasso \cite{rakitsch2013lasso}, MLMM \cite{segura2012efficient} , and r2VIM \cite{szymczak2016r2vim}.  
-Each takes a different approach to multiple-locus association mapping. A summary of the key attributes to the different computer programs/packages 
+Each takes a different approach to multiple-locus association mapping. A summary of the key attributes of the different computer programs/packages 
 is given in Supplementary Table \ref{suptabsummary}. 
-Further details about the computer programs/packages and how they were used in the analysis of the simulated data can be found in Methods.
+Further details about the methods being implemented in these computer programs/packages can be found in Methods.
 
@@
 
+ A replicate was analysed via the different methods as follows. All programs/packages were run at their default settings. 
+ The Eagle package and MLMM script were the easiest of the programs/packages to implement. 
+ Their results are immediately 
+ interpretable. Their findings are the set of SNPs in strongest association with the trait. Each 
+SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
+
+BigRR, LMM-Lasso, and glmnet  implement regression-based 
+regularisation methods. Coefficient estimates of the SNP from fitting all loci simultaneously are reported. However, for these results to be useable, the statistical significance of a
+SNP in the model is also required.  The statistical significance rather than the coefficient size of a SNP is a better measure of the strength of evidence for a SNP-trait association. 
+Calculating the significance of effects within these regularisation methods is challenging. We resort to 
+stability selection, a sampling approach that allows us to calculate 
+the statistical significance of the 
+SNP empirically (see Methods for details). This does increase the computational burden of implementing these methods for genome-wide association mapping.
+
+R2VIM implements random forests. As part of analysing data with r2VIM, we were required to specify the minimum size of a terminal node, the number of trees, and number of potential variables 
+for building a tree.   HOW TO CHOSE THESE - REASONABLE VALUES ???? LOOK AT PAPERS ...
+
 \subsection{Memory Usage and Run Times}
@@
 implementation, required more than 128 gigabytes of memory for the analysis of the larger data sets 
-when all the marker data was used to calculate the relationship matrix. Of the multiple-locus programs, only Eagle,  
+when all the marker data was used to calculate the similarity matrix. Of the multiple-locus programs, only Eagle,  
 with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
-for data from our largest scenario 10000 x 1.5M.
+for data from our largest scenario, 10000 x 1.5M.
 
@@
  were thousands of individuals and 1.5 million SNPs, Eagle took just over two hours for the analysis of data from 
- 4000 x 1.5M and just under 12 hours for the analysis of data from 10000 x 1.5M. 
+ 4000 x 1.5M and  12 hours for the analysis of data from 10000 x 1.5M. 
  Towards the final stages of writing this paper, 
@@
   scenario 10000 x 1.5M to measure the impact on run time. The median run time dropped by more than 70\% 
-  from just under 12 hours to 3.31 hours.
+  from 12 hours to 3.31 hours.
 
- It is also worth noting that an increase in the number of SNP genotypes in a study does not necessitate, automatically,  
- an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
- increasing behaviour of the curves in the memory (Supplementary Figure \ref{supfigmem}) and run-time plots (Figure \ref{fig_time}).  It 
-is the study dimensions, not number of genotypes, that drives the computing resources required by 
-association mapping programs.
+ %%It is also worth noting that an increase in the number of SNP genotypes in a study does not necessitate, automatically,  
+ %%an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
+ %%increasing behaviour of the curves in the memory (Supplementary Figure \ref{supfigmem}) and run-time plots (Figure \ref{fig_time}).  It 
+%%is the study dimensions, not number of genotypes, that drives the computing resources required by
+%%association mapping programs.
 
@@
 \subsection{Power and False Discovery Rates}
+=============================
+How did we define a finding ?????
+============================
 
@@
 the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these SNPs. By knowing which SNPs are in true 
-association with the trait, we were able to assess the validity of a  method's findings. When a replicate was analysed, 
- a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of SNP in true 
+association with the trait, we were able to assess the validity of a  method's findings.
+
+==>
+
+Of the methods, the findings from Eagle and MLMM are immediately usable. 
+Eagle and MLMM, return as their findings
+ the set of SNPs in strongest association with the trait. Each 
+SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
+For the other methods, FaST-LMM and GEMMA report a P value for each SNP as their primary finding. BigRR, LMM-Lasso, and glmnet, after 
+stability selection has been performed (see Methods),  report a probability of importance for each SNP.
+
+FaST-LMM and GEMMA report P values for the statistical significance of a SNP,
+
+For the other methods, either a P value (.....), relative importance measure (r2VIM), or probability of importance (...,, ..., ...) are reported as the 
+findings. Before the results can be interpreted, the snp are placed in map order, peak regions (based on the significance measure) are identified, 
+a threshold is set, and a set of SNP findings obtained from identifying the SNP with the largest significance measure in those peak regions above the 
+threshold.
+
+findings were reported as either a P value (FaST-LMM and GEMMA), importance
+
+Eagle and MLMM, return as their findings
+ the set of SNPs in strongest association with the trait. Each 
+SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
+BigRR, LMM-Lasso, and glment, after stability selection has been performed (see Methods),
+
+Eagle and MLMM return as their findings the set of SNPs in strongest association with the trait. Each 
+SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP. 
+The other methods required the identification of SNP in 
+BigRR, XX, and XX return as their findings
+
+Interpreting the findings of Eagle and MLMM is trivial. They return as their findings the set of SNPs in strongest association with the trait. Each 
+SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP. 
+Interpreting the findings of the other methods is more difficult
+
+Only those SNP found to be in strongest association
+
+Eagle and MLMM, return as their findings, those SNPs found to be in strongest association with the trait. Each SNP identifies a separate 
+genomic region of interest via it's map location.
+
+The results from Eagle and MLMM required no further processing. Each SNP denotes a separate region of interest where the SNP
+
+To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
+in creating the trait data. We then averaged across the 100 replicates.  
+Similarly, to calculate the false discovery rate of a method, 
+for each replicate, 
+we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.  
+When a replicate was analysed, 
+ a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of a SNP in true 
+association with the trait.
+
+<======
+
+When a replicate was analysed, 
+ a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of a SNP in true 
 association with the trait. To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
@@
 
+\subsection{Analyses of Simulated Data}
+
+\subsection{Analyses of Mouse Data}
+
 \subsection{Implementation}
@@
 
+Much of the computational cost for bigRR, lmmlasso, ,,,, lies not in the analysis but in calculating the significance of the effects. We used 
+stability selection that involved 100 further analyses, though these analyses were based on only 50\% of the original data. 
+These methods give parameter estimates efficiently.
+
+Since we consider the goal of associaiton mapping is to identify those genomic regions housing genes that are influencing a trait, then
+
 %%\bibliographystyle{naturemag}

### 2018-07-13 AM+draft.tex
13July v1

@@
 
- A replicate was analysed via the different methods as follows. All programs/packages were run at their default settings. 
+ A replicate was analysed as follows. All programs/packages were run at their default settings. 
  The Eagle package and MLMM script were the easiest of the programs/packages to implement. 
@@
 SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.  
-
-BigRR, LMM-Lasso, and glmnet  implement regression-based 
-regularisation methods. Coefficient estimates of the SNP from fitting all loci simultaneously are reported. However, for these results to be useable, the statistical significance of a
-SNP in the model is also required.  The statistical significance rather than the coefficient size of a SNP is a better measure of the strength of evidence for a SNP-trait association. 
-Calculating the significance of effects within these regularisation methods is challenging. We resort to 
-stability selection, a sampling approach that allows us to calculate 
-the statistical significance of the 
-SNP empirically (see Methods for details). This does increase the computational burden of implementing these methods for genome-wide association mapping.
+Implementing BigRR, LMM-Lasso, and glmnet required more work. 
+The computer programs/packages are based on regularisation methods. Being able to fit all the SNPs simultaneously within a regression 
+framework is their strength. The analytical challenges of calculating p-values due to the adaptive nature of the estimation procedure is their 
+weakness.
+To solve this problem, at the cost of increased computation, we calculated the significance of a SNP empirically via stability selection (see Methods).
+R2VIM is an R package that implements random forests.
+We were required to 
+specify the minimum size of a terminal node, the number of trees, and number of potential variables 
+for building a tree.
 

### 2018-07-18 AM+draft.tex
19/07/18 working on discussion v1

@@
 the publicly available 1000 Genome Project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
-by selecting, randomly, a set of SNP loci, assigning additive allelic effects to these loci, and then aggregating these effects for 
+by selecting, randomly, a set of SNPs, assigning additive allelic effects to these loci, and then aggregating their effects for 
 each individual along with a random error.  
@@
  A replicate was analysed as follows. All programs/packages were run at their default settings. 
- The Eagle package and MLMM script were the easiest of the programs/packages to implement. 
+ Eagle and MLMM were the easiest of the programs/packages to implement. 
  Their results are immediately 
@@
 SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.  
-Implementing BigRR, LMM-Lasso, and glmnet required more work. 
-The computer programs/packages are based on regularisation methods. Being able to fit all the SNPs simultaneously within a regression 
-framework is their strength. The analytical challenges of calculating p-values due to the adaptive nature of the estimation procedure is their 
-weakness.
-To solve this problem, at the cost of increased computation, we calculated the significance of a SNP empirically via stability selection (see Methods).
-R2VIM is an R package that implements random forests.
-We were required to 
-specify the minimum size of a terminal node, the number of trees, and number of potential variables 
-for building a tree.
-
-R2VIM implements random forests. As part of analysing data with r2VIM, we were required to specify the minimum size of a terminal node, the number of trees, and number of potential variables 
-for building a tree.   HOW TO CHOSE THESE - REASONABLE VALUES ???? LOOK AT PAPERS ...
+The other programs/packages were not as easy to implement  BigRR, LMM-Lasso, and glmnet make use of regularlisation methods. 
+All SNPs are fitted simultaneously in a regression framework.  However, due to the adaptive nature of the estimation procedure, it is challenging, 
+analytically, to calculate P values for the effects in the model. 
+To solve this problem, at the cost of increased computation, we calculated the significance of a SNP empirically 
+via stability selection (see Methods).  R2VIM, a random forest approach, required three parameters to be specified. These are the
+ minimum size of a terminal node, the number of trees, and number of potential variables 
+for building a tree.  It is not clear how "best" to set these parameters so we followed ref. The minimum size of a node was set to 10\% of the sample size. 
+The number of trees was set to 1000. The number of variables was set to 20\% of the number of SNP. A relative importance measure is calculated 
+for each SNP measuring the strength of association with the trait. 
+FaST-LMM and GEMMA implement single-locus association mapping. The P values of the SNP are reported as their results. 
+For all but Eagle and MLMM, the set of SNPs in strongest and significant association with the trait were found 
+by placing the SNP in map order, setting a threshold, identifying peak regions with 
+significance measures above the threshold, and recording the SNP with the largest significance measure in these peak regions.
 
@@
 \subsection{Power and False Discovery Rates}
-=============================
-How did we define a finding ?????
-============================
 
@@
 
-==>
-
-Of the methods, the findings from Eagle and MLMM are immediately usable. 
-Eagle and MLMM, return as their findings
- the set of SNPs in strongest association with the trait. Each 
-SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
-For the other methods, FaST-LMM and GEMMA report a P value for each SNP as their primary finding. BigRR, LMM-Lasso, and glmnet, after 
-stability selection has been performed (see Methods),  report a probability of importance for each SNP.
-
-FaST-LMM and GEMMA report P values for the statistical significance of a SNP,
-
-For the other methods, either a P value (.....), relative importance measure (r2VIM), or probability of importance (...,, ..., ...) are reported as the 
-findings. Before the results can be interpreted, the snp are placed in map order, peak regions (based on the significance measure) are identified, 
-a threshold is set, and a set of SNP findings obtained from identifying the SNP with the largest significance measure in those peak regions above the 
-threshold.
-
-findings were reported as either a P value (FaST-LMM and GEMMA), importance
-
-Eagle and MLMM, return as their findings
- the set of SNPs in strongest association with the trait. Each 
-SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
-BigRR, LMM-Lasso, and glment, after stability selection has been performed (see Methods),
-
-Eagle and MLMM return as their findings the set of SNPs in strongest association with the trait. Each 
-SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP. 
-The other methods required the identification of SNP in 
-BigRR, XX, and XX return as their findings
-
-Interpreting the findings of Eagle and MLMM is trivial. They return as their findings the set of SNPs in strongest association with the trait. Each 
-SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP. 
-Interpreting the findings of the other methods is more difficult
-
-Only those SNP found to be in strongest association
-
-Eagle and MLMM, return as their findings, those SNPs found to be in strongest association with the trait. Each SNP identifies a separate 
-genomic region of interest via it's map location.
-
-The results from Eagle and MLMM required no further processing. Each SNP denotes a separate region of interest where the SNP
-
 To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
@@
 we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.  
-When a replicate was analysed, 
- a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of a SNP in true 
+A finding was counted as true if the SNP is located within 40 kilobase pairs of a SNP in true 
 association with the trait.
-
-<======
-
-When a replicate was analysed, 
- a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of a SNP in true 
-association with the trait. To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
-in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the false discovery rate of a method, 
-for each replicate, 
-we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.
 
@@
 
+Eagle is a new method and R package for multiple locus association mapping.
+It differs from other multiple locus methods.
+It can handle large datasets due to dimension reduction It treats association mapping as a model selection instead of a variable selection problem so we don?t have to deal with multiple testing issues.
+Its results are immediately interpretable.
+As we saw in the simulation study, Eagle has considerably higher power than single-locus methods and is comparable in run time.
+However, It was in the analysis of the mouse data that the true worth of Eagle is evident.
+With Eagle, we found more than double the results than with single locus association mapping, the method of choice and these extra results were all validated as true.
+
 How to deal with more complex models.
@@
 Since we consider the goal of associaiton mapping is to identify those genomic regions housing genes that are influencing a trait, then
+
+======
+
+General introductory para - gwas growing in size.
+
+Epistatic effects
+
+More complex models
+
+Last paragraph
+==============
+We see what we are presented here as the first step in the development of Eagle. GWAS are only growing in size, both in terms of number of 
+samples and number of genotypes. Our goal here was to develop the methods and implement those methods in package that could be run on a desktop. 
+However, the computational power of desktop will not suffice for datasets where there are hundreds of thousands of samples and millions of loci. 
+Here, computer clusters will be needed to meet the computational challenges and we have written the Eagle package with this in mind. We do not 
+keep the marker data in memory. Also, we perform block multiplication of matrices that would otherwise be too large to fit into memory. These are all 
+preemptive steps to make the porting of Eagle onto a cluster environment easier.
 

### 2018-07-19 AM+draft.tex
19/07/18 working on discussion v2

@@
 
-Eagle is a new method and R package for multiple locus association mapping.
+Eagle is a new method (and R package) for multiple locus association mapping.
 It differs from other multiple locus methods.
@@
 However, It was in the analysis of the mouse data that the true worth of Eagle is evident.
-With Eagle, we found more than double the results than with single locus association mapping, the method of choice and these extra results were all validated as true.
+With Eagle, we found more than double the results than with single locus association mapping, the method of choice.
+The truth of these extra results were all validated.
+
+What we present in this paper is a first step in the development of Eagle. GWAS are only becoming larger. 
+Ever larger GWAS are becoming feasible, as the cost of collecting genetic data from 
+array-based WGS technologies continues to drop.  We are already experimenting with a GPU-based version of Eagle. Early results suggest 
+that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance. However for larger study sizes, 
+we are seeing a 50\% increase in speed, and this is with prototype code. We also have plans to develop Eagle for clusters, and our larger than 
+memory calculations were a preemptive step on our part in this direction.  Interestingly, as were were in the final stages of drafting this paper, 
+we became aware of Microsofts further advancement of FaST-LMM where they have implemented it for the cloud. This is an exciting development, 
+enabling huge datasets to be analysed quickly. However, it still has the inherint weaknesses association with single-locus association mapping.
+
+How we decided on stability selection for the empirical calculation fo signfi of SNP effects for ..., ..., and ... deserves some disucssion.
+
+In implementing the regularisation methods, ..., ..., and ..., we opted to calculate the significance of the SNP effects via stability selection. 
+However, there are other empirical approaches. 
+Permutation 
+Multi sample-splitting approach.
+
+Stability selection is not the only empirical approach. Permutation
+
+However, there are other empirical approaches.  Permutation
+
+The primary purpose of a GWAS is to identify the genomic regions housing genes that are influencing a heritable trait. 
+This is why extra computational effort is devoted to calculating the significance of the SNP effects in the regularisation methods bigRR, LMM-Lasso, and 
+glmnet.
+
+Much of the computational cost for the regularisation methods bigRR, LMM-Lasso, and glmnet, comes from the empirical calculation of the 
+significance of a SNP.
 

### 2018-07-20 AM+draft.tex
12/07/18 working on discussion v1

--- 2018-07-18 AM+draft.tex
@@
 
-Eagle is a new method and R package for multiple locus association mapping.
-It differs from other multiple locus methods.
-It can handle large datasets due to dimension reduction It treats association mapping as a model selection instead of a variable selection problem so we don?t have to deal with multiple testing issues.
-Its results are immediately interpretable.
+Eagle is a new method (and R package) for multiple-locus association mapping. It differs from other multiple-locus methods. Its results are 
+immediately interpretable,  it can handle much larger datasets due to dimension reduction, and it treats association mapping as a model selection instead of a variable selection problem which avoids multiple testing issues.
 As we saw in the simulation study, Eagle has considerably higher power than single-locus methods and is comparable in run time.
-However, It was in the analysis of the mouse data that the true worth of Eagle is evident.
-With Eagle, we found more than double the results than with single locus association mapping, the method of choice and these extra results were all validated as true.
+However, the true worth of Eagle is apparent from the analyses of the mouse data.
+It  found
+more than double the SNP-trait associations than with single-locus association mapping, the method of choice. Furthermore, these extra results 
+were not false positives but true verifiable results.  Eagle was built to make multiple-locus association mapping an attractive alternative 
+to single-locus association mapping.
+
+High-throughput array-based technologies 
+continue to drive down the cost of collecting SNP data. Also, access to whole-genome sequence is no longer a distant reality. The cost is still 
+prohibitive for large scale deployment but this will change. 
+The datasets being generated in GWASs are only going to increase in size over the coming years. 
+This is why we have structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
+Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
+computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times which is encouraging for prototype code. 
+We also have plans for Eagle to run on computer clusters. Structuring Eagle for larger-than-memory calculations was a 
+preemptive step on our part in this direction.
+
+In our simulation study, we compared the performance of Eagle against six multiple-locus methods.  We found that Eagle outperforms these methods. 
+However, we are cognisant that our deductions are contingent on several implementation choices. We chose to calculate the significance of the 
+SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
+Permutation (ref) and multi sample-splitting (ref) are also equally valid empirical approaches.
+Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
+data. Also, when analysing the (sub)samples, calculating a regularisation method's entire solution path is avoided.  Instead,  analyses are 
+performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation.
 

### 2018-07-20 AM+draft.tex
12/07/18 working on discussion v3

@@
 data. Also, when analysing the (sub)samples, calculating a regularisation method's entire solution path is avoided.  Instead,  analyses are 
-performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation.
+performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
+implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
+The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert,
+ given our resource constraints, the right balance between 
+computational and statistical performance could be reached. 
+We would like to think that the parameter settings used here are defensible since they match the values used in the original r2VIM publication.
 
-How to deal with more complex models.
+Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here.  
+Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
+is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
+including the extra random effects, is 
+captured in this first-stage model.
+In the second stage, Eagle is run not on the original but adjusted trait data which were obtained from the first stage analysis.
+Even though this is a well accepted practice, it is approximate (ref).
+A better solution is to fit a single model to the data.
+Although not specifically designed for association mapping,
+WGAIM, upon which Eagle is based, and RRGAIM are two R packages where this is possible. Here, the difficulty is that for large datasets, 
+run time and memory usage become limiting factors for analysis.
 
-Comment on cloud based FaST-LMM
-
-Performance could change with changes to default values.
-
-Much of the computational cost for bigRR, lmmlasso, ,,,, lies not in the analysis but in calculating the significance of the effects. We used 
-stability selection that involved 100 further analyses, though these analyses were based on only 50\% of the original data. 
-These methods give parameter estimates efficiently.
-
-Since we consider the goal of associaiton mapping is to identify those genomic regions housing genes that are influencing a trait, then
-
-======
-
-General introductory para - gwas growing in size.
-
-Epistatic effects
-
-More complex models
-
-Last paragraph
-==============
-We see what we are presented here as the first step in the development of Eagle. GWAS are only growing in size, both in terms of number of 
-samples and number of genotypes. Our goal here was to develop the methods and implement those methods in package that could be run on a desktop. 
-However, the computational power of desktop will not suffice for datasets where there are hundreds of thousands of samples and millions of loci. 
-Here, computer clusters will be needed to meet the computational challenges and we have written the Eagle package with this in mind. We do not 
-keep the marker data in memory. Also, we perform block multiplication of matrices that would otherwise be too large to fit into memory. These are all 
-preemptive steps to make the porting of Eagle onto a cluster environment easier.
+Over the past decade, GWASs have undergone great change. They have gone from being misunderstood and distrusted to garnering 
+wide acceptance.  Much of this acceptance has been earned through the generation of many new and reproducible discoveries. 
+However, it is this coming decade that we are most excited about. Whole-genome sequencing is already a reality and poised to become 
+affordable on a large scale. Our depth of understanding of the genome and its function is growing daily.  We already have huge data repositories 
+containing vast amounts of GWAS  and omics data inviting new ways of analysis. These all present new 
+methodological and computational challenges but they also hold great opportunity.
 

### 2018-07-22 AM+draft.tex
23/07 Discussion v1

@@
 
-Eagle is a new method (and R package) for multiple-locus association mapping. It differs from other multiple-locus methods. Its results are 
-immediately interpretable,  it can handle much larger datasets due to dimension reduction, and it treats association mapping as a model selection instead of a variable selection problem which avoids multiple testing issues.
-As we saw in the simulation study, Eagle has considerably higher power than single-locus methods and is comparable in run time.
-However, the true worth of Eagle is apparent from the analyses of the mouse data.
+Eagle is a new method for multple-locus association mapping.
+
+Eagle is a new method (and R package) for multiple-locus association mapping. It advances the state of multiple-locus association mapping 
+in several ways. First, the results from Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
+each SNP identifies a separate genomic region of interest. Second, Eagle treats association mapping as a model selection instead of a 
+variable selection problem. We do not have to content with thresholds and multiple testing issues.
+
+ It differs from other multiple-locus methods. Its results are 
+immediately interpretable,  it can handle much larger datasets due to dimension reduction, and it treats association mapping as a model selection instead of a variable selection problem avoiding threshold and  multiple testing issues.
+As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
+The true worth of Eagle is apparent from the analyses of the mouse data.
 It  found
@@
 
+Over the past decade, GWASs have undergone great change. From their humble beginnings, GWASs are now widely accepted, having 
+generated many new and reproducible discoveries.
+
+What began as a fledgling and poorly understood platform
+
+They have gone from a fledging, poorly understood
+
 Over the past decade, GWASs have undergone great change. They have gone from being misunderstood and distrusted to garnering 

### 2018-07-24 AM+draft.tex
24/07 Discussion v1

@@
 \section{Discussion}
+Eagle is a new linear mixed model based method (and R package) for multiple-locus association mapping. It advances the state of association mapping in several ways. 
+First, its computational footprint is much smaller than other multiple-locus implementations. Eagle makes multiple-locus analysis 
+practical, even when the datasets are large. Second, the results from
+ Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
+each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
+multiple testing issues and the need for significance thresholds. 
+ As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
+Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
+with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.
 
-Eagle is a new method for multple-locus association mapping.
+Eagle outperformed the other multiple-locus methods in our simulation study. However, we are aware/cognisant that we made several implementation 
+choices that impacts our conclusions.  For instance, we chose to calculate the significance of the 
+SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
+Permutation (ref) and multi sample-splitting (ref) are also equally valid empirical approaches.
+Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
+data. Also, when analysing the (sub)samples, it was not necessary to calculate the entire solution path for a method. 
+ Instead,  analyses are 
+performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
+implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
+The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
+r2VIM could of been fine-tuned for a better balance of computational and statistical performance because random forests are highly adjustable. 
+However, we would like to think that the parameter settings we used are sensible since they match the values in the original r2VIM publication (ref).
 
-Eagle is a new method (and R package) for multiple-locus association mapping. It advances the state of multiple-locus association mapping 
-in several ways. First, the results from Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
-each SNP identifies a separate genomic region of interest. Second, Eagle treats association mapping as a model selection instead of a 
-variable selection problem. We do not have to content with thresholds and multiple testing issues.
+Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here, although in different ways. 
+Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
+is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
+including the extra random effects, is 
+captured in this first-stage model.
+In the second stage, Eagle is run not on the original but adjusted trait data which were obtained from the first stage analysis.
+Even though this is a well accepted practice, it is approximate (ref).
+A better solution is to fit a single model to the data.
+Although not specifically designed for association mapping,
+WGAIM, upon which Eagle is based, and RRGAIM are two R packages where this is possible. Here, the difficulty is that for large datasets and/or complex 
+models,  run time and memory usage can become limiting factors for analysis.
 
- It differs from other multiple-locus methods. Its results are 
-immediately interpretable,  it can handle much larger datasets due to dimension reduction, and it treats association mapping as a model selection instead of a variable selection problem avoiding threshold and  multiple testing issues.
-As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
-The true worth of Eagle is apparent from the analyses of the mouse data.
-It  found
-more than double the SNP-trait associations than with single-locus association mapping, the method of choice. Furthermore, these extra results 
-were not false positives but true verifiable results.  Eagle was built to make multiple-locus association mapping an attractive alternative 
-to single-locus association mapping.
+Over the coming years, computationally, the demand placed on association mapping methods is going to increase. 
+High-throughput array-based technologies continue to decrease the cost of genotyping, permitting ever larger GWASs to be performed. 
+Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs (refs), 
+culminating in data on millions of SNPs. It is because of this growing demand that 
+we have purposely structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
+Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
+computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times.  
+We also have plans for Eagle to run on computer clusters. Structuring Eagle for larger-than-memory calculations was a 
+preemptive step in this direction. GWASs have changed significantly in the past decade but the size and complexity of GWASs is expected 
+to change even more in the coming decade.
 
@@
 preemptive step on our part in this direction.
-
-In our simulation study, we compared the performance of Eagle against six multiple-locus methods.  We found that Eagle outperforms these methods. 
-However, we are cognisant that our deductions are contingent on several implementation choices. We chose to calculate the significance of the 
-SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
-Permutation (ref) and multi sample-splitting (ref) are also equally valid empirical approaches.
-Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
-data. Also, when analysing the (sub)samples, calculating a regularisation method's entire solution path is avoided.  Instead,  analyses are 
-performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
-implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
-The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert,
- given our resource constraints, the right balance between 
-computational and statistical performance could be reached. 
-We would like to think that the parameter settings used here are defensible since they match the values used in the original r2VIM publication.
-
-Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here.  
-Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
-is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
-including the extra random effects, is 
-captured in this first-stage model.
-In the second stage, Eagle is run not on the original but adjusted trait data which were obtained from the first stage analysis.
-Even though this is a well accepted practice, it is approximate (ref).
-A better solution is to fit a single model to the data.
-Although not specifically designed for association mapping,
-WGAIM, upon which Eagle is based, and RRGAIM are two R packages where this is possible. Here, the difficulty is that for large datasets, 
-run time and memory usage become limiting factors for analysis.
 

### 2018-07-24 AM+draft.tex
24/07 Discussion v2

@@
 
-High-throughput array-based technologies 
-continue to drive down the cost of collecting SNP data. Also, access to whole-genome sequence is no longer a distant reality. The cost is still 
-prohibitive for large scale deployment but this will change. 
-The datasets being generated in GWASs are only going to increase in size over the coming years. 
-This is why we have structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
-Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
-computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times which is encouraging for prototype code. 
-We also have plans for Eagle to run on computer clusters. Structuring Eagle for larger-than-memory calculations was a 
-preemptive step on our part in this direction.
-
-Over the past decade, GWASs have undergone great change. From their humble beginnings, GWASs are now widely accepted, having 
-generated many new and reproducible discoveries.
-
-What began as a fledgling and poorly understood platform
-
-They have gone from a fledging, poorly understood
-
-Over the past decade, GWASs have undergone great change. They have gone from being misunderstood and distrusted to garnering 
-wide acceptance.  Much of this acceptance has been earned through the generation of many new and reproducible discoveries. 
-However, it is this coming decade that we are most excited about. Whole-genome sequencing is already a reality and poised to become 
-affordable on a large scale. Our depth of understanding of the genome and its function is growing daily.  We already have huge data repositories 
-containing vast amounts of GWAS  and omics data inviting new ways of analysis. These all present new 
-methodological and computational challenges but they also hold great opportunity.
-
 %%\bibliographystyle{naturemag}

### 2018-07-24 paper.tex
24/07 added paper.tex

--- 2018-07-24 AM+draft.tex
@@
-\documentclass{nature}
+\documentclass{article}
 \usepackage{xr}
@@
 \renewcommand{\baselinestretch}{1.3} 
+\usepackage{graphicx}
 
 %% external tex files
-\externaldocument{figure1_runtimes}
-\externaldocument{figure2_powerFDRmultiple}
-\externaldocument{figure3_mouse}
+%%\externaldocument{figure1_runtimes}
+%%\externaldocument{figure2_powerFDRmultiple}
+%%\externaldocument{figure3_mouse}
 
@@
 %%-------------------------
-%% 1. how real data was analysed
-%% 2. default settings for methods in simulation study.
-%% 3. real data analyses - how were they done fixed effects?
-%% 4. simulation study very simple in terms of design, fixed effects etc.
+%% 1. how real data were analysed
 
-%% Issue 
-%%====================
-%% Where does us doing stability selection and adding time to jobs come in.
-
-\begin{affiliations}
-\item Data61, CSIRO, Australia.
-\item IM \&T, CSIRO, Australia.
-\end{affiliations}
+%%\begin{affiliations}
+%%\item Data61, CSIRO, Australia.
+%%\item IM \&T, CSIRO, Australia.
+%%\end{affiliations}
 
@@
 
-{\em Figure \ref{fig_time} goes around here}
+\begin{figure}
+\caption{Median run times, in minutes, for the analysis of association study data from the six simulation scenarios. 
+Eagle is compared against five other multiple-locus programs (top left plot) and two single-locus programs (top right plot). 
+The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
+programs, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
+The actual median run times for the programs across the scenarios are given in the table. The entries in a bold font 
+correspond to the lowest run-time for a scenario. 
+ FaST-LMM$^{all}$ is where the single-locus results are based on all the available 
+SNP data. FaST-LMM$^{few}$ is where the single-locus results as based on a subset of the available SNP data. }
+
+\label{fig_time}
+\begin{center}
+\includegraphics[width=18cm, height=14cm]{Figure2_time.jpg}
+\end{center}
+
+\end{figure}
+
+%%{\em Figure \ref{fig_time} goes around here}
 
@@
 
+\begin{figure}
+\caption{Genome-wide association mapping results from the analysis of the mouse data for the single-locus method FaST-LMM and the 
+multiple-locus method
+Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
+for a false discovery rate of 5\% (Eagle$^{optimal}$). 
+The number of SNP-trait associations found are reported. The bar plots with boxes around them are for those traits where using 
+Eagle$^{default}$ and Eagle$^{optimal}$ resulted in more findings than with single-locus analysis.} 
+\label{figmouse}
+\begin{center}
+%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
+\includegraphics[width=15cm, height=15cm]{mouseresults.eps}
+\end{center}
+
+\end{figure}
+
+%%{\em Figure \ref{figpowermultiple} goes here}
+
 In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
@@
 
-{\em Figure \ref{figpowermultiple} goes here}
+\begin{figure}
+\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
+scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
+scenarios but MLMM also performs well. }
+\label{figpowermultiple}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{power1main.eps}
+\end{center}
+
+\end{figure}
+
+%% {\em Figure \ref{figpowermultiple} goes here}
 
@@
 
-\subsection{Analyses of Simulated Data}
-
 \subsection{Analyses of Mouse Data}
@@
 
-\begin{addendum}
- \item Put acknowledgements here.
- \item[Competing Interests] The authors declare that they have no
-competing financial interests.
- \item[Correspondence] Correspondence and requests for materials
-should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
-\end{addendum}
+%%\begin{addendum}
+%% \item Put acknowledgements here.
+%% \item[Competing Interests] The authors declare that they have no
+%%competing financial interests.
+%% \item[Correspondence] Correspondence and requests for materials
+%%should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
+%%\end{addendum}
 

### 2018-07-25 paper.tex
26/7 refinement v1

@@
 associations better than completing single- and multiple-locus methods and does it faster.
-Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users unfamiliar with R.
+Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users whom may be unfamiliar with R.
 \end{abstract}
 
-Over the past decade,  genome-wide association studies have changed considerably in both their analysis and design. Early studies
+Over the past decade,  genome-wide association studies (GWASs) have changed considerably in both their analysis and design. Early studies
  followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
@@
 This prompted a shift towards family-based designs and score tests, such as the tdt test and its variants (refs).
-Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification.
+Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification (refs).
 This has meant that data can be collected from general populations, even if these populations are highly structured.
-Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine.
+Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine (refs).
 
@@
 The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
-that plot the -log$_{10}$ of the P value for each locus against the map position of the locus. The P value is obtained by testing the statistical 
+that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ valueue is obtained by testing the statistical 
 significance of a SNP when treated as an effect in an appropriate model. 
@@
 Third, many of the traits whose genetic secretes we are trying to discover are complex.
-There will be multiple SNPs in linkage disequilibrium with genes that influence the trait.
+There will be multiple SNPs in linkage disequilibrium with genes that are influencing the trait.
 Yet, a locus-by-locus mapping approach only assesses the evidence for association between a single marker locus and trait.
@@
 It is somewhat surprising then that multiple-locus association mapping methods haven't attracted more attention. Methods based on 
-regularisation techniques, such as ridge regression and lasso, measure all locus-trait associations simultaneously. 
+regularisation techniques, such as ridge regression (ref) and lasso (ref), measure all locus-trait associations simultaneously. 
 Here, multiple testing is not an issue.
@@
 Also, their results can be difficult to interpret.
-The strength of association is not measured by a P value but by the size of the regression coefficient for the SNP in the model.
+The strength of association is not measured by a $p$ value but by the size of the regression coefficient for the SNP in the model.
 More recently, associations have started to be mapped with random forests (refs).
@@
 The best multiple-locus model is built with simple forward selection.
-Results are immediately interpretable but here, computation can become challenging if the datasets become large.
+Results are immediately interpretable but here, computation becomes challenging for large datasets.
 
@@
 formats,  can handle data larger than a computer's  memory capacity, and makes heavy use of 
-distributed computing for computation when available.
+parallel computing for computation when available.
 
@@
 Each takes a different approach to multiple-locus association mapping. A summary of the key attributes of the different computer programs/packages 
-is given in Supplementary Table \ref{suptabsummary}. 
-Further details about the methods being implemented in these computer programs/packages can be found in Methods.
+is given in Supplementary Table \ref{suptabsummary} (See Methods for further details).
 
 \subsection{Simulation Study}
-We performed a large simulation study where we sought to  answer two questions. 
+A large simulation study was performed where we sought to  answer two questions. 
 First, how does Eagle compare, in terms of run time and memory usage, to 
 competing implementations? Second, how well does Eagle find true associations (power) and avoid 
-false associations (type 1 errors)? We generated data under five different scenarios; a study of size 150 individuals 
+false associations (type 1 errors)? Data were generated under five different scenarios; a study of size 150 individuals 
 and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
@@
 1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).   
-We chose these scenarios to mirror some of the different sized GWAS being performed in animals, plants, and humans.
+These scenarios reflect, at least in some cases, the sizes of study being performed in animals, plants, and humans.
 
-For each scenario, we generated 100 replicates of data. A single replicate consists of SNP data and quantitative trait data. 
-To introduce some of the complexities of dealing with real genotypes into the study, we obtained the SNP genotypes from 
-the publicly available 1000 Genome Project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
-by selecting, randomly, a set of SNPs, assigning additive allelic effects to these loci, and then aggregating their effects for 
-each individual along with a random error.  
- The number of SNPs selected per replicate follows a Poisson distribution with mean 30. The sizes of the allelic effects 
- across the selected loci are equal.  A heritability of  50\% is assumed for the trait.
+For each scenario, 100 replicates were generated. A single replicate consisted of SNP and quantitative trait data. 
+Extra realism was introduced into the simulation study through the drawing of the SNP genotypes from the 1000 Genome Project, phase 3  \cite{10002010map}.
+ The quantitative trait data were generated 
+by selecting, randomly, a set of SNPs and assigning these loci additive allelic effects.  Random errors were then drawn from a normal distribution 
+with variance set to give a heritability of 50\% for the trait. 
+For each individual, a quantitative trait value was obtained by summing its random error and additive allelic effects. 
+The number of randomly selected SNPs follows a Poisson distribution with mean 30. The size of the allelic effects 
+ across the selected loci are equal.
 
- A replicate was analysed as follows. All programs/packages were run at their default settings. 
+ %%=== Tense/flow  ==>>
+
+ Analyses by the seven programs/packages of a replicate proceeded as follows. They were all run at their default settings. 
  Eagle and MLMM were the easiest of the programs/packages to implement. 
- Their results are immediately 
- interpretable. Their findings are the set of SNPs in strongest association with the trait. Each 
-SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.  
-The other programs/packages were not as easy to implement  BigRR, LMM-Lasso, and glmnet make use of regularlisation methods. 
-All SNPs are fitted simultaneously in a regression framework.  However, due to the adaptive nature of the estimation procedure, it is challenging, 
-analytically, to calculate P values for the effects in the model. 
-To solve this problem, at the cost of increased computation, we calculated the significance of a SNP empirically 
-via stability selection (see Methods).  R2VIM, a random forest approach, required three parameters to be specified. These are the
- minimum size of a terminal node, the number of trees, and number of potential variables 
-for building a tree.  It is not clear how "best" to set these parameters so we followed ref. The minimum size of a node was set to 10\% of the sample size. 
-The number of trees was set to 1000. The number of variables was set to 20\% of the number of SNP. A relative importance measure is calculated 
-for each SNP measuring the strength of association with the trait. 
-FaST-LMM and GEMMA implement single-locus association mapping. The P values of the SNP are reported as their results. 
+ The only parameters requiring specification were the amount of available memory and number of CPU for 
+ Eagle and the number of chunks for MLMM. 
+ Their results were also immediately 
+ interpretable. Their findings were the set of SNPs in strongest association with the trait. Each 
+SNP in this set identified a separate genomic region of interest, whose position was given by the map location of the SNP.
+
+BigRR, LMM-Lasso, and glmnet were not as easy to implement. They are based on regularisation methods and as such, all the SNPs were fitted simultaneously in a regression 
+framework. The difficulty was in calculating the significance of the SNP effects. To do this analytically is challenging. We instead opted for stability selection (see Methods),  
+an empirical approach for calculating significance.
+
+R2VIM is different from the rest in that it is a nonparametric approach for association mapping. It is based on random forests. Three important parameters needed to be  set. 
+These were the number of trees, the number of variables for building a tree, and the minimum size of a terminal node. Ideally, these parameters would be "tuned" on a replicate-by-replicate 
+basis (ref). However, this is not feasible here. We instead used the same settings as in \cite{szymczak2016r2vim} where 
+the number of trees was set to 1000, the number of variables was set to 20\% of the number of SNPs, and 
+  the minimum size of a node was set to 10\% of the sample size.
+A relative importance measure was calculated 
+for each SNP measuring its strength of association with the trait.
+
+FaST-LMM and GEMMA implement single-locus association mapping. The $p$ values of the SNP were reported as their results.
+
 For all but Eagle and MLMM, the set of SNPs in strongest and significant association with the trait were found 

### 2018-07-26 paper.tex
26/7 refinement end-of-day

@@
 competing implementations? Second, how well does Eagle find true associations (power) and avoid 
-false associations (type 1 errors)? Data were generated under five different scenarios; a study of size 150 individuals 
+false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
 and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
@@
 
- %%=== Tense/flow  ==>>
-
  Analyses by the seven programs/packages of a replicate proceeded as follows. They were all run at their default settings. 
@@
 
-BigRR, LMM-Lasso, and glmnet were not as easy to implement. They are based on regularisation methods and as such, all the SNPs were fitted simultaneously in a regression 
+BigRR, LMM-Lasso, and glmnet required more effort to implement. They are based on regularisation methods and as such, all the SNPs were fitted simultaneously in a regression 
 framework. The difficulty was in calculating the significance of the SNP effects. To do this analytically is challenging. We instead opted for stability selection (see Methods),  
@@
 These were the number of trees, the number of variables for building a tree, and the minimum size of a terminal node. Ideally, these parameters would be "tuned" on a replicate-by-replicate 
-basis (ref). However, this is not feasible here. We instead used the same settings as in \cite{szymczak2016r2vim} where 
+basis \cite{boulesteix2012overview}. However, this was not practical here. We instead used the same settings as in \cite{szymczak2016r2vim} where 
 the number of trees was set to 1000, the number of variables was set to 20\% of the number of SNPs, and 
@@
 
-FaST-LMM and GEMMA implement single-locus association mapping. The $p$ values of the SNP were reported as their results.
+FaST-LMM and GEMMA implement single-locus association mapping. 
+FaST-LMM was run in two ways. One way was where a subset of the SNP data were used in calculating the similarity (or relationship) matrix. Here, FaST-LMM is 
+highly efficient, computationally. 
+The other was where calculation of the 
+similarity matrix was based on all the SNP data. 
+The $p$ values of the SNP were reported as their results.
 
-For all but Eagle and MLMM, the set of SNPs in strongest and significant association with the trait were found 
-by placing the SNP in map order, setting a threshold, identifying peak regions with 
-significance measures above the threshold, and recording the SNP with the largest significance measure in these peak regions.
+The results from all but Eagle and MLMM required post-processing before the findings were interpretable.  The SNPs were placed in map order, 
+a significance threshold was set, peak regions containing SNPs with significance measures above the threshold were identified, and the SNP with the largest 
+significance measure in each of the peak regions was recorded as a finding.
 
@@
 
-We analysed the simulated data with Eagle and the other computer programs, recording their memory usage and 
-run (or elapse) times.
-The analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM.
-Not all data generated under the five scenarios could be analysed by all implementations.
+Memory usage and run (or elapse) times were recorded for Eagle and the other computer programs/packages across the simulation scenarios. 
+Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM.
+Not all data generated under the six scenarios could be analysed by all implementations.
 Memory usage
-for many of the computer programs was the limiting factor (See {\bf Supplementary Figure \ref{supfigmem}}).  The single-locus program GEMMA was by 
+for many of the computer programs/packages was the limiting factor (See {\bf Supplementary Figure \ref{supfigmem}}).  The single-locus program GEMMA was by 
 far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
 excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
-Even FaST-LMM, a single-locus
-implementation, required more than 128 gigabytes of memory for the analysis of the larger data sets 
-when all the marker data was used to calculate the similarity matrix. Of the multiple-locus programs, only Eagle,  
+Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
+Of the multiple-locus programs/packages, only Eagle,  
 with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
@@
 The median run times for Eagle and the other computer programs across the six scenarios are shown in Figure \ref{fig_time}. 
-The x- and y-axes are on a log scale.
-This means a unit change on the x- or y-axis is equivalent to a change in the order of magnitude.
+The x- and y-axes are on a log scale.  A unit change on the x- or y-axis is equivalent to a change in the order of magnitude.  
 In answer to our question of how does Eagle compare in terms of run time to competing implementations, 
@@
  implementations and is comparable to the single-locus implementations. For a simulation study with 150 individuals and 
- 5000 SNPs, Eagle produced results in seconds.  For the larger simulation scenarios 1500 x  50K and 350 x4 00K, 
+ 5000 SNPs, Eagle produced results in seconds.  For the larger simulation scenarios of 1500 x  50K and 350 x4 00K, 
  analyses with Eagle took under two minutes. Even for data from a couple of thousand individuals and half a million 
@@
  Towards the final stages of writing this paper, 
- we gained access to a desktop computer with 14-core Xeon processors with 256 gigabytes of RAM. We reran Eagle on data from the largest
+ we gained access to a desktop computer with 14-core Xeon processors and 256 gigabytes of RAM. We reran Eagle on data from the largest
   scenario 10000 x 1.5M to measure the impact on run time. The median run time dropped by more than 70\% 
@@
 \begin{figure}
-\caption{Median run times, in minutes, for the analysis of association study data from the six simulation scenarios. 
-Eagle is compared against five other multiple-locus programs (top left plot) and two single-locus programs (top right plot). 
+\caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
+Eagle is compared against five other multiple-locus programs (top left) and two single-locus programs (top right). 
 The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
@@
 correspond to the lowest run-time for a scenario. 
- FaST-LMM$^{all}$ is where the single-locus results are based on all the available 
-SNP data. FaST-LMM$^{few}$ is where the single-locus results as based on a subset of the available SNP data. }
+ FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.  
+ FaST-LMM$^{few}$ is where calculation  the similarity matrix is based on a subset of the SNP data. }
 
@@
 \begin{center}
-\includegraphics[width=18cm, height=14cm]{Figure2_time.jpg}
+\includegraphics[width=14cm, height=12cm]{Figure2_time.jpg}
 \end{center}
@@
 
-We calculated, empirically, the statistical power and false discovery rates of Eagle and the other methods across the six scenarios. 
-We were interested in answering the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations. 
-For each replicate, we knew which SNPs had been used in creating the quantitative trait data. These SNPs are in true association with 
-the trait. It is the goal of the single- and multiple-locus association mapping methods to discover these SNPs. By knowing which SNPs are in true 
-association with the trait, we were able to assess the validity of a  method's findings.
+We were interested in answering the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.  To 
+answer this question, we calculated, empirically, 
+the statistical power and false discovery rates of Eagle and the other methods across the six scenarios
+For each replicate, we knew which SNPs  had been used to create the quantitative trait data. These SNPs are in true association with 
+the trait. By knowing these SNPs, we were able to assess the validity of a  method's findings.
 
-To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
-in creating the trait data. We then averaged across the 100 replicates.  
-Similarly, to calculate the false discovery rate of a method, 
-for each replicate, 
-we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.  
-A finding was counted as true if the SNP is located within 40 kilobase pairs of a SNP in true 
+The power (false discovery rate) of a method, under a particular scenario,  was calculated by taking the median of its power (false discovery rate) 
+values across the 100 replicates of the scenario.
+To calculate the power of a method from its analysis of a replicate, we divided the number of findings for that method that were true by the number of SNP in true association with the trait.
+To calculate the false discovery rate of a method from its analysis of a replicate,
+we instead divided by the number of  findings for that method. 
+A finding was counted as true if the SNP was located within 40 kilobase pairs of a SNP in true 
 association with the trait.

### 2018-07-26 paper.tex
27/7 refinement v1

@@
 
-We were interested in answering the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.  To 
-answer this question, we calculated, empirically, 
-the statistical power and false discovery rates of Eagle and the other methods across the six scenarios
-For each replicate, we knew which SNPs  had been used to create the quantitative trait data. These SNPs are in true association with 
-the trait. By knowing these SNPs, we were able to assess the validity of a  method's findings.
+We answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations. We do this by estimating the 
+power and false discovery rates of Eagle and the other methods for the six scenarios.  Since, for a replicate, we knew which SNP were assigned additive effects in 
+creating its trait data, we knew the SNP that were in true association with the trait. This allowed us to assess if a method's findings were true or 
+false. A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a SNP in true association. 
+When a replicate was analysed, we obtained an estimate of power, the number of findings that were true for the method divided by the 
+number of SNPs in true association. We also obtained an estimate of the false discovery rate, the number of findings that were false for the method 
+divided by the number of (true and false) findings for the method.  Both these estimates varied with replicate. 
+The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates that were calculated from the analyses of 
+the 100 replicates.
 
-The power (false discovery rate) of a method, under a particular scenario,  was calculated by taking the median of its power (false discovery rate) 
-values across the 100 replicates of the scenario.
-To calculate the power of a method from its analysis of a replicate, we divided the number of findings for that method that were true by the number of SNP in true association with the trait.
-To calculate the false discovery rate of a method from its analysis of a replicate,
-we instead divided by the number of  findings for that method. 
-A finding was counted as true if the SNP was located within 40 kilobase pairs of a SNP in true 
-association with the trait.
-
-The power and false discovery rate of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
+The power and false discovery rates of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
 We restricted our attention to these scenarios for the multiple-locus methods because for scenario 4000 x 1.5M,
@@
 These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
-The power curves are for those methods that treat association mapping as a parameter estimation problem. Here, the 
+The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
 significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 

### 2018-07-27 paper.tex
27/7 refinement end-of-day

@@
 
-We answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations. We do this by estimating the 
-power and false discovery rates of Eagle and the other methods for the six scenarios.  Since, for a replicate, we knew which SNP were assigned additive effects in 
-creating its trait data, we knew the SNP that were in true association with the trait. This allowed us to assess if a method's findings were true or 
-false. A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a SNP in true association. 
-When a replicate was analysed, we obtained an estimate of power, the number of findings that were true for the method divided by the 
-number of SNPs in true association. We also obtained an estimate of the false discovery rate, the number of findings that were false for the method 
-divided by the number of (true and false) findings for the method.  Both these estimates varied with replicate. 
-The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates that were calculated from the analyses of 
-the 100 replicates.
+Here, we answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.
+We do this by estimating the power and false discovery rates of Eagle and the other methods for the six scenarios.
+Since, for a replicate, we knew which SNPs were assigned additive effects, we knew the SNPs that were in true association with the trait.
+We will refer to these SNPs as being
+true SNPs.
+By knowing the true SNPS, we were able to assess the validity of a method's findings.
+A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a true SNP.
+When a replicate was analysed, we obtained an estimate of the power of the method by taking  the number of findings that were found to be  true and dividing by the
+number of true SNPs.
+We also obtained an estimate of a method's false discovery rate.
+It is the number of findings that were found to be false divided by the number of (true and false) findings found by the method.
+Both these estimates varied with replicate.
+The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates over the 100 replicates.
 
 The power and false discovery rates of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
-We restricted our attention to these scenarios for the multiple-locus methods because for scenario 4000 x 1.5M,
-the data could only be analysed by Eagle, MLMM, and r2VIM. For scenario 10000 x 1.5M, the data could only be analysed by Eagle. 
-The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
-Supplementary Figure \ref{supfigpowersingle}.  Each plot contains 
+We restricted our attention to these scenarios because not all
+multiple-locus methods could handle the size of data in the other scenarios. 
+  Each plot contains 
 single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
@@
 the false discovery rate as the significance threshold  is adjusted. 
+The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
+Supplementary Figure \ref{supfigpowersingle}.
+
+\begin{figure}
+\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
+scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
+scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
+upper limit of the x-axis is 0.1.}
+\label{figpowermultiple}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{power1main.eps}
+\end{center}
+
+\end{figure}
+
+%%{\em Figure \ref{figpowermultiple} goes here}
+
+In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
+well.  Of the multiple-locus methods, Eagle had the highest power
+while keeping its false discovery rate low (Figure \ref{figpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
+that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
+association with a trait while avoiding false associations (Supplementary Figure \ref{supfigpowersingle}).
+
+%%At first glance, Eagle appears underpowered.
+Several of the other methods have higher power than Eagle.
+The goal of association mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
+For those methods with higher power than Eagle, the cost is much higher false discovery rates.
+In other words, these methods are finding more true results  but they are also finding many more false results.
+If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than  5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest  fdr.
+This is especially true when the power and fdr of Eagle is compared to the single-locus methods.
+For single-locus methods to achieve the same level of power as Eagle, a threshold has to be set where the fdr is extremely high.
+
+\subsection{Mouse Data Analysis}
+
+=====> Up to here ====>
+
+We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
+ We chose to focus on data from a large outbred mouse study \cite{nicod2016genome}.
+This study was a little unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
+Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \cite{zheng2011comparison}.
+By converting the dosages into genotypes and analysing the data with the single-locus program FaST-LMM, we obtained a subset of those findings reported in the original study.
+We then analysed the data with Eagle.
+Due to Eagles increased power, we found SNP-trait associations not found with  FaST-LMM.
+However, we were
+ able to confirm the validity of these new findings as they matched what was found in the original study. Having the ability to confirm new findings  in a real study was 
+ one of the primary motivators for choosing these data for analysis.
+
+We repeated the single-locus analyses, as first performed  \cite{nicod2016genome}, except 
+that we focused on autosomal SNP and our analyses were based on SNP genotype  rather than SNP dosage. 
+In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation. We followed the 
+same empirical procedure but had to increase the number of permutations from 100 to 500.
+
+Eagle was run in two ways; under its default settings (Eagle$^{default}$) and where we specified the regularisation parameter for model selection (Eagle$^{optimal}$ ).
+Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}.
+  The conservativeness of the extBIC can be adjusted by a single regularisation parameter that ranges from zero to one.
+In the simulation study, this parameter was set to one, its most conservative and default setting.
+However, there is also opportunity to set the parameter to a value less than one.
+This increases power but also increases the false discovery rate.
+For each trait, we used permutation to set the regularisation parameter to give a false discovery rate of 5\% .
+
+The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}. The mouse study recorded
+measurements on 200 traits. When these traits were first analysed in the original study, findings for 45 of these traits were able to be 
+corroborated by prior published evidence. We focused our analyses here on these same 45 traits. For 39 traits, SNP-trait associations 
+were found. For the other six, neither FaST-LMM nor Eagle found any associations. 
+Each plot contains the number of SNP-trait associations that were found and in agreement with the original findings. 
+Neither method found SNPs that had not been first identified in the original mouse study so neither method found false positives. 
+As we saw in the simulation study, there was a notable difference between the two methods capacity to discover SNP-trait associations.
+Eagle$^{default}$, under its default settings, for two traits found fewer findings than FaST-LMM, for eight traits found the same number
+of findings as FaST-LMM, and for 28 traits found more findings. Eagle$^{optimal}$, 
+with its regularisation parameter fine tuned to the trait, for seven traits found the same number of findings as FaST-LMM and for 32 traits 
+found more findings. Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, snp-trait findings respectively. 
+Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
+associations than what is the established way of analysing these data. Furthermore, these are all findings that were confirmed in the original study.
 
@@
 \end{figure}
-
-%%{\em Figure \ref{figpowermultiple} goes here}
-
-In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
-well.  Of the multiple-locus methods, Eagle has the highest power
-while keeping its false discovery rate low (Figure \ref{figpowermultiple}). MLMM also performed well. However, it is when Eagle is compared against single-locus methods 
-that the difference in power is most noticeable.  Eagle has much greater power than single-locus methods for finding SNP in true 
-association with a trait while avoiding false associations (Supplementary Figure \ref{supfigpowersingle}).
-
-%%At first glance, Eagle appears underpowered.
-Several of the other methods have higher power than Eagle.
-The goal of association mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
-For those methods with higher power than Eagle, the cost is much higher false discovery rates.
-In other words, these methods are finding more true results  but they are also finding many more false results.
-If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than  5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest  fdr.
-This is especially true when the power and fdr of Eagle is compared to the single-locus methods.
-For single-locus methods to achieve the same level of power as Eagle, a threshold has to be set where the fdr is extremely high.
-
-\begin{figure}
-\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
-scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
-scenarios but MLMM also performs well. }
-\label{figpowermultiple}
-\begin{center}
-\includegraphics[width=15cm, height=15cm]{power1main.eps}
-\end{center}
-
-\end{figure}
-
-%% {\em Figure \ref{figpowermultiple} goes here}
-
-\subsection{Mouse Data Analysis}
-
-We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
- We chose to focus on data from a large outbred mouse study \cite{nicod2016genome}.
-This study was unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
-Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \cite{zheng2011comparison}.
-By converting the dosages into genotypes and analysing the data with the single-locus program FaST-LMM, we obtained a subset of those findings reported in the original study.
-We then analysed the data with Eagle.
-Due to Eagles increased power, we found SNP-trait associations not found with the FaST-LMM.
-However, we were
- able to confirm the validity of these new findings as they matched what was found in the original study. Having the ability to confirm new findings 
- in a real study was 
- one of the primary motivators for choosing these data for analysis. 
-For the single-locus analysis of the data, we followed the same procedure as originally followed for the analysis of the mouse data \cite{nicod2016genome}.
-The only differences were that we focused on the autosomal SNP and it was necessary to
-increase the number of permutations for the controlling of the false discovery rate from 100 to 500.
-
-Eagle was run in two ways; under its default settings (Eagle$^{default}$) and where we specified the regularisation parameter for model selection (Eagle$^{optimal}$ ) Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}.
-  The conservativeness of the extBIC can be adjusted by a single regularisation parameter that ranges from zero to one.
-In the simulation study, this parameter was set to one, its most conservative and default setting.
-However, there is also opportunity to set the parameter to a value less than one.
-This increases power but also increases the false discovery rate.
-For each trait, we used permutation to set the regularisation parameter to give a false discovery rate of 5\% .
-
-The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}. The mouse study took
-measurements on 200 traits. When these traits were first analysed in the original study, findings for 45 of these traits were able to be 
-corroborated by prior published evidence. We focused our analyses here on these same 45 traits. For 39 traits, we found SNP-trait associations. 
-For the other six, neither FaST-LMM nor Eagle found any associations. 
-Each plot contains the number of SNP-trait associations that were found and in agreement with the original findings. 
-Neither method found SNP not identified in the original mouse study so neither method found false positives. 
-As we saw in the simulation study, there is a notable difference in the two methods capacity to discover SNP-trait associations.
-Eagle$^{default}$, under its default settings, for eight traits found the same number of findings as FaST-LMM and for 28 traits found more findings.
-Eagle$^{optimal}$,
-with its regularisation parameter fine tuned to the trait, for six traits found the same number of findings as FaST-LMM and for 32 traits 
-found more findings. Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, snp-trait findings respectively. 
-Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
-associations than what is the established way of analysing these data. Furthermore, these are all findings that were confirmed in the original study.
 

### 2018-07-27 paper.tex
28/7 refinement v1

@@
 
-=====> Up to here ====>
-
 We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
@@
 of findings as FaST-LMM, and for 28 traits found more findings. Eagle$^{optimal}$, 
-with its regularisation parameter fine tuned to the trait, for seven traits found the same number of findings as FaST-LMM and for 32 traits 
-found more findings. Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, snp-trait findings respectively. 
+with its regularisation parameter fine tuned to the trait data, for seven traits found the same number of findings as FaST-LMM and for 32 traits found more findings.
+Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, snp-trait findings respectively.
 Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
-associations than what is the established way of analysing these data. Furthermore, these are all findings that were confirmed in the original study.
+associations than what is the established way of analysing these data. Furthermore, these extra results found by Eagle were not 
+false positives but could be confirmed from the original study.
 
 \begin{figure}
-\caption{Genome-wide association mapping results from the analysis of the mouse data for the single-locus method FaST-LMM and the 
+\caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
 multiple-locus method
@@
 for a false discovery rate of 5\% (Eagle$^{optimal}$). 
-The number of SNP-trait associations found are reported. The bar plots with boxes around them are for those traits where using 
-Eagle$^{default}$ and Eagle$^{optimal}$ resulted in more findings than with single-locus analysis.} 
+The number of SNP-trait associations found are reported.
+The bar plots with boxes around them are for those traits where using Eagle (default and optimal)   resulted in more findings than with single-locus analysis.}
 \label{figmouse}
@@
 
-The data were obtained from a large genome-wide association study which was performed in outbred mice \cite{nicod2016genome}. 
+The data were obtained from a large genome-wide association study that was performed in outbred mice \cite{nicod2016genome}. 
 Phenotypic and genotypic data were available on 1,887 adult mice. 
-The phenotypic data consisted of measurements from 200 behavioural, tissue, and physiological traits.  Of these traits, 
+The phenotypic data included raw and adjusted (for fixed effects) measurements from 200 behavioural, tissue, and physiological traits.  
+Of these traits, 
 43 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
-43 traits that were the focus of our real data analyses. Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
+43 traits that were the focus of our real data analyses. As in the original study  \cite{nicod2016genome}, our analyses 
+were based on the adjusted traits.
+Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
 form of marker dosages (expected allele counts that ranged from zero to one). All missing data had been imputed. 
 We converted the dosages into discrete genotypes 
-by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively.
+by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively. We focused on autosomal SNPs.
+
+=====> Up to here ====>
 

### 2018-07-31 paper.tex
31/07/18 revision v1

@@
 \usepackage{graphicx}
+
+%% To Do
+%% 1. add bold font to matrices/vectors
+%% 2. change mouse results plot - use superscripts for traits of the same class to reduce space of titles.
 
@@
 \maketitle
-
-%% To Add
-%%-------------------------
-%% 1. how real data were analysed
 
@@
 with its regularisation parameter fine tuned to the trait data, for seven traits found the same number of findings as FaST-LMM and for 32 traits found more findings.
-Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, snp-trait findings respectively.
+Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, SNP-trait findings respectively.
 Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
@@
 
-=====> Up to here ====>
-
 \subsection{Eagle Approach for Multiple-locus Association Mapping}
@@
 Eagle also 
-reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest, to the genes influencing a trait. 
+reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest to the genes influencing a trait. 
 The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
@@
 
-Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
-selected in the $k$th iteration. Suppose three iterations of our model building procedure 
-have been performed and say the 500023rd, 15th, and 420th, 
-SNP were selected, then $S=\{500023, 15, 420\}$. 
+Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of $s$ ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
+selected in the $k$th iteration of the model building process. Suppose three iterations  $(s=3)$
+have been performed and say the 500023rd, 15th, and 420th
+SNP were selected. Then $S=\{500023, 15, 420\}$. 
 Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
@@
 However, 
-it is how we specify $u_g$ that distinguishes our model from others.
+it is how we specify $u_g$ that distinguishes our model from the others.
 
@@
 \end{equation}
-where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the $S_k$th SNP locus which is the $k$th selected SNP, 
-$a_{S_k}$ is the additive effect of the $S_k$th SNP locus, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
-with the data for the selected SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
+where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
+$a_{S_k}$ is the additive effect of the $k$th selected SNP, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
+with the data for the SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
 $a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
-The terms in the summation on the left hand side are fixed effects.  The other term is a random effect.  The terms 
-in the summation account 
-for the additive effects of the selected snp that are in linkage disequilibrium with genes that are influencing the trait. The other term models 
-snp-trait associations along the entire genome, simultaneously, except for those snp that have already been selected. 
+The terms in the summation on the left hand side are fixed effects.   They account 
+for the additive effects of those SNPs that have been found to be in association with the trait. The other term is a random effect. 
+It accounts for the joint effect of the yet-to-be-identified SNP that are in association with the trait. 
 This is a simple genetic model but it 
-is effective for discovering snp-trait associations.
+is effective for discovering SNP-trait associations.
 
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood estimation.
+Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood.  For complex models, residual maximum likelihood
+can be computationally demanding. However, our model only contains a single random effect ($a_{-S}$). Here, highly efficient single-dimension 
+optimisation via spectral decomposition is possible (EMMA ref).
 
-Third, we identify the $(s+1)$th snp that is in strongest association with the trait, based on the maximum score statistic
+Third, we identify the $(s+1)$th SNP that is in strongest association with the trait, based on the maximum score statistic
 $t_j^2 = \frac{ \widetilde{a} _j^2}{\textrm{var}(\widetilde{a}_j)}$ where $\widetilde{a}_j$ is the best linear unbiased predictor (BLUP), 
 and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
-identify a snp based on its (random) effect size and accuracy, but is theoretically justified.  It follows from outlier detection 
+identify a SNP based on its (random) effect size and accuracy, but is theoretically justified.  It follows from outlier detection 
 in linear and linear mixed models (ref).
 
-Fourth, we determine the importance of the $(s+1)$th selected snp via a model selection strategy. 
-We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected snp.  We then fit this new model to the data
+Fourth, we determine the importance of the $(s+1)$th selected SNP via a model selection strategy. 
+We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected SNP.  We then fit this new model to the data
 via maximum likelihood and calculate its extended Bayesian information criteria (extBIC) \cite{chen2008extended}.  The 
@@
 of the model space.
-It is especially well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
-If this new model has a larger  than the current model, then the $s+1$th selected snp is added to
+It is well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
+If this new model has a larger extBIC than the current model, then the $s+1$th selected SNP is added to
 the current model and the above process is repeated. If this new model has a smaller extBIC than the current model, then the 
-model building process is complete. The set of snp in strongest association with the trait is the $s$ snp previously identified.
+model building process is complete. The set of SNP in strongest association with the trait is the $s$ SNPs previously identified.
 
@@
 The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
-$L$, the number of snp, can be extremely large.  An alternative model is given by 
+$L$, the number of SNPs, can be extremely large.  An alternative model is given by 
 Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
@@
 \end{equation}
-These values are needed in order to calculate the score statistic $t_j^2$ for identifying the snp in strongest association with the trait. 
-Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation. 
+These values are needed in order to calculate the score statistic $t_j^2$ for identifying the SNP in strongest association with the trait. 
+Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation 
 of (\ref{eq4}).
+
+\iffalse
 
@@
 association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
-These five methods were selected because of their demonstrated value 
-for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
-either stand-alone computer programs or as  packages.
-
 BigRR, LMM-Lasso, and glmnet  implement three different regression-based regularisation methods; generalized ridge regression, 
@@
  far greater than the number of  samples. This is made possible through a penalty function that balances bias against variance. 
-The application of regularisation methods to association mapping has appeal because it means that the snp effects across an 
-entire genome can be estimated, simultaneously. 
-All three methods are sophisticated pieces of statistical machinery but the significance of an effect, which 
-is needed in association mapping,  can not be calculated easily. Permutation has been suggest as an empirical solution to this 
-limitation \cite{shen2013novel} 
-but we have instead opted for stability selection (see below).
+The application of regularisation methods to association mapping has appeal because it means that the SNP effects across an 
+entire genome can be estimated, simultaneously.  The downside is that it is challenging, analytically, to calculate  
+the statistical significance of an SNP effect. We instead calculate significance, empirically, via stability selection (see below). 
+%% Permutation has been suggest as an empirical solution to this 
+%% limitation \cite{shen2013novel} 
+%% but we have instead opted for stability selection (see below).
 
-MLMM, of the five multiple-locus methods, is the closest in philosophy to Eagle. It too is based on  building the 
-best  linear mixed model via forward selection using the ExtBIC criterion. 
+Of the five multiple-locus methods, MLMM is the closest in philosophy to Eagle. It too is based on  building the 
+best  linear mixed model via forward selection using the extBIC criterion. 
 However, there are several differences between the two methods.
- MLMM does not make use of dimension reduction and they are implemented quite differently leading to substantial 
+ MLMM does not make use of dimension reduction, which gives it a larger  computational footprint than Eagle. 
+  It also builds its model differently to Eagle. Eagle uses a score statistic in which to identify SNPs to enter the model. 
+  MLMM uses the statistical significance of a SNP when it is treated as a fixed effect in the model. 
+ This necessitates fitting a separate linear mixed model for each unidentified SNP. Both are R packages but much of Eagle is written in 
+ C++ code. There are significant differences in run time and memory usage (see Results).
+
+ MLMM and Eagle are both R packages. However, much of Eagle is written in C++ code.
+
+ and they are implemented quite differently leading to substantial 
  differences  in run time and memory usage (see Results).  Also, they differ in how a model is built. 
- Eagle uses a score statistic formed from parameter estimates of the random genetic effect in which to identify the next snp to enter 
- the model.  MLMM uses the statistical significance of a snp when it is treated as a fixed effect in the model. 
- This necessitates fitting a separate linear mixed model for each unidentified snp. Fortunately,  MLMM does do this 
+ Eagle uses a score statistic formed from parameter estimates of the random genetic effect in which to identify the next SNP to enter 
+ the model.  MLMM uses the statistical significance of a SNP when it is treated as a fixed effect in the model. 
+ This necessitates fitting a separate linear mixed model for each unidentified SNP. Fortunately,  MLMM does do this 
  efficiently.
@@
 R2VIM is quite different to the other four methods. Here, a non-parametric model-free rather than model-based statistical approach is adopted. 
-R2VIM implements random forests and a new way of measuring the importance of a snp.  
+R2VIM implements random forests and a new way of measuring the importance of a SNP.  
 In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
-It is from these importance scores that snp can be ordered in terms of their strength of association with a trait. 
+It is from these importance scores that SNP can be ordered in terms of their strength of association with a trait. 
 The difficulty is in knowing what proportion of the 
-highest ordered snp are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
+highest ordered SNP are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
 random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score (f) is calculated for 
-each snp. The greater f is than one, the more evidence there is that the snp is in association with a trait.  A snp-trait association is 
-declared if f for the snp is above some threshold value across all parallel runs. The difficulty is that the relationship between 
+each SNP. The greater f is than one, the more evidence there is that the SNP is in association with a trait.  A SNP-trait association is 
+declared if f for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
 this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
@@
 
+\fi
+
 \subsection{Stability Selection}
 When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 
-BigRR has an approximate way of setting each snp's separate regularization parameter (ref). However, the optimal value of 
+BigRR has an approximate way of setting each SNP's separate regularization parameter (ref). However, the optimal value of 
 $\lambda$ must be found when using LMM-Lasso and glmnet. Also, 
-all three applications yield the effect sizes of the snp across the entire genome but not their statistical significance. To address 
+all three applications yield the effect sizes of the SNP across the entire genome but not their statistical significance. To address 
 these issues, we made use of stability selection \cite{meinshausen2010stability}. 
 Stability selection is a resampling strategy.  With stability selection, we were able to avoid having to optimise $\lambda$ while 
-still being able to calculate, empirically, the statistical significance of the snp effects.
+still being able to calculate, empirically, the statistical significance of the SNP effects.
 
@@
 appropriate value for the regularization parameter.  We adjusted $\lambda$ so that LMM-Lasso and glmnet yielded 
-between 10 to 30 snp with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
+between 10 to 30 SNPs with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
  We then repeatedly subsampled, 

### 2018-08-01 paper.tex
1August refining v1

@@
 
-Methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
+We present Eagle, a new method for multiple-locus association mapping. The
+motivation for developing Eagle was to make multiple locus association mapping ?easy? and the method-of-choice. 
+Eagle's strengths are that it a.
+is considerably more powerful than single-locus association mapping b.
+doesn?t suffer from multiple testing issues c.
+is threshold free d.
+gives results that are immediately interpretable and e.
+has a computational footprint comparable to single-locus association mapping.
+By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and 
+multiple-locus methods. We also analyse data from a published mouse study. Eagle found over two times more validated findings than the state-of-the-art 
+single-locus method. 
+Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
+
+%% Old version of abstract
+%%Methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
 Yet a well know limitation has persisted, the strength of association between a locus and trait is assessed only on a locus-by-locus basis.
@@
 Our goal was to develop a method that shifts mainstream association mapping from being single-locus to multiple-locus.
-We will show through the analysis of simulated and mouse data that Eagle finds true and avoids false
-associations better than completing single- and multiple-locus methods and does it faster.
+We will show through the analysis of simulated and mouse data that Eagle finds true and avoids false associations better than completing single- and multiple-locus methods and does it faster.
 Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users whom may be unfamiliar with R.
@@
 FaST-LMM and GEMMA implement single-locus association mapping.
-FaST-LMM was run in two ways. One way was where a subset of the SNP data were used in calculating the similarity (or relationship) matrix. Here, FaST-LMM is 
-highly efficient, computationally. 
-The other was where calculation of the 
-similarity matrix was based on all the SNP data. 
+FaST-LMM was run in two ways.
+One way was where a subset of the SNP data were used in calculating the similarity (or relationship) matrix.
+Here, FaST-LMM is highly efficient, computationally.
+The other was where calculation of the similarity matrix was based on all the SNP data.
 The $p$ values of the SNP were reported as their results.
@@
 
-\iffalse
+%%\iffalse
 
@@
 
-We compared the computational and statistical performance of Eagle against five other multiple-locus methods for 
-association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
-BigRR, LMM-Lasso, and glmnet  implement three different regression-based regularisation methods; generalized ridge regression, 
-lasso, and elastic net, respectively. Regularisation methods can handle parameter estimation problems where the number of predictors is
- far greater than the number of  samples. This is made possible through a penalty function that balances bias against variance. 
-The application of regularisation methods to association mapping has appeal because it means that the SNP effects across an 
-entire genome can be estimated, simultaneously.  The downside is that it is challenging, analytically, to calculate  
-the statistical significance of an SNP effect. We instead calculate significance, empirically, via stability selection (see below). 
+We purposely chose  methods that differed from one another. BigRR, LMM-Lasso, and glmnet are all regression-based regularisation 
+methods. However, BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
+Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
+They allow the strength of association between all the SNPs and trait to be measured within a single model, simultaneously. 
+A limitation of these methods though is that the statistical significance of the SNP effects cannot be easily determined. 
+ Due to the adaptive nature of the estimation procedures, to do this 
+analytically is challenging and is an area of active research \ref{lockhart2014significance}.
+
+The application of regularisation methods to association mapping is appealing because the strength of association between all the SNPs and 
+trait can be measured simultaneously. 
+ The downside is that it is challenging to calculate  
+the statistical significance of a SNP effect, analytically. We instead calculate significance, empirically, via stability selection (see below). 
 %% Permutation has been suggest as an empirical solution to this 
@@
 
-Of the five multiple-locus methods, MLMM is the closest in philosophy to Eagle. It too is based on  building the 
+MLMM is close in philosophy to Eagle. It too is based on  building the 
 best  linear mixed model via forward selection using the extBIC criterion. 
@@
 
- MLMM and Eagle are both R packages. However, much of Eagle is written in C++ code.
-
- and they are implemented quite differently leading to substantial 
- differences  in run time and memory usage (see Results).  Also, they differ in how a model is built. 
- Eagle uses a score statistic formed from parameter estimates of the random genetic effect in which to identify the next SNP to enter 
- the model.  MLMM uses the statistical significance of a SNP when it is treated as a fixed effect in the model. 
- This necessitates fitting a separate linear mixed model for each unidentified SNP. Fortunately,  MLMM does do this 
- efficiently.
-
-R2VIM is quite different to the other four methods. Here, a non-parametric model-free rather than model-based statistical approach is adopted. 
-R2VIM implements random forests and a new way of measuring the importance of a SNP.  
+R2VIM is different to the other four methods in that it is a non-parametric model-free approach. It implements  
+random forests and a new way of measuring the importance of a SNP.  
 In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
-It is from these importance scores that SNP can be ordered in terms of their strength of association with a trait. 
+It is from these importance scores that SNPs can be ordered in terms of their strength of association with a trait. 
 The difficulty is in knowing what proportion of the 
-highest ordered SNP are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
-random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score (f) is calculated for 
-each SNP. The greater f is than one, the more evidence there is that the SNP is in association with a trait.  A SNP-trait association is 
-declared if f for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
+highest ordered SNPs are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
+random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score ($f$) is calculated for 
+each SNP. The greater $f$ is than one, the more evidence there is that the SNP is in association with the trait.  A SNP-trait association is 
+declared if $f$ for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
 this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
@@
 
-We were also interested in comparing Eagle to single-locus methods to measure the difference in run times and statistical power.
-Even by limiting our focus to LMM-based methods, there were a number of efficient applications to choose from.
-We decided on GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
-These two applications have the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.
-They were developed at the same time, independently, but are similar theoretically.
-Both perform a single spectral decomposition of the relationship matrix $K$.
-Both use the eigenvector matrix to rotate the data.
-Both reformulate the log likelihood and
-restricted/residual maximum-likelihood (REML) 
+GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast} are of  the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.
+Both perform a single spectral decomposition of the relationship matrix $K$, use  eigenvector matrix to rotate the data,
+and reformulate the (reml) log likelihood for easier computation.  They differ though in their estimation procedure. 
+FaST-LMM implement's the Brent's algorithm. GEMMA instead implement's the Newton-Raphson algorithm. The  Newton-Raphson is more complicated  but has superior 
+convergence properties.
+
+===>
+
 log likelihood into a sum of $n$ terms that are easier to compute.
@@
 
-\fi
+%%\fi
 

### 2018-08-01 paper.tex
1August refining v2

@@
 
-The application of regularisation methods to association mapping is appealing because the strength of association between all the SNPs and 
-trait can be measured simultaneously. 
- The downside is that it is challenging to calculate  
-the statistical significance of a SNP effect, analytically. We instead calculate significance, empirically, via stability selection (see below). 
-%% Permutation has been suggest as an empirical solution to this 
-%% limitation \cite{shen2013novel} 
-%% but we have instead opted for stability selection (see below).
-
-MLMM is close in philosophy to Eagle. It too is based on  building the 
-best  linear mixed model via forward selection using the extBIC criterion. 
-However, there are several differences between the two methods.
- MLMM does not make use of dimension reduction, which gives it a larger  computational footprint than Eagle. 
-  It also builds its model differently to Eagle. Eagle uses a score statistic in which to identify SNPs to enter the model. 
-  MLMM uses the statistical significance of a SNP when it is treated as a fixed effect in the model. 
- This necessitates fitting a separate linear mixed model for each unidentified SNP. Both are R packages but much of Eagle is written in 
- C++ code. There are significant differences in run time and memory usage (see Results).
+MLMM is closest in philosophy to Eagle. It too is based on building the best model via forward selection, 
+within  a linear mixed model framework, using the 
+extBIC. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
+model differently to Eagle.  Eagle uses a score statistic in which to identify SNPs to enter the model.
+MLMM uses the statistical significance of the SNPs. This involves fitting a separate linear mixed model for each unidentified SNP. 
+Both are R packages but there is a significant difference in computational performance (see Results).
 
@@
 random forests and a new way of measuring the importance of a SNP.  
+Here, multiple parallel runs are performed. Within a run, a relative importance score is calculated for each SNP and 
+calculated by dividing the importance score for a SNP by a run's minimum importance score. 
+Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
+The relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
+but the computation cost is high.
+
+ The greater the score is than one, 
+the stronger the evidence is that the SNP is in association with the trait.  Only those SNPs with scores above a certain threshold are 
+deemed as being significant. The difficulty is that the relationship between this threshold and false positive rate is unknown. The threshold could be 
+found empirically, via permutation but this is computationally demanding.
+
 In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 

### 2018-08-01 paper.tex
1August refining v3

@@
 
-We purposely chose  methods that differed from one another. BigRR, LMM-Lasso, and glmnet are all regression-based regularisation 
-methods. However, BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
+We compared the computational and statistical performance of Eagle against five multiple-locus methods. 
+BigRR, LMM-Lasso, and glmnet are all regression-based regularisation 
+methods. BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
 Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
@@
  Due to the adaptive nature of the estimation procedures, to do this 
-analytically is challenging and is an area of active research \ref{lockhart2014significance}.
+analytically is challenging and is an area of active research \ref{lockhart2014significance}. Instead, we calculate significance empirically via 
+stability selection (see below).
 
@@
 
-R2VIM is different to the other four methods in that it is a non-parametric model-free approach. It implements  
-random forests and a new way of measuring the importance of a SNP.  
-Here, multiple parallel runs are performed. Within a run, a relative importance score is calculated for each SNP and 
-calculated by dividing the importance score for a SNP by a run's minimum importance score. 
+R2VIM differs to the other four methods in that it is a non-parametric model-free approach. It implements  
+random forests but where multiple parallel runs are performed. Each run begins with a different random number seed, leading to 
+different random forests being formed.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
+importance score by the minimum importance score observed across all the SNPs within a run. 
 Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
-The relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
-but the computation cost is high.
-
- The greater the score is than one, 
-the stronger the evidence is that the SNP is in association with the trait.  Only those SNPs with scores above a certain threshold are 
-deemed as being significant. The difficulty is that the relationship between this threshold and false positive rate is unknown. The threshold could be 
-found empirically, via permutation but this is computationally demanding.
-
-In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
-It is from these importance scores that SNPs can be ordered in terms of their strength of association with a trait. 
-The difficulty is in knowing what proportion of the 
-highest ordered SNPs are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
-random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score ($f$) is calculated for 
-each SNP. The greater $f$ is than one, the more evidence there is that the SNP is in association with the trait.  A SNP-trait association is 
-declared if $f$ for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
-this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
-this would be computationally challenging for all but the smallest of association studies.
+Unfortunately, the relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
+but the computational cost is high, restricting the size of data that can be analysed.
 
@@
 Both perform a single spectral decomposition of the relationship matrix $K$, use  eigenvector matrix to rotate the data,
-and reformulate the (reml) log likelihood for easier computation.  They differ though in their estimation procedure. 
-FaST-LMM implement's the Brent's algorithm. GEMMA instead implement's the Newton-Raphson algorithm. The  Newton-Raphson is more complicated  but has superior 
-convergence properties.
-
-===>
-
-log likelihood into a sum of $n$ terms that are easier to compute.
-The difference lies in their estimation procedure.
-FaST-LMM implement's the Brent's algorithm to optimise $\delta$.
+and reformulate the (reml) log likelihood for easier computation.  They differ in their estimation procedure. 
+FaST-LMM implement's the Brent's algorithm.
 GEMMA instead implement's the Newton-Raphson algorithm.
-Newton-Raphson is more complicated in that it also requires the first and second derivatives of a function to be calculated.
-However, it is superior in terms of its convergence properties to the Brent algorithm.
-Both applications are stand-alone computer programs, popular, and in common use.
-
-%%\fi
+The  Newton-Raphson is more complicated  but has superior convergence properties to Brent's algorithm.
+Both applications are stand-alone computer programs and in common use.
 

### 2018-08-02 paper.tex
2Aug refinement v1

@@
 
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood.  For complex models, residual maximum likelihood
+Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (reml).  For complex models, residual maximum likelihood
 can be computationally demanding. However, our model only contains a single random effect ($a_{-S}$). Here, highly efficient single-dimension 
@@
 
-We compared the computational and statistical performance of Eagle against five multiple-locus methods. 
-BigRR, LMM-Lasso, and glmnet are all regression-based regularisation 
+We compare the computational and statistical performance of Eagle against five multiple-locus methods. They are BigRR, LMM-Lasso, glmnet, 
+MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
+BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
 methods. BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
@@
  Due to the adaptive nature of the estimation procedures, to do this 
-analytically is challenging and is an area of active research \ref{lockhart2014significance}. Instead, we calculate significance empirically via 
+analytically is challenging and is an area of active research \cite{lockhart2014significance}. Instead, we calculate significance empirically via 
 stability selection (see below).
@@
 R2VIM differs to the other four methods in that it is a non-parametric model-free approach. It implements  
-random forests but where multiple parallel runs are performed. Each run begins with a different random number seed, leading to 
-different random forests being formed.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
+random forests but where multiple parallel runs are performed. Each run leads to 
+different random forests being created.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
 importance score by the minimum importance score observed across all the SNPs within a run. 
@@
 
-GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast} are of  the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.
-Both perform a single spectral decomposition of the relationship matrix $K$, use  eigenvector matrix to rotate the data,
-and reformulate the (reml) log likelihood for easier computation.  They differ in their estimation procedure. 
-FaST-LMM implement's the Brent's algorithm.
-GEMMA instead implement's the Newton-Raphson algorithm.
-The  Newton-Raphson is more complicated  but has superior convergence properties to Brent's algorithm.
-Both applications are stand-alone computer programs and in common use.
+We also compared the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
+Both are based on linear mixed models.
+The models have a single fixed effect for the SNP,  other fixed effects,
+a single random effect to account for familial relatedness (or polygenic background), and an error. The significance of the SNP effect in the model 
+is a measure of the strength of association.  They are of the same computational complexity \cite{zhou2012genome}, and produce exact results.  
+Both perform a single spectral decomposition of the relationship (or similarity) matrix $K$, use  an eigenvector matrix to rotate the data, 
+and reformulate the (residual) log likelihood for easier computation. They do differ in their estimation procedure. GEMMA implements Newton-Raphson. 
+FaST-LMM implements Brent's algorithm. Newton-Raphson is more complicated but has better convergence properties than Brent's algorithm. 
+Both methods are state-of-the-art and have been implemented in highly efficient computer programs.
 
 \subsection{Stability Selection}
+Stability selection is a subsampling strategy with a wide range of applications.  In the context of regularisation methods, it is a way of estimating, 
+empirically, the significance of the SNP effects. Compared to permutation the amount of computation is significantly less.  
+Analyses are based on only a fraction of the data  and the regularisation parameter need not be optimised.
+
+.
+We use it for calculating, empirically,  the statistical significance of the SNP effects for the regularisation methods.
+Unlike permutation, only subsets of the data are drawn for (repeated) analysis.
+Also, here,
+the regularisation parameter need not be optimised. Instead it can be fixed to a "suitable" value, greatly reducing the amount of computation.
+
+We performed stability selection as follows. For LMM-Lasso and glmnet, we conducted a preliminary analysis to find an 
+appropriate value for the regularisation parameter.  We adjusted the parameter so that the analysis yielded 
+between 10 to 30 SNPs with non-zero effects. We knew for the simulated data that the number of SNPs in association with the trait followed 
+a Poisson distribution with mean 30. In practice, this is obviously unknown. However, stability selection is robust to the number of non-zero effects. 
+Any reasonable number here would suffice. 
+ We then repeatedly subsampled, without replacement, the data.
+As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$.
+Each subsample was analysed with LMM-Lasso and glmnet, with their regularisation parameters fixed.
+
+ We analysed a subsample, with the regularisation parameter  fixed for LMM-Lasso and glmnet.  
+ To calculate, empirically,  the statistical significance of each 
+SNP across the genome, we counted the number of times a SNP had a non-zero effect size over all the replicates. We then 
+divided this number by the number of replicates (which was 100) to form an empirical probability estimate of the significance of a SNP.
+
+====>
+
 When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 

### 2018-08-02 paper.tex
2Aug refinement v2

@@
 Stability selection is a subsampling strategy with a wide range of applications.  In the context of regularisation methods, it is a way of estimating, 
-empirically, the significance of the SNP effects. Compared to permutation the amount of computation is significantly less.  
+empirically, the significance of the effects. Compared to permutation, the amount of computation is significantly less.  
 Analyses are based on only a fraction of the data  and the regularisation parameter need not be optimised.  
+To perform stability selection, we did the following.  
+For LMM-Lasso and glmnet, the regularisation parameter was fixed to a "suitable" value. This was done  
+via a preliminary analysis of the data where we adjusted the regularisation parameter  to yield
+between 20 to 30 SNPs with non-zero effect sizes.  Here, the number of non-zero effects is rather arbitrary and any reasonable 
+number of non-zero effects would suffice. 
+ We then repeatedly subsampled, without replacement, the data.
+As recommended \cite{meinshausen2010stability}, we drew 100 subsamples of size $n/2$ from the data.
+Each subsample was analysed with LMM-Lasso and glmnet, with their regularisation parameters fixed.
+A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
+number of times the SNP had a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
 
-.
-We use it for calculating, empirically,  the statistical significance of the SNP effects for the regularisation methods.
-Unlike permutation, only subsets of the data are drawn for (repeated) analysis.
-Also, here,
-the regularisation parameter need not be optimised. Instead it can be fixed to a "suitable" value, greatly reducing the amount of computation.
-
-We performed stability selection as follows. For LMM-Lasso and glmnet, we conducted a preliminary analysis to find an 
-appropriate value for the regularisation parameter.  We adjusted the parameter so that the analysis yielded 
-between 10 to 30 SNPs with non-zero effects. We knew for the simulated data that the number of SNPs in association with the trait followed 
-a Poisson distribution with mean 30. In practice, this is obviously unknown. However, stability selection is robust to the number of non-zero effects. 
-Any reasonable number here would suffice. 
- We then repeatedly subsampled, without replacement, the data.
-As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$.
-Each subsample was analysed with LMM-Lasso and glmnet, with their regularisation parameters fixed.
-
- We analysed a subsample, with the regularisation parameter  fixed for LMM-Lasso and glmnet.  
  To calculate, empirically,  the statistical significance of each 

### 2018-08-02 paper.tex
3Aug refinement v1

@@
 \subsection{Stability Selection}
-Stability selection is a subsampling strategy with a wide range of applications.  In the context of regularisation methods, it is a way of estimating, 
-empirically, the significance of the effects. Compared to permutation, the amount of computation is significantly less.  
-Analyses are based on only a fraction of the data  and the regularisation parameter need not be optimised.  
+Stability selection is a subsampling strategy with a range of applications, including the estimation of significance in regression models where 
+the number of parameters ($p$) is much larger than the number of samples ($n$). We chose stability selection over permutation and other 
+sampling procedures because of its low computational cost. 
+Stability selection proceeds as follows. For LMM-Lasso and 
+glmnet, we begin by performing a preliminary analysis of the data where  we set the  regularisation parameter to a value that yields 20 to 30 non-zero SNP effects.
+This was based on the number of SNP-trait associations we thought was realistic for the data.
+Fortunately, the significance estimates from
+stablity selection are robust (ref) so any reasonable number here would suffice.
+
+ selection is a subsampling strategy with a wide range of applications,  one of which is estimating the 
+statistical significance of effects in a regression model. Stability selection only requires analyses to take place on a fraction of the data. 
+Also, for these analyses, there is no need to find the entire solution path for the regularisation method. Instead, the regularisation method 
+is run at a fixed value of the regularisation parameter. This gives stability selection a much lower computational cost than other 
+sampling approaches such as permutation and multi-sampling.
+
+For the analysis of the simulated data, stability selection for LMM-Lasso and glmnet proceeds as follows. 
+We began by finding the value of the regularisation parameter 
+that gave 20 to 30 SNPs with non-zero effects.  This number was based on our a priori knowledge of the expected 
+number of SNPs in association with the (simulated) trait. Obviously, this is unknown when analysing real data.
+
+Stability selection for LMM-Lasso and glmnet was performed as follows. We began by finding the value of the regularisation parameter 
+that gave 20 to 30 SNPs with non-zero effects. This number was based on our a priori belief of the number of SNPs in association with the trait.
+
+ of 20 to 30 was based upon the number of SNP-trait associations we thought could 
+be found from the data.
+
+A preliminary analysis of the data was conducted to find the value 
+of the regularisation parameter that yields 20 to 30 SNPs with non-zero effect sizes.
+
 To perform stability selection, we did the following.  
-For LMM-Lasso and glmnet, the regularisation parameter was fixed to a "suitable" value. This was done  
-via a preliminary analysis of the data where we adjusted the regularisation parameter  to yield
-between 20 to 30 SNPs with non-zero effect sizes.  Here, the number of non-zero effects is rather arbitrary and any reasonable 
-number of non-zero effects would suffice. 
- We then repeatedly subsampled, without replacement, the data.
-As recommended \cite{meinshausen2010stability}, we drew 100 subsamples of size $n/2$ from the data.
-Each subsample was analysed with LMM-Lasso and glmnet, with their regularisation parameters fixed.
+We began by fixing the regularisation parameter  of LMM-Lasso and glment to a "suitable" value. Through preliminary analyses of the data, 
+we were able to adjust the regularisation parameter to yield 20 to 30 SNPs with non-zero effect sizes. This number of 20 to 30 was based on 
+the number of findings we were expecting from the analysis of the simulated data. Obviously, in practice, this is unknown. However,  the estimates 
+from stability selection are  robust (ref) so fixing the regularisation parameter to return any reasonable number of non-zero effect sizes would suffice. 
+ We repeatedly subsampled, without replacement, the data. We drew 
+ 100 subsamples of size $n/2$ from the data \cite{meinshausen2010stability}. Each subsample was analysed with LMM-Lasso and glmnet. 
+ The analyses were performed with the regularisation parameter fixed to the value found from the preliminary analysis.
 A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
 number of times the SNP had a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
+
+====>
 

### 2018-08-02 paper.tex
3Aug refinement v2

--- 2018-08-02 paper.tex
@@
 \subsection{Stability Selection}
-Stability selection is a subsampling strategy with a wide range of applications.  In the context of regularisation methods, it is a way of estimating, 
-empirically, the significance of the effects. Compared to permutation, the amount of computation is significantly less.  
-Analyses are based on only a fraction of the data  and the regularisation parameter need not be optimised.  
-To perform stability selection, we did the following.  
-For LMM-Lasso and glmnet, the regularisation parameter was fixed to a "suitable" value. This was done  
-via a preliminary analysis of the data where we adjusted the regularisation parameter  to yield
-between 20 to 30 SNPs with non-zero effect sizes.  Here, the number of non-zero effects is rather arbitrary and any reasonable 
-number of non-zero effects would suffice. 
- We then repeatedly subsampled, without replacement, the data.
-As recommended \cite{meinshausen2010stability}, we drew 100 subsamples of size $n/2$ from the data.
-Each subsample was analysed with LMM-Lasso and glmnet, with their regularisation parameters fixed.
+Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of significance in regression models where 
+the number of parameters ($p$) can be much larger than the number of samples ($n$).
+We chose stability selection over permutation and other sampling procedures because of its low computational cost.
+We employed stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
+
+To obtain significance estimates via stability selection, we did the following.
+For LMM-Lasso and glmnet, we began by performing a preliminary analysis of a replicate where  we tuned the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
+This tuning of the regularisation parameter was done for data under each simulation scenario but it was not necessary to do it for every replicate within a scenario.
+It was sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
+Here, we had the luxury of knowing the genetic conditions 
+under which the simulated trait data were generated. We knew that 20 to 30 SNP-trait associations was a reasonable number of findings to expect. 
+ However, stability selection estimates are robust  \cite{meinshausen2010stability} 
+ and the regularisation parameter could have been tuned to any reasonable number of 
+ non-zero effects. We sampled repeatedly, without replacement, from the replicate. We drew 100 replicate subsets of size $n/2$.  
+Each replicate subset was then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
 A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
@@
 
- To calculate, empirically,  the statistical significance of each 
-SNP across the genome, we counted the number of times a SNP had a non-zero effect size over all the replicates. We then 
-divided this number by the number of replicates (which was 100) to form an empirical probability estimate of the significance of a SNP.
-
-====>
-
-When using BigRR, LMM-Lasso, and glmnet, the results are affected by the amount of regularization ($\lambda$). 
-BigRR has an approximate way of setting each SNP's separate regularization parameter (ref). However, the optimal value of 
-$\lambda$ must be found when using LMM-Lasso and glmnet. Also, 
-all three applications yield the effect sizes of the SNP across the entire genome but not their statistical significance. To address 
-these issues, we made use of stability selection \cite{meinshausen2010stability}. 
-Stability selection is a resampling strategy.  With stability selection, we were able to avoid having to optimise $\lambda$ while 
-still being able to calculate, empirically, the statistical significance of the SNP effects.
-
-We performed stability selection as follows. For LMM-Lasso and glmnet, we performed a preliminary analysis to find an 
-appropriate value for the regularization parameter.  We adjusted $\lambda$ so that LMM-Lasso and glmnet yielded 
-between 10 to 30 SNPs with non-zero effects. Fortunately, with stability selection, the setting of $\lambda$ does not have to be exact 
- We then repeatedly subsampled, 
-without replacement, the data. As recommended \cite{meinshausen2010stability}, we draw 100 subsamples of size $n/2$. We 
-analysed the subsample, with $\lambda$ fixed for LMM-Lasso and glmnet.  To calculate, empirically,  the statistical significance of each 
-SNP across the genome, we counted the number of times a SNP had a non-zero effect size over all the replicates. We then 
-divided this number by the number of replicates (which was 100) to form an empirical probability estimate of the significance of a SNP.
-
 For BigRR, we modified our stability selection procedure slightly. 
-There was no need to find an appropriate value for the regularization 
-parameters as the BigRR package has an internal procedure for their 
-calculation. We draw subsamples as above and analyzed the data with BigRR. However, within an analysis, we ordered 
-the SNPs according to the absolute size of their SNP effects and recorded the top 20 SNPs. We then measured the significance 
-of the SNPs across the entire genome as above.
+There was no need to tune the regularisation parameter for BigRR as an optimal value is found as part of its analysis. 
+As described above, we drew 100 replicate subsets of size $n/2$ and analysed these data with BigRR. We then ordered the 
+SNPs according to the absolute size of their SNP effects and recorded the top 20 SNPs. 
+A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
+number of times the SNP was in the top 20 SNPs divided by the number of replicates.
 

### 2018-08-03 paper.tex
3Aug refinement v3

@@
 We present Eagle, a new method for multiple-locus association mapping. The
-motivation for developing Eagle was to make multiple locus association mapping ?easy? and the method-of-choice. 
+motivation for developing Eagle was to make multiple locus association mapping "easy" and the method-of-choice. 
 Eagle's strengths are that it a.
 is considerably more powerful than single-locus association mapping b.
-doesn?t suffer from multiple testing issues c.
+doesn't suffer from multiple testing issues c.
 is threshold free d.
@@
 \subsection{Stability Selection}
-Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of significance in regression models where 
+Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of the significance of effects in regression models where 
 the number of parameters ($p$) can be much larger than the number of samples ($n$).
@@
 A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
-number of times the SNP was in the top 20 SNPs divided by the number of replicates.
+number of times the SNP was recorded divided by the number of replicates.
 
@@
 
-To generate the trait data $y$, first, $q$, the number of SNP that are to be assigned a quantitative value is drawn from a Poisson distribution with 
-mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNP. The SNP genotypes AA, AB, and BB 
+To generate the trait data $y$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
+mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
 are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
@@
 not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach (ref) 
-is often adopted to deal with this situation, be we chose not to introduce this complexity to the analyses.
-
-\subsection{Analyses of Mouse Data}
+is often adopted to deal with this situation, be we chose not to introduce this complexity into the analyses.
 
@@
 
-Eagle has been implemented as an R package of the same name. Much of the heavy computation though is performed outside of R 
-via C++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
+Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
+via C++ functions that utilise Eigen C++ routines (ref). Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
 mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
@@
 a link to our server for demonstrating Eagle on some test data are available on the Eagle website (ref).  
-Eagle is available for download from the CRAN website.
+Eagle is available for download from the CRAN website (ref).
 
@@
 
-Eagle outperformed the other multiple-locus methods in our simulation study. However, we are aware/cognisant that we made several implementation 
-choices that impacts our conclusions.  For instance, we chose to calculate the significance of the 
+Eagle outperformed the other multiple-locus methods in our simulation study. However, we are aware of the fact  that we made several implementation 
+choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
 SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
@@
 captured in this first-stage model.
-In the second stage, Eagle is run not on the original but adjusted trait data which were obtained from the first stage analysis.
+In the second stage, Eagle is run not on the original trait data but adjusted trait data which were obtained from the first stage analysis.
 Even though this is a well accepted practice, it is approximate (ref).
@@
 Although not specifically designed for association mapping,
-WGAIM, upon which Eagle is based, and RRGAIM are two R packages where this is possible. Here, the difficulty is that for large datasets and/or complex 
+WGAIM, upon which Eagle is based, and RRGAIM are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
 models,  run time and memory usage can become limiting factors for analysis.

### 2018-08-05 supmaterials.tex
6Aug18 about to add Ari files v1

@@
 \begin{table}
-\caption{{\bf Implementation and methodology attributes of eight computer programs/packages for genome-wide association mapping.} For the different types of model, 
-LMM is linear mixed model.  GLM is generalised linear model., GLMM is generalised linear mixed model, and RF is random forests. 
-}
+\caption{Implementation and methodology attributes of eight computer programs/packages for genome-wide association mapping.  }
 \label{suptabsummary}
@@
                                                    Attributes                  & {\bf Eagle}                                 & {\bf bigRR}             & {\bf glmnet}            & {\bf LMM-Lasso}                    & {\bf MLMM} & {\bf r2VIM}      & {\bf FaST-LMM} & {\bf GEMMA} \\  \hline
-{\em Implementation}    &         &            &             &                   &            &                &      &      \\ 
+{\em Implementation}    &         &            &             &                   &            &                &      &      \\ [0.35cm]
 \hspace{1mm}  Purpose built$^a$    &   Yes     &    Yes      &  No   &   Yes  &  Yes  &  Yes  & Yes  & Yes          \\ [0.35cm]
@@
 
-{\em Methodology}    &         &            &             &                   &            &                &      &      \\  
-\hspace{1mm}  Model                     & LMM   &   &  GLMM &   GLM  & LMM  & RF & LMM   & LMM \\ [0.15cm]
+{\em Methodology}    &         &            &             &                   &            &                &      &      \\  [0.35cm]
+\hspace{1mm}  Model $^d$                    & LMM   &   &  GLMM &   GLM  & LMM  & RF & LMM   & LMM \\ [0.15cm]
 
-\hspace{1mm} SNPs fitted       &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.35cm]
+\hspace{1mm} SNPs fitted $^e$      &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.35cm]
 
@@
 
-\hspace{1mm} Threshold free$^5$        &    Yes     &  No    &  No         &    No               &     Yes       &  No              &  No   & No      \\  [0.15cm]  \hline
+\hspace{1mm} Threshold free      &    Yes     &  No    &  No         &    No               &     Yes       &  No              &  No   & No      \\  [0.15cm]  \hline
 
@@
 {$^c$ \scriptsize{Capacity for additional fixed effects (such as  age, sex, and/or population structure effects) to be included directly in the model.}} \\
-\end{table}
-
-\end{landscape}
-
-\begin{landscape}
-
-\begin{table}
-\caption{Some of the key features possessed by Eagle and the other seven computer programs/packages for association mapping}
-\label{suptabsummary}
-
-\begin{tabular}{lccccccccc} 
-                                &  \multicolumn{9}{c}{Computer Programs/Packages for Association Mapping} \\ \cline{2-10}
-                                 & \multicolumn{6}{c}{Multiple-locus}  & & \multicolumn{2}{c}{Single-locus} \\  \cline{2-7}   \cline{9-10}
- Features                 & Eagle  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
-Purpose built$^1$    &   Yes     &    Yes      &  No   &  Yes  &  Yes  &  Yes && Yes & Yes           \\ [0.2cm]
-Well documented$^2$           & Yes &   No   & Yes   & No     &   Yes     & No          && Yes    &    Yes    \\  [0.2cm]
-Simultaneous   &    Yes     &    Yes        &   Yes          &      Yes             &  No          &  No              &&   No   &  No     \\  
-fitting of SNPs     &         &            &             &                   &            &                &&      &      \\  [0.2cm]
-Additional fixed effects$^3$          &   Yes   &      Yes      &     Yes    &  No                 &      Yes      &      Yes          &&      Yes & Yes     \\   [0.2cm]
-Data larger than RAM$^4$                   &   Yes    &      No      & No          &  No    &  No    &   No     && Yes      & No    \\  [0.2cm]
-Threshold free$^5$        &    Yes     &  No    &  No         &    No               &     Yes       &  No              &&  No   & No      \\  [0.2cm]
-Informative error checking           &  Yes      &     No        &      No     &           No        &      No       &         No       && Yes  & No \\ \hline
-
-\end{tabular}
-{$^1$ \scriptsize{Specifically created for the analysis of GWAS data.}}\\
-{$^2$ \scriptsize{More than just a readme file or comments in an example file. Programs/packages with a tick had a detailed user manual.}}\\
-{$^3$ \scriptsize{Ability to handle additional fixed effects such as  age, sex, and/or population structure effects.}} \\
-{$^4$ \scriptsize{Able to deal with data larger than the memory capacity of the computer.}} \\
-{$^5$ \scriptsize{Results reported as the set of SNP closest to the genes influencing a trait. No need to construct thresholds to determine 
-significance of the findings.}}\\
-{$^6$  \scriptsize{All the programs terminated on errors. However, not all the programs informed the user of the cause and how to fix the errors.}}\\
+{$^d$ \scriptsize{For the different types of model, 
+LMM is linear mixed model.  GLM is generalised linear model., GLMM is generalised linear mixed model, and RF is random forests.}} \\
+{$^e$ \scriptsize{Association is assessed a SNP at a time (single), for multiple SNPs (multiple), or for all SNPs (all).
+Eagle fits all SNPs but also identifies multiple SNPs (All/multiple) in association with the trait.} } \\
 \end{table}

### 2018-10-02 paper.tex
adding refs v1

@@
 These designs though had a tendency to yield spurious findings if there was unrecognised population stratification.
-This prompted a shift towards family-based designs and score tests, such as the tdt test and its variants (refs).
-Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification (refs).
+This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \cite{spielman1996tdt}.
+Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification \cite{price2010new}.
 This has meant that data can be collected from general populations, even if these populations are highly structured.
-Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine (refs).
+Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine \cite{purcell2007plink, bradbury2007tassel}.
 
@@
 The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
-that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ valueue is obtained by testing the statistical 
+that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained by testing the statistical 
 significance of a SNP when treated as an effect in an appropriate model. 
@@
 Second, when multiple statistical tests are performed, the probability of wrongly accepting a result (type 1 error) is inflated.
-This is known as the multiple testing problem (refs).
-Many different solutions have been offered (refs).
+This is known as the multiple testing problem.
+Many different solutions have been offered \cite{storey2003statistical, li2005adjusting, de2005efficiency}.
 Yet, there is still no well accepted way of correcting for multiple testing in the context of genome-wide association mapping.
@@
 It is somewhat surprising then that multiple-locus association mapping methods haven't attracted more attention. Methods based on 
-regularisation techniques, such as ridge regression (ref) and lasso (ref), measure all locus-trait associations simultaneously. 
+regularisation techniques, such as ridge regression \cite{shen2013novel}  and lasso \cite{rakitsch2013lasso}, measure all locus-trait associations simultaneously. 
 Here, multiple testing is not an issue.
@@
 The strength of association is not measured by a $p$ value but by the size of the regression coefficient for the SNP in the model.
-More recently, associations have started to be mapped with random forests (refs).
-Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings (refs).
-A multiple-locus method that does show promise is the multiple-locus linear mixed model method (ref).
+More recently, associations have started to be mapped with random forests \cite{szymczak2016r2vim}.
+Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings.
+A multiple-locus method that does show promise is the multiple-locus linear mixed model method \cite{segura2012efficient}.
 The best multiple-locus model is built with simple forward selection.

### 2018-10-03 paper.tex
refinement of paper - adding references

@@
 \maketitle
+
+%%      To Do
+%%    1. is tense correct in methods - multiple and single locus mehtods
+%%    2. add bold font to matrices and vectors
+%%    3. change plot titles to be more condense
+%%    4. fix box bounding problem with figure 1.
 
@@
  followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
-linear regression.
-These designs though had a tendency to yield spurious findings if there was unrecognised population stratification.
+linear regression. These designs though had a tendency to yield spurious findings if there was unrecognised population stratification 
+\cite{cardon2003population}.
 This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \cite{spielman1996tdt}.
@@
 
- Analyses by the seven programs/packages of a replicate proceeded as follows. They were all run at their default settings. 
+ Analyses by the eight programs/packages of a replicate proceeded as follows. They were all run at their default settings. 
  Eagle and MLMM were the easiest of the programs/packages to implement. 
@@
 
-The median run times for Eagle and the other computer programs across the six scenarios are shown in Figure \ref{fig_time}. 
+The median run times for Eagle and the other computer programs/packages across the six scenarios are shown in Figure \ref{fig_time}. 
 The x- and y-axes are on a log scale.  A unit change on the x- or y-axis is equivalent to a change in the order of magnitude.  
@@
 \caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
-Eagle is compared against five other multiple-locus programs (top left) and two single-locus programs (top right). 
+Eagle is compared against five other multiple-locus programs/packages (top left) and two single-locus programs (top right). 
 The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
-programs, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
-The actual median run times for the programs across the scenarios are given in the table. The entries in a bold font 
+programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
+The actual median run times for the programs/packages across the scenarios are given in the table. The entries in a bold font 
 correspond to the lowest run-time for a scenario. 
  FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.  
- FaST-LMM$^{few}$ is where calculation  the similarity matrix is based on a subset of the SNP data. }
+ FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
 
@@
 We restricted our attention to these scenarios because not all
-multiple-locus methods could handle the size of data in the other scenarios. 
+multiple-locus methods could cope with the size of data in the other scenarios. 
   Each plot contains 
@@
 were found. For the other six, neither FaST-LMM nor Eagle found any associations. 
-Each plot contains the number of SNP-trait associations that were found and in agreement with the original findings. 
+Each plot in Figure  \ref{figmouse} contains the number of SNP-trait associations that were found and in agreement with the original findings. 
 Neither method found SNPs that had not been first identified in the original mouse study so neither method found false positives. 
@@
 with its regularisation parameter fine tuned to the trait data, for seven traits found the same number of findings as FaST-LMM and for 32 traits found more findings.
-Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, SNP-trait findings respectively.
+Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, SNP-trait findings, respectively.
 Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
@@
 We converted the dosages into discrete genotypes 
-by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively. We focused on autosomal SNPs.
+by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively. 
+We focused our analyses on the autosomal SNPs.
 
@@
 The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
-quantitative trait loci in experimental crosses \cite{verbyla2007analysis}
+quantitative trait loci in experimental crosses \cite{verbyla2007analysis}.
 
@@
 $e^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \; \times \; n)})$. 
-So far,  this model differs little from standard linear mixed models for association mapping (refs). 
+So far,  this model differs little from standard linear mixed models for association mapping \cite{yu2006unified,zhao2007arabidopsis} 
 However, 
@@
 
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (reml).  For complex models, residual maximum likelihood
+Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
 can be computationally demanding. However, our model only contains a single random effect ($a_{-S}$). Here, highly efficient single-dimension 
-optimisation via spectral decomposition is possible (EMMA ref).
+optimisation via spectral decomposition is possible \cite{kang2008efficient}.
 
@@
 and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
-identify a SNP based on its (random) effect size and accuracy, but is theoretically justified.  It follows from outlier detection 
-in linear and linear mixed models (ref).
+identify a SNP based on its (random) effect size and accuracy, but is theoretically justified \cite{verbyla2012rwgaim}.
 
@@
 where $a^* \sim N(0, \sigma_a^2 I^{(n_g \; \times \;  n_g)})$, and 
-$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via single value decomposition (ref).  
+$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
 Although it may not be obvious, the two models are equivalent, 

### 2018-10-03 supmaterials.tex
refinement of paper - adding references

@@
                                                    Attributes                  & {\bf Eagle}                                 & {\bf bigRR}             & {\bf glmnet}            & {\bf LMM-Lasso}                    & {\bf MLMM} & {\bf r2VIM}      & {\bf FaST-LMM} & {\bf GEMMA} \\  \hline
-{\em Implementation}    &         &            &             &                   &            &                &      &      \\ [0.35cm]
-\hspace{1mm}  Purpose built$^a$    &   Yes     &    Yes      &  No   &   Yes  &  Yes  &  Yes  & Yes  & Yes          \\ [0.35cm]
+{\bf {\em Implementation}}    &         &            &             &                   &            &                &      &      \\ [0.15cm]
+\hspace{1mm}  Purpose built$^a$    &   Yes     &    Yes      &  No   &   Yes  &  Yes  &  Yes  & Yes  & Yes          \\ [0.15cm]
 
 \hspace{1mm}  Language                 &  R/C++       &    R        &      R       &     Python     &  R          &    R         &  C++ and     &   C++   \\  
-                                                          &         &            &             &                   &            &                                     &   Python$^b$       &      \\  [0.35cm]
+                                                          &         &            &             &                   &            &                                     &   Python$^b$       &      \\  [0.15cm]
 
-\hspace{1mm} GUI                            & Yes &    No      & No          &  No    &  No    &   No     & No     & No    \\  [0.35cm]
+\hspace{1mm} GUI                            & Yes &    No      & No          &  No    &  No    &   No     & No     & No    \\  [0.15cm]
 
-\hspace{1mm} Documentation        &  Videos,                &    R help       &  Vignettes,    &   Readme.txt,       &   Vignette,          &     R help           &   Videos,              &  User-manual,     \\
-                                                        &   user-manuals,    &                      &  R help         &    test script           &  R help              &                           &    user-manuals,  &   website  \\
-                                                        &   website,    R help          &                      &                     &                              &                           &                           &    website            &    \\   [0.35cm]
+\hspace{1mm} Documentation        &  Videos,  user-manuals,              &    R help       &  Vignettes,    &   Readme.txt,       &   Vignette,          &     R help           &   Videos, website             &  User-manual,     \\
+                                                        &  website,    R help     &                      &  R help         &    test script           &  R help              &                           &    user-manuals,  &   website  \\ [0.15cm]
 
-\hspace{1mm} Additional fixed        &   Yes   &      Yes      &     Yes    &  No                 &      Yes      &      Yes          &      Yes & Yes     \\  
-\hspace{1mm} effects$^c$               &           &                   &                 &                   &                   &                     &              &   adsf    \\  [0.35cm]
+\hspace{1mm} Additional fixed        &   Yes   &      Yes      &     Yes    &  No                 &      Yes      &      No          &      Yes & Yes     \\  
+\hspace{1mm} effects$^c$               &           &                   &                 &                   &                   &                     &              &      \\  [0.15cm]
 
-\hspace{1mm} Types of  &  Cont.      &   Cont.,          &   Cont.,           &   Cont.                 &   Cont.          &     Cont.           &   Cont.   &  Cont.    \\  
-\hspace{1mm} trait data  &                &    binary,      &      binary,         &                   &            &                &      &      \\  
-                                      &                &   count         &       count          &                   &            &                &      &      \\  [0.35cm]
+\hspace{1mm} Types of  &  Cont.      &   Cont.,          &   Cont.,           &   Cont.                 &   Cont.          &     Cont.,          &   Cont.   &  Cont.,    \\  
+\hspace{1mm} trait data  &                &    binary,      &      binary,         &                   &            &     binary           &      &   binary    \\  
+                                      &                &   count         &       count          &                   &            &                &      &      \\  [0.15cm]
 
@@
 
-{\em Methodology}    &         &            &             &                   &            &                &      &      \\  [0.35cm]
-\hspace{1mm}  Model $^d$                    & LMM   &   &  GLMM &   GLM  & LMM  & RF & LMM   & LMM \\ [0.15cm]
+{\bf {\em Methodology} }   &         &            &             &                   &            &                &      &      \\  [0.15cm]
+\hspace{1mm}  Model $^d$                    & LMM   & HEM  &  GLMM &   LMM  & LMM  & RF & LMM   & LMM, mvLMM\\ 
+                      					    &             &   &              &             &          &        &           &   Bayesian Sparse LMM      \\  [0.15cm]
 
-\hspace{1mm} SNPs fitted $^e$      &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.35cm]
+\hspace{1mm} SNPs fitted $^e$      &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.15cm]
 
-\hspace{1mm}  Selection type             & Model   & Variable  &  Variable &  Variable  &  Model  &  Variable & Variable  & Variable \\  [0.35cm]
+\hspace{1mm}  Selection type             & Model   & Variable  &  Variable &  Variable  &  Model  &  Variable & Variable  & Variable \\  [0.15cm]
 
@@
 \begin{figure}
-\caption{Memory usage (in gigabytes) of Eagle and the other association mapping programs across 
+\caption{Memory usage (in gigabytes) of Eagle and the other association mapping programs/packages across 
 the six simulation scenarios. The maximum amount of memory on the computer is 128 gigabytes. 
@@
 
-\begin{table}
-\caption{The median run times (in minutes) of Eagle and the other association mapping programs across the six simulation scenarios. }
-\begin{tabular}{llcccccc}
-              &           &  \multicolumn{6}{c}{Simulation Scenarios} \\ \cline{3-8}
- Method & Name & 150 x 5K & 1500 x 50K & 350 x 400K & 2000 x 500K & 4000 x 1.5M & 10000 x 1.5M \\ \hline
-  Multiple & Eagle 	&	   0.08  &   1.62   &  \bf{2.7}1  &   \bf{13.65}   & \bf{127.63}  &   699.55  \\
-               & MLMM 	&	   0.15    &  2.91    & 19.04  &  143.01  &  870.84  &     \\
-               & glmnet 	&	  0.11     & 3.95    & 14.06    & 74.03    &        &    \\
-               & r2VIM 	&	   0.09    &  3.66    &  5.51    & 50.59    & 380.52  &   \\ 
-               & bigRR 	&	    1.01   & 113.35   &  54.99   & 1030.61  &        &     \\
-               & LMM-Lasso 	&     0.57  &   52.08 &    92.20  & 1031.85 &           &     \\ \\
-Single     &  GEMMA 	&      0.02  &   5.02   &   6.17  &   84.83   & 723.33  & 4071.60 \\
-               & FaST-LMM$^{few}$ 	&    \bf{0.01}   &  \bf{0.80}   &   7.07   &  20.16   & 193.90   & \bf{346.19} \\ 
-               & FaST-LMM$^{all}$   	&   0.03   & 2.96  &    7.90   &  41.27  &            &   \\ \hline
-\end{tabular}
-
-\end{table}
+%% This table appears in Figure 1 in paper. It's included here in case I need to regenerate the table. 
+%%
+%%\begin{table}
+%%\caption{The median run times (in minutes) of Eagle and the other association mapping programs across the six simulation scenarios. }
+%%\begin{tabular}{llcccccc}
+%%              &           &  \multicolumn{6}{c}{Simulation Scenarios} \\ \cline{3-8}
+%% Method & Name & 150 x 5K & 1500 x 50K & 350 x 400K & 2000 x 500K & 4000 x 1.5M & 10000 x 1.5M \\ \hline
+%% Multiple & Eagle 	&	   0.08  &   1.62   &  \bf{2.7}1  &   \bf{13.65}   & \bf{127.63}  &   699.55  \\
+%%               & MLMM 	&	   0.15    &  2.91    & 19.04  &  143.01  &  870.84  &     \\
+%%              & glmnet 	&	  0.11     & 3.95    & 14.06    & 74.03    &        &    \\
+%%               & r2VIM 	&	   0.09    &  3.66    &  5.51    & 50.59    & 380.52  &   \\ 
+%%               & bigRR 	&	    1.01   & 113.35   &  54.99   & 1030.61  &        &     \\
+%%               & LMM-Lasso 	&     0.57  &   52.08 &    92.20  & 1031.85 &           &     \\ \\
+%%Single     &  GEMMA 	&      0.02  &   5.02   &   6.17  &   84.83   & 723.33  & 4071.60 \\
+%%               & FaST-LMM$^{few}$ 	&    \bf{0.01}   &  \bf{0.80}   &   7.07   &  20.16   & 193.90   & \bf{346.19} \\ 
+%%               & FaST-LMM$^{all}$   	&   0.03   & 2.96  &    7.90   &  41.27  &            &   \\ \hline
+%%\end{tabular}
+%%
+%%\end{table}
 

### 2018-10-04 paper.tex
getting close to finishing YAY

@@
 
-We compare the computational and statistical performance of Eagle against five multiple-locus methods. They are BigRR, LMM-Lasso, glmnet, 
+We compare the computational and statistical performance of Eagle against five multiple-locus methods. They are bigRR, LMM-Lasso, glmnet, 
 MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
 BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
-methods. BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
+methods. BigRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
 Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
@@
 
-We also compared the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
+We also compare the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
 Both are based on linear mixed models.
@@
 We chose stability selection over permutation and other sampling procedures because of its low computational cost.
-We employed stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
+We employe stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
 
-To obtain significance estimates via stability selection, we did the following.
-For LMM-Lasso and glmnet, we began by performing a preliminary analysis of a replicate where  we tuned the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
-This tuning of the regularisation parameter was done for data under each simulation scenario but it was not necessary to do it for every replicate within a scenario.
-It was sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
-Here, we had the luxury of knowing the genetic conditions 
-under which the simulated trait data were generated. We knew that 20 to 30 SNP-trait associations was a reasonable number of findings to expect. 
+To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, we begin
+ by performing a preliminary analysis of a replicate where  we tune the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
+This tuning of the regularisation parameter is done for data under each simulation scenario but it is not necessary to do it for every replicate within a scenario.
+It is sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
+Here, we have the luxury of knowing the genetic conditions 
+under which the simulated trait data are generated. We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
  However, stability selection estimates are robust  \cite{meinshausen2010stability} 
- and the regularisation parameter could have been tuned to any reasonable number of 
- non-zero effects. We sampled repeatedly, without replacement, from the replicate. We drew 100 replicate subsets of size $n/2$.  
-Each replicate subset was then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
+ and the regularisation parameter can be tuned to any reasonable number of 
+ non-zero effects. We sample repeatedly, without replacement, from the replicate. We draw 100 replicate subsets of size $n/2$.  
+Each replicate subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
 A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
-number of times the SNP had a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
+number of times the SNP have a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
 
-For BigRR, we modified our stability selection procedure slightly. 
-There was no need to tune the regularisation parameter for BigRR as an optimal value is found as part of its analysis. 
-As described above, we drew 100 replicate subsets of size $n/2$ and analysed these data with BigRR. We then ordered the 
-SNPs according to the absolute size of their SNP effects and recorded the top 20 SNPs. 
+For bigRR, we modify our stability selection procedure slightly. 
+There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
+As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
+SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
 A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
-number of times the SNP was recorded divided by the number of replicates.
+number of times the SNP is recorded divided by the number of replicates.
 
@@
 
-The data were generated via data perturbation (Zhao et al. 2007). Data perturbation amalgamates real with simulated data to generate replicates. 
+The data are generated via data perturbation (Zhao et al. 2007). Data perturbation amalgamates real with simulated data to generate replicates. 
 It is a way of introducing greater realism into a simulation study. 
@@
 quantitative trait data are simulated. 
- The SNP genotypes were drawn, 
+ The SNP genotypes are drawn, 
 according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \cite{10002010map}. 
@@
 In forming $y$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
-not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach (ref) 
-is often adopted to deal with this situation, be we chose not to introduce this complexity into the analyses.
+not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
+is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.
 
@@
 calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
-a link to our server for demonstrating Eagle on some test data are available on the Eagle website (ref).  
-Eagle is available for download from the CRAN website (ref).
+a link to our server for demonstrating Eagle on some test data are available on the Eagle website (http://eagle.r-forge.r-project.org).  
+Eagle is available for download from the CRAN website (https://cran.r-project.org).
 
@@
 
-Eagle outperformed the other multiple-locus methods in our simulation study. However, we are aware of the fact  that we made several implementation 
+Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
 choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
 SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
-Permutation (ref) and multi sample-splitting (ref) are also equally valid empirical approaches.
+Permutation and its variants \cite{ browning2008presto, pahl2010permory  } are also equally valid empirical approaches.
 Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
@@
 The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
-r2VIM could of been fine-tuned for a better balance of computational and statistical performance because random forests are highly adjustable. 
-However, we would like to think that the parameter settings we used are sensible since they match the values in the original r2VIM publication (ref).
+r2VIM could be fine-tuned for a better balance of computational and statistical performance. 
+However, we would like to think that the parameter settings we used are sensible 
+since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.
 
@@
 captured in this first-stage model.
-In the second stage, Eagle is run not on the original trait data but adjusted trait data which were obtained from the first stage analysis.
+In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis.
 Even though this is a well accepted practice, it is approximate (ref).
@@
 Although not specifically designed for association mapping,
-WGAIM, upon which Eagle is based, and RRGAIM are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
+WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RRGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
 models,  run time and memory usage can become limiting factors for analysis.
 
-Over the coming years, computationally, the demand placed on association mapping methods is going to increase. 
+Over the coming years, computationally, the demand placed upon association mapping methods is going to increase. 
 High-throughput array-based technologies continue to decrease the cost of genotyping, permitting ever larger GWASs to be performed. 
-Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs (refs), 
+Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs \cite{gudbjartsson2015large, long2017whole}
 culminating in data on millions of SNPs. It is because of this growing demand that 

### 2018-10-04 paper.tex
revisions

@@
 \usepackage{graphicx}
+\usepackage{bm}
 
@@
 %%      To Do
-%%    1. is tense correct in methods - multiple and single locus mehtods
+%%   1. make ARi's changes to paper
 %%    2. add bold font to matrices and vectors
 %%    3. change plot titles to be more condense
-%%    4. fix box bounding problem with figure 1.
+%%    4.tidy up doc - reformat - remove comments make nice
+%%    5 write cover letter
+%%    6 send cover letter to dan
+%%    7 send paper to Josh
+%%    8 find reviewer
+
+\newcommand{\bwidetildea}{\widetilde{\bm{a}}}
+
+\newcommand{\bu}{\bm{u}}
+\newcommand{\be}{\bm{e}}
+\newcommand{\btau}{\bm{\tau}}
+\newcommand{\ba}{\bm{a}}
+\newcommand{\bzero}{\bm{0}}
+\newcommand{\bI}{\bm{I}}
+
+\newcommand{\bX}{\bm{X}}
+\newcommand{\by}{\bm{y}}
+\newcommand{\bZ}{\bm{Z}}
+\newcommand{\bM}{\bm{M}}
+\newcommand{\bmm}{\bm{m}}
 
@@
 false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
-and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
+and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 x 400K),  1,500 individuals and 
 50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
@@
 Memory usage
-for many of the computer programs/packages was the limiting factor (See {\bf Supplementary Figure \ref{supfigmem}}).  The single-locus program GEMMA was by 
+for many of the computer programs/packages was the limiting factor (See Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
 far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
@@
 \label{fig_time}
-\begin{center}
-\includegraphics[width=14cm, height=12cm]{Figure2_time.jpg}
-\end{center}
-
+\centering
+    \includegraphics[width=1.0\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
 \end{figure}
@@
 SNP were selected. Then $S=\{500023, 15, 420\}$. 
-Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
-Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
+Let $\by^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
+Let $\bM^{(n_g \; \times \; L)} = [\bmm_1 \bmm_2 \ldots \bmm_L]$ be a matrix containing the genotype data which have been collected 
 from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
 may be taken of the same group/line/strain. 
- It is common for the columns of $M$ to be in map order but this is not a requirement. 
-The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
+ It is common for the columns of $\bM$ to be in map order but this is not a requirement. 
+The vector $\bmm_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
 The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.
@@
 \label{eq1}
-y = X \tau + Z u_g + e
+\by = \bX \btau + \bZ \bu_g + \be
 \end{equation}
 where 
-$X^{(n \; \times \; p)}$ and $Z^{( n \; \times \; n_g)}$ are known design matrices with $X$ being of full rank and $Z$ 
+$\bX^{(n \; \times \; p)}$ and $\bZ^{( n \; \times \; n_g)}$ are known design matrices with $\bX$ being of full rank and $\bZ$ 
 containing zeros and ones that assign the appropriate genetic effect to each measurement. 
 The vector 
-$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
-$u_g^{(n_g \; \times \; 1)}$ contains the 
+$\btau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
+$\bu_g^{(n_g \; \times \; 1)}$ contains the 
 genetic effects. The vector of residuals is 
-$e^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \; \times \; n)})$. 
+$\be^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(\bzero, \sigma^2_e \bI^{(n \; \times \; n)})$. 
 So far,  this model differs little from standard linear mixed models for association mapping \cite{yu2006unified,zhao2007arabidopsis} 
 However, 
-it is how we specify $u_g$ that distinguishes our model from the others.
+it is how we specify $\bu_g$ that distinguishes our model from the others.
 
-The genetic effects $u_g$ are modelled as 
+The genetic effects $\bu_g$ are modelled as 
 \begin{equation}
 \label{eq2}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + M_{-S} a_{-S}
+\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} +\bM_{-S} \ba_{-S}
 \end{equation}
-where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
-$a_{S_k}$ is the additive effect of the $k$th selected SNP, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
-with the data for the SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
-$a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
+where $\bmm_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
+$a_{S_k}$ is the additive effect of the $k$th selected SNP, $\bM_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
+with the data for the SNP in $S$ removed,  and $\ba_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
+$\ba_{-S} \sim N(\bzero, \sigma_a^2 \bI^{(L-s \; \times \;  L-s)})$. 
 The terms in the summation on the left hand side are fixed effects.   They account 
@@
 Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
-can be computationally demanding. However, our model only contains a single random effect ($a_{-S}$). Here, highly efficient single-dimension 
+can be computationally demanding. However, our model only contains a single random effect ($\ba_{-S}$). Here, highly efficient single-dimension 
 optimisation via spectral decomposition is possible \cite{kang2008efficient}.
@@
 In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
-The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
+The vector $\ba_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
 $L$, the number of SNPs, can be extremely large.  An alternative model is given by 
@@
 \label{eq3}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + (M_{-S} M_{-S}^T)^{1/2} a^*_{-S}
+\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} + (\bM_{-S} \bM_{-S}^T)^{1/2} \ba^*_{-S}
 \end{equation}
-where $a^* \sim N(0, \sigma_a^2 I^{(n_g \; \times \;  n_g)})$, and 
-$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
+where $\ba^* \sim N(\bzero, \sigma_a^2 \bI^{(n_g \; \times \;  n_g)})$, and 
+$(\bM_{-S} \bM_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
 Although it may not be obvious, the two models are equivalent, 
@@
 
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model  (\ref{eq3})  with 
+Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\bwidetildea$ from estimates from model  (\ref{eq3})  with 
 \begin{equation}
-\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
+\bwidetildea = \left [ \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \right ] \bwidetildea^*
 \end{equation}
@@
 \label{eq4}
-\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
+\textrm{var}(\bwidetildea) = \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \textrm{var}(\bwidetildea^*) (\bM_{-S} \bM_{-S}^T)^{-1/2} \bM_{-S}
 \end{equation}
@@
 
-The data are generated via data perturbation (Zhao et al. 2007). Data perturbation amalgamates real with simulated data to generate replicates. 
+The data are generated via data perturbation \cite{zhao2007arabidopsis}. 
+Data perturbation amalgamates real with simulated data to generate replicates. 
 It is a way of introducing greater realism into a simulation study. 
@@
 
-To generate the trait data $y$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
+To generate the trait data $\by$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
 mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
 are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
-generate a $g^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
-Fifth, $e^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
-the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $y =  g + e$.  
-In forming $y$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
+generate a $\bm{g}^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
+Fifth, $\bm{e}^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
+the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $\by =  \bm{g} + \bm{e}$.  
+In forming $\by$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
 not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
@@
 Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
-via C++ functions that utilise Eigen C++ routines (ref). Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
+via C++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
 mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
@@
 SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
-Permutation and its variants \cite{ browning2008presto, pahl2010permory  } are also equally valid empirical approaches.
+Permutation and its variants \cite{ browning2008presto,pahl2010permory} are also equally valid empirical approaches.
 Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
@@
 Although not specifically designed for association mapping,
-WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RRGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
+WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RWGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
 models,  run time and memory usage can become limiting factors for analysis.

### 2018-10-08 paper.tex
revisions

@@
 %%   1. make ARi's changes to paper
-%%    2. add bold font to matrices and vectors
+%%    X 2. add bold font to matrices and vectors
 %%    3. change plot titles to be more condense
@@
 that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained by testing the statistical 
-significance of a SNP when treated as an effect in an appropriate model. 
+significance of a single nucleotide polymorphisms (SNP) when treated as an effect in an appropriate model. 
 The location of peaks in this plot identify genomic 
@@
 Yet, there is still no well accepted way of correcting for multiple testing in the context of genome-wide association mapping.
-Third, many of the traits whose genetic secretes we are trying to discover are complex.
+Third, many of the traits whose genetic secrets we are trying to discover are complex.
 There will be multiple SNPs in linkage disequilibrium with genes that are influencing the trait.
@@
 false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
-and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 x 400K),  1,500 individuals and 
+and 5,000 single SNPs (150 x 5K),  350 individuals and 400,000 SNPs (350 x 400K),  1,500 individuals and 
 50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
@@
 and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
-identify a SNP based on its (random) effect size and accuracy, but is theoretically justified \cite{verbyla2012rwgaim}.
+identify a SNP based on its (random) effect size and accuracy, but is justifiable, theoretically \cite{verbyla2012rwgaim}.
 
-Fourth, we determine the importance of the $(s+1)$th selected SNP via a model selection strategy. 
+Fourth, we determine the importance of the $(s+1)$th selected SNP via a model selection strategy  \cite{verbyla2007analysis}. 
 We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected SNP.  We then fit this new model to the data
@@
 extBIC is a model selection measure that takes into account the number of unknown parameters and the complexity 
-of the model space.
-It is well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
+of the model space.  It is well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}. 
+It is different to the model selection measure used in  \cite{verbyla2007analysis}.
 If this new model has a larger extBIC than the current model, then the $s+1$th selected SNP is added to 
@@
 In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis.
-Even though this is a well accepted practice, it is approximate (ref).
+Even though this is a well accepted practice, it is approximate \cite{gogel2018comparison}.
 A better solution is to fit a single model to the data.

### 2018-10-13 figure1_runtimes.tex
put paper in correct format

--- 2018-10-04 paper.tex
@@
-\documentclass{article}
+\documentclass[10pt]{article}
 \usepackage{xr}
-\bibliographystyle{naturemag}
-\renewcommand{\baselinestretch}{1.3} 
+
 \usepackage{graphicx}
+\usepackage{amsfonts}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
-%% To Do
-%% 1. add bold font to matrices/vectors
-%% 2. change mouse results plot - use superscripts for traits of the same class to reduce space of titles.
-
-%% external tex files
-%%\externaldocument{figure1_runtimes}
-%%\externaldocument{figure2_powerFDRmultiple}
-%%\externaldocument{figure3_mouse}
-
-\externaldocument{supmaterials}
-
+\linespread{1}
+\addtolength{\oddsidemargin}{-2.5cm}
+\addtolength{\textwidth}{2.5cm}
 \begin{document}
-
-\title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George$^1$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
-
-\maketitle
-
-%%      To Do
-%%    1. is tense correct in methods - multiple and single locus mehtods
-%%    2. add bold font to matrices and vectors
-%%    3. change plot titles to be more condense
-%%    4. fix box bounding problem with figure 1.
-
-%%\begin{affiliations}
-%%\item Data61, CSIRO, Australia.
-%%\item IM \&T, CSIRO, Australia.
-%%\end{affiliations}
-
-\begin{abstract}
-
-We present Eagle, a new method for multiple-locus association mapping. The
-motivation for developing Eagle was to make multiple locus association mapping "easy" and the method-of-choice. 
-Eagle's strengths are that it a.
-is considerably more powerful than single-locus association mapping b.
-doesn't suffer from multiple testing issues c.
-is threshold free d.
-gives results that are immediately interpretable and e.
-has a computational footprint comparable to single-locus association mapping.
-By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and 
-multiple-locus methods. We also analyse data from a published mouse study. Eagle found over two times more validated findings than the state-of-the-art 
-single-locus method. 
-Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
-
-%% Old version of abstract
-%%Methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
-Yet a well know limitation has persisted, the strength of association between a locus and trait is assessed only on a locus-by-locus basis.
-Statistically superior multiple-locus methods that detect multiple SNP-trait associations, simultaneously, have been available for some time but they have failed to become the method-of-choice.
-Here, we present Eagle for multiple-locus association mapping.
-Our goal was to develop a method that shifts mainstream association mapping from being single-locus to multiple-locus.
-We will show through the analysis of simulated and mouse data that Eagle finds true and avoids false associations better than completing single- and multiple-locus methods and does it faster.
-Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users whom may be unfamiliar with R.
-\end{abstract}
-
-Over the past decade,  genome-wide association studies (GWASs) have changed considerably in both their analysis and design. Early studies
- followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
-linear regression. These designs though had a tendency to yield spurious findings if there was unrecognised population stratification 
-\cite{cardon2003population}.
-This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \cite{spielman1996tdt}.
-Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification \cite{price2010new}.
-This has meant that data can be collected from general populations, even if these populations are highly structured.
-Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine \cite{purcell2007plink, bradbury2007tassel}.
-
-What has not changed is that it remains common practice to analyse genome-wide association study (GWAS) data on a locus-by-locus basis.
-This is despite there being several significant problems with analysing data in this way.
-First, the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
-The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
-that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained by testing the statistical 
-significance of a SNP when treated as an effect in an appropriate model. 
-The location of peaks in this plot identify genomic 
-regions of interest.
-Inferring the exact number of regions though can be difficult if the peaks are not well separated.
-Second, when multiple statistical tests are performed, the probability of wrongly accepting a result (type 1 error) is inflated.
-This is known as the multiple testing problem.
-Many different solutions have been offered \cite{storey2003statistical, li2005adjusting, de2005efficiency}.
-Yet, there is still no well accepted way of correcting for multiple testing in the context of genome-wide association mapping.
-Third, many of the traits whose genetic secretes we are trying to discover are complex.
-There will be multiple SNPs in linkage disequilibrium with genes that are influencing the trait.
-Yet, a locus-by-locus mapping approach only assesses the evidence for association between a single marker locus and trait.
-
-It is somewhat surprising then that multiple-locus association mapping methods haven't attracted more attention. Methods based on 
-regularisation techniques, such as ridge regression \cite{shen2013novel}  and lasso \cite{rakitsch2013lasso}, measure all locus-trait associations simultaneously. 
-Here, multiple testing is not an issue.
-These techniques though are computationally demanding.
-Also, their results can be difficult to interpret.
-The strength of association is not measured by a $p$ value but by the size of the regression coefficient for the SNP in the model.
-More recently, associations have started to be mapped with random forests \cite{szymczak2016r2vim}.
-Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings.
-A multiple-locus method that does show promise is the multiple-locus linear mixed model method \cite{segura2012efficient}.
-The best multiple-locus model is built with simple forward selection.
-Results are immediately interpretable but here, computation becomes challenging for large datasets.
-
-In this paper, we present our new multiple-locus method for genome-wide association mapping, which we are calling Eagle. 
-Eagle combines the strength of regularisation techniques (being able to fit all SNP-trait associations jointly), with forward selection giving easy-to-interpret threshold-free results.
-We are able to achieve a computational performance similar to the fastest single-locus linear mixed model implementations
-through a dimension reduction step.
-Our aim was to make multiple-locus association mapping on a genome-wide scale routine. To this end, we have implemented Eagle 
-within an R package of the same name. 
-Our package accepts marker data of different 
-formats,  can handle data larger than a computer's  memory capacity, and makes heavy use of 
-parallel computing for computation when available.
-
-\section{Results}
-
-\subsection{Association Mapping Methods}
-
-We compared Eagle, in terms of computational and statistical performance, against seven other association mapping methods. 
-We chose methods that almost all had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs or packages.
-Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multiple-locus models.
-Of the many ways of performing single-locus association mapping, we chose
-GEMMA \cite{zhou2012genome}  and FaST-LMM \cite{lippert2011fast} because of their popularity and computational speed. 
-For multiple-locus association mapping, we chose bigRR \cite{shen2013novel}, glmnet \cite{Friedman2010glmnet}, 
-LMM-Lasso \cite{rakitsch2013lasso}, MLMM \cite{segura2012efficient} , and r2VIM \cite{szymczak2016r2vim}.  
-Each takes a different approach to multiple-locus association mapping. A summary of the key attributes of the different computer programs/packages 
-is given in Supplementary Table \ref{suptabsummary} (See Methods for further details).
-
-\subsection{Simulation Study}
-A large simulation study was performed where we sought to  answer two questions. 
-First, how does Eagle compare, in terms of run time and memory usage, to 
-competing implementations? Second, how well does Eagle find true associations (power) and avoid 
-false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
-and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
-50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
-1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).   
-These scenarios reflect, at least in some cases, the sizes of study being performed in animals, plants, and humans.
-
-For each scenario, 100 replicates were generated. A single replicate consisted of SNP and quantitative trait data. 
-Extra realism was introduced into the simulation study through the drawing of the SNP genotypes from the 1000 Genome Project, phase 3  \cite{10002010map}.
- The quantitative trait data were generated 
-by selecting, randomly, a set of SNPs and assigning these loci additive allelic effects.  Random errors were then drawn from a normal distribution 
-with variance set to give a heritability of 50\% for the trait. 
-For each individual, a quantitative trait value was obtained by summing its random error and additive allelic effects. 
-The number of randomly selected SNPs follows a Poisson distribution with mean 30. The size of the allelic effects 
- across the selected loci are equal.
-
- Analyses by the eight programs/packages of a replicate proceeded as follows. They were all run at their default settings. 
- Eagle and MLMM were the easiest of the programs/packages to implement. 
- The only parameters requiring specification were the amount of available memory and number of CPU for 
- Eagle and the number of chunks for MLMM. 
- Their results were also immediately 
- interpretable. Their findings were the set of SNPs in strongest association with the trait. Each 
-SNP in this set identified a separate genomic region of interest, whose position was given by the map location of the SNP.
-
-BigRR, LMM-Lasso, and glmnet required more effort to implement. They are based on regularisation methods and as such, all the SNPs were fitted simultaneously in a regression 
-framework. The difficulty was in calculating the significance of the SNP effects. To do this analytically is challenging. We instead opted for stability selection (see Methods),  
-an empirical approach for calculating significance.
-
-R2VIM is different from the rest in that it is a nonparametric approach for association mapping. It is based on random forests. Three important parameters needed to be  set. 
-These were the number of trees, the number of variables for building a tree, and the minimum size of a terminal node. Ideally, these parameters would be "tuned" on a replicate-by-replicate 
-basis \cite{boulesteix2012overview}. However, this was not practical here. We instead used the same settings as in \cite{szymczak2016r2vim} where 
-the number of trees was set to 1000, the number of variables was set to 20\% of the number of SNPs, and 
-  the minimum size of a node was set to 10\% of the sample size.
-A relative importance measure was calculated 
-for each SNP measuring its strength of association with the trait.
-
-FaST-LMM and GEMMA implement single-locus association mapping.
-FaST-LMM was run in two ways.
-One way was where a subset of the SNP data were used in calculating the similarity (or relationship) matrix.
-Here, FaST-LMM is highly efficient, computationally.
-The other was where calculation of the similarity matrix was based on all the SNP data.
-The $p$ values of the SNP were reported as their results.
-
-The results from all but Eagle and MLMM required post-processing before the findings were interpretable.  The SNPs were placed in map order, 
-a significance threshold was set, peak regions containing SNPs with significance measures above the threshold were identified, and the SNP with the largest 
-significance measure in each of the peak regions was recorded as a finding.
-
-\subsection{Memory Usage and Run Times}
-
-Memory usage and run (or elapse) times were recorded for Eagle and the other computer programs/packages across the simulation scenarios. 
-Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM.
-Not all data generated under the six scenarios could be analysed by all implementations.
-Memory usage
-for many of the computer programs/packages was the limiting factor (See {\bf Supplementary Figure \ref{supfigmem}}).  The single-locus program GEMMA was by 
-far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
-excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
-Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
-Of the multiple-locus programs/packages, only Eagle,  
-with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
-for data from our largest scenario, 10000 x 1.5M.
-
-The median run times for Eagle and the other computer programs/packages across the six scenarios are shown in Figure \ref{fig_time}. 
-The x- and y-axes are on a log scale.  A unit change on the x- or y-axis is equivalent to a change in the order of magnitude.  
-In answer to our question of how does Eagle compare in terms of run time to competing implementations, 
-Eagle was significantly faster, sometimes by orders of magnitude,  than the other multiple-locus
- implementations and is comparable to the single-locus implementations. For a simulation study with 150 individuals and 
- 5000 SNPs, Eagle produced results in seconds.  For the larger simulation scenarios of 1500 x  50K and 350 x4 00K, 
- analyses with Eagle took under two minutes. Even for data from a couple of thousand individuals and half a million 
- SNPs (2000 x 500K), the median run time of Eagle was under 14 minutes. For our scenarios where there 
- were thousands of individuals and 1.5 million SNPs, Eagle took just over two hours for the analysis of data from 
- 4000 x 1.5M and  12 hours for the analysis of data from 10000 x 1.5M. 
- Towards the final stages of writing this paper, 
- we gained access to a desktop computer with 14-core Xeon processors and 256 gigabytes of RAM. We reran Eagle on data from the largest
-  scenario 10000 x 1.5M to measure the impact on run time. The median run time dropped by more than 70\% 
-  from 12 hours to 3.31 hours.
-
- %%It is also worth noting that an increase in the number of SNP genotypes in a study does not necessitate, automatically,  
- %%an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
- %%increasing behaviour of the curves in the memory (Supplementary Figure \ref{supfigmem}) and run-time plots (Figure \ref{fig_time}).  It 
-%%is the study dimensions, not number of genotypes, that drives the computing resources required by
-%%association mapping programs.
 
@@
 
+%% plot formed via PowerPoint 
 \label{fig_time}
-\begin{center}
-\includegraphics[width=14cm, height=12cm]{Figure2_time.jpg}
-\end{center}
-
+\centering
+    \includegraphics[width=1.0\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
 \end{figure}
 
-%%{\em Figure \ref{fig_time} goes around here}
-
-\subsection{Power and False Discovery Rates}
-
-Here, we answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.
-We do this by estimating the power and false discovery rates of Eagle and the other methods for the six scenarios.
-Since, for a replicate, we knew which SNPs were assigned additive effects, we knew the SNPs that were in true association with the trait.
-We will refer to these SNPs as being
-true SNPs.
-By knowing the true SNPS, we were able to assess the validity of a method's findings.
-A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a true SNP.
-When a replicate was analysed, we obtained an estimate of the power of the method by taking  the number of findings that were found to be  true and dividing by the
-number of true SNPs.
-We also obtained an estimate of a method's false discovery rate.
-It is the number of findings that were found to be false divided by the number of (true and false) findings found by the method.
-Both these estimates varied with replicate.
-The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates over the 100 replicates.
-
-The power and false discovery rates of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
-We restricted our attention to these scenarios because not all
-multiple-locus methods could cope with the size of data in the other scenarios. 
-  Each plot contains 
-single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
-These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
-The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
-significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
-the false discovery rate as the significance threshold  is adjusted. 
-The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
-Supplementary Figure \ref{supfigpowersingle}.
-
-\begin{figure}
-\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
-scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
-scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
-upper limit of the x-axis is 0.1.}
-\label{figpowermultiple}
-\begin{center}
-\includegraphics[width=15cm, height=15cm]{power1main.eps}
-\end{center}
-
-\end{figure}
-
-%%{\em Figure \ref{figpowermultiple} goes here}
-
-In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
-well.  Of the multiple-locus methods, Eagle had the highest power
-while keeping its false discovery rate low (Figure \ref{figpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
-that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
-association with a trait while avoiding false associations (Supplementary Figure \ref{supfigpowersingle}).
-
-%%At first glance, Eagle appears underpowered.
-Several of the other methods have higher power than Eagle.
-The goal of association mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
-For those methods with higher power than Eagle, the cost is much higher false discovery rates.
-In other words, these methods are finding more true results  but they are also finding many more false results.
-If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than  5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest  fdr.
-This is especially true when the power and fdr of Eagle is compared to the single-locus methods.
-For single-locus methods to achieve the same level of power as Eagle, a threshold has to be set where the fdr is extremely high.
-
-\subsection{Mouse Data Analysis}
-
-We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
- We chose to focus on data from a large outbred mouse study \cite{nicod2016genome}.
-This study was a little unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
-Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \cite{zheng2011comparison}.
-By converting the dosages into genotypes and analysing the data with the single-locus program FaST-LMM, we obtained a subset of those findings reported in the original study.
-We then analysed the data with Eagle.
-Due to Eagles increased power, we found SNP-trait associations not found with  FaST-LMM.
-However, we were
- able to confirm the validity of these new findings as they matched what was found in the original study. Having the ability to confirm new findings  in a real study was 
- one of the primary motivators for choosing these data for analysis.
-
-We repeated the single-locus analyses, as first performed  \cite{nicod2016genome}, except 
-that we focused on autosomal SNP and our analyses were based on SNP genotype  rather than SNP dosage. 
-In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation. We followed the 
-same empirical procedure but had to increase the number of permutations from 100 to 500.
-
-Eagle was run in two ways; under its default settings (Eagle$^{default}$) and where we specified the regularisation parameter for model selection (Eagle$^{optimal}$ ).
-Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}.
-  The conservativeness of the extBIC can be adjusted by a single regularisation parameter that ranges from zero to one.
-In the simulation study, this parameter was set to one, its most conservative and default setting.
-However, there is also opportunity to set the parameter to a value less than one.
-This increases power but also increases the false discovery rate.
-For each trait, we used permutation to set the regularisation parameter to give a false discovery rate of 5\% .
-
-The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}. The mouse study recorded
-measurements on 200 traits. When these traits were first analysed in the original study, findings for 45 of these traits were able to be 
-corroborated by prior published evidence. We focused our analyses here on these same 45 traits. For 39 traits, SNP-trait associations 
-were found. For the other six, neither FaST-LMM nor Eagle found any associations. 
-Each plot in Figure  \ref{figmouse} contains the number of SNP-trait associations that were found and in agreement with the original findings. 
-Neither method found SNPs that had not been first identified in the original mouse study so neither method found false positives. 
-As we saw in the simulation study, there was a notable difference between the two methods capacity to discover SNP-trait associations.
-Eagle$^{default}$, under its default settings, for two traits found fewer findings than FaST-LMM, for eight traits found the same number
-of findings as FaST-LMM, and for 28 traits found more findings. Eagle$^{optimal}$, 
-with its regularisation parameter fine tuned to the trait data, for seven traits found the same number of findings as FaST-LMM and for 32 traits found more findings.
-Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, SNP-trait findings, respectively.
-Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
-associations than what is the established way of analysing these data. Furthermore, these extra results found by Eagle were not 
-false positives but could be confirmed from the original study.
-
-\begin{figure}
-\caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
-multiple-locus method
-Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
-for a false discovery rate of 5\% (Eagle$^{optimal}$). 
-The number of SNP-trait associations found are reported.
-The bar plots with boxes around them are for those traits where using Eagle (default and optimal)   resulted in more findings than with single-locus analysis.}
-\label{figmouse}
-\begin{center}
-%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
-\includegraphics[width=15cm, height=15cm]{mouseresults.eps}
-\end{center}
-
-\end{figure}
-
-\section{Methods}
-
-\subsection{Mouse Data}
-
-The data were obtained from a large genome-wide association study that was performed in outbred mice \cite{nicod2016genome}. 
-Phenotypic and genotypic data were available on 1,887 adult mice. 
-The phenotypic data included raw and adjusted (for fixed effects) measurements from 200 behavioural, tissue, and physiological traits.  
-Of these traits, 
-43 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
-43 traits that were the focus of our real data analyses. As in the original study  \cite{nicod2016genome}, our analyses 
-were based on the adjusted traits.
-Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
-form of marker dosages (expected allele counts that ranged from zero to one). All missing data had been imputed. 
-We converted the dosages into discrete genotypes 
-by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively. 
-We focused our analyses on the autosomal SNPs.
-
-\subsection{Eagle Approach for Multiple-locus Association Mapping}
-
-Eagle is a method for multiple-locus association mapping on a genome-wide scale.
-It is based on linear mixed models.
-It differs from most other single- and multiple-locus association mapping methods.
-Eagle treats association mapping as a model selection instead of variable selection problem.
-Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
-Eagle also 
-reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest to the genes influencing a trait. 
-The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
-quantitative trait loci in experimental crosses \cite{verbyla2007analysis}.
-
-Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of $s$ ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
-selected in the $k$th iteration of the model building process. Suppose three iterations  $(s=3)$
-have been performed and say the 500023rd, 15th, and 420th
-SNP were selected. Then $S=\{500023, 15, 420\}$. 
-Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
-Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
-may be taken of the same group/line/strain. 
- It is common for the columns of $M$ to be in map order but this is not a requirement. 
-The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
-The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.
-
-The specifics of the Eagle method are as follows. 
-Eagle builds the "best" model iteratively, via forward selection. 
-Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
-associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
-To perform the $s+1$th  iteration, we first fit the current model to the data. 
-The (current) model is of the form 
-\begin{equation}
-\label{eq1}
-y = X \tau + Z u_g + e
-\end{equation}
-where 
-$X^{(n \; \times \; p)}$ and $Z^{( n \; \times \; n_g)}$ are known design matrices with $X$ being of full rank and $Z$ 
-containing zeros and ones that assign the appropriate genetic effect to each measurement. 
-The vector 
-$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
-$u_g^{(n_g \; \times \; 1)}$ contains the 
-genetic effects. The vector of residuals is 
-$e^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \; \times \; n)})$. 
-So far,  this model differs little from standard linear mixed models for association mapping \cite{yu2006unified,zhao2007arabidopsis} 
-However, 
-it is how we specify $u_g$ that distinguishes our model from the others.
-
-The genetic effects $u_g$ are modelled as 
-\begin{equation}
-\label{eq2}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + M_{-S} a_{-S}
-\end{equation}
-where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
-$a_{S_k}$ is the additive effect of the $k$th selected SNP, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
-with the data for the SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
-$a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
-The terms in the summation on the left hand side are fixed effects.   They account 
-for the additive effects of those SNPs that have been found to be in association with the trait. The other term is a random effect. 
-It accounts for the joint effect of the yet-to-be-identified SNP that are in association with the trait. 
-This is a simple genetic model but it 
-is effective for discovering SNP-trait associations.
-
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
-can be computationally demanding. However, our model only contains a single random effect ($a_{-S}$). Here, highly efficient single-dimension 
-optimisation via spectral decomposition is possible \cite{kang2008efficient}.
-
-Third, we identify the $(s+1)$th SNP that is in strongest association with the trait, based on the maximum score statistic
-$t_j^2 = \frac{ \widetilde{a} _j^2}{\textrm{var}(\widetilde{a}_j)}$ where $\widetilde{a}_j$ is the best linear unbiased predictor (BLUP), 
-and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
-identify a SNP based on its (random) effect size and accuracy, but is theoretically justified \cite{verbyla2012rwgaim}.
-
-Fourth, we determine the importance of the $(s+1)$th selected SNP via a model selection strategy. 
-We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected SNP.  We then fit this new model to the data
-via maximum likelihood and calculate its extended Bayesian information criteria (extBIC) \cite{chen2008extended}.  The 
-extBIC is a model selection measure that takes into account the number of unknown parameters and the complexity 
-of the model space.
-It is well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}.
-If this new model has a larger extBIC than the current model, then the $s+1$th selected SNP is added to
-the current model and the above process is repeated. If this new model has a smaller extBIC than the current model, then the 
-model building process is complete. The set of SNP in strongest association with the trait is the $s$ SNPs previously identified.
-
-\subsubsection{Reducing the dimension of the model}
-In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
-The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
-$L$, the number of SNPs, can be extremely large.  An alternative model is given by 
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
-They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
-\begin{equation}
-\label{eq3}
-u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + (M_{-S} M_{-S}^T)^{1/2} a^*_{-S}
-\end{equation}
-where $a^* \sim N(0, \sigma_a^2 I^{(n_g \; \times \;  n_g)})$, and 
-$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
-Although it may not be obvious, the two models are equivalent, 
-having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
-model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
-effects needing estimating.
-
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model  (\ref{eq3})  with 
-\begin{equation}
-\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
-\end{equation}
-where its variance matrix is
-\begin{equation}
-\label{eq4}
-\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
-\end{equation}
-These values are needed in order to calculate the score statistic $t_j^2$ for identifying the SNP in strongest association with the trait. 
-Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation 
-of (\ref{eq4}).
-
-%%\iffalse
-
-\subsection{Comparison Methods}
-
-\subsubsection{Multiple-locus methods}
-
-We compare the computational and statistical performance of Eagle against five multiple-locus methods. They are bigRR, LMM-Lasso, glmnet, 
-MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
-BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
-methods. BigRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
-Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
-They allow the strength of association between all the SNPs and trait to be measured within a single model, simultaneously. 
-A limitation of these methods though is that the statistical significance of the SNP effects cannot be easily determined. 
- Due to the adaptive nature of the estimation procedures, to do this 
-analytically is challenging and is an area of active research \cite{lockhart2014significance}. Instead, we calculate significance empirically via 
-stability selection (see below).
-
-MLMM is closest in philosophy to Eagle. It too is based on building the best model via forward selection, 
-within  a linear mixed model framework, using the 
-extBIC. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
-model differently to Eagle.  Eagle uses a score statistic in which to identify SNPs to enter the model.
-MLMM uses the statistical significance of the SNPs. This involves fitting a separate linear mixed model for each unidentified SNP. 
-Both are R packages but there is a significant difference in computational performance (see Results).
-
-R2VIM differs to the other four methods in that it is a non-parametric model-free approach. It implements  
-random forests but where multiple parallel runs are performed. Each run leads to 
-different random forests being created.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
-importance score by the minimum importance score observed across all the SNPs within a run. 
-Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
-Unfortunately, the relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
-but the computational cost is high, restricting the size of data that can be analysed.
-
-\subsubsection{Single-locus methods}
-
-We also compare the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
-Both are based on linear mixed models.
-The models have a single fixed effect for the SNP,  other fixed effects,
-a single random effect to account for familial relatedness (or polygenic background), and an error. The significance of the SNP effect in the model 
-is a measure of the strength of association.  They are of the same computational complexity \cite{zhou2012genome}, and produce exact results.  
-Both perform a single spectral decomposition of the relationship (or similarity) matrix $K$, use  an eigenvector matrix to rotate the data, 
-and reformulate the (residual) log likelihood for easier computation. They do differ in their estimation procedure. GEMMA implements Newton-Raphson. 
-FaST-LMM implements Brent's algorithm. Newton-Raphson is more complicated but has better convergence properties than Brent's algorithm. 
-Both methods are state-of-the-art and have been implemented in highly efficient computer programs.
-
-\subsection{Stability Selection}
-Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of the significance of effects in regression models where 
-the number of parameters ($p$) can be much larger than the number of samples ($n$).
-We chose stability selection over permutation and other sampling procedures because of its low computational cost.
-We employe stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
-
-To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, we begin
- by performing a preliminary analysis of a replicate where  we tune the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
-This tuning of the regularisation parameter is done for data under each simulation scenario but it is not necessary to do it for every replicate within a scenario.
-It is sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
-Here, we have the luxury of knowing the genetic conditions 
-under which the simulated trait data are generated. We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
- However, stability selection estimates are robust  \cite{meinshausen2010stability} 
- and the regularisation parameter can be tuned to any reasonable number of 
- non-zero effects. We sample repeatedly, without replacement, from the replicate. We draw 100 replicate subsets of size $n/2$.  
-Each replicate subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
-A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
-number of times the SNP have a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
-
-For bigRR, we modify our stability selection procedure slightly. 
-There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
-As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
-SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
-A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
-number of times the SNP is recorded divided by the number of replicates.
-
-\subsection{Generation of Simulation Data}
-
-The data are generated via data perturbation (Zhao et al. 2007). Data perturbation amalgamates real with simulated data to generate replicates. 
-It is a way of introducing greater realism into a simulation study. 
- Here, the genotype data are real, the 
-quantitative trait data are simulated. 
- The SNP genotypes are drawn, 
-according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \cite{10002010map}. 
-Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.
-
-To generate the trait data $y$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
-mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
-are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
-generate a $g^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
-Fifth, $e^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
-the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $y =  g + e$.  
-In forming $y$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
-not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
-is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.
-
-\subsection{Implementation}
-
-Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
-via C++ functions that utilise Eigen C++ routines (ref). Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
-mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
-calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
-a link to our server for demonstrating Eagle on some test data are available on the Eagle website (http://eagle.r-forge.r-project.org).  
-Eagle is available for download from the CRAN website (https://cran.r-project.org).
-
-\section{Discussion}
-Eagle is a new linear mixed model based method (and R package) for multiple-locus association mapping. It advances the state of association mapping in several ways. 
-First, its computational footprint is much smaller than other multiple-locus implementations. Eagle makes multiple-locus analysis 
-practical, even when the datasets are large. Second, the results from
- Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
-each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
-multiple testing issues and the need for significance thresholds. 
- As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
-Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
-with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.
-
-Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
-choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
-SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
-Permutation and its variants \cite{ browning2008presto, pahl2010permory  } are also equally valid empirical approaches.
-Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
-data. Also, when analysing the (sub)samples, it was not necessary to calculate the entire solution path for a method. 
- Instead,  analyses are 
-performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
-implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
-The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
-r2VIM could be fine-tuned for a better balance of computational and statistical performance. 
-However, we would like to think that the parameter settings we used are sensible 
-since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.
-
-Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here, although in different ways. 
-Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
-is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
-including the extra random effects, is 
-captured in this first-stage model.
-In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis.
-Even though this is a well accepted practice, it is approximate (ref).
-A better solution is to fit a single model to the data.
-Although not specifically designed for association mapping,
-WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RRGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
-models,  run time and memory usage can become limiting factors for analysis.
-
-Over the coming years, computationally, the demand placed upon association mapping methods is going to increase. 
-High-throughput array-based technologies continue to decrease the cost of genotyping, permitting ever larger GWASs to be performed. 
-Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs \cite{gudbjartsson2015large, long2017whole}
-culminating in data on millions of SNPs. It is because of this growing demand that 
-we have purposely structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
-Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
-computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times.  
-We also have plans for Eagle to run on computer clusters. Structuring Eagle for larger-than-memory calculations was a 
-preemptive step in this direction. GWASs have changed significantly in the past decade but the size and complexity of GWASs is expected 
-to change even more in the coming decade.
-
-%%\bibliographystyle{naturemag}
-\bibliography{biblibrary}
-
-%%\begin{addendum}
-%% \item Put acknowledgements here.
-%% \item[Competing Interests] The authors declare that they have no
-%%competing financial interests.
-%% \item[Correspondence] Correspondence and requests for materials
-%%should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
-%%\end{addendum}
-
 \end{document}

### 2018-10-13 figure3_mouse.tex
put paper in correct format

--- 2018-06-29 ForAri/figure3_mouse.tex
@@
 \begin{figure}
-\caption{Genome-wide association mapping results from the analysis of the mouse data for the single-locus method FaST-LMM and the 
+\caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
 multiple-locus method
@@
 for a false discovery rate of 5\% (Eagle$^{optimal}$). 
-The number of SNP-trait associations found are reported. The bar plots with boxes around them are for those traits where using 
-Eagle$^{default}$ and Eagle$^{optimal}$ resulted in more findings than with single-locus analysis.} 
+The number of SNP-trait associations found are reported.
+The bar plots framed by boxes are for those traits where using Eagle (default and optimal)   resulted in more findings than with single-locus analysis.}
 \label{figmouse}

### 2018-10-13 paper.tex
put paper in correct format

@@
-\documentclass{article}
+\documentclass{nature}
 \usepackage{xr}
@@
 \usepackage{bm}
-
-%% To Do
-%% 1. add bold font to matrices/vectors
-%% 2. change mouse results plot - use superscripts for traits of the same class to reduce space of titles.
+\usepackage{lineno}
+\linenumbers
 
 %% external tex files
-%%\externaldocument{figure1_runtimes}
-%%\externaldocument{figure2_powerFDRmultiple}
-%%\externaldocument{figure3_mouse}
+\externaldocument{figure1_runtimes}
+\externaldocument{figure2_powerFDRmultiple}
+\externaldocument{figure3_mouse}
 
@@
 %%      To Do
-%%   1. make ARi's changes to paper
-%%    X 2. add bold font to matrices and vectors
-%%    3. change plot titles to be more condense
 %%    4.tidy up doc - reformat - remove comments make nice
-%%    5 write cover letter
-%%    6 send cover letter to dan
-%%    7 send paper to Josh
-%%    8 find reviewer
 
+%% New commands ... 
 \newcommand{\bwidetildea}{\widetilde{\bm{a}}}
-
 \newcommand{\bu}{\bm{u}}
@@
 \newcommand{\bI}{\bm{I}}
-
 \newcommand{\bX}{\bm{X}}
@@
 
-%%\begin{affiliations}
-%%\item Data61, CSIRO, Australia.
-%%\item IM \&T, CSIRO, Australia.
-%%\end{affiliations}
+\begin{affiliations}
+\item Data61, CSIRO, Australia.
+\item IM \&T, CSIRO, Australia.
+\end{affiliations}
 
 \begin{abstract}
-
 We present Eagle, a new method for multiple-locus association mapping. The
@@
 Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
-
-%% Old version of abstract
-%%Methods for genome-wide association mapping have become progressively more sophisticated, statistically and computationally.
-Yet a well know limitation has persisted, the strength of association between a locus and trait is assessed only on a locus-by-locus basis.
-Statistically superior multiple-locus methods that detect multiple SNP-trait associations, simultaneously, have been available for some time but they have failed to become the method-of-choice.
-Here, we present Eagle for multiple-locus association mapping.
-Our goal was to develop a method that shifts mainstream association mapping from being single-locus to multiple-locus.
-We will show through the analysis of simulated and mouse data that Eagle finds true and avoids false associations better than completing single- and multiple-locus methods and does it faster.
-Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users whom may be unfamiliar with R.
 \end{abstract}
@@
 
- %%It is also worth noting that an increase in the number of SNP genotypes in a study does not necessitate, automatically,  
- %%an increase in the memory usage and run times for its analysis. This is  evidenced by the non-monotonically 
- %%increasing behaviour of the curves in the memory (Supplementary Figure \ref{supfigmem}) and run-time plots (Figure \ref{fig_time}).  It 
-%%is the study dimensions, not number of genotypes, that drives the computing resources required by
-%%association mapping programs.
-
-\begin{figure}
-\caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
-Eagle is compared against five other multiple-locus programs/packages (top left) and two single-locus programs (top right). 
-The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
-programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
-The actual median run times for the programs/packages across the scenarios are given in the table. The entries in a bold font 
-correspond to the lowest run-time for a scenario. 
- FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.  
- FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
-
-\label{fig_time}
-\centering
-    \includegraphics[width=1.0\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
-\end{figure}
-
-%%{\em Figure \ref{fig_time} goes around here}
+ {\em Figure \ref{fig_time} goes around here}
 
@@
 
-\begin{figure}
-\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
-scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
-scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
-upper limit of the x-axis is 0.1.}
-\label{figpowermultiple}
-\begin{center}
-\includegraphics[width=15cm, height=15cm]{power1main.eps}
-\end{center}
-
-\end{figure}
-
-%%{\em Figure \ref{figpowermultiple} goes here}
+{\em Figure \ref{figpowermultiple} goes here}
 
@@
 association with a trait while avoiding false associations (Supplementary Figure \ref{supfigpowersingle}).
-
-%%At first glance, Eagle appears underpowered.
-Several of the other methods have higher power than Eagle.
-The goal of association mapping though is not only to find true SNP-trait associations but to also avoid false SNP-trait associations.
-For those methods with higher power than Eagle, the cost is much higher false discovery rates.
-In other words, these methods are finding more true results  but they are also finding many more false results.
-If we restrict our attention to that part of the plot where the (genome-wide) fdr is less than  5\% (see inset plots in Figure X and Sup Figure X),  we see that Eagle, closely followed by MLMM,  has the highest power for the lowest  fdr.
-This is especially true when the power and fdr of Eagle is compared to the single-locus methods.
-For single-locus methods to achieve the same level of power as Eagle, a threshold has to be set where the fdr is extremely high.
 
@@
 
-\begin{figure}
-\caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
-multiple-locus method
-Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
-for a false discovery rate of 5\% (Eagle$^{optimal}$). 
-The number of SNP-trait associations found are reported.
-The bar plots with boxes around them are for those traits where using Eagle (default and optimal)   resulted in more findings than with single-locus analysis.}
-\label{figmouse}
-\begin{center}
-%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
-\includegraphics[width=15cm, height=15cm]{mouseresults.eps}
-\end{center}
+{\em Figure \ref{figmouse} goes here}
 
-\end{figure}
+\section{Discussion}
+Eagle is a new linear mixed model based method (and R package) for multiple-locus association mapping. It advances the state of association mapping in several ways. 
+First, its computational footprint is much smaller than other multiple-locus implementations. Eagle makes multiple-locus analysis 
+practical, even when the datasets are large. Second, the results from
+ Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
+each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
+multiple testing issues and the need for significance thresholds. 
+ As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
+Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
+with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.
+
+Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
+choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
+SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
+Permutation and its variants \cite{ browning2008presto,pahl2010permory} are also equally valid empirical approaches.
+Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
+data. Also, when analysing the (sub)samples, it was not necessary to calculate the entire solution path for a method. 
+ Instead,  analyses are 
+performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
+implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
+The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
+r2VIM could be fine-tuned for a better balance of computational and statistical performance. 
+However, we would like to think that the parameter settings we used are sensible 
+since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.
+
+Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here, although in different ways. 
+Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
+is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
+including the extra random effects, is 
+captured in this first-stage model.
+In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis.
+Even though this is a well accepted practice, it is approximate \cite{gogel2018comparison}.
+A better solution is to fit a single model to the data.
+Although not specifically designed for association mapping,
+WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RWGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
+models,  run time and memory usage can become limiting factors for analysis.
+
+Over the coming years, computationally, the demand placed upon association mapping methods is going to increase. 
+High-throughput array-based technologies continue to decrease the cost of genotyping, permitting ever larger GWASs to be performed. 
+Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs \cite{gudbjartsson2015large, long2017whole}
+culminating in data on millions of SNPs. It is because of this growing demand that 
+we have purposely structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
+Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
+computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times.  
+We also have plans for Eagle to run on computer clusters. Structuring Eagle for larger-than-memory calculations was a 
+preemptive step in this direction. GWASs have changed significantly in the past decade but the size and complexity of GWASs is expected 
+to change even more in the coming decade.
 
@@
 
-\subsubsection{Reducing the dimension of the model}
+\subsubsection*{Reducing the dimension of the model}
 In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
@@
 
-\subsubsection{Multiple-locus methods}
+\subsubsection*{Multiple-locus methods}
 
@@
 
-\subsubsection{Single-locus methods}
+\subsubsection*{Single-locus methods}
 
@@
 
-\section{Discussion}
-Eagle is a new linear mixed model based method (and R package) for multiple-locus association mapping. It advances the state of association mapping in several ways. 
-First, its computational footprint is much smaller than other multiple-locus implementations. Eagle makes multiple-locus analysis 
-practical, even when the datasets are large. Second, the results from
- Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
-each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
-multiple testing issues and the need for significance thresholds. 
- As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
-Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
-with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.
-
-Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
-choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
-SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
-Permutation and its variants \cite{ browning2008presto,pahl2010permory} are also equally valid empirical approaches.
-Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
-data. Also, when analysing the (sub)samples, it was not necessary to calculate the entire solution path for a method. 
- Instead,  analyses are 
-performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
-implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
-The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
-r2VIM could be fine-tuned for a better balance of computational and statistical performance. 
-However, we would like to think that the parameter settings we used are sensible 
-since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.
-
-Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here, although in different ways. 
-Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
-is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
-including the extra random effects, is 
-captured in this first-stage model.
-In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis.
-Even though this is a well accepted practice, it is approximate \cite{gogel2018comparison}.
-A better solution is to fit a single model to the data.
-Although not specifically designed for association mapping,
-WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RWGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
-models,  run time and memory usage can become limiting factors for analysis.
-
-Over the coming years, computationally, the demand placed upon association mapping methods is going to increase. 
-High-throughput array-based technologies continue to decrease the cost of genotyping, permitting ever larger GWASs to be performed. 
-Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs \cite{gudbjartsson2015large, long2017whole}
-culminating in data on millions of SNPs. It is because of this growing demand that 
-we have purposely structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
-Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
-computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times.  
-We also have plans for Eagle to run on computer clusters. Structuring Eagle for larger-than-memory calculations was a 
-preemptive step in this direction. GWASs have changed significantly in the past decade but the size and complexity of GWASs is expected 
-to change even more in the coming decade.
-
 %%\bibliographystyle{naturemag}

### 2018-10-15 paper.tex
refinement

@@
 %% external tex files
-\externaldocument{figure1_runtimes}
-\externaldocument{figure2_powerFDRmultiple}
-\externaldocument{figure3_mouse}
+%%\externaldocument{figure1_runtimes}
+%%\externaldocument{figure2_powerFDRmultiple}
+%%\externaldocument{figure3_mouse}
 
@@
 \title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George$^1$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
+\author{Andrew W. George$^{1,*}$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
 
@@
  Eagle and MLMM were the easiest of the programs/packages to implement. 
- The only parameters requiring specification were the amount of available memory and number of CPU for 
+ The only parameters requiring specification were the amount of available memory and number of CPUs for 
  Eagle and the number of chunks for MLMM. 
@@
  Towards the final stages of writing this paper, 
- we gained access to a desktop computer with 14-core Xeon processors and 256 gigabytes of RAM. We reran Eagle on data from the largest
+ we gained access to a high-end sever with 14-core Xeon processors and 256 gigabytes of RAM. We reran Eagle on data from the largest
   scenario 10000 x 1.5M to measure the impact on run time. The median run time dropped by more than 70\% 
@@
 
- {\em Figure \ref{fig_time} goes around here}
+\begin{figure}
+\caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
+Eagle is compared against five other multiple-locus programs/packages (top left) and two single-locus programs (top right). 
+The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
+programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
+The median run times for the programs/packages across the scenarios are given in the table. The entries in a bold font 
+correspond to the lowest run-time for a scenario. 
+ FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.  
+ FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
+
+%% plot formed via PowerPoint - ~/AM-Paper/Figure_power.pptx
+%% script on mac
+%% ~/AM-Paper/AM+/Plots_for_Paper/time_and_memory_plot_for_paper.R
+\label{fig_time}
+\centering
+    \includegraphics[width=1.0\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
+\end{figure}
 
@@
 
-{\em Figure \ref{figpowermultiple} goes here}
+\begin{figure}
+\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
+scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
+scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
+upper limit of the x-axis is 0.1.}
+\label{figpowermultiple}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{power1main.eps}
+\end{center}
+
+\end{figure}
 
@@
 
-We repeated the single-locus analyses, as first performed  \cite{nicod2016genome}, except 
+We repeated the single-locus analyses, as first performed in \cite{nicod2016genome}, except 
 that we focused on autosomal SNP and our analyses were based on SNP genotype  rather than SNP dosage. 
@@
 
-{\em Figure \ref{figmouse} goes here}
+\begin{figure}
+\caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
+multiple-locus method
+Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
+for a false discovery rate of 5\% (Eagle$^{optimal}$). 
+The number of SNP-trait associations found are reported.
+The bar plots framed by boxes are for those traits where using Eagle (default and optimal)   resulted in more findings than with single-locus analysis.}
+\label{figmouse}
+\begin{center}
+%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
+\includegraphics[width=15cm, height=15cm]{mouseresults.eps}
+\end{center}
+
+\end{figure}
 
@@
 calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
-a link to our server for demonstrating Eagle on some test data are available on the Eagle website (http://eagle.r-forge.r-project.org).  
-Eagle is available for download from the CRAN website (https://cran.r-project.org).
+a link to our server for demonstrating Eagle on some test data are available on the Eagle website. Eagle is available for download from the CRAN website.
+
+\section*{Data Availability}
+The input files for Eagle containing the mouse GWAS data are available for download from 
+https://doi.org/10.25919/5bc08287717dd. The original data were obtained from the Heterogeneous Stock Mice website 
+http://wp.cs.ucl.ac.uk/outbredmice/heterogeneous-stock-mice/.
 
 %%\bibliographystyle{naturemag}
+\section*{References}
 \bibliography{biblibrary}
 
-%%\begin{addendum}
-%% \item Put acknowledgements here.
-%% \item[Competing Interests] The authors declare that they have no
-%%competing financial interests.
-%% \item[Correspondence] Correspondence and requests for materials
-%%should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
-%%\end{addendum}
+\begin{addendum}
+ \item We would like to thank the IM\&T Scientific Computing group at the Commonwealth Scientific and Industrial Research Organisation (CSIRO) for providing access and support to the CSIRO Accelerator Cluster upon which all computation was performed.
+\item[Competing Interests] The authors declare that they have no
+competing financial interests.
+ \item[Correspondence] Correspondence and requests for materials
+should be addressed to Andrew George ~(email: andrew.george@csiro.au).
+\item[URLs]  Eagle website http://eagle.r-forge.r-project.org;   
+CRAN website https://cran.r-project.org.
+
+\end{addendum}
 

### 2018-10-15 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -174,6 +174,10 @@ p <- p + theme(legend.position = c(0.82, 0.8),
  p <- p + theme(plot.margin = grid::unit(c(1,5,1,1), "lines"))
  
 
+## increase font size of labels
+ p <- p +theme(axis.title.y=element_text(size=16),
+          axis.title.x=element_text(size=16))
+ 
 p_time <- p
 
 p_time
@@ -308,11 +312,19 @@ p_time
   p <- p + theme(plot.margin = grid::unit(c(1,5,1,1), "lines"))
   
    
+  ## increase font size of labels
+  p <- p +theme(axis.title.y=element_text(size=16),
+                axis.title.x=element_text(size=16))
+  
+  
   p_abstime <- p
   
   p_abstime
 
-
+  
+stop()
+  
+  
 ##-----------------------------
 ## Multiple plots on single page
 ##------------------------------

### 2018-10-22 paper.tex
Nature Methods submission Oct

@@
 
-\title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
+\title{Eagle: Making multi-locus association mapping on a genome-wide scale routine}
 \author{Andrew W. George$^{1,*}$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
@@
 \begin{abstract}
-We present Eagle, a new method for multiple-locus association mapping. The
-motivation for developing Eagle was to make multiple locus association mapping "easy" and the method-of-choice. 
+We present Eagle, a new method for multi-locus association mapping. The
+motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice. 
 Eagle's strengths are that it a.
@@
 By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and 
-multiple-locus methods. We also analyse data from a published mouse study. Eagle found over two times more validated findings than the state-of-the-art 
+multi-locus methods. We also analyse data from a published mouse study. Eagle found over two times more validated findings than the state-of-the-art 
 single-locus method. 
@@
 
-It is somewhat surprising then that multiple-locus association mapping methods haven't attracted more attention. Methods based on 
+It is somewhat surprising then that multi-locus association mapping methods haven't attracted more attention. Methods based on 
 regularisation techniques, such as ridge regression \cite{shen2013novel}  and lasso \cite{rakitsch2013lasso}, measure all locus-trait associations simultaneously. 
@@
 Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings.
-A multiple-locus method that does show promise is the multiple-locus linear mixed model method \cite{segura2012efficient}.
-The best multiple-locus model is built with simple forward selection.
+A multi-locus method that does show promise is the multiple-locus linear mixed model method \cite{segura2012efficient}.
+The best multi-locus model is built with simple forward selection.
 Results are immediately interpretable but here, computation becomes challenging for large datasets.
 
-In this paper, we present our new multiple-locus method for genome-wide association mapping, which we are calling Eagle. 
+In this paper, we present our new multi-locus method for genome-wide association mapping, which we are calling Eagle. 
 Eagle combines the strength of regularisation techniques (being able to fit all SNP-trait associations jointly), with forward selection giving easy-to-interpret threshold-free results.
@@
 through a dimension reduction step.
-Our aim was to make multiple-locus association mapping on a genome-wide scale routine. To this end, we have implemented Eagle 
+Our aim was to make multi-locus association mapping on a genome-wide scale routine. To this end, we have implemented Eagle 
 within an R package of the same name. 
@@
 We chose methods that almost all had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs or packages.
-Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multiple-locus models.
+Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multi-locus models.
 Of the many ways of performing single-locus association mapping, we chose
 GEMMA \cite{zhou2012genome}  and FaST-LMM \cite{lippert2011fast} because of their popularity and computational speed. 
-For multiple-locus association mapping, we chose bigRR \cite{shen2013novel}, glmnet \cite{Friedman2010glmnet}, 
+For multi-locus association mapping, we chose bigRR \cite{shen2013novel}, glmnet \cite{Friedman2010glmnet}, 
 LMM-Lasso \cite{rakitsch2013lasso}, MLMM \cite{segura2012efficient} , and r2VIM \cite{szymczak2016r2vim}.  
-Each takes a different approach to multiple-locus association mapping. A summary of the key attributes of the different computer programs/packages 
+Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 
 is given in Supplementary Table \ref{suptabsummary} (See Methods for further details).
@@
 for many of the computer programs/packages was the limiting factor (See Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
-far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
+far the most memory efficient. Not surprisingly, the multi-locus programs were memory intensive. Most required in 
 excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
 Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
-Of the multiple-locus programs/packages, only Eagle,  
+Of the multi-locus programs/packages, only Eagle,  
 with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
@@
 In answer to our question of how does Eagle compare in terms of run time to competing implementations, 
-Eagle was significantly faster, sometimes by orders of magnitude,  than the other multiple-locus
+Eagle was significantly faster, sometimes by orders of magnitude,  than the other multi-locus
  implementations and is comparable to the single-locus implementations. For a simulation study with 150 individuals and 
@@
 \caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
-Eagle is compared against five other multiple-locus programs/packages (top left) and two single-locus programs (top right). 
-The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multiple-locus 
+Eagle is compared against five other multi-locus programs/packages (top left) and two single-locus programs (top right). 
+The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multi-locus 
 programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
@@
 
-The power and false discovery rates of Eagle and the other multiple-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
+The power and false discovery rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
 We restricted our attention to these scenarios because not all
-multiple-locus methods could cope with the size of data in the other scenarios. 
+multi-locus methods could cope with the size of data in the other scenarios. 
   Each plot contains 
@@
 \begin{figure}
-\caption{Power verse false discovery rates for Eagle and the multiple-locus methods. Plots for only those simulation 
-scenarios where all multiple-locus methods could be implemented are shown.  Eagle has the highest power across the four 
+\caption{Power verse false discovery rates for Eagle and the multi-locus methods. Plots for only those simulation 
+scenarios where all multi-locus methods could be implemented are shown.  Eagle has the highest power across the four 
 scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
@@
 In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
-well.  Of the multiple-locus methods, Eagle had the highest power
+well.  Of the multi-locus methods, Eagle had the highest power
 while keeping its false discovery rate low (Figure \ref{figpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
@@
 \caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
-multiple-locus method
+multi-locus method
 Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
@@
 \section{Discussion}
-Eagle is a new linear mixed model based method (and R package) for multiple-locus association mapping. It advances the state of association mapping in several ways. 
-First, its computational footprint is much smaller than other multiple-locus implementations. Eagle makes multiple-locus analysis 
+Eagle is a new linear mixed model based method (and R package) for multi-locus association mapping. It advances the state of association mapping in several ways. 
+First, its computational footprint is much smaller than other multi-locus implementations. Eagle makes multi-locus analysis 
 practical, even when the datasets are large. Second, the results from
@@
 
-Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
+Eagle outperformed the other multi-locus methods in our simulation study. However, we are cognisant of the fact  that we made several implementation 
 choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
@@
 
-\subsection{Eagle Approach for Multiple-locus Association Mapping}
+\subsection{Eagle Approach for Multi-locus Association Mapping}
 
-Eagle is a method for multiple-locus association mapping on a genome-wide scale.
+Eagle is a method for multi-locus association mapping on a genome-wide scale.
 It is based on linear mixed models.
-It differs from most other single- and multiple-locus association mapping methods.
+It differs from most other single- and multi-locus association mapping methods.
 Eagle treats association mapping as a model selection instead of variable selection problem.
@@
 
-\subsubsection*{Multiple-locus methods}
+\subsubsection*{Multi-locus methods}
 
-We compare the computational and statistical performance of Eagle against five multiple-locus methods. They are bigRR, LMM-Lasso, glmnet, 
+We compare the computational and statistical performance of Eagle against five multi-locus methods. They are bigRR, LMM-Lasso, glmnet, 
 MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 

### 2018-10-22 supmaterials.tex
Nature Methods submission Oct

@@
 
-\title{Eagle: Making multiple-locus association mapping on a genome-wide scale routine}
+\title{Eagle: Making multi-locus association mapping on a genome-wide scale routine}
 \author[1]{Andrew W. George}
@@
 The x-axis is on the log scale. GEMMA, a single-locus implementation, had the lowest memory usage. 
-Of the multiple-locus implementations, Eagle had the lowest memory usage. Also, it 
-was the only multiple-locus 
+Of the multi-locus implementations, Eagle had the lowest memory usage. Also, it 
+was the only multi-locus 
 implementation able to produce results for data under  scenario 10000 x 1.5M. This is due to its ability 

### 2018-11-13 paper.tex
Refining figures

@@
 
-%%      To Do
-%%    4.tidy up doc - reformat - remove comments make nice
-
 %% New commands ... 
@@
 \begin{abstract}
-We present Eagle, a new method for multi-locus association mapping. The
-motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice. 
+We present Eagle, a new method for multi-locus association mapping.
+The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice.
 Eagle's strengths are that it a.
@@
 has a computational footprint comparable to single-locus association mapping.
-By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and 
-multi-locus methods. We also analyse data from a published mouse study. Eagle found over two times more validated findings than the state-of-the-art 
-single-locus method. 
+By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods.
+We also analyse data from a published mouse study.
+Eagle found over two times more validated findings than the state-of-the-art single-locus method.
 Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
@@
 Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 
-is given in Supplementary Table \ref{suptabsummary} (See Methods for further details).
+is given in Supplementary Table \ref{suptabsummary} (see Methods for further details).
 
@@
 A large simulation study was performed where we sought to  answer two questions. 
-First, how does Eagle compare, in terms of run time and memory usage, to 
-competing implementations? Second, how well does Eagle find true associations (power) and avoid 
-false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
+First, how well does Eagle find true associations (power) and avoid 
+false associations (type 1 errors)? Second, how does Eagle compare, in terms of run time and memory usage, to 
+competing implementations? Data were generated under six different scenarios; a study of size 150 individuals 
 and 5,000 single SNPs (150 x 5K),  350 individuals and 400,000 SNPs (350 x 400K),  1,500 individuals and 
@@
 
+\subsection{Power and False Discovery Rates}
+
+Here, we answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.
+We do this by estimating the power and false discovery rates of Eagle and the other methods for the six scenarios.
+Since, for a replicate, we knew which SNPs were assigned additive effects, we knew the SNPs that were in true association with the trait.
+We will refer to these SNPs as being
+true SNPs.
+By knowing the true SNPS, we were able to assess the validity of a method's findings.
+A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a true SNP.
+When a replicate was analysed, we obtained an estimate of the power of the method by taking  the number of findings that were found to be  true and dividing by the
+number of true SNPs.
+We also obtained an estimate of a method's false discovery rate.
+It is the number of findings that were found to be false divided by the number of (true and false) findings found by the method.
+Both these estimates varied with replicate.
+The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates over the 100 replicates.
+
+The power and false discovery rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure \ref{supfigpowermultiple}.
+We restricted our attention to these scenarios because not all
+multi-locus methods could cope with the size of data in the other scenarios. 
+  Each plot contains 
+single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
+These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
+The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
+significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
+the false discovery rate as the significance threshold  is adjusted. 
+The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
+ Figure \ref{figpowersingle}.
+
+\begin{figure}
+\caption{Power verse false discovery rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
+FaST-LMM was run where all the SNP data are used 
+to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
+estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
+
+\label{figpowersingle}
+\begin{center}
+\includegraphics[width=15cm, height=20cm]{power2main.eps}
+\end{center}
+
+\end{figure}
+
+In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
+well.  Of the multi-locus methods, Eagle had the highest power
+while keeping its false discovery rate low (Supplementary Figure \ref{supfigpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
+that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
+association with a trait while avoiding false associations (Figure \ref{figpowersingle}).
+
 \subsection{Memory Usage and Run Times}
@@
 Memory usage
-for many of the computer programs/packages was the limiting factor (See Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
+for many of the computer programs/packages was the limiting factor (see Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
 far the most memory efficient. Not surprisingly, the multi-locus programs were memory intensive. Most required in 
@@
 \end{figure}
-
-\subsection{Power and False Discovery Rates}
-
-Here, we answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.
-We do this by estimating the power and false discovery rates of Eagle and the other methods for the six scenarios.
-Since, for a replicate, we knew which SNPs were assigned additive effects, we knew the SNPs that were in true association with the trait.
-We will refer to these SNPs as being
-true SNPs.
-By knowing the true SNPS, we were able to assess the validity of a method's findings.
-A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a true SNP.
-When a replicate was analysed, we obtained an estimate of the power of the method by taking  the number of findings that were found to be  true and dividing by the
-number of true SNPs.
-We also obtained an estimate of a method's false discovery rate.
-It is the number of findings that were found to be false divided by the number of (true and false) findings found by the method.
-Both these estimates varied with replicate.
-The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates over the 100 replicates.
-
-The power and false discovery rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Figure \ref{figpowermultiple}.
-We restricted our attention to these scenarios because not all
-multi-locus methods could cope with the size of data in the other scenarios. 
-  Each plot contains 
-single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
-These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
-The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
-significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
-the false discovery rate as the significance threshold  is adjusted. 
-The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
-Supplementary Figure \ref{supfigpowersingle}.
-
-\begin{figure}
-\caption{Power verse false discovery rates for Eagle and the multi-locus methods. Plots for only those simulation 
-scenarios where all multi-locus methods could be implemented are shown.  Eagle has the highest power across the four 
-scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
-upper limit of the x-axis is 0.1.}
-\label{figpowermultiple}
-\begin{center}
-\includegraphics[width=15cm, height=15cm]{power1main.eps}
-\end{center}
-
-\end{figure}
-
-In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
-well.  Of the multi-locus methods, Eagle had the highest power
-while keeping its false discovery rate low (Figure \ref{figpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
-that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
-association with a trait while avoiding false associations (Supplementary Figure \ref{supfigpowersingle}).
 

### 2018-11-13 supmaterials.tex
Refining figures

@@
 \begin{figure}
+\caption{Power verse false discovery rates for Eagle and the multi-locus methods. Plots for only those simulation 
+scenarios where all multi-locus methods could be implemented are shown.  Eagle has the highest power across the four 
+scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
+upper limit of the x-axis is 0.1.}
+\label{supfigpowermultiple}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{power1main.eps}
+\end{center}
+
+\end{figure}
+
+\begin{figure}
 \caption{Memory usage (in gigabytes) of Eagle and the other association mapping programs/packages across 
@@
 \end{center}
-\end{figure}
-
-\begin{figure}
-\caption{Power verse false discovery rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
-FaST-LMM was run where all the SNP data are used 
-to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
-estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
-
-\label{supfigpowersingle}
-\begin{center}
-\includegraphics[width=15cm, height=15cm]{power2main.eps}
-\end{center}
-
 \end{figure}

### 2018-11-13 AM+/Plots_for_Paper/time_and_mem_plot_for_paper.R

@@ -110,7 +110,8 @@ p  <- p + scale_x_continuous(trans="log10", limits=c(500000, 1e11),
 
 
 ##log tick marks
- p <- p +annotation_logticks(scale=TRUE, side="bl")
+ p <- p +annotation_logticks(scale=TRUE, side="bl", size=2, short=unit(0.3, "cm"),mid=unit(0.6, "cm"), 
+                             long=unit(0.9, "cm") )
 
 
 
@@ -126,8 +127,8 @@ p <- p  + ylab(bquote("Median of elapse times (in minutes)\n")) +
 
 
 ##  change x and y labels size and bold
-p <- p + theme(axis.title.x = element_text(angle=0, vjust=1, size=12)) 
-p <- p + theme(axis.title.y = element_text(angle=90, vjust=1, size=12))
+#p <- p + theme(axis.title.x = element_text(angle=0, vjust=1, size=12)) 
+#p <- p + theme(axis.title.y = element_text(angle=90, vjust=1, size=12))
 
 # alter x and y axis labels 
 p <- p + 
@@ -149,16 +150,16 @@ p <- p +  annotate("text", x=pps, y=rep(9000,length(pps)), label= c("150 x 5K" ,
                                                                     "",  
                                                                     "4000 x 1.5M",  
                                                                     ""),
-                   size=4)
+                   size=11)
 
 
-p <- p +  annotate("text", x=pps, y=rep(6000,length(pps)), label= c("" , 
+p <- p +  annotate("text", x=pps, y=rep(5000,length(pps)), label= c("" , 
                                                                     "1500 x 50K",
                                                                     "" ,  
                                                                     "2000 x 500K", 
                                                                     "",  
                                                                     "10000 x 1.5M"),
-                   size=4)
+                   size=11)
 
 
 
@@ -171,18 +172,37 @@ p <- p + theme(legend.position = c(0.82, 0.8),
  #p <- p + theme(axis.text.x=element_blank(),axis.title.x=element_blank())
  
  ## putting some space around the plot
- p <- p + theme(plot.margin = grid::unit(c(1,5,1,1), "lines"))
+ p <- p + theme(plot.margin = grid::unit(c(1,5,2,1), "lines"))
  
 
 ## increase font size of labels
- p <- p +theme(axis.title.y=element_text(size=16),
-          axis.title.x=element_text(size=16))
+ p <- p +theme(axis.title.y=element_text(size=38),
+          axis.title.x=element_text(size=38))
+ 
+## increase width of axis lines
[…]

### 2018-11-17 paper.tex
nice plots

@@
 \caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
-Eagle is compared against five other multi-locus programs/packages (top left) and two single-locus programs (top right). 
+Eagle is compared against five other multi-locus programs/packages (A) and two single-locus programs (B). 
 The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multi-locus 
 programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
-The median run times for the programs/packages across the scenarios are given in the table. The entries in a bold font 
+The median run times for the programs/packages across the scenarios are given in the table (C). The entries in a blue font 
 correspond to the lowest run-time for a scenario. 

### 2018-11-17 AM+/Plots_for_Paper/power_plots.R

@@ -186,17 +186,10 @@ library(RColorBrewer)
 #  theme(aspect.ratio = 1) # try with and without
 
 require(mgcv)
-p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_smooth(size=1, se=FALSE, method="loess")  +
-  geom_point(data=df2, aes(FDR, recall), size=1.5) +
+p <- ggplot(data=df1, aes(FDR, recall, color=method)) + geom_smooth(size=1.8, se=FALSE, method="loess")  +
+  geom_point(data=df2, aes(FDR, recall), size=4) +
   facet_wrap(~fam, ncol=2) + 
   theme(aspect.ratio = 1) # try with and without
-
-
-ggplot(data=df1, aes(FDR, recall, color=method)) + geom_line(size=1)  +
-  geom_point(data=df2, aes(FDR, recall), size=1.5) +
-  facet_wrap(~fam, ncol=2) + 
-  theme(aspect.ratio = 1) # try with and without
-
 
 
 p <- p + scale_color_manual(
@@ -209,37 +202,32 @@ guide = guide_legend(override.aes = list(
   linetype = c("blank", "blank", "solid", "solid", "solid", "solid"),
   shape=c(16, 16, NA, NA, NA, NA))))
 
-
-
-
-
-
                   
 ## set theme
 p <- p + theme_hc()
 
 ## increase spacing between facet plots
-p <- p + theme(panel.spacing = unit(3, "lines"))
+p <- p + theme(panel.spacing = unit(2, "lines"))
 
 ## specify xlab and ylab
 p <- p  + ylab(bquote("Power")) + 
-  xlab(bquote('False discovery rate'))
+  xlab(bquote('False Discovery Rate'))
 
 
 ##  change x and y labels size and bold
-p <- p + theme(axis.title.x = element_text(angle=0, vjust=1, size=14)) 
-p <- p + theme(axis.title.y = element_text(angle=90, vjust=1, size=14))
+p <- p + theme(axis.title.x = element_text(angle=0, vjust=1, size=24)) 
+p <- p + theme(axis.title.y = element_text(angle=90, vjust=1, size=24))
 
 # alter x and y axis labels 
 p <- p + 
-  theme(axis.text.x = element_text(size=11,  angle=0)) +
-  theme(axis.text.y=element_text(size=11, hjust=0.5)) +
-  theme(strip.text = element_text(size=14))
[…]

### 2018-11-29 paper.tex
working on refinining methods - mouse

@@
 
-We repeated the single-locus analyses, as first performed in \cite{nicod2016genome}, except 
-that we focused on autosomal SNP and our analyses were based on SNP genotype  rather than SNP dosage. 
+We repeated the single-locus analyses, as first performed \cite{nicod2016genome}, except 
+that we focused on autosomal SNP and our analyses were based on SNP genotypes  rather than SNP dosages. 
 In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation. We followed the 
-same empirical procedure but had to increase the number of permutations from 100 to 500.
+same empirical procedure but increased the number of permutations from 100 to 250.
 
-Eagle was run in two ways; under its default settings (Eagle$^{default}$) and where we specified the regularisation parameter for model selection (Eagle$^{optimal}$ ).
-Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}.
-  The conservativeness of the extBIC can be adjusted by a single regularisation parameter that ranges from zero to one.
-In the simulation study, this parameter was set to one, its most conservative and default setting.
-However, there is also opportunity to set the parameter to a value less than one.
-This increases power but also increases the false discovery rate.
-For each trait, we used permutation to set the regularisation parameter to give a false discovery rate of 5\% .
+Eagle was run in three ways.  Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}. 
+  The conservativeness of the extBIC can be adjusted by a single regularisation parameter  $\gamma$ that ranges from zero to one. 
+In the simulation study, this parameter was set to one, its most conservative and default setting. The mouse data were also analysed 
+under this setting (Eagle$^{default}$).
+An alternate  \cite{chen2008extended} , less conservative way of setting $\gamma$ is to  let $\gamma = 1 - \frac{1}{(2\kappa)}$ with $\kappa=\frac{log(L)}{log(n_g)}$ where
+ $L$ is the number of loci that span the genome, and $n_g$ is the number of individuals/groups/lines/strains in the study
+  (Eagle$^{alt}$).
+However, our preferred way is to set the $\gamma$ parameter for each trait via permutation (Eagle$^{optimal}$).
+We used 100 permutations to set $\gamma$ to give a false  positive rate of 5\%.
+This only took six times as long as a single analysis of the data.
+This is because the marker data need only be read  once,
+ and only the trait data changes across permutations leading to other computational efficiencies.  
+  This permutation method has been implemented within the
+ Eagle package.
 
 The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}. The mouse study recorded
-measurements on 200 traits. When these traits were first analysed in the original study, findings for 45 of these traits were able to be 
-corroborated by prior published evidence. We focused our analyses here on these same 45 traits. For 39 traits, SNP-trait associations 
-were found. For the other six, neither FaST-LMM nor Eagle found any associations. 
-Each plot in Figure  \ref{figmouse} contains the number of SNP-trait associations that were found and in agreement with the original findings. 
-Neither method found SNPs that had not been first identified in the original mouse study so neither method found false positives. 
-As we saw in the simulation study, there was a notable difference between the two methods capacity to discover SNP-trait associations.
-Eagle$^{default}$, under its default settings, for two traits found fewer findings than FaST-LMM, for eight traits found the same number
-of findings as FaST-LMM, and for 28 traits found more findings. Eagle$^{optimal}$, 
-with its regularisation parameter fine tuned to the trait data, for seven traits found the same number of findings as FaST-LMM and for 32 traits found more findings.
-Overall, FaST-LMM, Eagle$^{default}$, and Eagle$^{optimal}$ found 26, 65, and 95, SNP-trait findings, respectively.
-Eagle$^{default}$ and Eagle$^{optimal}$ found two-and-a-half times and over three-and-a-half times, respectively, more SNP-trait 
-associations than what is the established way of analysing these data. Furthermore, these extra results found by Eagle were not 
-false positives but could be confirmed from the original study.
+measurements on 200 traits. Of these, in the original study, 45 were able to have their findings  corroborated by previously published work. We focused 
+our analyses here on these same 45 traits.
+Overall,  FaST-LMM, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 50, 37, 67, and 106, SNP-trait findings, respectively, across 39 traits.
+No associations were found by FaST-LMM and Eagle for the other six traits.
+Eagle$^{alt}$ and Eagle$^{optimal}$ also found SNP-trait associations not found in the original study. This is despite their analyses being based 
+on the SNP genotype data and the original study being based on SNP dosage data. Eagle$^{alt}$ found two  and 
+Eagle$^{optimal}$ found seven new findings (Supplementary Table \ref{suptabnew}).  These new findings all involved SNPs whose association had been confirmed for other related traits in the original 
+study.
+
+It is worth noting that the mouse data is similar in size to data generated under the simulation scenario 2000x500K. In the simulation study, Eagle had 
+far greater power than the single-locus methods.
+Yet here, Eagle$^{default}$ found only 37 associations compared to 50 associations found by FaST-LMM.
+The reason for the difference is due to Eagle, when run under its default setting, is far more conservative for the real data than the simulation data.
+Also, here FaST-LMM has been fine-tuned, through permutation,  to have a false discovery rate of 5\%.
+A fairer comparison is to compare Eagle$^{optimal}$ to FaST-LMM since they have similar false discovery rates.
+Here, Eagle clearly outperforms FaST-LMM.
 
@@
 multi-locus method
-Eagle. Eagle was run under two settings; its default setting (Eagle$^{default}$) and where the model selection criterion had been optimised 
-for a false discovery rate of 5\% (Eagle$^{optimal}$). 
-The number of SNP-trait associations found are reported.
-The bar plots framed by boxes are for those traits where using Eagle (default and optimal)   resulted in more findings than with single-locus analysis.}
+Eagle. Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on 
+the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate 
+of 5\% (Eagle$^{optimal}$).  The number of SNP-trait associations found are reported in the cells. }
 \label{figmouse}

### 2018-11-29 supmaterials.tex
working on refinining methods - mouse

@@
 
+\begin{table}[h]
+\scriptsize
+\caption{New findings for the analysis of the mouse data. Given is the trait for which the new SNP-trait association was found (Trait), 
+the chromosome (Chrm) and map position in megabases (Position) for the SNP, and the list of other traits for which this SNP is known to be 
+in association (Known Associations). }
+\label{suptabnew}
+\vspace{0.5cm}
+\begin{tabular}{llccp{6cm}} \hline
+Method           &  Trait  & Chrm  &  Position (Mb)  & Known Associations \\  \hline  
+Eagle$^{alt}$  & Bioch.HDL   & 6   & 17.53         & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+     SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin \\ 
+                       & Bioch.LDL   & 6    &  17.53        & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g SPPI.ln.pa Bioch.Calcium  Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin \\  [0.45cm]
+ Eagle$^{optimal}$  &        Bioch.Tot.Cholesterol  & 4      &  134.58  & FACS.CD3posCD44posCD4CD8Ratio  \\
+                        & Bioch.HDL                  & 4      &  134.58  & FACS.CD3posCD44posCD4CD8Ratio   \\
+                        &                                    & 6      &    17.53   & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+     SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin  \\
+                        & BMC.Mean                 & 15    &     86.57  & BMC.Median \\
+                        & BMC.Mean.N             &  5     &     24.64   &  BMC.Mean BMC.StdDev BMC.StdDev.N  BMC.Max.N  BMC.Median BMC.Kurt
+                        BMC.Max BMC.Kurt.N  \\
+                        & BMC.Max.N               & 15    &      86.57  & BMC.Median \\
+                        & Bioch.LDL                  & 6      &   17.53   & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+     SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin  \\  \hline
+
+\end{tabular}
+\end{table}
+
 %% This table appears in Figure 1 in paper. It's included here in case I need to regenerate the table. 

### 2018-11-30 paper.tex
getting close to final product

@@
  We chose to focus on data from a large outbred mouse study \cite{nicod2016genome}.
-This study was a little unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
+This study was unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
 Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \cite{zheng2011comparison}.
@@
 We then analysed the data with Eagle.
-Due to Eagles increased power, we found SNP-trait associations not found with  FaST-LMM.
+Due to Eagle's increased power, we found SNP-trait associations not found with  FaST-LMM.
 However, we were
@@
 
-We repeated the single-locus analyses, as first performed \cite{nicod2016genome}, except 
-that we focused on autosomal SNP and our analyses were based on SNP genotypes  rather than SNP dosages. 
+We repeated the single-locus analyses as first performed \cite{nicod2016genome}, except 
+that we focused on autosomal SNPs and our analyses were based on SNP genotypes  rather than SNP dosages. 
 In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation. We followed the 
-same empirical procedure but increased the number of permutations from 100 to 250.
+same empirical procedure but increased the number of permutations from 100 to 500 for more accurate thresholds.
 
-Eagle was run in three ways.  Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}. 
+We ran Eagle in three ways. 
+Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}. 
   The conservativeness of the extBIC can be adjusted by a single regularisation parameter  $\gamma$ that ranges from zero to one. 
@@
 
-It is worth noting that the mouse data is similar in size to data generated under the simulation scenario 2000x500K. In the simulation study, Eagle had 
-far greater power than the single-locus methods.
-Yet here, Eagle$^{default}$ found only 37 associations compared to 50 associations found by FaST-LMM.
-The reason for the difference is due to Eagle, when run under its default setting, is far more conservative for the real data than the simulation data.
-Also, here FaST-LMM has been fine-tuned, through permutation,  to have a false discovery rate of 5\%.
-A fairer comparison is to compare Eagle$^{optimal}$ to FaST-LMM since they have similar false discovery rates.
-Here, Eagle clearly outperforms FaST-LMM.
+In the simulation study, Eagle outperforms single-locus association mapping. Here, Eagle$^{default}$, where $\gamma=1$, finds less associations 
+than FaST-LMM.
+Why the discrepancy in performance?
+The answer lies in the conservativeness of Eagle.
+With the added genetic complexity implicit within the mouse data, Eagle is more conservative when $\gamma$ is set to one than in the simulation study.
+However, the relative results of the simulation study remain true.
+For similar false discover rates, Eagle is superior to single-locus association mapping.
+As a case in point, here FaST-LMM found 50 SNP-trait associations with a false discovery rate of 5\%.
+Eagle, with the same false discovery rate (Eagle$^{optimal}$) found 106 SNP-trait associations, more
+than a 100\% increase in findings.
 
@@
 we have purposely structured the Eagle package for continued development. We are already experimenting with a GPU-based version of Eagle. 
-Early results suggest that for small to moderate sized datasets (<10,000 samples), there is little improvement in performance over CPU-based 
+Early results suggest that for small to moderate sized datasets ($<$10,000 samples), there is little improvement in performance over CPU-based 
 computation.  However, for larger study sizes, we are seeing up to a 40\% decrease in run times.  
@@
 
-\subsubsection*{Reducing the dimension of the model}
+\subsubsection {Reducing the dimension of the model:}
+
 In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
@@
 
-\subsubsection*{Multi-locus methods}
-
-We compare the computational and statistical performance of Eagle against five multi-locus methods. They are bigRR, LMM-Lasso, glmnet, 
+ \subsubsection{ Multi-locus methods:} We compare the computational and statistical performance of Eagle against five multi-locus methods. They are bigRR, LMM-Lasso, glmnet, 
 MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
@@
 
-\subsubsection*{Single-locus methods}
-
-We also compare the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
+\subsubsection{Single-locus methods:} We also compare the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
 Both are based on linear mixed models.
@@
 \subsection{Stability Selection}
+
  Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of the significance of effects in regression models where 

### 2018-11-30 supmaterials.tex
getting close to final product

--- 2018-11-13 supmaterials.tex
@@
 
+\begin{table}[h]
+\scriptsize
+\caption{New SNP-trait associations found from the analysis of the mouse data with Eagle.  Due to many of the traits in the study 
+being correlated, genetically,  these new findings could be confirmed through their prior discovery with other traits in the original 
+mouse study (Prior Associations). }
+\label{suptabnew}
+\vspace{0.5cm}
+\begin{tabular}{lcccp{6cm}}  \hline
+     & \multicolumn{3}{c}{New SNP-trait Association}  &  \\  \cline{2-4}
+Method           &  Trait  & Chrm  &  Position (Mb)  & \hfil Prior Associations  \\  \hline  
+Eagle$^{alt}$  & Bioch.HDL   & 6   & 17.53         & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+     SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin \\ 
+                       & Bioch.LDL   & 6    &  17.53        & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g SPPI.ln.pa Bioch.Calcium  Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin \\
+ Eagle$^{optimal}$  &        Bioch.Tot.Cholesterol  & 4      &  134.58  & FACS.CD3posCD44posCD4CD8Ratio  \\
+                        & Bioch.HDL                  & 4      &  134.58  & FACS.CD3posCD44posCD4CD8Ratio   \\
+                        &                                    & 6      &    17.53   & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+     SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin  \\
+                        & BMC.Mean                 & 15    &     86.57  & BMC.Median \\
+                        & BMC.Mean.N             &  5     &     24.64   &  BMC.Mean BMC.StdDev BMC.StdDev.N  BMC.Max.N  BMC.Median BMC.Kurt
+                        BMC.Max BMC.Kurt.N  \\
+                        & BMC.Max.N               & 15    &      86.57  & BMC.Median \\
+                        & Bioch.LDL                  & 6      &   17.53   & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+     SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin  \\  \hline
+
+\end{tabular}
+\end{table}
+
 %% This table appears in Figure 1 in paper. It's included here in case I need to regenerate the table. 

### 2019-01-08 AM+/Plots_for_Paper/power_plots.R
08Jan18 Nature Communications Submission Version 1

@@ -252,6 +252,7 @@ p
 #dev.off()
 
 
+
 postscript("~/Papers/AM-Paper/power1main.eps", width=10, height=10, fonts=c("sans"),
            horizontal=FALSE)
 # square plot
@@ -393,9 +394,9 @@ p <- p +
   theme(strip.text = element_text(size=24))
 
 ## increase font of lengend + remove legend title
-p <- p +  theme(legend.text=element_text(size=22))
+p <- p +  theme(legend.text=element_text(size=18))
 p <- p +  theme(legend.title=element_blank())
-p <- p+ theme(legend.key.width=grid:::unit(2.5,"cm"))
+p <- p+ theme(legend.key.width=grid:::unit(1,"cm"))
 
 #p + theme_base()
 #p + theme_economist_white()
@@ -424,13 +425,13 @@ p
 
 
 
-
-postscript("~/Papers/AM-Paper/power2main.eps", width=15, height=20,  fonts=c("sans"),
+postscript("~/Papers/AM-Paper/power2main.eps", width=20, height=20,  fonts=c("sans"),
            horizontal=FALSE)
 # square plot
 print(p)
 dev.off()
 
+stop()
 
 
 ##----------- Subplots -- insets  ----------  ##

### 2019-02-08 bioinformaticspaper.tex
Just submitted to Bioinformatics

--- 2018-11-30 paper.tex
@@
-\documentclass{nature}
+%%\documentclass{nature}
+\documentclass{article}
 \usepackage{xr}
-\bibliographystyle{naturemag}
-\renewcommand{\baselinestretch}{1.3} 
+\bibliographystyle{natbib}
+\renewcommand{\baselinestretch}{1.4} 
 \usepackage{graphicx}
@@
 \usepackage{lineno}
+\usepackage{natbib}
+\usepackage{setspace}
+\usepackage{authblk}
+
 \linenumbers
@@
 
-\title{Eagle: Making multi-locus association mapping on a genome-wide scale routine}
-\author{Andrew W. George$^{1,*}$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
+\title{Eagle: multi-locus association mapping on a genome-wide scale made routine}
+\author[1]{Andrew W. George}
+\author[2]{Arunas Verbyla}
+\author[3]{Joshua Bowden}
+\affil[1]{Data61, CSIRO, Brisbane, 4102, Australia.}
+\affil[2]{Data61, CSIRO, Atherton, 4883, Australia.}
+\affil[3]{IM \&T, CSIRO, Brisbane, 4067, Australia}
+\date{}
 
@@
 
-\begin{affiliations}
-\item Data61, CSIRO, Australia.
-\item IM \&T, CSIRO, Australia.
-\end{affiliations}
-
-\begin{abstract}
-We present Eagle, a new method for multi-locus association mapping.
+\begin{abstract} \noindent
+\setstretch{1.6}{
+{\bf Motivation:} We present Eagle, a new method for multi-locus association mapping.
 The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice.
@@
 has a computational footprint comparable to single-locus association mapping.
-By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods.
+\\
+{\bf Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods.
 We also analyse data from a published mouse study.
-Eagle found over two times more validated findings than the state-of-the-art single-locus method.
-Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
+Eagle found over 100\% more validated findings than the state-of-the-art single-locus method.
+\\
+{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
+It is freely available via the CRAN website  at https://cran.r-project.org.
+\\
+{\bf Contact:} andrew.george@csiro.au
+}
 \end{abstract}
 
+\section{Introduction}
 Over the past decade,  genome-wide association studies (GWASs) have changed considerably in both their analysis and design. Early studies
@@
 linear regression. These designs though had a tendency to yield spurious findings if there was unrecognised population stratification 
-\cite{cardon2003population}.
-This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \cite{spielman1996tdt}.
-Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification \cite{price2010new}.
+\citep{cardon2003population}.
+This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \citep{spielman1996tdt}.
+Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification \citep{price2010new}.
 This has meant that data can be collected from general populations, even if these populations are highly structured.
-Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine \cite{purcell2007plink, bradbury2007tassel}.
+Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine \citep{purcell2007plink, bradbury2007tassel}.
 
@@
 This is known as the multiple testing problem.
-Many different solutions have been offered \cite{storey2003statistical, li2005adjusting, de2005efficiency}.
+Many different solutions have been offered \citep{storey2003statistical, li2005adjusting, de2005efficiency}.
 Yet, there is still no well accepted way of correcting for multiple testing in the context of genome-wide association mapping.
@@
 It is somewhat surprising then that multi-locus association mapping methods haven't attracted more attention. Methods based on 
-regularisation techniques, such as ridge regression \cite{shen2013novel}  and lasso \cite{rakitsch2013lasso}, measure all locus-trait associations simultaneously. 
+regularisation techniques, such as ridge regression \citep{shen2013novel}  and lasso \citep{rakitsch2013lasso}, measure all locus-trait associations simultaneously. 
 Here, multiple testing is not an issue.
@@
 The strength of association is not measured by a $p$ value but by the size of the regression coefficient for the SNP in the model.
-More recently, associations have started to be mapped with random forests \cite{szymczak2016r2vim}.
+More recently, associations have started to be mapped with random forests \citep{szymczak2016r2vim}.
 Similar to regularisation techniques though, it is not clear how to infer genomic regions of interest from their findings.
-A multi-locus method that does show promise is the multiple-locus linear mixed model method \cite{segura2012efficient}.
+A multi-locus method that does show promise is the multiple-locus linear mixed model method \citep{segura2012efficient}.
 The best multi-locus model is built with simple forward selection.
@@
 
+\section{Methods}
+
+\subsection{Mouse Data}
+
+The data were obtained from a large genome-wide association study that was performed in outbred mice \citep{nicod2016genome}. 
+Phenotypic and genotypic data were available on 1,887 adult mice. 
+The phenotypic data included raw and adjusted (for fixed effects) measurements from 200 behavioural, tissue, and physiological traits.  
+Of these traits, 
+43 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
+43 traits that were the focus of our real data analyses. As in the original study  \citep{nicod2016genome}, our analyses 
+were based on the adjusted traits.
+Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
+form of marker dosages (expected allele counts that ranged from zero to one). All missing data had been imputed. 
+We converted the dosages into discrete genotypes 
+by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively. 
+We focused our analyses on the autosomal SNPs.
+
+\subsection{Eagle Approach for Multi-locus Association Mapping}
+
+Eagle is a method for multi-locus association mapping on a genome-wide scale.
+It is based on linear mixed models.
+It differs from most other single- and multi-locus association mapping methods.
+Eagle treats association mapping as a model selection instead of variable selection problem.
+Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
+Eagle also 
+reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest to the genes influencing a trait. 
+The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
+quantitative trait loci in experimental crosses \citep{verbyla2007analysis}.
+
+Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of $s$ ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
+selected in the $k$th iteration of the model building process. Suppose three iterations  $(s=3)$
+have been performed and say the 500023rd, 15th, and 420th
+SNP were selected. Then $S=\{500023, 15, 420\}$. 
+Let $\by^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
+Let $\bM^{(n_g \; \times \; L)} = [\bmm_1 \bmm_2 \ldots \bmm_L]$ be a matrix containing the genotype data which have been collected 
+from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
+may be taken of the same group/line/strain. 
+ It is common for the columns of $\bM$ to be in map order but this is not a requirement. 
+The vector $\bmm_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
+The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.
+
+The specifics of the Eagle method are as follows. 
+Eagle builds the "best" model iteratively, via forward selection. 
+Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
+associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
+To perform the $s+1$th  iteration, we first fit the current model to the data. 
+The (current) model is of the form 
+\begin{equation}
+\label{eq1}
+\by = \bX \btau + \bZ \bu_g + \be
+\end{equation}
+where 
+$\bX^{(n \; \times \; p)}$ and $\bZ^{( n \; \times \; n_g)}$ are known design matrices with $\bX$ being of full rank and $\bZ$ 
+containing zeros and ones that assign the appropriate genetic effect to each measurement. 
+The vector 
+$\btau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
+$\bu_g^{(n_g \; \times \; 1)}$ contains the 
+genetic effects. The vector of residuals is 
+$\be^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(\bzero, \sigma^2_e \bI^{(n \; \times \; n)})$. 
+So far,  this model differs little from standard linear mixed models for association mapping \citep{yu2006unified,zhao2007arabidopsis} 
+However, 
+it is how we specify $\bu_g$ that distinguishes our model from the others.
+
+The genetic effects $\bu_g$ are modelled as 
+\begin{equation}
+\label{eq2}
+\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} +\bM_{-S} \ba_{-S}
+\end{equation}
+where $\bmm_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
+$a_{S_k}$ is the additive effect of the $k$th selected SNP, $\bM_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
+with the data for the SNP in $S$ removed,  and $\ba_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
+$\ba_{-S} \sim N(\bzero, \sigma_a^2 \bI^{(L-s \; \times \;  L-s)})$. 
+The terms in the summation on the left hand side are fixed effects.   They account 
+for the additive effects of those SNPs that have been found to be in association with the trait. The other term is a random effect. 
+It accounts for the joint effect of the yet-to-be-identified SNP that are in association with the trait. 
+This is a simple genetic model but it 
+is effective for discovering SNP-trait associations.
+
+Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
+can be computationally demanding. However, our model only contains a single random effect ($\ba_{-S}$). Here, highly efficient single-dimension 
+optimisation via spectral decomposition is possible \citep{kang2008efficient}.
+
+Third, we identify the $(s+1)$th SNP that is in strongest association with the trait, based on the maximum score statistic
+$t_j^2 = \frac{ \widetilde{a} _j^2}{\textrm{var}(\widetilde{a}_j)}$ where $\widetilde{a}_j$ is the best linear unbiased predictor (BLUP), 
+and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
+identify a SNP based on its (random) effect size and accuracy, but is justifiable, theoretically \citep{verbyla2012rwgaim}.
+
+Fourth, we determine the importance of the $(s+1)$th selected SNP via a model selection strategy  \citep{verbyla2007analysis}. 
+We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected SNP.  We then fit this new model to the data
+via maximum likelihood and calculate its extended Bayesian information criteria (extBIC) \citep{chen2008extended}.  The 
+extBIC is a model selection measure that takes into account the number of unknown parameters and the complexity 
+of the model space.  It is well suited to the model selection problem in genome-wide association studies \citep{chen2008extended}. 
+It is different to the model selection measure used in  \citep{verbyla2007analysis}.
+If this new model has a larger extBIC than the current model, then the $s+1$th selected SNP is added to 
+the current model and the above process is repeated. If this new model has a smaller extBIC than the current model, then the 
+model building process is complete. The set of SNP in strongest association with the trait is the $s$ SNPs previously identified.
+
+\subsubsection {Reducing the dimension of the model:}
+
+In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
+The vector $\ba_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
+$L$, the number of SNPs, can be extremely large.  An alternative model is given by 
+Verbyla \citep{verbyla2012rwgaim,verbyla2014whole}. 
+They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
+\begin{equation}
+\label{eq3}
+\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} + (\bM_{-S} \bM_{-S}^T)^{1/2} \ba^*_{-S}
+\end{equation}
+where $\ba^* \sim N(\bzero, \sigma_a^2 \bI^{(n_g \; \times \;  n_g)})$, and 
+$(\bM_{-S} \bM_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \citep{golub2012matrix}.  
+Although it may not be obvious, the two models are equivalent, 
+having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
+model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
+effects needing estimating.
+
+Verbyla \citep{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\bwidetildea$ from estimates from model  (\ref{eq3})  with 
+\begin{equation}
+\bwidetildea = \left [ \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \right ] \bwidetildea^*
+\end{equation}
+where its variance matrix is
+\begin{equation}
+\label{eq4}
+\textrm{var}(\bwidetildea) = \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \textrm{var}(\bwidetildea^*) (\bM_{-S} \bM_{-S}^T)^{-1/2} \bM_{-S}
+\end{equation}
+These values are needed in order to calculate the score statistic $t_j^2$ for identifying the SNP in strongest association with the trait. 
+Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation 
+of (\ref{eq4}).
+
+%%\iffalse
+
+\subsection{Comparison Methods}
+
+ \subsubsection{ Multi-locus methods:} We compare the computational and statistical performance of Eagle against five multi-locus methods. They are bigRR, LMM-Lasso, glmnet, 
+MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
+BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
+methods. BigRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
+Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
+They allow the strength of association between all the SNPs and trait to be measured within a single model, simultaneously. 
+A limitation of these methods though is that the statistical significance of the SNP effects cannot be easily determined. 
+ Due to the adaptive nature of the estimation procedures, to do this 
+analytically is challenging and is an area of active research \citep{lockhart2014significance}. Instead, we calculate significance empirically via 
+stability selection (see below).
+
+MLMM is closest in philosophy to Eagle. It too is based on building the best model via forward selection, 
+within  a linear mixed model framework, using the 
+extBIC. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
+model differently to Eagle.  Eagle uses a score statistic in which to identify SNPs to enter the model.
+MLMM uses the statistical significance of the SNPs. This involves fitting a separate linear mixed model for each unidentified SNP. 
+Both are R packages but there is a significant difference in computational performance (see Results).
+
+R2VIM differs to the other four methods in that it is a non-parametric model-free approach. It implements  
+random forests but where multiple parallel runs are performed. Each run leads to 
+different random forests being created.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
+importance score by the minimum importance score observed across all the SNPs within a run. 
+Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
+Unfortunately, the relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
+but the computational cost is high, restricting the size of data that can be analysed.
+
+\subsubsection{Single-locus methods:} We also compare the performance of Eagle against two single-locus methods, GEMMA \citep{zhou2012genome} and FaST-LMM \citep{lippert2011fast}.
+Both are based on linear mixed models.
+The models have a single fixed effect for the SNP,  other fixed effects,
+a single random effect to account for familial relatedness (or polygenic background), and an error. The significance of the SNP effect in the model 
+is a measure of the strength of association.  They are of the same computational complexity \citep{zhou2012genome}, and produce exact results.  
+Both perform a single spectral decomposition of the relationship (or similarity) matrix $K$, use  an eigenvector matrix to rotate the data, 
+and reformulate the (residual) log likelihood for easier computation. They do differ in their estimation procedure. GEMMA implements Newton-Raphson. 
+FaST-LMM implements Brent's algorithm. Newton-Raphson is more complicated but has better convergence properties than Brent's algorithm. 
+Both methods are state-of-the-art and have been implemented in highly efficient computer programs.
+
+\subsection{Stability Selection}
+
+ Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of the significance of effects in regression models where 
+the number of parameters ($p$) can be much larger than the number of samples ($n$).
+We chose stability selection over permutation and other sampling procedures because of its low computational cost.
+We employe stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
+
+To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, we begin
+ by performing a preliminary analysis of a replicate where  we tune the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
+This tuning of the regularisation parameter is done for data under each simulation scenario but it is not necessary to do it for every replicate within a scenario.
+It is sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
+Here, we have the luxury of knowing the genetic conditions 
+under which the simulated trait data are generated. We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
+ However, stability selection estimates are robust  \citep{meinshausen2010stability} 
+ and the regularisation parameter can be tuned to any reasonable number of 
+ non-zero effects. We sample repeatedly, without replacement, from the replicate. We draw 100 replicate subsets of size $n/2$.  
+Each replicate subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
+A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
+number of times the SNP have a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
+
+For bigRR, we modify our stability selection procedure slightly. 
+There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
+As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
+SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
+A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
+number of times the SNP is recorded divided by the number of replicates.
+
+\subsection{Generation of Simulation Data}
+
+The data are generated via data perturbation \citep{zhao2007arabidopsis}. 
+Data perturbation amalgamates real with simulated data to generate replicates. 
+It is a way of introducing greater realism into a simulation study. 
+ Here, the genotype data are real, the 
+quantitative trait data are simulated. 
+ The SNP genotypes are drawn, 
+according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \citep{10002010map}. 
+Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.
+
+To generate the trait data $\by$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
+mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
+are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
+generate a $\bm{g}^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
+Fifth, $\bm{e}^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
+the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $\by =  \bm{g} + \bm{e}$.  
+In forming $\by$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
+not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
+is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.
+
+\subsection{Implementation}
+
+Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
+via C++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
+mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
+calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
+a link to our server for demonstrating Eagle on some test data are available on the Eagle website. Eagle is available for download from the CRAN website.
+
 \section{Results}
@@
 Of the many ways of performing single-locus association mapping, we chose
-GEMMA \cite{zhou2012genome}  and FaST-LMM \cite{lippert2011fast} because of their popularity and computational speed. 
-For multi-locus association mapping, we chose bigRR \cite{shen2013novel}, glmnet \cite{Friedman2010glmnet}, 
-LMM-Lasso \cite{rakitsch2013lasso}, MLMM \cite{segura2012efficient} , and r2VIM \cite{szymczak2016r2vim}.  
+GEMMA \citep{zhou2012genome}  and FaST-LMM \citep{lippert2011fast} because of their popularity and computational speed. 
+For multi-locus association mapping, we chose bigRR \citep{shen2013novel}, glmnet \citep{Friedman2010glmnet}, 
+LMM-Lasso \citep{rakitsch2013lasso}, MLMM \citep{segura2012efficient} , and r2VIM \citep{szymczak2016r2vim}.  
 Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 
@@
 For each scenario, 100 replicates were generated. A single replicate consisted of SNP and quantitative trait data. 
-Extra realism was introduced into the simulation study through the drawing of the SNP genotypes from the 1000 Genome Project, phase 3  \cite{10002010map}.
+Extra realism was introduced into the simulation study through the drawing of the SNP genotypes from the 1000 Genome Project, phase 3  \citep{10002010map}.
  The quantitative trait data were generated 
@@
 These were the number of trees, the number of variables for building a tree, and the minimum size of a terminal node. Ideally, these parameters would be "tuned" on a replicate-by-replicate 
-basis \cite{boulesteix2012overview}. However, this was not practical here. We instead used the same settings as in \cite{szymczak2016r2vim} where 
+basis \citep{boulesteix2012overview}. However, this was not practical here. We instead used the same settings as in \citep{szymczak2016r2vim} where 
 the number of trees was set to 1000, the number of variables was set to 20\% of the number of SNPs, and 
@@
 \begin{figure}
-\caption{Power verse false discovery rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
+\caption{
+\setstretch{1.6}{
+Power verse false discovery rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
 FaST-LMM was run where all the SNP data are used 
@@
 estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
-
+}
+\vspace{-3cm}
 \label{figpowersingle}
-\begin{center}
+%%\begin{center}
 \includegraphics[width=15cm, height=20cm]{power2main.eps}
-\end{center}
+%%\end{center}
 
@@
 \begin{figure}
-\caption{Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
+\caption{
+\setstretch{1.6}{
+Median run times, in minutes, for the analysis of simulation study data from the six scenarios. 
 Eagle is compared against five other multi-locus programs/packages (A) and two single-locus programs (B). 
@@
  FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
-
+ }
+\vspace{1cm}
 %% plot formed via PowerPoint - ~/AM-Paper/Figure_power.pptx
@@
 \centering
-    \includegraphics[width=1.0\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
+    \includegraphics[width=1.2\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
 \end{figure}
@@
 We were interested in comparing results from Eagle with those from single-locus association mapping for a real data set.
- We chose to focus on data from a large outbred mouse study \cite{nicod2016genome}.
+ We chose to focus on data from a large outbred mouse study \citep{nicod2016genome}.
 This study was unusual in that it collected and analysed SNP dosages (continuous values from zero to one of expected allele counts)  instead of the more common SNP genotypes.
-Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \cite{zheng2011comparison}.
+Analyses based on dosages rather than discrete genotypes have been shown to have greater power for the detection of genes that are influencing a trait  \citep{zheng2011comparison}.
 By converting the dosages into genotypes and analysing the data with the single-locus program FaST-LMM, we obtained a subset of those findings reported in the original study.
@@
 
-We repeated the single-locus analyses as first performed \cite{nicod2016genome}, except 
+We repeated the single-locus analyses as first performed \citep{nicod2016genome}, except 
 that we focused on autosomal SNPs and our analyses were based on SNP genotypes  rather than SNP dosages. 
@@
 We ran Eagle in three ways. 
-Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \cite{chen2008extended}. 
+Eagle chooses the best model via the extended Bayesian information criteria (extBIC) \citep{chen2008extended}. 
   The conservativeness of the extBIC can be adjusted by a single regularisation parameter  $\gamma$ that ranges from zero to one. 
@@
 under this setting (Eagle$^{default}$).
-An alternate  \cite{chen2008extended} , less conservative way of setting $\gamma$ is to  let $\gamma = 1 - \frac{1}{(2\kappa)}$ with $\kappa=\frac{log(L)}{log(n_g)}$ where
+An alternate  \citep{chen2008extended} , less conservative way of setting $\gamma$ is to  let $\gamma = 1 - \frac{1}{(2\kappa)}$ with $\kappa=\frac{log(L)}{log(n_g)}$ where
  $L$ is the number of loci that span the genome, and $n_g$ is the number of individuals/groups/lines/strains in the study
@@
 \begin{figure}
-\caption{Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
+\caption{
+\setstretch{1.6}{
+Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
 multi-locus method
@@
 of 5\% (Eagle$^{optimal}$).  The number of SNP-trait associations found are reported in the cells. }
+}
 \label{figmouse}
@@
 
-\section{Discussion}
+\section{Discussion/Conclusion}
 Eagle is a new linear mixed model based method (and R package) for multi-locus association mapping. It advances the state of association mapping in several ways. 
@@
 SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.
-Permutation and its variants \cite{ browning2008presto,pahl2010permory} are also equally valid empirical approaches.
+Permutation and its variants \citep{ browning2008presto,pahl2010permory} are also equally valid empirical approaches.
 Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the
@@
 However, we would like to think that the parameter settings we used are sensible 
-since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.
+since they match the values in the original r2VIM publication \citep{szymczak2016r2vim}.
 
@@
 In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis.
-Even though this is a well accepted practice, it is approximate \cite{gogel2018comparison}.
+Even though this is a well accepted practice, it is approximate \citep{gogel2018comparison}.
 A better solution is to fit a single model to the data.
 Although not specifically designed for association mapping,
-WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RWGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
+WGAIM \citep{verbyla2007analysis}, upon which Eagle is based, and RWGAIM \citep{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
 models,  run time and memory usage can become limiting factors for analysis.
@@
 High-throughput array-based technologies continue to decrease the cost of genotyping, permitting ever larger GWASs to be performed. 
-Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs \cite{gudbjartsson2015large, long2017whole}
+Whole-genome sequencing is also now a reality. Already sequence across entire genomes are being collected for GWASs \citep{gudbjartsson2015large, long2017whole}
 culminating in data on millions of SNPs. It is because of this growing demand that 
@@
 
-\section{Methods}
-
-\subsection{Mouse Data}
-
-The data were obtained from a large genome-wide association study that was performed in outbred mice \cite{nicod2016genome}. 
-Phenotypic and genotypic data were available on 1,887 adult mice. 
-The phenotypic data included raw and adjusted (for fixed effects) measurements from 200 behavioural, tissue, and physiological traits.  
-Of these traits, 
-43 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
-43 traits that were the focus of our real data analyses. As in the original study  \cite{nicod2016genome}, our analyses 
-were based on the adjusted traits.
-Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
-form of marker dosages (expected allele counts that ranged from zero to one). All missing data had been imputed. 
-We converted the dosages into discrete genotypes 
-by clustering around 0, 0.5, and 1, corresponding to SNP genotypes AA, AB, and BB, respectively. 
-We focused our analyses on the autosomal SNPs.
-
-\subsection{Eagle Approach for Multi-locus Association Mapping}
-
-Eagle is a method for multi-locus association mapping on a genome-wide scale.
-It is based on linear mixed models.
-It differs from most other single- and multi-locus association mapping methods.
-Eagle treats association mapping as a model selection instead of variable selection problem.
-Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
-Eagle also 
-reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest to the genes influencing a trait. 
-The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
-quantitative trait loci in experimental crosses \cite{verbyla2007analysis}.
-
-Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of $s$ ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
-selected in the $k$th iteration of the model building process. Suppose three iterations  $(s=3)$
-have been performed and say the 500023rd, 15th, and 420th
-SNP were selected. Then $S=\{500023, 15, 420\}$. 
-Let $\by^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
-Let $\bM^{(n_g \; \times \; L)} = [\bmm_1 \bmm_2 \ldots \bmm_L]$ be a matrix containing the genotype data which have been collected 
-from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
-may be taken of the same group/line/strain. 
- It is common for the columns of $\bM$ to be in map order but this is not a requirement. 
-The vector $\bmm_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
-The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.
-
-The specifics of the Eagle method are as follows. 
-Eagle builds the "best" model iteratively, via forward selection. 
-Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
-associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
-To perform the $s+1$th  iteration, we first fit the current model to the data. 
-The (current) model is of the form 
-\begin{equation}
-\label{eq1}
-\by = \bX \btau + \bZ \bu_g + \be
-\end{equation}
-where 
-$\bX^{(n \; \times \; p)}$ and $\bZ^{( n \; \times \; n_g)}$ are known design matrices with $\bX$ being of full rank and $\bZ$ 
-containing zeros and ones that assign the appropriate genetic effect to each measurement. 
-The vector 
-$\btau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
-$\bu_g^{(n_g \; \times \; 1)}$ contains the 
-genetic effects. The vector of residuals is 
-$\be^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(\bzero, \sigma^2_e \bI^{(n \; \times \; n)})$. 
-So far,  this model differs little from standard linear mixed models for association mapping \cite{yu2006unified,zhao2007arabidopsis} 
-However, 
-it is how we specify $\bu_g$ that distinguishes our model from the others.
-
-The genetic effects $\bu_g$ are modelled as 
-\begin{equation}
-\label{eq2}
-\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} +\bM_{-S} \ba_{-S}
-\end{equation}
-where $\bmm_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
-$a_{S_k}$ is the additive effect of the $k$th selected SNP, $\bM_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
-with the data for the SNP in $S$ removed,  and $\ba_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
-$\ba_{-S} \sim N(\bzero, \sigma_a^2 \bI^{(L-s \; \times \;  L-s)})$. 
-The terms in the summation on the left hand side are fixed effects.   They account 
-for the additive effects of those SNPs that have been found to be in association with the trait. The other term is a random effect. 
-It accounts for the joint effect of the yet-to-be-identified SNP that are in association with the trait. 
-This is a simple genetic model but it 
-is effective for discovering SNP-trait associations.
-
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
-can be computationally demanding. However, our model only contains a single random effect ($\ba_{-S}$). Here, highly efficient single-dimension 
-optimisation via spectral decomposition is possible \cite{kang2008efficient}.
-
-Third, we identify the $(s+1)$th SNP that is in strongest association with the trait, based on the maximum score statistic
-$t_j^2 = \frac{ \widetilde{a} _j^2}{\textrm{var}(\widetilde{a}_j)}$ where $\widetilde{a}_j$ is the best linear unbiased predictor (BLUP), 
-and $\textrm{var}(\widetilde{a}_j)$ is its variance. This statistic is not only appealing intuitively, where we 
-identify a SNP based on its (random) effect size and accuracy, but is justifiable, theoretically \cite{verbyla2012rwgaim}.
-
-Fourth, we determine the importance of the $(s+1)$th selected SNP via a model selection strategy  \cite{verbyla2007analysis}. 
-We begin by reforming (\ref{eq2}) where $S$ now contains the $s + 1$ selected SNP.  We then fit this new model to the data
-via maximum likelihood and calculate its extended Bayesian information criteria (extBIC) \cite{chen2008extended}.  The 
-extBIC is a model selection measure that takes into account the number of unknown parameters and the complexity 
-of the model space.  It is well suited to the model selection problem in genome-wide association studies \cite{chen2008extended}. 
-It is different to the model selection measure used in  \cite{verbyla2007analysis}.
-If this new model has a larger extBIC than the current model, then the $s+1$th selected SNP is added to 
-the current model and the above process is repeated. If this new model has a smaller extBIC than the current model, then the 
-model building process is complete. The set of SNP in strongest association with the trait is the $s$ SNPs previously identified.
-
-\subsubsection {Reducing the dimension of the model:}
-
-In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
-The vector $\ba_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
-$L$, the number of SNPs, can be extremely large.  An alternative model is given by 
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
-They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
-\begin{equation}
-\label{eq3}
-\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} + (\bM_{-S} \bM_{-S}^T)^{1/2} \ba^*_{-S}
-\end{equation}
-where $\ba^* \sim N(\bzero, \sigma_a^2 \bI^{(n_g \; \times \;  n_g)})$, and 
-$(\bM_{-S} \bM_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
-Although it may not be obvious, the two models are equivalent, 
-having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
-model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
-effects needing estimating.
-
-Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\bwidetildea$ from estimates from model  (\ref{eq3})  with 
-\begin{equation}
-\bwidetildea = \left [ \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \right ] \bwidetildea^*
-\end{equation}
-where its variance matrix is
-\begin{equation}
-\label{eq4}
-\textrm{var}(\bwidetildea) = \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \textrm{var}(\bwidetildea^*) (\bM_{-S} \bM_{-S}^T)^{-1/2} \bM_{-S}
-\end{equation}
-These values are needed in order to calculate the score statistic $t_j^2$ for identifying the SNP in strongest association with the trait. 
-Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation 
-of (\ref{eq4}).
-
-%%\iffalse
-
-\subsection{Comparison Methods}
-
- \subsubsection{ Multi-locus methods:} We compare the computational and statistical performance of Eagle against five multi-locus methods. They are bigRR, LMM-Lasso, glmnet, 
-MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
-BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
-methods. BigRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
-Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
-They allow the strength of association between all the SNPs and trait to be measured within a single model, simultaneously. 
-A limitation of these methods though is that the statistical significance of the SNP effects cannot be easily determined. 
- Due to the adaptive nature of the estimation procedures, to do this 
-analytically is challenging and is an area of active research \cite{lockhart2014significance}. Instead, we calculate significance empirically via 
-stability selection (see below).
-
-MLMM is closest in philosophy to Eagle. It too is based on building the best model via forward selection, 
-within  a linear mixed model framework, using the 
-extBIC. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
-model differently to Eagle.  Eagle uses a score statistic in which to identify SNPs to enter the model.
-MLMM uses the statistical significance of the SNPs. This involves fitting a separate linear mixed model for each unidentified SNP. 
-Both are R packages but there is a significant difference in computational performance (see Results).
-
-R2VIM differs to the other four methods in that it is a non-parametric model-free approach. It implements  
-random forests but where multiple parallel runs are performed. Each run leads to 
-different random forests being created.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
-importance score by the minimum importance score observed across all the SNPs within a run. 
-Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
-Unfortunately, the relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
-but the computational cost is high, restricting the size of data that can be analysed.
-
-\subsubsection{Single-locus methods:} We also compare the performance of Eagle against two single-locus methods, GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast}.
-Both are based on linear mixed models.
-The models have a single fixed effect for the SNP,  other fixed effects,
-a single random effect to account for familial relatedness (or polygenic background), and an error. The significance of the SNP effect in the model 
-is a measure of the strength of association.  They are of the same computational complexity \cite{zhou2012genome}, and produce exact results.  
-Both perform a single spectral decomposition of the relationship (or similarity) matrix $K$, use  an eigenvector matrix to rotate the data, 
-and reformulate the (residual) log likelihood for easier computation. They do differ in their estimation procedure. GEMMA implements Newton-Raphson. 
-FaST-LMM implements Brent's algorithm. Newton-Raphson is more complicated but has better convergence properties than Brent's algorithm. 
-Both methods are state-of-the-art and have been implemented in highly efficient computer programs.
-
-\subsection{Stability Selection}
-
- Stability selection \cite{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of the significance of effects in regression models where 
-the number of parameters ($p$) can be much larger than the number of samples ($n$).
-We chose stability selection over permutation and other sampling procedures because of its low computational cost.
-We employe stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
-
-To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, we begin
- by performing a preliminary analysis of a replicate where  we tune the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
-This tuning of the regularisation parameter is done for data under each simulation scenario but it is not necessary to do it for every replicate within a scenario.
-It is sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
-Here, we have the luxury of knowing the genetic conditions 
-under which the simulated trait data are generated. We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
- However, stability selection estimates are robust  \cite{meinshausen2010stability} 
- and the regularisation parameter can be tuned to any reasonable number of 
- non-zero effects. We sample repeatedly, without replacement, from the replicate. We draw 100 replicate subsets of size $n/2$.  
-Each replicate subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
-A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
-number of times the SNP have a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
-
-For bigRR, we modify our stability selection procedure slightly. 
-There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
-As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
-SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
-A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
-number of times the SNP is recorded divided by the number of replicates.
-
-\subsection{Generation of Simulation Data}
-
-The data are generated via data perturbation \cite{zhao2007arabidopsis}. 
-Data perturbation amalgamates real with simulated data to generate replicates. 
-It is a way of introducing greater realism into a simulation study. 
- Here, the genotype data are real, the 
-quantitative trait data are simulated. 
- The SNP genotypes are drawn, 
-according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \cite{10002010map}. 
-Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.
-
-To generate the trait data $\by$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
-mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
-are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
-generate a $\bm{g}^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
-Fifth, $\bm{e}^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
-the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $\by =  \bm{g} + \bm{e}$.  
-In forming $\by$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
-not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
-is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.
-
-\subsection{Implementation}
-
-Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
-via C++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
-mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
-calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
-a link to our server for demonstrating Eagle on some test data are available on the Eagle website. Eagle is available for download from the CRAN website.
-
 \section*{Data Availability}
@@
 
+\section*{Acknowledgements}
+We would like to thank the IM\&T Scientific Computing group at the Commonwealth Scientific and Industrial Research Organisation (CSIRO) for providing access and support to the CSIRO Accelerator Cluster upon which all computation was performed.
+
+\section*{Funding}
+This research received no specific grant from any funding agency in the public, commercial, or not-for-profit sectors.
+
+\vspace{1cm}
+\noindent
+{\em Conflict of interest:} The authors declare that they have no
+competing financial interests.
+
 %%\bibliographystyle{naturemag}
-\section*{References}
 \bibliography{biblibrary}
 
-\begin{addendum}
- \item We would like to thank the IM\&T Scientific Computing group at the Commonwealth Scientific and Industrial Research Organisation (CSIRO) for providing access and support to the CSIRO Accelerator Cluster upon which all computation was performed.
-\item[Competing Interests] The authors declare that they have no
-competing financial interests.
- \item[Correspondence] Correspondence and requests for materials
-should be addressed to Andrew George ~(email: andrew.george@csiro.au).
-\item[URLs]  Eagle website http://eagle.r-forge.r-project.org;   
-CRAN website https://cran.r-project.org.
-
-\end{addendum}
-
 \end{document}

### 2019-02-08 paper.tex
08Jan18 Nature Communications Submission Version 1
Just submitted to Bioinformatics

@@
 
-\title{Eagle: Making multi-locus association mapping on a genome-wide scale routine}
+\title{Eagle makes multi-locus association mapping on a genome-wide scale routine}
 \author{Andrew W. George$^{1,*}$, Arunas Verbyla$^1$, Joshua Bowden$^2$}
@@
 We also analyse data from a published mouse study.
-Eagle found over two times more validated findings than the state-of-the-art single-locus method.
+Eagle found over 100\% more validated findings than the state-of-the-art single-locus method.
 Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.

### 2019-07-24 BioinformaticsRevison/bioinformaticspaper.tex
Revesion 1

--- 2019-02-08 bioinformaticspaper.tex
@@
 \usepackage{authblk}
+\usepackage{xcolor}
 
@@
 is considerably more powerful than single-locus association mapping b.
-doesn't suffer from multiple testing issues c.
-is threshold free d.
-gives results that are immediately interpretable and e.
+does not suffer from multiple testing issues c.
+gives results that are immediately interpretable and d.
 has a computational footprint comparable to single-locus association mapping.
@@
 This has meant that data can be collected from general populations, even if these populations are highly structured.
-Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine \citep{purcell2007plink, bradbury2007tassel}.
+Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine  \citep{yu2006unified,zhao2007arabidopsis}.
 
@@
 This is despite there being several significant problems with analysing data in this way.
-First, the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
+First, \textcolor{red}{ for each SNP, a hypothesis test is performed. The null hypothesis is that  there is no association between the SNP and trait. 
+The alternative is that the SNP is in association with the trait. It is straight forward to guard against wrongly rejecting the null 
+hypothesis (or making a type 1 error) if only a single hypothesis test is being performed. However, the analysis of GWAS data 
+with locus-by-locus methods necessitates  conducting a large number of correlated hypothesis tests, simultaneously. 
+ This leads to an increased risk of type 1 errors.  To deal with this challenge, many different solutions have been offered 
+ \citep{storey2003statistical, li2005adjusting, de2005efficiency}}.  Second, 
+the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
 The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
-that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained by testing the statistical 
-significance of a single nucleotide polymorphisms (SNP) when treated as an effect in an appropriate model. 
+that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained from the hypothesis test. 
 The location of peaks in this plot identify genomic 
-regions of interest.
-Inferring the exact number of regions though can be difficult if the peaks are not well separated.
-Second, when multiple statistical tests are performed, the probability of wrongly accepting a result (type 1 error) is inflated.
-This is known as the multiple testing problem.
-Many different solutions have been offered \citep{storey2003statistical, li2005adjusting, de2005efficiency}.
-Yet, there is still no well accepted way of correcting for multiple testing in the context of genome-wide association mapping.
+regions of interest. Inferring the exact number of regions though can be difficult if the peaks are not well separated. 
  Third, many of the traits whose genetic secrets we are trying to discover are complex.
@@
 regularisation techniques, such as ridge regression \citep{shen2013novel}  and lasso \citep{rakitsch2013lasso}, measure all locus-trait associations simultaneously.
-Here, multiple testing is not an issue.
 These techniques though are computationally demanding.
-Also, their results can be difficult to interpret.
-The strength of association is not measured by a $p$ value but by the size of the regression coefficient for the SNP in the model.
+\textcolor{red}{Also, the strength of association is not measured by a $p$ value but by the size of the regression coefficient for the SNP in the model.
+Further processing is required before the results can be interpreted \citep{cho2010joint, rakitsch2013lasso}}.
 More recently, associations have started to be mapped with random forests \citep{szymczak2016r2vim}.
@@
 A multi-locus method that does show promise is the multiple-locus linear mixed model method \citep{segura2012efficient}.
-The best multi-locus model is built with simple forward selection.
-Results are immediately interpretable but here, computation becomes challenging for large datasets.
+The best multi-locus model is built with forward and backward stepwise selection.
+Results are immediately interpretable \textcolor{red}{in that the SNP closest to the genes underlying the trait are identified} 
+but computation does become challenging for large datasets.
 
@@
 were based on the adjusted traits.
-Genotypic data were available on 359, 559 (353,697 autosomal) SNPs in the 
+Genotypic data were available on \textcolor{red}{359,559} (353,697 autosomal) SNPs in the 
 form of marker dosages (expected allele counts that ranged from zero to one). All missing data had been imputed. 
@@
 It is based on linear mixed models.
-It differs from most other single- and multi-locus association mapping methods.
-Eagle treats association mapping as a model selection instead of variable selection problem.
-Consequently,  we do not have to contend with multiple testing issues or having to construct significance thresholds.
+It differs from most other single- and multi-locus association mapping methods in that Eagle treats association mapping as a model selection problem (ball2001bayesian,broman2002model,yi2005bayesian).
+The "best" model is found via forward selection. It makes use of a modified form of the Bayesian information criterion, BIC, for model selection.
+
+Eagle does not suffer from multiple testing issues. A "best" model is built iteratively. At each iteration, a hypothesis test is performed. 
+Only a small number of iterations are needed in building the "best" model.  In contrast,  for single-locus methods, multiple testing is an 
+issue because each SNP is assessed separately, culminating in a large number of hypothesis tests being performed. 
 Eagle also 
-reports as its findings only those SNPs that are in strongest linkage disequilibrium, and hence closest to the genes influencing a trait. 
+reports as its findings only those SNPs that are in strongest linkage disequilibrium with  the genes influencing a trait. 
 The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
@@
 each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
-multiple testing issues and the need for significance thresholds. 
+ the need for significance thresholds. 
  As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.

### 2019-07-24 BioinformaticsRevison/bioinformaticspaper.tex
Revision working through Review 1 changes

@@
  This leads to an increased risk of type 1 errors.  To deal with this challenge, many different solutions have been offered 
- \citep{storey2003statistical, li2005adjusting, de2005efficiency}}.  Second, 
+ \citep{storey2003statistical, li2005adjusting, de2005efficiency}.  Second, 
 the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
@@
 The location of peaks in this plot identify genomic 
-regions of interest. Inferring the exact number of regions though can be difficult if the peaks are not well separated. 
+regions of interest. Inferring the exact number of regions though can be difficult if the peaks are not well separated. }
  Third, many of the traits whose genetic secrets we are trying to discover are complex.
@@
 
+\textcolor{red}{
 Eagle is a method for multi-locus association mapping on a genome-wide scale.
 It is based on linear mixed models.
-It differs from most other single- and multi-locus association mapping methods in that Eagle treats association mapping as a model selection problem (ball2001bayesian,broman2002model,yi2005bayesian).
+It differs from most other single- and multi-locus association mapping methods in that Eagle treats association mapping as a model selection problem \citep{ball2001bayesian,broman2002model,yi2005bayesian}.
 The "best" model is found via forward selection. It makes use of a modified form of the Bayesian information criterion, BIC, for model selection.
-
-Eagle does not suffer from multiple testing issues. A "best" model is built iteratively. At each iteration, a hypothesis test is performed. 
-Only a small number of iterations are needed in building the "best" model.  In contrast,  for single-locus methods, multiple testing is an 
-issue because each SNP is assessed separately, culminating in a large number of hypothesis tests being performed. 
-Eagle also 
+A "best" model is built iteratively. At each iteration, a hypothesis test is performed. 
+Only a small number of iterations are needed in building the "best" model.  Consequently, 
+Eagle does not suffer from multiple testing issues.
+In contrast,  for single-locus methods, multiple testing is an 
+issue because each SNP is assessed separately, culminating in the need for a large number of hypothesis tests to be  performed. }
+Eagle 
 reports as its findings only those SNPs that are in strongest linkage disequilibrium with  the genes influencing a trait. 
@@
 each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
- the need for significance thresholds. 
+ multiple testing issues. 
  As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.

### 2019-07-24 BioinformaticsRevison/bioinformaticspaper.tex
Working of Reviewers 1 changes

@@
 In contrast,  for single-locus methods, multiple testing is an 
-issue because each SNP is assessed separately, culminating in the need for a large number of hypothesis tests to be  performed. }
+issue because each SNP is assessed separately, culminating in the need for a large number of hypothesis tests to be  performed. 
 Eagle 
-reports as its findings only those SNPs that are in strongest linkage disequilibrium with  the genes influencing a trait. 
+reports as its findings only those SNPs that are in strongest linkage disequilibrium with  the genes influencing a trait. }
 The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
@@
 
-Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
+Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via \textcolor{red}{restricted} maximum likelihood (REML).  For complex models, REML
 can be computationally demanding. However, our model only contains a single random effect ($\ba_{-S}$). Here, highly efficient single-dimension 
@@
 
- \subsubsection{ Multi-locus methods:} We compare the computational and statistical performance of Eagle against five multi-locus methods. They are bigRR, LMM-Lasso, glmnet, 
-MLMM, and r2VIM.  All but glmnet have been purposely designed for genome-wide association mapping. 
+ \subsubsection{ Multi-locus methods:} We compare the computational and statistical performance of Eagle against five multi-locus methods.
+\textcolor{red}{They are bigRR  \citep{shen2013novel}, LMM-Lasso \citep{rakitsch2013lasso}, glmnet \citep{Friedman2010glmnet},
+MLMM \citep{segura2012efficient}, and r2VIM \citep{szymczak2016r2vim}}.  All but glmnet have been purposely designed for genome-wide association mapping. 
 BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
@@
 
-MLMM is closest in philosophy to Eagle. It too is based on building the best model via forward selection, 
+MLMM is closest in philosophy to Eagle. It too is based on building the best model via \textcolor{red}{stepwise} selection, 
 within  a linear mixed model framework, using the 
-extBIC. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
-model differently to Eagle.  Eagle uses a score statistic in which to identify SNPs to enter the model.
-MLMM uses the statistical significance of the SNPs. This involves fitting a separate linear mixed model for each unidentified SNP. 
+extBIC as one of its model selection criterion. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
+model differently to Eagle.  Eagle chooses a SNP to enter the model by finding the SNP with the largest score statistic. 
+This score statistic has been used, typically, to detect outliers  but we are using it here to identify SNP with 
+unusually large random effect sizes. 
+MLMM uses the statistical significance of the SNPs,  when treated as a fixed effect in the model. This involves fitting a separate linear mixed model for each unidentified SNP, a 
+potentially computationally expensive exercise. However, MLMM does this in a clever and efficient way via the Gram-Schmit process. 
+Note, even though a hypothesis test is being performed for each SNP by MLMM, it does not suffer from multiple testing issues. 
+Neither the null nor the alternative hypothesis is being accepted or rejected. Only the hypothesis yielding the most significant association is 
+of interest. 
 Both are R packages but there is a significant difference in computational performance (see Results).
@@
 Of the many ways of performing single-locus association mapping, we chose
-GEMMA \citep{zhou2012genome}  and FaST-LMM \citep{lippert2011fast} because of their popularity and computational speed. 
-For multi-locus association mapping, we chose bigRR \citep{shen2013novel}, glmnet \citep{Friedman2010glmnet}, 
-LMM-Lasso \citep{rakitsch2013lasso}, MLMM \citep{segura2012efficient} , and r2VIM \citep{szymczak2016r2vim}.  
+GEMMA and FaST-LMM  because of their popularity and computational speed. 
+For multi-locus association mapping, we chose bigRR, glmnet, 
+LMM-Lasso, MLMM, and r2VIM.  
 Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 

### 2019-07-25 BioinformaticsRevison/bioinformaticspaper.tex
Working of Reviewers 1 changes v2

@@
 \subsection{Eagle Approach for Multi-locus Association Mapping}
+\label{subsection:Eagle}
 
@@
 MLMM is closest in philosophy to Eagle. It too is based on building the best model via \textcolor{red}{stepwise} selection, 
-within  a linear mixed model framework, using the 
-extBIC as one of its model selection criterion. However, there are differences between the two methods.  MLMM does not make use of dimension reduction. Also, it builds its "best" 
-model differently to Eagle.  Eagle chooses a SNP to enter the model by finding the SNP with the largest score statistic. 
-This score statistic has been used, typically, to detect outliers  but we are using it here to identify SNP with 
-unusually large random effect sizes. 
-MLMM uses the statistical significance of the SNPs,  when treated as a fixed effect in the model. This involves fitting a separate linear mixed model for each unidentified SNP, a 
-potentially computationally expensive exercise. However, MLMM does this in a clever and efficient way via the Gram-Schmit process. 
+within  a linear mixed model framework, and uses the 
+extBIC as one of its model selection criterion. However, there are differences between the two methods. 
+MLMM does not make use of dimension reduction. \textcolor{red}{Also, how SNP are selected to enter the model differs between the 
+two methods. 
+Eagle identifies a SNP of interest from its score statistic (see Section \ref{subsection:Eagle} for details).  This score statistic was originally developed for 
+outlier detection in linear (mixed) models but it is being used by Eagle to identify 
+SNP with unusually large random effects. MLMM instead uses 
+the statistical significance of a SNP,  when treated as a fixed effect in the model. This involves fitting a separate linear mixed model for each candidate SNP, a potentially
+computationally expensive exercise. However, MLMM does this in a clever and efficient way via the Gram-Schmidt process. 
+Both are R packages but there is a significant difference in computational performance (see Results). 
 Note, even though a hypothesis test is being performed for each SNP by MLMM, it does not suffer from multiple testing issues. 
 Neither the null nor the alternative hypothesis is being accepted or rejected. Only the hypothesis yielding the most significant association is 
-of interest. 
-Both are R packages but there is a significant difference in computational performance (see Results).
+of interest. }
 
@@
 
-\subsection{Stability Selection}
-
- Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications, including the estimation of the significance of effects in regression models where 
-the number of parameters ($p$) can be much larger than the number of samples ($n$).
-We chose stability selection over permutation and other sampling procedures because of its low computational cost.
-We employe stability selection to estimate, empirically, the significance of the SNP effects from the regularisation method's analyses of the simulated data.
-
-To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, we begin
- by performing a preliminary analysis of a replicate where  we tune the  regularisation parameter to yield 20 to 30 non-zero SNP effects.
-This tuning of the regularisation parameter is done for data under each simulation scenario but it is not necessary to do it for every replicate within a scenario.
-It is sufficient to pick a single replicate at random upon which to tune the regularisation parameter.
-Here, we have the luxury of knowing the genetic conditions 
-under which the simulated trait data are generated. We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
- However, stability selection estimates are robust  \citep{meinshausen2010stability} 
- and the regularisation parameter can be tuned to any reasonable number of 
- non-zero effects. We sample repeatedly, without replacement, from the replicate. We draw 100 replicate subsets of size $n/2$.  
-Each replicate subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
-A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
-number of times the SNP have a non-zero effect size over all the replicates divided by the number of replicates (which was 100).
-
-For bigRR, we modify our stability selection procedure slightly. 
-There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
-As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
-SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
-A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
-number of times the SNP is recorded divided by the number of replicates.
-
 \subsection{Generation of Simulation Data}
@@
  The SNP genotypes are drawn, 
-according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \citep{10002010map}. 
-Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.
+according to the specifications of the scenario , from data collected from the 1000 Genome Project, version 3   \citep{10002010map}. 
+Across scenarios (see Results for details), the SNP data differs. Across replicates within a scenario, the SNP data are the same. For each scenario, 100 replicates are generated.
 
@@
 is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.
+
+\subsection{Stability Selection}
+
+ Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
+ It is used here to assign statistical significance to the regression coefficient estimates of
+ LMM-Lasso, glmnet, and bigRR.   LMM-Lasso, glmnet, and bigRR are regularisation 
+ methods. As such,  the size of a SNP effect in the model is estimated but not its significance.  
+ Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost.
+
+The Stability selection procedure for LMM-Lasso and glmnet is as follows. Remembering that for each simulation scenario we generate 100 
+replicates of data under that scenario, we begin by choosing one of these replicates at random. 
+We then analyse these data with LMM-Lasso (glmnet) with different values of its  regularisation parameter. 
+Our goal is to find the parameter value that  yields 20 to 30 non-zero SNP effects. 
+This can be done via a simple binary search approach. 
+ Then, for each replicate, we subsampled repeatedly, 100 subsets of size $n/2$.  A replicate is subsampled without replacement.  
+ These subsets are then analysed with LMM-Lasso (glmnet) with its regularisation parameter fixed to the value found previously. 
+A binary vector, of length the number of SNP, is recorded as the result. A one (zero) means the SNP had a non-zero (zero) effect size. 
+Calculating a SNP's statistical significance is now a simple task. We take the vector sum of the binary vectors over all 100 subsets. 
+This vector sum will contain elements in the range of 0 to 100. By dividing each element in this vector sum by the number of 
+subsets upon which the sum was calculated, we obtain empirical probabilities.  
+ These 
+probabilities measure the strength of evidence for the SNPs to be in association with the trait.
+
+By taking the vector sum over all 100 subsets and dividing each element in the sum by 100, we obtain the probability 
+estimates of the significance of the SNPs.
+
+A probability estimate of the significance of a SNP is now just a simple vector sum, over the 100 subsets,  of the binary vectors
+
+By taking the vector sum over all 100 subsets, for a particular replicate, and dividing each element sum by 100, we obtain a probability 
+estimate of significance.
+
+Now, for a replicate,  to calculate the statistical significance of the coefficient estimates obtained with LMM-Lasso (glmnet),
+
+estimate the statistical significance of a SNP, for a particular replicate, is obtained by taking the
+
+Now, for a particular replicate, we count the number of times LMM-Lasso (glmnet) gave a non-zero
+
+Now, for a particular replicate which was analysed with LMM-Lasso (glmnet), 
+we count the number of times
+
+we calculate the statistical significance of a SNP by counting the number of times its 
+coefficient was non-zero in the LMM-Lasso (glmnet) analyses of the 100 subsets of data.  A probability estimate of significance is obtained 
+by dividing this count by the number of subsetted analyses, which was 100.  Note that it is not 
+necessary to tune the regularisation   parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. Also, 
+we have the luxury of knowing the genetic conditions 
+under which the simulated trait data are generated. we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
+ However, stability selection is a subsampling procedure that is robust to parameter misspecification \citep{meinshausen2010stability}. 
+ The regularisation parameter can be tuned to any reasonable number of 
+ non-zero effects.
+
+Then, for the LMM-Lasso (glmnet) analysis of a given replicate, to estimate the statistical significance of a SNP effect, we count 
+the number of times the SNP effect has non-zero effect size across the 100 subset LMM-Lasso (glmnet) analyses.  This number is then 
+divided by the number of subsets (which is 100 for the replicate)
+
+ to estimate the statistical significance of the SNP effects for a given replicate, the number of times
+
+For a replicate, a (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
+number of times a SNP has a non-zero effect size over all the subsets divided by the number of subsets (which was 100).
+
+Here, the link between 
+trait and genotype is preserved, with only the sample size being reduced.
+
+\textcolor{red}{
+To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, for a given scenario, 
+we begin by performing a preliminary analysis of a randomly chosen  replicate (i.e. trait and SNP genotype data)  from the 
+ 100 available replicates for that scenario.  Our goal is to find a value of the regularisation parameter used by LMM-Lasso and glmnet 
+ that yields 20 to 30 non-zero SNP effects. It is not necessary to tune the regularisation 
+  parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. 
+  Since we have the luxury of knowing the genetic conditions 
+under which the simulated trait data are generated, we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
+ However, stability selection is a subsampling procedure that is robust to misspecification of parameters  \citep{meinshausen2010stability}. 
+ The regularisation parameter can be tuned to any reasonable number of 
+ non-zero effects.  Once we have a value for the regularisation parameter,
+
+ We subsample repeatedly, without replacement, from a replicate. We draw 100 subsets of size $n/2$.  
+Each subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
+A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
+number of times the SNP have a non-zero effect size over all the replicates divided by the number of subsets (which was 100). }
+
+For bigRR, we modify our stability selection procedure slightly. 
+There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
+As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
+SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
+A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
+number of times the SNP is recorded divided by the number of replicates.
 

### 2019-07-26 BioinformaticsRevison/bioinformaticspaper.tex
Working of stability selection section git push

@@
  Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
- It is used here to assign statistical significance to the regression coefficient estimates of
- LMM-Lasso, glmnet, and bigRR.   LMM-Lasso, glmnet, and bigRR are regularisation 
- methods. As such,  the size of a SNP effect in the model is estimated but not its significance.  
- Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost.
+ It is used here to estimate, empirically, the statistical significance of the results from LMM-Lasso, glment, and bigRR 
+ of the simulated data. These three regularisation methods give the effect sizes of the SNPs, but not their significance, as their 
+ results.   Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost.
 
-The Stability selection procedure for LMM-Lasso and glmnet is as follows. Remembering that for each simulation scenario we generate 100 
+The Stability selection procedure for LMM-Lasso and glmnet is as follows. 
+For a particular scenario, we begin by finding the value of the regularisation parameter that yields 20 to 30 non-zero SNP effects. 
+We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect from the analysis of a replicate in the simulation study.
+The regularisation parameter though could have been tuned to give any reasonable number of non-zero SNP effects.
+ This tuning was done for each of the six scenarios 
+ but only for a single replicate within a scenario. It is not necessary to tune the regularisation parameter on every replicate 
+ when replicates are generated under the same (trait, sample size, and number of SNP) conditions within a scenario.
+
+We did this for each  scenarios but not for every replicate within a scenario. It is unnecessary to set the regularisation parameter 
+for every replicate when replicates within a scenario are generated under the same conditions. 
+For a given scenario, we chose a replicate at random.  We then used a binary search procedure to set the regularisation parameter. 
+Then, for a replicate, we subsample repeatedly, 100 subsets of size $n/2$. These subsamples are drawn without replacement.  
+Here, the matching of trait to genotype is preserved in the subsamples. A subsample differs to the replicate in size only.
+
+Remembering that for each simulation scenario we generate 100 
 replicates of data under that scenario, we begin by choosing one of these replicates at random. 
-We then analyse these data with LMM-Lasso (glmnet) with different values of its  regularisation parameter. 
+We then analyse this replicate with LMM-Lasso (glmnet) with different values of its  regularisation parameter. 
 Our goal is to find the parameter value that  yields 20 to 30 non-zero SNP effects. 
@@
  These subsets are then analysed with LMM-Lasso (glmnet) with its regularisation parameter fixed to the value found previously. 
-A binary vector, of length the number of SNP, is recorded as the result. A one (zero) means the SNP had a non-zero (zero) effect size. 
+A binary vector, of length the number of SNP, is recorded as the result. 
+A one (zero) means the SNP had a non-zero (zero) effect size. 
 Calculating a SNP's statistical significance is now a simple task. We take the vector sum of the binary vectors over all 100 subsets. 
@@
 probabilities measure the strength of evidence for the SNPs to be in association with the trait. 
-
-By taking the vector sum over all 100 subsets and dividing each element in the sum by 100, we obtain the probability 
-estimates of the significance of the SNPs.
-
-A probability estimate of the significance of a SNP is now just a simple vector sum, over the 100 subsets,  of the binary vectors
-
-By taking the vector sum over all 100 subsets, for a particular replicate, and dividing each element sum by 100, we obtain a probability 
-estimate of significance.
-
-Now, for a replicate,  to calculate the statistical significance of the coefficient estimates obtained with LMM-Lasso (glmnet),
-
-estimate the statistical significance of a SNP, for a particular replicate, is obtained by taking the
-
-Now, for a particular replicate, we count the number of times LMM-Lasso (glmnet) gave a non-zero
-
-Now, for a particular replicate which was analysed with LMM-Lasso (glmnet), 
-we count the number of times
-
-we calculate the statistical significance of a SNP by counting the number of times its 
-coefficient was non-zero in the LMM-Lasso (glmnet) analyses of the 100 subsets of data.  A probability estimate of significance is obtained 
-by dividing this count by the number of subsetted analyses, which was 100.  Note that it is not 
+ Note that it is not 
 necessary to tune the regularisation   parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. Also, 
@@
 
-Then, for the LMM-Lasso (glmnet) analysis of a given replicate, to estimate the statistical significance of a SNP effect, we count 
-the number of times the SNP effect has non-zero effect size across the 100 subset LMM-Lasso (glmnet) analyses.  This number is then 
-divided by the number of subsets (which is 100 for the replicate)
+For bigRR, we modify our stability selection procedure slightly.  Where LMM-Lasso and glmnet shrink non-important effects to zero, 
+bigRR is a ridge regression method where all SNP are given an effect size.  XXXXX expand on this. ...
 
- to estimate the statistical significance of the SNP effects for a given replicate, the number of times
-
-For a replicate, a (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
-number of times a SNP has a non-zero effect size over all the subsets divided by the number of subsets (which was 100).
-
-Here, the link between 
-trait and genotype is preserved, with only the sample size being reduced.
-
-\textcolor{red}{
-To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, for a given scenario, 
-we begin by performing a preliminary analysis of a randomly chosen  replicate (i.e. trait and SNP genotype data)  from the 
- 100 available replicates for that scenario.  Our goal is to find a value of the regularisation parameter used by LMM-Lasso and glmnet 
- that yields 20 to 30 non-zero SNP effects. It is not necessary to tune the regularisation 
-  parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. 
-  Since we have the luxury of knowing the genetic conditions 
-under which the simulated trait data are generated, we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
- However, stability selection is a subsampling procedure that is robust to misspecification of parameters  \citep{meinshausen2010stability}. 
- The regularisation parameter can be tuned to any reasonable number of 
- non-zero effects.  Once we have a value for the regularisation parameter,
-
- We subsample repeatedly, without replacement, from a replicate. We draw 100 subsets of size $n/2$.  
-Each subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
-A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
-number of times the SNP have a non-zero effect size over all the replicates divided by the number of subsets (which was 100). }
-
-For bigRR, we modify our stability selection procedure slightly. 
 There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 

### 2019-07-27 BioinformaticsRevison/bioinformaticspaper.tex
Working of stability selection section v2

@@
  Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
- It is used here to estimate, empirically, the statistical significance of the results from LMM-Lasso, glment, and bigRR 
- of the simulated data. These three regularisation methods give the effect sizes of the SNPs, but not their significance, as their 
+ It is used here to estimate, empirically, the statistical significance of the results from LMM-Lasso, glment, and bigRR analyses 
+ of the simulated data. These three regularisation methods give the effect sizes of the SNPs, but not their significance as their 
  results.   Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost.
 
-The Stability selection procedure for LMM-Lasso and glmnet is as follows. 
-For a particular scenario, we begin by finding the value of the regularisation parameter that yields 20 to 30 non-zero SNP effects. 
+The stability selection procedure for LMM-Lasso and glmnet is as follows. 
+For a particular scenario, we begin by finding, via a binary search,  
+the value of the regularisation parameter that yields 20 to 30 non-zero SNP effects. 
 We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect from the analysis of a replicate in the simulation study.
@@
  This tuning was done for each of the six scenarios 
- but only for a single replicate within a scenario. It is not necessary to tune the regularisation parameter on every replicate 
+ but only for a single replicate, selected at random,  from within a scenario. It is not necessary to tune the regularisation parameter on every replicate 
  when replicates are generated under the same (trait, sample size, and number of SNP) conditions within a scenario.
 
-We did this for each  scenarios but not for every replicate within a scenario. It is unnecessary to set the regularisation parameter 
-for every replicate when replicates within a scenario are generated under the same conditions. 
-For a given scenario, we chose a replicate at random.  We then used a binary search procedure to set the regularisation parameter. 
-Then, for a replicate, we subsample repeatedly, 100 subsets of size $n/2$. These subsamples are drawn without replacement.  
-Here, the matching of trait to genotype is preserved in the subsamples. A subsample differs to the replicate in size only.
-
-Remembering that for each simulation scenario we generate 100 
-replicates of data under that scenario, we begin by choosing one of these replicates at random. 
-We then analyse this replicate with LMM-Lasso (glmnet) with different values of its  regularisation parameter. 
-Our goal is to find the parameter value that  yields 20 to 30 non-zero SNP effects. 
-This can be done via a simple binary search approach. 
- Then, for each replicate, we subsampled repeatedly, 100 subsets of size $n/2$.  A replicate is subsampled without replacement.  
- These subsets are then analysed with LMM-Lasso (glmnet) with its regularisation parameter fixed to the value found previously. 
-A binary vector, of length the number of SNP, is recorded as the result. 
-A one (zero) means the SNP had a non-zero (zero) effect size. 
-Calculating a SNP's statistical significance is now a simple task. We take the vector sum of the binary vectors over all 100 subsets. 
+ Once a suitable value for the regularisation parameter had been found, for the replicate  whose 
+ SNP results are to be assigned statistical significance, we subsample repeatedly, 100 data sets of size $n/2$. A larger number of 
+ data sets  and/or larger sized data sets could have been chosen here but we found these changes to have 
+ little impact on the final significance estimates. The subsamples are drawn without replacement. Also, 
+ the matching of trait to genotype is preserved in the subsamples. A subsample differs to the replicate in size only. 
+ The subsamples are analysed with LMM-Lasso (glmnet) with its regularisation parameter fixed to the tuned value found previously. 
+From the analysis of a subsample, a binary vector, of length the number of SNP, is recorded as the result where a one (zero) means the SNP had a non-zero (zero) effect size. 
+Calculating a SNP's statistical significance is now a simple task. We calculate the vector sum of the binary vectors over all 100 subsamples. 
 This vector sum will contain elements in the range of 0 to 100. By dividing each element in this vector sum by the number of 
-subsets upon which the sum was calculated, we obtain empirical probabilities.  
+subsamples upon which the sum is calculated (which is 100), we obtain empirical probabilities.  
  These 
 probabilities measure the strength of evidence for the SNPs to be in association with the trait.
- Note that it is not 
-necessary to tune the regularisation   parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. Also, 
-we have the luxury of knowing the genetic conditions 
-under which the simulated trait data are generated. we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
- However, stability selection is a subsampling procedure that is robust to parameter misspecification \citep{meinshausen2010stability}. 
- The regularisation parameter can be tuned to any reasonable number of 
- non-zero effects.
 
-For bigRR, we modify our stability selection procedure slightly.  Where LMM-Lasso and glmnet shrink non-important effects to zero, 
-bigRR is a ridge regression method where all SNP are given an effect size.  XXXXX expand on this. ...
+For bigRR, we implement stability selection in a different way. Unlike LMM-Lasso and glmnet, bigRR yields non-zero SNP effects 
+for all the SNPs. Also, there is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis 
+procedure. We still draw 100 subsamples of size $n/2$, without replacement and each subsample is analysed with bigRR. 
+However, from each analysis, we order the SNPs according to the absolute size of their SNP effect estimates from bigRR. 
+ A binary vector, of length the number of SNPs is then formed 
+where a one (zero) means the SNP is (not) in the top 20 ordered SNPs.  Calculating the significance of the SNPs then proceeds as described 
+above.
 

### 2019-07-28 BioinformaticsRevison/bioinformaticspaper.tex
Working of Generating simulated data section git push

@@
  The SNP genotypes are drawn, 
-according to the specifications of the scenario , from data collected from the 1000 Genome Project, version 3   \citep{10002010map}. 
-Across scenarios (see Results for details), the SNP data differs. Across replicates within a scenario, the SNP data are the same. For each scenario, 100 replicates are generated.
+according to the specifications of a particular simulation scenario , from data collected from the 1000 Genome Project, version 3   \citep{10002010map}.
+Six different scenarios are considered.
+These scenarios differ in  their sample size and number of
+SNPs (see Results for details).  Here, across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same. For each scenario, 100 replicates are generated.
 
@@
 above.
-
-There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
-As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
-SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
-A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
-number of times the SNP is recorded divided by the number of replicates.
 

### 2019-07-30 BioinformaticsRevison/bioinformaticspaper.tex
working on revisions - manhattan plots

@@
 \subsection{Generation of Simulation Data}
+\label{subsection:simdata}
 
@@
  The SNP genotypes are drawn, 
-according to the specifications of a particular simulation scenario , from data collected from the 1000 Genome Project, version 3   \citep{10002010map}.
+according to the specifications of a particular simulation scenario, from data collected from the 1000 Genome Project, version 3   \citep{10002010map}.
 Six different scenarios are considered.
@@
 
- Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
+ \textcolor{red}{Stability selection  \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
  It is used here to estimate, empirically, the statistical significance of the results from LMM-Lasso, glment, and bigRR analyses 
@@
  results.   Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost. 
+ }
 
+ \textcolor{red}{
 The stability selection procedure for LMM-Lasso and glmnet is as follows. 
@@
  when replicates are generated under the same (trait, sample size, and number of SNP) conditions within a scenario. 
+ }
 
+ \textcolor{red}{
  Once a suitable value for the regularisation parameter had been found, for the replicate  whose 
@@
 probabilities measure the strength of evidence for the SNPs to be in association with the trait. 
+}
 
-For bigRR, we implement stability selection in a different way. Unlike LMM-Lasso and glmnet, bigRR yields non-zero SNP effects 
+\textcolor{red}{
+For bigRR, stability selection is implemented in a different way. Unlike LMM-Lasso and glmnet, bigRR yields non-zero SNP effects 
 for all the SNPs. Also, there is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis 
-procedure. We still draw 100 subsamples of size $n/2$, without replacement and each subsample is analysed with bigRR. 
+procedure. We still draw 100 subsamples of size $n/2$, without replacement, and each subsample is analysed with bigRR. 
 However, from each analysis, we order the SNPs according to the absolute size of their SNP effect estimates from bigRR. 
@@
 above. 
+}
 
@@
 
-For each scenario, 100 replicates were generated. A single replicate consisted of SNP and quantitative trait data. 
+For each scenario, 100 replicates were generated. A single replicate consisted of SNP and quantitative trait data \textcolor{red}{(see Section 
+\ref{subsection:simdata})}. 
 Extra realism was introduced into the simulation study through the drawing of the SNP genotypes from the 1000 Genome Project, phase 3  \citep{10002010map}.
@@
 The number of randomly selected SNPs follows a Poisson distribution with mean 30. The size of the allelic effects 
- across the selected loci are equal.
+ across the selected loci are equal, \textcolor{red}{because the SNP genotypes AA, AB, and BB 
+are assigned the values -1, 0, and 1, respectively (Section \ref{subsection:simdata})} .
 
@@
  The only parameters requiring specification were the amount of available memory and number of CPUs for 
- Eagle and the number of chunks for MLMM. 
- Their results were also immediately 
+ Eagle and the number of chunks for MLMM.  \textcolor{red}{MLMM breaks its matrices into blocks or chunks, 
+ reducing its memory 
+ footprint but at the cost of increased computation}.  Their results were also immediately 
  interpretable. Their findings were the set of SNPs in strongest association with the trait. Each 
@@
 A  finding was counted as true if it was positioned within 40 kilobase pairs of the location of a true SNP.
+\textcolor{red}{This was the average (confidence interval) distance used by
+\cite{nicod2016genome} for determining if their findings were close to candidate genes and whose mouse data we analyse below}.
+
 When a replicate was analysed, we obtained an estimate of the power of the method by taking  the number of findings that were found to be  true and dividing by the 
@@
 that we focused on autosomal SNPs and our analyses were based on SNP genotypes  rather than SNP dosages. 
-In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation. We followed the 
+In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation  
+\textcolor{red}{\citep{nicod2016genome}} . We followed the 
 same empirical procedure but increased the number of permutations from 100 to 500 for more accurate thresholds.

### 2019-07-30 BioinformaticsRevison/supmaterials.tex
Revesion 1
working on revisions - manhattan plots

--- 2018-11-30 supmaterials.tex
@@
 
-\title{Eagle: Making multi-locus association mapping on a genome-wide scale routine}
+\title{Eagle: multi-locus association mapping on a genome-wide scale made routine}
 \author[1]{Andrew W. George}
-\author[1]{Arunas Verbyla}
-\author[2]{Joshua Bowden}
-
-\affil[1]{Data61, CSIRO, Australia.}
-\affil[2]{IM \&T, CSIRO, Australia.}
+\author[2]{Arunas Verbyla}
+\author[3]{Joshua Bowden}
+\affil[1]{Data61, CSIRO, Brisbane, 4102, Australia.}
+\affil[2]{Data61, CSIRO, Atherton, 4883, Australia.}
+\affil[3]{IM \&T, CSIRO, Brisbane, 4067, Australia}
+\date{}
 
@@
 
+%% Manhattan plots of real mouse data analysis
+
 %% This table appears in Figure 1 in paper. It's included here in case I need to regenerate the table. 
@@
 
+\begin{figure}
+\caption{Manhattan plot of single-locus analysis of the mouse data with FaST-LMM for the traits BMC Kurt,  BMC Kurt N, BMC Max, and BMC Max N}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{1.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\caption{Manhattan plot of single-locus analysis of the mouse data with FaST-LMM for the traits BMC Mean,  BMC Mean N, BMC Median, and BMC Perim}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{5.jpeg}
+\end{center}
+\end{figure}
+
 \end{document}

### 2019-08-01 BioinformaticsRevison/bioinformaticspaper.tex
significant changes - FDR -> FPR. Revising heat map, adding Manhattan plots to sup materials, added Bonf correction to FaST-LMM etc

@@
 Of these traits, 
-43 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
-43 traits that were the focus of our real data analyses. As in the original study  \citep{nicod2016genome}, our analyses 
+45 yielded SNP-trait associations that could be corroborated through other independent published work. It was these 
+45 traits that were the focus of our real data analyses. As in the original study  \citep{nicod2016genome}, our analyses 
 were based on the adjusted traits.
@@
 
-\subsection{Power and False Discovery Rates}
+\subsection{Power and False Positive Rates}
 
 Here, we answer the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations.
-We do this by estimating the power and false discovery rates of Eagle and the other methods for the six scenarios.
+We do this by estimating the power and false positve rates of Eagle and the other methods for the six scenarios.
 Since, for a replicate, we knew which SNPs were assigned additive effects, we knew the SNPs that were in true association with the trait.
@@
 number of true SNPs.
-We also obtained an estimate of a method's false discovery rate.
-It is the number of findings that were found to be false divided by the number of (true and false) findings found by the method.
+We also obtained an estimate of a method's false positive rate.
+It is the number of findings that were found to be false divided by the number of true SNPs.
 Both these estimates varied with replicate.
-The power (false discovery rate) of a method, for a scenario, was found by taking the median of the power (false discovery rate) estimates over the 100 replicates.
+The power (false positive rate) of a method, for a scenario, was found by taking the median of the power (false positive rate) estimates over the 100 replicates.
 
-The power and false discovery rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure \ref{supfigpowermultiple}.
+The power and false positive rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure \ref{supfigpowermultiple}.
 We restricted our attention to these scenarios because not all
@@
   Each plot contains 
-single points and power curves. The single points are the power and false discovery rates for Eagle and MLMM.
+single points and power curves. The single points are the power and false positive rates for Eagle and MLMM.
 These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
@@
 significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
-the false discovery rate as the significance threshold  is adjusted. 
-The power and false discovery rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
+the false positive rate as the significance threshold  is adjusted. 
+The power and false positive rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
  Figure \ref{figpowersingle}.
@@
 \setstretch{1.6}{
-Power verse false discovery rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
+Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
 FaST-LMM was run where all the SNP data are used 
@@
 well.  Of the multi-locus methods, Eagle had the highest power
-while keeping its false discovery rate low (Supplementary Figure \ref{supfigpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
+while keeping its false positive rate low (Supplementary Figure \ref{supfigpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
 that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
@@
 
-We repeated the single-locus analyses as first performed \citep{nicod2016genome}, except 
-that we focused on autosomal SNPs and our analyses were based on SNP genotypes  rather than SNP dosages. 
-In the original analysis, a genome-wide threshold that gave a false discovery rate of 5\%, was found via permutation  
-\textcolor{red}{\citep{nicod2016genome}} . We followed the 
-same empirical procedure but increased the number of permutations from 100 to 500 for more accurate thresholds.
+We repeated the single-locus analyses as first performed \citep{nicod2016genome} but with a few exceptions. We 
+focused on autosomal SNPs, our analyses were based on SNP genotypes  rather than SNP dosages, 
+we sought to control the false positive rate not false discovery rate of the methods, and significance thresholds were 
+found empirically via permutation \citep{doerge1996permutation}.
 
@@
 our analyses here on these same 45 traits.
-Overall,  FaST-LMM, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 50, 37, 67, and 106, SNP-trait findings, respectively, across 39 traits.
+Overall,  FaST-LMM$^{Bonf}$,  FaST-LMM$^{perm}$, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 50, 37, 67, and 106, SNP-trait findings, respectively, across 39 traits.
 No associations were found by FaST-LMM and Eagle for the other six traits.
@@
 However, the relative results of the simulation study remain true.
-For similar false discover rates, Eagle is superior to single-locus association mapping.
-As a case in point, here FaST-LMM found 50 SNP-trait associations with a false discovery rate of 5\%.
-Eagle, with the same false discovery rate (Eagle$^{optimal}$) found 106 SNP-trait associations, more
+For similar false positve rates, Eagle is superior to single-locus association mapping.
+As a case in point, here FaST-LMM found 50 SNP-trait associations with a false positive rate of 5\%.
+Eagle, with the same false positive rate (Eagle$^{optimal}$) found 106 SNP-trait associations, more
 than a 100\% increase in findings.
@@
 \caption{
-\setstretch{1.6}{
+\setstretch{1.1}{
 Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
-multi-locus method
-Eagle. Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on 
+multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated via the Bonferroni correction 
+(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$). 
+Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on 
 the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate 
@@
 %% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
-\includegraphics[width=15cm, height=15cm]{mouseresults.eps}
+\includegraphics[width=12cm, height=12cm]{mouseresults.eps}
 \end{center}

### 2019-08-02 BioinformaticsRevison/bioinformaticspaper.tex
Review 1 revisions almost complete - waiting on analyses

@@
 %%\begin{center}
-\includegraphics[width=15cm, height=20cm]{power2main.eps}
+\includegraphics[width=10cm, height=15cm]{power2main.eps}
 %%\end{center}
@@
 
-We repeated the single-locus analyses as first performed \citep{nicod2016genome} but with a few exceptions. We 
+We repeated the single-locus analyses as first performed \citep{nicod2016genome} but some exceptions. We 
 focused on autosomal SNPs, our analyses were based on SNP genotypes  rather than SNP dosages, 
@@
 under this setting (Eagle$^{default}$).
-An alternate  \citep{chen2008extended} , less conservative way of setting $\gamma$ is to  let $\gamma = 1 - \frac{1}{(2\kappa)}$ with $\kappa=\frac{log(L)}{log(n_g)}$ where
+An alternate  \citep{chen2008extended}, less conservative way of setting $\gamma$ is to  let $\gamma = 1 - \frac{1}{(2\kappa)}$ with $\kappa=\frac{log(L)}{log(n_g)}$ where
  $L$ is the number of loci that span the genome, and $n_g$ is the number of individuals/groups/lines/strains in the study
@@
 
-The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}. The mouse study recorded
+The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}, 
+\textcolor{red}{
+with the Manhattan plots of the 
+single-locus analysis shown in Supplementary Figure \ref{supmanhattan}.} 
+The mouse study recorded
 measurements on 200 traits. Of these, in the original study, 45 were able to have their findings  corroborated by previously published work. We focused 
 our analyses here on these same 45 traits.
-Overall,  FaST-LMM$^{Bonf}$,  FaST-LMM$^{perm}$, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 50, 37, 67, and 106, SNP-trait findings, respectively, across 39 traits.
+Overall,  FaST-LMM$^{Bonf}$,  FaST-LMM$^{perm}$, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 47, 68, 37, 67, 106, SNP-trait findings, respectively, across 39 traits.
 No associations were found by FaST-LMM and Eagle for the other six traits.
@@
 For similar false positve rates, Eagle is superior to single-locus association mapping.
-As a case in point, here FaST-LMM found 50 SNP-trait associations with a false positive rate of 5\%.
+As a case in point, here FaST-LMM$^{perm}$ found 68 SNP-trait associations with a false positive rate of 5\%.
 Eagle, with the same false positive rate (Eagle$^{optimal}$) found 106 SNP-trait associations, more
-than a 100\% increase in findings.
+than a 50\% increase in findings.
 
@@
 Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
-multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated via the Bonferroni correction 
-(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$). 
+multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
+(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$), at the 5\% significance level. 
 Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on 
@@
 Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
-with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.
+with single-locus association mapping, the method of choice. \textcolor{red}{Furthermore, because we converted the SNP dosages of the 
+original study into genotypes and focusing our analyses on these data, the validity of the extra findings can be confirmed 
+against the original findings}. These extra findings were all true.
 
@@
 We would like to thank the IM\&T Scientific Computing group at the Commonwealth Scientific and Industrial Research Organisation (CSIRO) for providing access and support to the CSIRO Accelerator Cluster upon which all computation was performed.
+We also gratefully acknowledge the comments of two anonymous reviewers whose suggestions have helped "polish" this paper.
 

### 2019-08-02 BioinformaticsRevison/supmaterials.tex
significant changes - FDR -> FPR. Revising heat map, adding Manhattan plots to sup materials, added Bonf correction to FaST-LMM etc
Review 1 revisions almost complete - waiting on analyses

@@
 \begin{figure}
-\caption{Manhattan plot of single-locus analysis of the mouse data with FaST-LMM for the traits BMC Kurt,  BMC Kurt N, BMC Max, and BMC Max N}
+\caption{Manhattan plots of the single-locus analysis of the mouse data with FaST-LMM. 
+Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
+(pink horizontal line) and permutation (blue horizontal line), at a 5\% significance level. 
+}
+\label{supmanhattan}
 \begin{center}
@@
 \begin{figure}
-\caption{Manhattan plot of single-locus analysis of the mouse data with FaST-LMM for the traits BMC Mean,  BMC Mean N, BMC Median, and BMC Perim}
 \begin{center}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{9.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{13.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{17.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{21.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{25.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{29.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{33.jpeg}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=15cm, height=15cm]{37.jpeg}
+\end{center}
+\end{figure}
+
 \end{document}

### 2019-08-02 AM+/Plots_for_Paper/power_plots.R

@@ -211,7 +211,7 @@ p <- p + theme(panel.spacing = unit(2, "lines"))
 
 ## specify xlab and ylab
 p <- p  + ylab(bquote("Power")) + 
-  xlab(bquote('False Discovery Rate'))
+  xlab(bquote('False Positive Rate'))
 
 
 ##  change x and y labels size and bold
@@ -380,7 +380,7 @@ p <- p + theme(panel.spacing = unit(2, "lines"))
 
 ## specify xlab and ylab
 p <- p  + ylab(bquote("Power")) + 
-  xlab(bquote('False Discovery Rate'))
+  xlab(bquote('False Positive Rate'))
 
 
 ##  change x and y labels size and bold

### 2019-08-07 BioinformaticsRevison/bioinformaticspaper.tex
Adding FarmCPU discussion point

@@
 
+FarmCPU (ref) is a multi-locus association mapping method that was brought to our attention 
+by a reviewer.  Unlike other LMM-based methods, it makes use of  two separate models for analysis, a fixed effect model and 
+a random effect model.  The fixed effect model measures the strength of association between the SNPs and trait. 
+This is done on a locus-by-locus basis but the model includes psuedo quantitative trait nucleotides (QTN), making it a multi-locus model. 
+The random effect model is used to search for the optimal number and  size of "bins".  By "binning" the SNP data and identifying the 
+SNP within a bin with the highest p-value (which is obtained from the fixed effect model), psuedo QTN are identified and feed back into the fixed effect model. The procedure stops 
+once the number and size of the bins stabilises.
+
+We analysed the mouse data with FarmCPU.
+
+By "binning" the SNP data based on map location,
+
+The 
+random effect model identifies the optimal set of psuedo quantitiative trait nucleotides. The method is cyclic. Findings from the fixed effect model
+
+This is done on a locus-by-locus basis but the model includes psuedo quantitative trait nucleotides, making it a multi-locus model. 
+The random effect model contains a random effect and residual. The covariance matrix for the random effect is formed from
+
+After completing our analyses, a new and different multi-locus method became available, FarmCPU (ref). 
+FarmCPU is an iterative procedure, built around the repeated fitting of two models, a fixed effect model and a random effect model.
+
+The fixed effect model measures the strength of association of a SNP with a trait, on a locus by locus basis. It controls for false positives 
+by including QTNs as additional fixed effects.  The random effects model
+
+is then used to optimize the binning of the genome
+
+One a p value has been obtained for the SNPs,
+
+The fixed effect model measures the statistical significance of the association between a SNP and trait. 
+ It does this on a locus-by-locus basis.  Based on the SNPs significance and genome LD structure ,
+
+The fixed effect model is fitted on a locus-by-locus basis. It contains the SNP effect and any identified QTNs.
+
+The fixed effect model contains any identified QTNs
+
+It is an iterative approach. At each iteration, a fixed effect and random effect model are fitted. The fixed effect model is fitted on a 
+locus-by-locus basis. This can be done efficiently because the model only contains fixed effects.
+
+ It is based on the fitting of two models, a fixed effect model and a random effect model. 
+The fixed effect model is used to identify SNP in association with the trait. The random effect model is used to fine-tune the 
+number of QTN.
+
+The fixed effect model is fitted on a locus-by-locus basis. It contains quantitative trait nucleotides (QTNs) and
+
+ It is build around two models, a fixed effect model and a random effect model.
+
+FarmCPU is an iterative algorithm where for each iteration, both a fixed effect and random effect model are fitted. 
+Single marker tests are performed with the fixed effect model.
+
+The fixed effect model
+
+While preparing our manuscript, a promising new multi-locus association mapping method became available, FarmCPU (ref).
+
+During the course of preparing our manscript, three new multi-locus associaiton mapping methods became available. These are
+
+While preparing our manuscript, three new multi-locus association mapping methods became available. These are 
+mrMLM, FarmCPU, and BLINK.  mrMLM is an R package and available from CRAN, FarmCPU is written in R but is not in package form,  and 
+BLINK is available in R or as stand alone C code.
+
 Over the coming years, computationally, the demand placed upon association mapping methods is going to increase. 

### 2019-08-08 BioinformaticsRevison/bioinformaticspaper.tex
Adding FarmCPU discussion point v2

--- 2019-08-02 BioinformaticsRevison/bioinformaticspaper.tex
@@
 
+FarmCPU (ref) is a multi-locus association mapping method that was brought to our attention 
+by a reviewer. It is different to other linear mixed model approaches in that it makes use of two separate models for analysis, 
+a fixed effects model and a random effects model. FarmCPU iterates over these two models. 
+The measures of association, obtained from the fixed 
+effect model are used to define pseudo-QTN from the fitting of the random effect model. Conversely, by placing the SNP into "bins" and 
+optimising the random effects model for bin number and size, pseudo-QTN  are identified and passed back to the fixed effect model. 
+It is an intriguing approach.  We were interested in comparing Eagle to FarmCPU so we reanalysed the mouse data with FarmCPU. 
+Like Eagle, FarmCPU found 107 SNP-trait associations but the findings were the same for only 11 of the traits.
+For 14 (14) traits, Eagle (FarmCPU) found more associations.
+Furthermore, unlike for Eagle, FarmCPU found X associations that could not be confirmed from the original study's findings.
+Further work would be needed to determine the true validity of these new results.
+An area where the two
+methods differ significantly is in their implementation.  Eagle has been developed for ease-of-use. Eagle is an R package but 
+its browser-based GUI makes it accessible to all users.  In contrast,  FarmCPU exists as a set of downloadable R scripts.
+
 Over the coming years, computationally, the demand placed upon association mapping methods is going to increase. 

### 2019-08-12 BioinformaticsRevison/bioinformaticspaper.tex
STill making changes

--- 2019-08-02 BioinformaticsRevison/bioinformaticspaper.tex
@@
 with single-locus association mapping, the method of choice. \textcolor{red}{Furthermore, because we converted the SNP dosages of the 
-original study into genotypes and focusing our analyses on these data, the validity of the extra findings can be confirmed 
-against the original findings}. These extra findings were all true.
+original study into genotypes and focused our analyses on these data, the validity of the extra findings were able to  be confirmed 
+against the original findings}. These extra findings were all found to be true.
 
@@
 models,  run time and memory usage can become limiting factors for analysis.
+
+\textcolor{red}{
+Upon submitting our paper for review, a more recent multi-locus association mapping method, FarmCPU(), was brought to our attention. 
+It is an unorthodox approach. Instead of working with a single model, results are passed back and forth between two models, 
+a fixed effects model and a random effects model. Measures of association are obtained from the fixed effect model, which in turn 
+help define pseudo-QTN from the random effects model. Conversely, pseudo-QTN found from the random effects model are passed 
+back to the fixed effect model to better refine the measures of association.  The method involves multiple rounds of genome-wide 
+testing. Out of interest, we reanalysed the mouse data with FarmCPU.  Runtimes were around five times longer for FarmCPU than 
+Eagle.  Interestingly, FarmCPU found the same number of associations, 106,  as Eagle$^{optimal}$, but the findings were only the 
+same for 11 of the traits.  For 14 (14) traits, Eagle (FarmCPU) found more associations.  Where the new findings from Eagle could 
+be confirmed, of the 18 new associations found by FarmCPU, we were unable to confirm 10 using the results from the original study. 
+The two methods differ significantly in their implementation. FarmCPU is not an R package but a set of R scripts. We had 
+no problem in using the scripts but this may not be true for non-R users. In contrast, Eagle has been developed for ease-of-use.
+Eagle is an R package but its browser-based GUI makes it accessible to all.
+}
 

### 2019-08-12 BioinformaticsRevison/supmaterials.tex
STill making changes

@@
 
-\hspace{1mm} SNPs fitted $^e$      &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.15cm]
-
-\hspace{1mm}  Selection type             & Model   & Variable  &  Variable &  Variable  &  Model  &  Variable & Variable  & Variable \\  [0.15cm]
-
-\hspace{1mm} Threshold free      &    Yes     &  No    &  No         &    No               &     Yes       &  No              &  No   & No      \\  [0.15cm]  \hline
+\hspace{1mm} SNPs fitted $^e$      &    All/multiple      &    All       &   All        &      All          &  Multiple          &  Multiple              &   Single   &  Single   \\  [0.15cm]  \hline
 

### 2019-08-12 BioinformaticsRevison/bioinformaticspaper.tex
getting close to final revisions

@@
 we sought to control the false positive rate not false discovery rate of the methods, and significance thresholds were 
-found empirically via permutation \citep{doerge1996permutation}.
+found empirically via permutation \textcolor{red}{\citep{doerge1996permutation}}.
 
@@
  As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
-Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
+Also, when analysing the mouse data, Eagle found more than 50\% the SNP-trait associations than 
 with single-locus association mapping, the method of choice. \textcolor{red}{Furthermore, because we converted the SNP dosages of the 

### 2019-08-15 BioinformaticsRevison/bioinformaticspaper.tex
Bioinfo revisions complete

@@
 \textcolor{red}{
-Upon submitting our paper for review, a more recent multi-locus association mapping method, FarmCPU(), was brought to our attention. 
-It is an unorthodox approach. Instead of working with a single model, results are passed back and forth between two models, 
+Upon submitting our paper for review, a more recent multi-locus association mapping method, FarmCPU \citep{liu2016iterative}, was brought to our attention. 
+It is an unorthoﬂdox approach. Instead of working with a single model, results are passed back and forth between two models, 
 a fixed effects model and a random effects model. Measures of association are obtained from the fixed effect model, which in turn 
@@
 be confirmed, of the 18 new associations found by FarmCPU, we were unable to confirm 10 using the results from the original study. 
-The two methods differ significantly in their implementation. FarmCPU is not an R package but a set of R scripts. We had 
+Also, the two methods differ significantly in their implementation. FarmCPU is not an R package but a set of R scripts. We had 
 no problem in using the scripts but this may not be true for non-R users. In contrast, Eagle has been developed for ease-of-use.
-Eagle is an R package but its browser-based GUI makes it accessible to all.
+Its browser-based GUI makes it accessible to all.
 }

### 2019-08-15 BIOINF-2019-0261-Revision/bioinformaticspaper.tex
Final version of revised paper v2

--- 2019-08-15 BioinformaticsRevison/bioinformaticspaper.tex
@@
 \\
-{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a web-based Graphical User Interface (GUI) for users less familiar with R.
+{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R.
 It is freely available via the CRAN website  at https://cran.r-project.org.
-\\
+Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
 {\bf Contact:} andrew.george@csiro.au
@@
 calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
-a link to our server for demonstrating Eagle on some test data are available on the Eagle website. Eagle is available for download from the CRAN website.
+a link to our server for demonstrating Eagle on some test data are available on the Eagle website http://eagle.r-forge.r-project.org. Eagle is available for download from the CRAN website.
 
@@
 Eagle.  Interestingly, FarmCPU found the same number of associations, 106,  as Eagle$^{optimal}$, but the findings were only the 
-same for 11 of the traits.  For 14 (14) traits, Eagle (FarmCPU) found more associations.  Where the new findings from Eagle could 
+same for 11 of the traits.  For 14 traits, Eagle found more associations. For the other 14 traits, FarmCPU found more associations. 
+Where the new findings from Eagle could 
 be confirmed, of the 18 new associations found by FarmCPU, we were unable to confirm 10 using the results from the original study. 

### 2019-08-19 BIOINF-2019-0261-Revision/bioinformaticspaper.tex
still tring to get bioinformatics in the right latex format

@@
 \usepackage{xr}
-\bibliographystyle{natbib}
+
 \renewcommand{\baselinestretch}{1.4} 
@@
 \linenumbers
-
-%% external tex files
-%%\externaldocument{figure1_runtimes}
-%%\externaldocument{figure2_powerFDRmultiple}
-%%\externaldocument{figure3_mouse}
-
-\externaldocument{supmaterials}
 
@@
 We also analyse data from a published mouse study.
-Eagle found over 100\% more validated findings than the state-of-the-art single-locus method.
+Eagle found over 50\% more validated findings than the state-of-the-art single-locus method.
 \\
@@
 Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 
-is given in Supplementary Table \ref{suptabsummary} (see Methods for further details).
+is given in Supplementary Table 1 (see Methods for further details).
 
@@
 
-The power and false positive rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure \ref{supfigpowermultiple}.
+The power and false positive rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure 1.
 We restricted our attention to these scenarios because not all
@@
 well.  Of the multi-locus methods, Eagle had the highest power
-while keeping its false positive rate low (Supplementary Figure \ref{supfigpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
+while keeping its false positive rate low (Supplementary Figure 1). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
 that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
@@
 Memory usage
-for many of the computer programs/packages was the limiting factor (see Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
+for many of the computer programs/packages was the limiting factor (see Supplementary Figure 2).  The single-locus program GEMMA was by 
 far the most memory efficient. Not surprisingly, the multi-locus programs were memory intensive. Most required in 
@@
 Eagle is compared against five other multi-locus programs/packages (A) and two single-locus programs (B).
-The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multi-locus 
-programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. 
-The median run times for the programs/packages across the scenarios are given in the table (C). The entries in a blue font 
-correspond to the lowest run-time for a scenario. 
+The x- and y-axes are on a log scale for improved aesthetics.
+Eagle has the lowest run-times of the multi-locus programs/packages, sometimes by orders of magnitude.
+Eagle can even produce results faster than single-locus programs.
+The median run times for the programs/packages across the scenarios are given in the table (C).
+The entries in a blue font  correspond to the lowest run-time for a scenario.
 FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.
@@
 with the Manhattan plots of the 
-single-locus analysis shown in Supplementary Figure \ref{supmanhattan}.} 
+single-locus analysis shown in Supplementary Figure 3.} 
 The mouse study recorded
@@
 on the SNP genotype data and the original study being based on SNP dosage data. Eagle$^{alt}$ found two  and 
-Eagle$^{optimal}$ found seven new findings (Supplementary Table \ref{suptabnew}).  These new findings all involved SNPs whose association had been confirmed for other related traits in the original 
+Eagle$^{optimal}$ found seven new findings (Supplementary Table 2).  These new findings all involved SNPs whose association had been confirmed for other related traits in the original 
 study.
@@
 (FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$), at the 5\% significance level.
-Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on 
-the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate 
-of 5\% (Eagle$^{optimal}$).   The number of SNP-trait associations found are reported in the cells. }
+Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate of 5\% (Eagle$^{optimal}$).
+The number of SNP-trait associations found are reported in the cells.
+}
 }
@@
 
+\bibliographystyle{natbib}
 %%\bibliographystyle{naturemag}

### 2019-08-19 BIOINF-2019-0261-RevisionV2/bioinformaticspaper.tex
still tring to get bioinformatics in the right latex format

--- 2019-08-19 BIOINF-2019-0261-Revision/bioinformaticspaper.tex
@@
-%%\documentclass{nature}
-\documentclass{article}
-\usepackage{xr}
+\documentclass{bioinfo}
+\copyrightyear{2019} \pubyear{2019}
 
-\renewcommand{\baselinestretch}{1.4} 
-\usepackage{graphicx}
-\usepackage{bm}
-\usepackage{lineno}
-\usepackage{natbib}
-\usepackage{setspace}
-\usepackage{authblk}
-\usepackage{xcolor}
+%%\usepackage{xr}
+%%\renewcommand{\baselinestretch}{1.4} 
+%%\usepackage{graphicx}
+%%\usepackage{bm}
+%%\usepackage{lineno}
+%%\usepackage{natbib}
+%%\usepackage{setspace}
+%%\usepackage{authblk}
+%%\usepackage{xcolor}
 
-\linenumbers
+\access{Advance Access Publication Date: Day Month Year}
+\appnotes{Original Paper}
 
 \begin{document}
+\firstpage{1}
 
-\title{Eagle: multi-locus association mapping on a genome-wide scale made routine}
-\author[1]{Andrew W. George}
-\author[2]{Arunas Verbyla}
-\author[3]{Joshua Bowden}
-\affil[1]{Data61, CSIRO, Brisbane, 4102, Australia.}
-\affil[2]{Data61, CSIRO, Atherton, 4883, Australia.}
-\affil[3]{IM \&T, CSIRO, Brisbane, 4067, Australia}
-\date{}
+\subtitle{Genetics and Population Analysis}
 
-\maketitle
+\title[Eagle: making multi-locus association mapping routine]{Eagle: multi-locus association mapping on a genome-wide scale made routine}
+\author[Sample \textit{et~al}.]{Andrew W. George\,$^{\text{\sfb 1,}*}$, Arunas Verbyla\,$^{\text{\sfb 2}}$ and Joshua Bowden\,$^{\text{\sfb 3,}}$}
+\address{$^{\text{\sf 1}}$Data61, CSIRO, Brisbane, 4102, Australia and \\
+$^{\text{\sf 2}}$Data61, CSIRO, Atherton, 4883, Australia and \\
+$^{\text{\sf 3}}$IM\&T, CSIRO, Brisbane, 4067, Australia.
+}
+
+\corresp{$^\ast$To whom correspondence should be addressed}
+
+\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
+
+\editor{Associate Editor: Alfonso Valencia}
+
+%%\linenumbers
 
@@
 
-\begin{abstract} \noindent
-\setstretch{1.6}{
-{\bf Motivation:} We present Eagle, a new method for multi-locus association mapping.
+\abstract{\textbf{Motivation:} We present Eagle, a new method for multi-locus association mapping.
 The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice.
@@
 \\
-{\bf Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods.
+\textbf{Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods.
 We also analyse data from a published mouse study.
@@
 \\
-{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R.
+\textbf{Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R.
 It is freely available via the CRAN website  at https://cran.r-project.org.
 Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
-{\bf Contact:} andrew.george@csiro.au
-}
-\end{abstract}
+\textbf{Contact:} \href{andrew.george@csiro.au}\\
+\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
+online.}
+
+\maketitle
 
 \section{Introduction}
+
+\citep{Boffelli03}
+
 Over the past decade,  genome-wide association studies (GWASs) have changed considerably in both their analysis and design. Early studies
@@
 
-\begin{figure}
-\caption{
-\setstretch{1.6}{
-Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
+%%\begin{figure}
+%%\caption{
+%%\setstretch{1.6}{
+%%Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
+%%FaST-LMM was run where all the SNP data are used 
+%%to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
+%%estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
+%%}
+%%\vspace{-3cm}
+%%\label{figpowersingle}
+%%\begin{center}
+%%\includegraphics[width=10cm, height=15cm]{power2main.eps}
+%%\end{center}
+
+%%\end{figure}
+
+\begin{figure}[!tpb]%figure2
+\centerline{\includegraphics[width=8cm, height=10cm]{power2main.eps}}
+\vspace{-2cm}
+\caption{Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
 FaST-LMM was run where all the SNP data are used 
@@
 estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods. }
-}
-\vspace{-3cm}
 \label{figpowersingle}
-%%\begin{center}
-\includegraphics[width=10cm, height=15cm]{power2main.eps}
-%%\end{center}
-
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[!tpb]
+ \centerline{\includegraphics[width=0.45\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}}
  \caption{
-\setstretch{1.6}{
 Median run times, in minutes, for the analysis of simulation study data from the six scenarios.
@@
  FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
- }
-\vspace{1cm}
-%% plot formed via PowerPoint - ~/AM-Paper/Figure_power.pptx
-%% script on mac
-%% ~/AM-Paper/AM+/Plots_for_Paper/time_and_memory_plot_for_paper.R
 \label{fig_time}
-\centering
-    \includegraphics[width=1.2\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[!tpb]
+%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
+\centerline{\includegraphics[width=8cm, height=8cm]{mouseresults.eps}}
 \caption{
-\setstretch{1.1}{
 Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
@@
 }
-}
 \label{figmouse}
-\begin{center}
-%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
-\includegraphics[width=12cm, height=12cm]{mouseresults.eps}
-\end{center}
 
@@
 
-\bibliographystyle{natbib}
-%%\bibliographystyle{naturemag}
-\bibliography{biblibrary}
+\begin{thebibliography}{}
+
+\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
+Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
+
+\bibitem[Ball, 2001]{ball2001bayesian}
+Ball,R.D. (2001) Bayesian methods for quantitative trait loci mapping based on model
+  selection: approximate analysis using the bayesian information criterion.
+\textit{Genetics}, \textbf{159}, 1351--1364.
+
+\bibitem[Boulesteix \textit{et~al.}, 2012]{boulesteix2012overview}
+Boulesteix,A.-L.  \textit{et~al}. (2012)
+\newblock Overview of random forest methodology and practical guidance with
+  emphasis on computational biology and bioinformatics.
+\newblock {\em Wiley Interdisciplinary Reviews: Data Mining and Knowledge
+  Discovery\/}, {\bf 2}, 493--507.
+
+\bibitem[Broman and Speed, 2002]{broman2002model}
+Broman,K.W. and Speed,T.P. (2002)
+\newblock A model selection approach for the identification of quantitative
+  trait loci in experimental crosses.
+\newblock {\em Journal of the Royal Statistical Society: Series B (Statistical
+  Methodology)\/}, {\bf 64}, 641--656.
+
+\bibitem[Browning, 2008]{browning2008presto}
+Browning,B.L. (2008)
+\newblock {PRESTO}: rapid calculation of order statistic distributions and
+  multiple-testing adjusted {P}-values via permutation for one and two-stage
+  genetic association studies.
+\newblock {\em BMC Bioinformatics\/}, {\bf 9}, 309.
+
+\bibitem[Cardon and Palmer, 2003]{cardon2003population}
+Cardon,L.R. and Palmer,L.J. (2003)
+\newblock Population stratification and spurious allelic association.
+\newblock {\em The Lancet\/}, {\bf 361}, 598--604.
+
+\bibitem[Chen and Chen(2008)]{chen2008extended}
+Chen,J. and Chen,Z. (2008)
+\newblock Extended {B}ayesian information criteria for model selection with
+  large model spaces.
+\newblock {\em Biometrika\/}, {\bf 95}, 759--771.
+
+\bibitem[Cho \textit{et~al}., 2010]{cho2010joint}
+Cho,S. \textit{et~al}. (2010) Joint identification of multiple genetic variants via elastic-net
+  variable selection in a genome-wide association analysis. \textit{Annals of Human Genetics}, {\bf 74}(5), 416--428.
+
+\bibitem[Consortium {\em et~al.}, 2010]{10002010map}
+Consortium,G.P. {\em et~al.} (2010)
+\newblock A map of human genome variation from population-scale sequencing.
+\newblock {\em Nature\/}, {\bf 467}, 1061--1073.
+
+\bibitem[de~Bakker {\em et~al.}, 2005]{de2005efficiency}
+de~Bakker,P.I. \textit{et al.} (2005)
+\newblock Efficiency and power in genetic association studies.
+\newblock {\em Nature Genetics\/}, {\bf 37}, 1217--1223.
+
+\bibitem[Doerge and Churchill, 1996]{doerge1996permutation}
+Doerge,R.W. and Churchill,G.A. (1996)
+\newblock Permutation tests for multiple loci affecting a quantitative
+  character.
+\newblock {\em Genetics\/}, {\bf 142}, 285--294.
+
+\bibitem[Friedman {\em et~al.}, 2010]{Friedman2010glmnet}
+Friedman,J. \textit{et al.} (2010)
+\newblock Regularization {P}aths for {G}eneralized {L}inear {M}odels via
+  {C}oordinate {D}escent.
+\newblock {\em Journal of Statistical Software\/}, {\bf 33}, 1--22.
+
+\bibitem[Gogel {\em et~al.}, 2018]{gogel2018comparison}
+Gogel, B. \textit{et al.} (2018)
+\newblock Comparison of a one-and two-stage mixed model analysis of
+  {A}ustralia's {N}ational {V}ariety {T}rial {S}outhern {R}egion wheat data.
+\newblock {\em Euphytica\/}, {\bf 214}, 44.
+
+\bibitem[Golub and Van~Loan, 2012]{golub2012matrix}
+Golub,G.H. and Van~Loan,C.F. (2012)
+\newblock {\em Matrix {C}omputations\/}, volume~3.
+\newblock JHU Press.
+
+\bibitem[Gudbjartsson {\em et~al.}, 2015]{gudbjartsson2015large}
+Gudbjartsson,D.F.  {\em et~al.} (2015)
+\newblock Large-scale whole-genome sequencing of the {I}celandic population.
+\newblock {\em Nature Genetics\/}, {\bf 47}, 435--444.
+
+\bibitem[Kang {\em et~al.}, 2008]{kang2008efficient}
+Kang,H.M. \textit{et al.} (2008)
+\newblock Efficient control of population structure in model organism
+  association mapping.
+\newblock {\em Genetics\/}, {\bf 178}, 1709--1723.
+
+\bibitem[Li and Ji, 2005]{li2005adjusting}
+Li,J. and Ji,L. (2005)
+\newblock Adjusting multiple testing in multilocus analyses using the
+  eigenvalues of a correlation matrix.
+\newblock {\em Heredity\/}, {\bf 95}, 221--227.
+
+\bibitem[Lippert {\em et~al.}, 2011]{lippert2011fast}
+Lippert,C. \textit{et al.} (2011)
+\newblock {FaST} linear mixed models for genome-wide association studies.
+\newblock {\em Nature Methods\/}, {\bf 8}, 833--835.
+
+\bibitem[Liu {\em et~al.}, 2016]{liu2016iterative}
+Liu, X. \textit{et al.} (2016)
+\newblock Iterative usage of fixed and random effect models for powerful and
+  efficient genome-wide association studies.
+\newblock {\em PLoS Genetics\/}, {\bf 12}, e1005767.
+
+\bibitem[Lockhart {\em et~al.}, 2014]{lockhart2014significance}
+Lockhart,R. \textit{et al.}  (2014)
+\newblock A significance test for the {L}asso.
+\newblock {\em Annals of Statistics\/}, {\bf 42}, 2138--2139.
+
+\bibitem[Long {\em et~al.}, 2017]{long2017whole}
+LongT.  {\em et~al.} (2017)
+\newblock Whole-genome sequencing identifies common-to-rare variants associated
+  with human blood metabolites.
+\newblock {\em Nature Genetics\/}, {\bf 49}, 568--578.
+
+\bibitem[Meinshausen and B{\"u}hlmann, 2010]{meinshausen2010stability}
+Meinshausen,N. and B{\"u}hlmann,P. (2010)
+\newblock Stability {S}election.
+\newblock {\em Journal of the Royal Statistical Society: Series B (Statistical
+  Methodology)\/}, {\bf 72}, 417--473.
+
+\bibitem[Nicod {\em et~al.}, 2016]{nicod2016genome}
+Nicod,J.  {\em et~al.} (2016)
+\newblock Genome-wide association of multiple complex traits in outbred mice by
+  ultra-low-coverage sequencing.
+\newblock {\em Nature Genetics\/},  {\bf 48}, 912--918.
+
+\bibitem[Pahl and Schafer, 2010]{pahl2010permory}
+Pahl,R. and Schafer,H. (2010)
+\newblock {PERMORY}: an {LD}-exploiting permutation test algorithm for powerful
+  genome-wide association testing.
+\newblock {\em Bioinformatics\/}, {\bf 26}, 2093--2100.
+
+\bibitem[Price {\em et~al.}, 2010]{price2010new}
+Price,A.L. \textit{et al.} (2010)
+\newblock New approaches to population stratification in genome-wide
+  association studies.
+\newblock {\em Nature Reviews Genetics\/}, {\bf 11}, 459--463.
+
+\bibitem[Rakitsch {\em et~al.}, 2013]{rakitsch2013lasso}
+Rakitsch,B. \textit{et al.} (2013)
+\newblock A {L}asso multi-marker mixed model for association mapping with
+  population structure correction.
+\newblock {\em Bioinformatics\/}, {\bf 29}, 206--214.
+
+\bibitem[Segura {\em et~al.}, 2012]{segura2012efficient}
+Segura,V. \textit{et al.} (2012)
+\newblock An efficient multi-locus mixed-model approach for genome-wide
+  association studies in structured populations.
+\newblock {\em Nature Genetics\/}, {\bf 44}, 825--830.
+
+\bibitem[Shen {\em et~al.}, 2013]{shen2013novel}
+Shen, X. \textit{et al.} (2013) 
+\newblock A novel generalized ridge regression method for quantitative
+  genetics.
+\newblock {\em Genetics\/}, {\bf 193}, 1255--1268.
+
+\bibitem[Spielman and Ewens, 1996]{spielman1996tdt}
+Spielman,R.S. and Ewens,W.J. (1996)
+\newblock The {TDT} and other family-based tests for linkage disequilibrium and
+  association.
+\newblock {\em American Journal of Human Genetics\/}, {\bf 59}, 983--989.
+
+\bibitem[Storey and Tibshirani, 2003]{storey2003statistical}
+Storey,J.D. and Tibshirani,R. (2003) 
+\newblock Statistical significance for genomewide studies.
+\newblock {\em Proceedings of the National Academy of Sciences\/}, {\bf
+  100}, 9440--9445.
+
+\bibitem[Szymczak {\em et~al.}, 2016]{szymczak2016r2vim}
+Szymczak,S. \textit{et al.} (2016)
+\newblock {r2VIM}: A new variable selection method for random forests in
+  genome-wide association studies.
+\newblock {\em BioData Mining\/}, {\bf 9}, 7.
+
+\bibitem[Verbyla {\em et al.}, 2007]{verbyla2007analysis}
+Verbyla,A.P. \textit{et al.} (2007)
+\newblock The analysis of {QTL} by simultaneous use of the full linkage map.
+\newblock {\em Theoretical and Applied Genetics\/}, {\bf 116}, 95--111.
+
+\bibitem[Verbyla {\em et al.}, 2012]{verbyla2012rwgaim}
+Verbyla,A.P. \textit{et al.} (2012)
+\newblock {RWGAIM}: an efficient high-dimensional random whole genome average
+  ({QTL}) interval mapping approach.
+\newblock {\em Genetics Research\/}, {\bf 94}, 291--306.
+
+\bibitem[Verbyla {\em et al.}, 2014]{verbyla2014whole}
+Verbyla,A.P. \textit{et al.} (2014)
+\newblock Whole-genome analysis of multienvironment or multitrait {QTL} in
+  {MAGIC}.
+\newblock {\em G3: Genes, Genomes, Genetics\/}, {\bf 4}, 1569--1584.
+
+\bibitem[Yi {\em et~al.}, 2005]{yi2005bayesian}
+Yi, N. \textit{et al.} (2005)
+\newblock Bayesian model selection for genome-wide epistatic quantitative trait
+  loci analysis.
+\newblock {\em Genetics\/}, {\bf 170}, 1333--1344.
+
+\bibitem[Yu {\em et~al.}, 2006]{yu2006unified}
+Yu,J.  {\em et al.} (2006) 
+\newblock A unified mixed-model method for association mapping that accounts
+  for multiple levels of relatedness.
+\newblock {\em Nature Genetics\/}, {\bf 38}, 203--208.
+
+\bibitem[Zhao {\em et al.}, 2007]{zhao2007arabidopsis}
+Zhao, K.  {\em et al.} (2007)
+\newblock An {A}rabidopsis example of association mapping in structured
+  samples.
+\newblock {\em PLoS Genetics\/}, {\bf 3}, e4.
+
+\bibitem[Zheng {\em et~al.}, 2011]{zheng2011comparison}
+Zheng,J. \textit{et al.} (2011)
+\newblock A comparison of approaches to account for uncertainty in analysis of
+  imputed genotypes.
+\newblock {\em Genetic Epidemiology\/}, {\bf 35}, 102--110.
+
+\bibitem[Zhou and Stephens, 2012]{zhou2012genome}
+Zhou,X. and Stephens,M. (2012)
+\newblock Genome-wide efficient mixed-model analysis for association studies.
+\newblock {\em Nature Genetics\/}, {\bf 44}, 821--824.
+
+\end{thebibliography}
 

### 2019-08-19 BIOINF-2019-0261-RevisionV2/supmaterials.tex
still tring to get bioinformatics in the right latex format

--- 2019-08-12 BioinformaticsRevison/supmaterials.tex
@@
 \begin{figure}
-\caption{Power verse false discovery rates for Eagle and the multi-locus methods. Plots for only those simulation 
+\caption{Power verse false positive rates for Eagle and the multi-locus methods. Plots for only those simulation 
 scenarios where all multi-locus methods could be implemented are shown.  Eagle has the highest power across the four 
-scenarios but MLMM also performs well. A false discovery rate of greater than 10\% is typically not employed in GWASs so the 
+scenarios but MLMM also performs well. A false posirive rate of greater than 10\% is typically not employed in GWASs so the 
 upper limit of the x-axis is 0.1.}
@@
                         & Bioch.HDL                  & 4      &  134.58  & FACS.CD3posCD44posCD4CD8Ratio   \\
-                        &                                    & 6      &    17.53   & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
+                        & Bioch.HDL                                   & 6      &    17.53   & Muscles.Sol.g Muscles.TA.g Muscles.Plant.g Muscles.EDL.g  Muscles.Gast.g 
      SPPI.ln.pa Bioch.Calcium    Bioch.Tot.Cholesterol Bioch.Tot.Protein Bioch.Albumin  \\
@@
 Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
-(pink horizontal line) and permutation (blue horizontal line), at a 5\% significance level. 
+(pink horizontal line) and permutation (purple horizontal line), at a 5\% significance level. 
 }

### 2019-08-20 BIOINF-2019-0261-RevisionV2/bioinformaticspaper.tex
Morning submit
Final Revision submission - 20Aug2019

@@
 
-%%\usepackage{xr}
-%%\renewcommand{\baselinestretch}{1.4} 
-%%\usepackage{graphicx}
-%%\usepackage{bm}
-%%\usepackage{lineno}
-%%\usepackage{natbib}
-%%\usepackage{setspace}
-%%\usepackage{authblk}
-%%\usepackage{xcolor}
+%% New commands ... 
+\newcommand{\bwidetildea}{\widetilde{\bm{a}}}
+\newcommand{\bu}{\bm{u}}
+\newcommand{\be}{\bm{e}}
+\newcommand{\btau}{\bm{\tau}}
+\newcommand{\ba}{\bm{a}}
+\newcommand{\bzero}{\bm{0}}
+\newcommand{\bI}{\bm{I}}
+\newcommand{\bX}{\bm{X}}
+\newcommand{\by}{\bm{y}}
+\newcommand{\bZ}{\bm{Z}}
+\newcommand{\bM}{\bm{M}}
+\newcommand{\bmm}{\bm{m}}
+
+\usepackage{lmodern}% http://ctan.org/pkg/lm
+\usepackage{bm}
+\usepackage{lineno}
 
@@
 
-%%\linenumbers
-
-%% New commands ... 
-\newcommand{\bwidetildea}{\widetilde{\bm{a}}}
-\newcommand{\bu}{\bm{u}}
-\newcommand{\be}{\bm{e}}
-\newcommand{\btau}{\bm{\tau}}
-\newcommand{\ba}{\bm{a}}
-\newcommand{\bzero}{\bm{0}}
-\newcommand{\bI}{\bm{I}}
-\newcommand{\bX}{\bm{X}}
-\newcommand{\by}{\bm{y}}
-\newcommand{\bZ}{\bm{Z}}
-\newcommand{\bM}{\bm{M}}
-\newcommand{\bmm}{\bm{m}}
+\linenumbers
 
@@
 Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
-\textbf{Contact:} \href{andrew.george@csiro.au}\\
+\textbf{Contact:} andrew.george@csiro.au\\
 \textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
@@
 \section{Introduction}
-
-\citep{Boffelli03}
 
@@
 
-\subsubsection {Reducing the dimension of the model:}
+\subsubsection {Reducing the dimension of the model}
 
@@
 
- \subsubsection{ Multi-locus methods:} We compare the computational and statistical performance of Eagle against five multi-locus methods.
+ \subsubsection{ Multi-locus methods}
+
+ We compare the computational and statistical performance of Eagle against five multi-locus methods.
 \textcolor{red}{They are bigRR  \citep{shen2013novel}, LMM-Lasso \citep{rakitsch2013lasso}, glmnet \citep{Friedman2010glmnet},
@@
 BigRR, LMM-Lasso, and glmnet are regression-based regularisation 
-methods. BigRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
+methods. BigRR is based on generalised ridge regression, \hbox{LMM-Lasso} is based on lasso, and glmnet is based on elastic net. 
 Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
@@
 
-\subsubsection{Single-locus methods:} We also compare the performance of Eagle against two single-locus methods, GEMMA \citep{zhou2012genome} and FaST-LMM \citep{lippert2011fast}.
+\subsubsection{Single-locus methods} 
+We also compare the performance of Eagle against two single-locus methods, GEMMA \citep{zhou2012genome} and FaST-LMM \citep{lippert2011fast}.
 Both are based on linear mixed models.
@@
 \subsection{Simulation Study}
+
 A large simulation study was performed where we sought to  answer two questions. 
@@
 
-\begin{figure}[!tpb]
+\begin{figure}
  \centerline{\includegraphics[width=0.45\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}}
@@
 
+\begin{figure}[!tpb]
+%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
+\centerline{\includegraphics[width=8cm, height=8cm]{mouseresults.eps}}
+\caption{
+Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
+multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
+(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$), at the 5\% significance level.
+Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate of 5\% (Eagle$^{optimal}$).
+The number of SNP-trait associations found are reported in the cells.
+}
+\label{figmouse}
+
+\end{figure}
+
 The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}, 
@@
 than a 50\% increase in findings.
-
-\begin{figure}[!tpb]
-%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
-\centerline{\includegraphics[width=8cm, height=8cm]{mouseresults.eps}}
-\caption{
-Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
-multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
-(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$), at the 5\% significance level.
-Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate of 5\% (Eagle$^{optimal}$).
-The number of SNP-trait associations found are reported in the cells.
-}
-\label{figmouse}
-
-\end{figure}
 
@@
 Upon submitting our paper for review, a more recent multi-locus association mapping method, FarmCPU \citep{liu2016iterative}, was brought to our attention. 
-It is an unorthoﬂdox approach. Instead of working with a single model, results are passed back and forth between two models, 
+It is a statistically unorthodox approach. Instead of working with a single model, results are passed back and forth between two models, 
 a fixed effects model and a random effects model. Measures of association are obtained from the fixed effect model, which in turn 
@@
 \begin{thebibliography}{}
-
-\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
-Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
 
